### 2015-06-09 main.tex
Initial ShareLaTeX Import

%                 Doubly-Stopped Binomial Distribution

%                        Kane and Zelterman

\typeout{}\typeout{}\typeout{}\typeout{}
\typeout{ Two-Stage, Phase II Designs }
\typeout{}\typeout{}
\typeout{ Michael Kane \& Dan Zelterman }
\typeout{}\typeout{}\typeout{}\typeout{}

\documentclass[12pt]{article}         %LaTeX preamble
\parskip   .4em
\jot  2ex
\parindent .3in
\textwidth      6.5in
\oddsidemargin    0in
\textheight 8.75in
\topmargin      -.5in
\baselineskip    26pt
\renewcommand{\baselinestretch}{1.5}
\renewcommand{\arraystretch}{.75}     % vertical spacing in tables
\usepackage{graphicx}                 % Needed to import pdf files
\usepackage{adjustbox}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{url}
\begin{document}

%\font\fiverm=cmr5                     % Font definition for PicTex
%\input{prepic.tex}                    % Input three PicTeX files
%\input{pictex.tex}
%\input{postpic.tex}

\newtheorem{prop}{Proposition}

% --------------- New command definitions ----------------------

%  These commands allow exercises listed as \item[]'s to be listed in
%  the .aux file and referenced in the text.

\newcounter{bibno}%            Bibliographic counter

\def\waux{1}%                  Write number for the .AUX file

\newcommand\writelabel[2]%     Write label[#1] and value[#2] to .aux file
   {%                          General command, called by \Aexer and \Texer
     \immediate\write\waux%         Write this information to the .aux file:
          {\noexpand\newlabel{#1}%  Arguments #1, #2, and the page number
             {{#2}{\thepage}}}%     Warning: the page number will be incorrect
   }%                               because I use the \immediate command

\newcommand\bibref[1]%              Reference name and my label (#1)
   {%
    \addtocounter{bibno}{1}%        Increment the counter
    \def\biblab{\thebibno}%         name is same as counter with dot
    \writelabel{#1}{\biblab}%       Write the label to .AUX file 
    \noindent\biblab.%              Write this exercise name for \item[ ]
   }%

% -------------------  global definitions  ------------------------

\newcommand{\0}{\mbox{\boldmath $0$}}         %  bold zero in math mode
\newcommand{\bbeta}{\mbox{\boldmath $\beta$}} %  bold beta in math mode
\newcommand{\bd}{\mbox{\boldmath $d$}}        %  bold d in math mode
\newcommand{\bg}{\mbox{\boldmath $\gamma$}}   %  bold gamma in math mode
\newcommand{\blam}{\mbox{\boldmath $\lambda$}}%  bold lambda in math mode
\newcommand{\bpi}{\mbox{\boldmath $\pi$}}     %  bold pi in math mode
\newcommand{\bsig}{\mbox{\boldmath $\sigma$}} %  bold sigma in math mode
\newcommand{\bth}{\mbox{\boldmath$\theta$}}   %  bold theta in math mode
\newcommand{\dt}{\widetilde{d}}               %  d tilde in math mode
\newcommand{\m}{\mbox{\boldmath $m$}}         %  bold m in math mode
\newcommand{\M}{\mbox{\boldmath $M$}}         %  bold M in math mode
\newcommand{\mi}{{$-$}}                       %  large math minus sign
\newcommand{\n}{\mbox{\boldmath $n$}}         %  bold n in math mode
\newcommand{\ndot}{n{\rm\bf .}}               %  n sub-dot
\newcommand{\Nh}{\widehat{N}}                 %  N-hat
\newcommand{\nsp}{\hspace*{-.3in}}            %  large negative space
\newcommand{\p}{\phantom{5}}                  %  space for one digit
\newcommand{\pih}{\widehat{\pi}}              %  pi-hat
\newcommand{\pmi}{\phantom{$-$}}              %  space for wide minus
\newcommand{\ps}{\phantom{$*$}}               %  space for * footnote
\newcommand{\pc}{{\scriptsize$\bullet$}}      %  plotting dot character
\newcommand{\q}{\phantom{55}}                 %  space for two digits
\newcommand{\qs}{\hspace*{.1in}}              %  wide space
\newcommand{\re}{{\rm e}}                     %  Roman e (...to the power)
\newcommand{\bt}{\mbox{\boldmath $t$}}        %  bold t in math mode
\newcommand{\bT}{\mbox{\boldmath $T$}}        %  bold T in math mode
\newcommand{\U}{\mbox{\boldmath $u$}}         %  bold u in math mode
\newcommand{\X}{\mbox{\boldmath $X$}}         %  bold X in math mode
\newcommand{\x}{\mbox{\boldmath $x$}}         %  bold x in math mode
\newcommand{\Y}{\mbox{\boldmath $Y$}}         %  bold Y in math mode
\newcommand{\y}{\mbox{\boldmath $y$}}         %  bold y in math mode
\newcommand{\z}{\mbox{\boldmath $z$}}         %  bold z in math mode

% ------------------- my hyphenation rules ---------------------

\hyphenation{general-iza-tions}
\hyphenation{distri-bution}
\hyphenation{Zel-ter-man}

% ---------------  cover page  --------------------------------

\thispagestyle{empty}
\begin{center}
\vspace*{\fill}

{\Large\bf A Stopped Negative Binomial Distribution} \\ [1ex]

\vspace*{1.25in}
\begin{tabular}{l} 
 {\bf Michael Kane${}^*$} \\
 {\bf Daniel Zelterman}  \\[.1in]

 Department of Biostatistics \\
 School of Epidemiology  \\
 \hspace*{.15in} and Public Health  \\
 Yale University       \\
 New Haven, CT  06520  \\[.2in]

\today
\end{tabular}
\end{center}

\vspace*{\fill}

\noindent${}^*$Email address for author: {\tt
michael.kane@yale.edu}.  This research was supported by grants 
	R01CA131301, R01CA157749, R01CA148996, R01CA168733 awarded by the National Cancer Institute, and support from the Yale Comprehensive Cancer Center.

% ------------------------ Abstract ----------------------------

\newpage
\thispagestyle{empty}

\section*    {\bf   Abstract}

This paper introduces a new discrete distribution suggested by curtailed sampling rules common in early-stage clinical trials. We derive the distribution of the smallest number of independent Bernoulli($p$) trials needed in order to observe either $s$ successes or $t$ failures. The closed form expression for the distribution as well as the compound distribution are derived. Properties of the distribution are shown and discussed. A case study is presented showing how the distribution can be used to monitor sequential enrollment of clinical trials with binary outcomes as well as providing post-hoc analysis of completed trials.

\bigskip

\noindent{\bf Keywords:}  discrete distribution; curtailed sampling

% ---------------------------------------------------------------

\thispagestyle{empty}
\setcounter{page}{1}

% --------------------  1. Introduction   ----------------------
\section            {Introduction and Motivation}
% --------------------------------------------------------------

In the classic Simon 2-stage Optimal Design, Phase II clinical trial [\ref{Simon 1989}] a small number of patients are enrolled onto a therapeutic protocol for an initial stage and assess either success according to a clinically useful criteria or else as treatment failure. Failure might include adverse reaction to the drug or withdrawal due to toxicity. We use a large number of these early successes as evidence for whether to additional patients for a second stage. A statistically small number of successes lead us to terminate the trial.

Consider the Phase 2 clinical trial that investigates the efficacy of iniparib (BSI-201/SAR240550) in patients with breast cancer gene-associated (BRCA) ovarian cancer [\ref{Sanofi 2013}]. In the first stage 12 patients were treated with iniparib (BSI-201/SAR240550). If at least two of those patients respond, as defined by Response Evaluation Criteria in Solid Tumor (RECIST), then the trial proceeds to the next stage where additional patients were treated.

If, in the first stage, two or more are successful then we are lead to believe that the therapy is useful and will proceed to the second stage. Otherwise, the trial will terminate. Our goal is to know the distribution of the sample size of the design for planning purposes.
If all 12 patients are enrolled at once, as in the classic design, then the sample size is 12. However, in most clinical trials, the patients are enrolled sequentially, often with one patient's outcome realized before the next one enters the trial. In the present example, observing two successful patients allows us to proceed to the next stage. Similarly 11 observed treatment failures also ends the stage. This sampling mechanism, in which the experiment ends as soon as any of the endpoints is reached, is call {\em curtailed sampling}. Under curtailed sampling the range of the sample size is between two and 11.

We will assume each of the patient enrollments can be modeled as independent, identically distributed Bernoulli($p$) random variables. The realization of the trial is a sequence of these random variables that stops when either a specified number of success or failures is reached. As an example, suppose in the previously described trial, two successes were reached after enrolling 10 patients (one in the third step and one at the $10^{th}$). The trial is visualized in Fig \ref{fig:kane_viz}. The vertical axis denotes the number of successful outcomes. The horizontal axis denotes the number of patients that have been enrolled. The horizontal and vertical line segments represent endpoints for the trial.

\begin{figure}[ht]
\includegraphics[width=\textwidth]{KanePlot.pdf}
\caption{
A hypothetical realization of the breast cancer trial.
}
\label{fig:kane_viz}
\end{figure}

%Our goals for this paper are three-fold. First, we derive the distribution for the number patients that need to be enrolled in the described sequential trial, which we refer to as the {\em Stopped Negative Binomial} (SNB) distribution. Using a sequential trial it is often possible to reduce the number of patients needed when compared to traditional design. Understanding of this distribution allows to create better, more {\em enrollment efficient} designs. Second, assuming a Beta distribution on the probability of success, we will derive the compound distribution of the SNB. Both the success probability and the SNB compound distribution can be updated as the individual outcome of patients are known and, as a result, better estimates of the distribution of patient-enrollment size can be derived as new data are received. Third, we will use the derived results to propose an enrollment design framework alternative to the Simon Two-stage Optimal trial design. We will show a hypothetical but representative case where the number of enrollments is less than those proposed by the Simon Two-stage and the uncertainty associated with both the individual success outcome as well as the trial is fully characterized.

The next section of this paper introduces our notation and basic results including the density of the SNB and compound SNB along with a description of it's relation to other distributions. Sections 2 and 3 derive the distribution based on a defined Bernoulli process and give some basic properties of the distribution. Section 4 derives the compound distribution using a Beta prior. Section 5 develops a post-hoc analysis of a completed trial. Section 6 is devoted to discussion and conclusions. The simulations, model fitting routines, and visualizations presented in this paper were generated using the snb package [\ref{Kane 2015}] for the R programming environment [\ref{R Core Team 2015}]

%Their individual outcomes are summarized as independent Bernoulli, %with either success according to a clinically useful criteria or else %as treatment failure. Failure might include adverse reaction to the %drug or withdrawal due to toxicity. Successes occur with probability %$\,p\; (0\leq p\leq 1)$.

%More generally, suppose we have a sequence of independent Bernoulli$(p)\,$random variables $\,Z_1, Z_2, \ldots$. For positive integers $\,s\,$ and $\,t,$  we study the distribution of the smallest number $\,Y\,$ such that $\,Z_1, Z_2, \ldots, Z_Y\,$ %contains either $\,s\,$ successes of $\,t\,$ failures.

% ---------------------------  2. Notation ----------------------
\section   {Notation and summary of distributional results}
\label{notation.section}
% --------------------------------------------------------------

Let $\,B_1, B_2, \ldots \,$ denote a sequence of independent, identically distributed, Bernoulli random variables with 
$\mathbb{P}\{B_i=1\}=p$, for probability parameter $p \in \{0\leq p\leq 1\}$. Let $s$ and $t$ be positive integers. Define the Stopped Negative Binomial (SNB) random variable $Y$ as the smallest integer value such that $\,\{B_1, \ldots , B_Y\}\,$ contains either $\,s\,$ successes or else $\,t\,$ failures. The distribution of $\,Y\,$ has support on integer values in the range 
\begin{equation*}                                     %   (3)
     \min(s,t) \leq \; Y \;\leq s+t-1  \label{range.y.eq}
\end{equation*}
and it is distributed as
\begin{equation} \label{eqn:pmf}
\mathbb{P}\{Y=k\} = S(k, p, s)  \{s \leq k \leq s+t-1\} + T(k, p, t) \{t \leq k \leq s+t-1 \} 
\end{equation}
where
\begin{equation} \label{eqn:N}
S(k, p, s) = {k-1 \choose s-1} p^s (1-p)^{k-s} 
\end{equation}
is the negative binomial probability of $s$ successes and
\begin{equation} \label{eqn:R}
T(k, p, t) = {k-1 \choose k-t} p^{k-t} (1-p)^t
\end{equation}
is the negative binomial probability of $t$ failures.

% ------------   Section 3: Random Walk  --------------
%\section {Deriving the SNB with a Random Walk}
%\label{random.walk.section}
% --------------------------------------------------------------

To show that (\ref{eqn:pmf}) is the distribution of $Y$, consider the process $\mathbf{X} = \{X(k) : k \in \mathbb{Z}_{\ge 0} \}$ with $X(0)=0$ and
\begin{equation*} \label{eqn:proc}
X_{k+1} = X_k + B_{k+1} \{ s \leq X_k \leq k-t \}
\end{equation*}
where $B_{k+1}$ is once again distributed as Bernoulli($p$) for all $k$ in the domain of $\mathbf{X}$. The process can be conceptualized as a series of coin flips that stops when either $s$ heads or $t$ tails are reached. At each step a coin is flipped. If it is heads, the process advances one diagonally in the positive horizontal and vertical direction. Otherwise, it advances in the positive horizontal direction only. The process stops, $X_k$'s value does not change, after either $X_k = s$ or $X_k = k-t$. One example is given in Fig. \ref{fig:kane_viz}.

\begin{prop}
The distribution of the stopping time $\{k: s \leq X_k \leq k-t \}$ is equal to the SNB and is given in (\ref{eqn:pmf}). 
\end{prop}
\begin{proof}
The proof will proceed in two parts. First, a combinatorial justification will be given for the probability mass value on each element of the support. Second, it will be shown that the sum of the masses over the support sums to one.

The probability that a given realization of $\mathbf{X}$ reaches $s$ at time $k$ is  the probability that, at time $k-1$ there are $s-1$ successful coin flips and $k-s$ unsuccessful coin flips times the probability of a successful coin flip at time $k$ (\ref{eqn:N}). The probability $X_k = k-t$ is the probability that, at time $k-1$ there are $k-t$ successful coin flips and $t-1$ unsuccessful coin flips times the probability of an unsuccessful coin flip at time $k$ (\ref{eqn:R}).

Next, we need to show that the sum
\begin{equation} \label{eqn:sum_proof}
R = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s} + \sum_{k=t}^{s+t-1} {k-1 \choose k-t} p^{k-t} (1-p)^t
\end{equation}
is equal to one.
First, substitute $i=k-s$ in the first summation and
$j=k-t$ in the second. Then (\ref{eqn:sum_proof}) becomes:
\begin{equation*}
R = \sum_{i=0}^{t-1} {i+s-1 \choose s-1} p^s (1-p)^i +
\sum_{j=0}^{s-1} {j+t-1 \choose j} p^j (1-p)^t.
\end{equation*}
Write   
\begin{equation*}
{j+s-1 \choose s-1} = {j+s-1 \choose j}
\end{equation*}
and so $R$ can be written
as the cumulative distribution function of two 
negative binomial distributions
\begin{equation} \label{eqn:transformed_sum}
R = \sum_{i=0}^{t-1} {i+s-1 \choose i} p^s (1-p)^i +
\sum_{j=0}^{s-1} {j+t-1 \choose j} p^j (1-p)^t.
\end{equation}
Let $I_p(s, t)$ be the {\em regularized incomplete beta function}, which is also one minus the c.d.f. of the negative
binomial distribution, and $I_p(s, t) = I_{1-p}(t, s)$.
Then the result can be proven as:
\begin{align*}
R = \sum_{i=0}^{t-1} &{i+s-1 \choose i} p^s (1-p)^i +
\sum_{j=0}^{s-1}  {j+t-1 \choose j} p^j  (1-p)^t \\
   &= 1-I_p(s, t) + 1 - I_{1-p}(t, s) \\
   &= 1. 
\end{align*}
\end{proof}
From (\ref{eqn:pmf}) one can see that the mass at any point on the support of the distribution may come from either $X_k = s$, captured in the function $S$, or $X_k = k-t$, captured in the function $T$. The support of $S$ is defined by $s \leq S \leq s+t+1$ and the support of $T$ is defined by $t \leq T \leq s+t+1$.

\begin{figure}[p!]
\begin{center}
\includegraphics[scale=.9]{shapes.pdf}
\end{center}
\caption{Different shapes of the SNB distribution with parameters ($s$, $t$, $p$), as given. \label{shapes.fig}}
\end{figure}

The probability mass function of $\,Y\,$ has a variety of shapes for different choices of the parameters $(s,\, t,\,p)$.
These shapes are illustrated in Fig.~\ref{shapes.fig}.
The SNB is related to the negative binomial distribution. Specifically, if $t$ is infinitely large then the $Y-s$ has a negative binomial distribution with
\begin{equation*}                                    %   (1)
   \mathbb{P}\{Y=s+j \}        \label{nb1.eq}
          = {{s+j-1}\choose{s-1}} p^s (1-p)^j
\end{equation*}
for $\,j=0, 1,\ldots\,$.
Similarly, if $s$ is infinitely large then $\,Y - t\,$ has a negative binomial distribution with
\begin{equation*}                                    %   (2)
    \mathbb{P}\{Y=t+j \}        \label{nb2.eq} 
               = {{t+j-1}\choose{t-1}} p^j(1-p)^t.
\end{equation*}

For the special case of $\,s=t,$ the distribution of $\,Y\,$ is the riff-shuffle, or minimum negative binomial distribution~[\ref{Uppuluri 1970}].
Similar derivations of the closely-related maximum negative binomial discrete distributions also appear in~[\ref{Zhang 2000}, \ref{Zelterman 2004}].

%\subsection{Updating the SNB distribution}

%Under the presented random walk construction if the process is observed before reaching one of its endpoints the conditional distribution is a reparametrized SNB distribution whose parameters are determined by $X_k$ and $k$, the process value and time step. For example, in the breast cancer trial the distribution of the number of patients is parameterized as $SNB(2, 11, p)$. At the third enrollment there is one successful outcome. Conditioned on this information, the number of additional patient enrollments is $SNB(1, 9, p)$. As new outcomes are collected the conditional distribution can be re-derived to get better estimates of the total patient enrollment. Furthermore, if our task is to estimate $p$ along with the patient enrollment distribution, this parameter can also be updated as new data are received.

\section{The Compound Distribution}

%The previous section derives a new distribution for a sequential Bernoulli process where the process is stopped when either a total number of successes is reached or a total number of failure is reached. The process can be alternatively conceptualize as a series of coin flips that stops when either $s$ heads or $t$ tails are reached. At each step a coin is flipped. If it is heads then one step is taken in the positive direction on the vertical axis. Otherwise, a step is taken in the positive direction on the horizontal axis. The walk stops when either $s$ is reached on the vertical axis or $t$ is reached on the horizontal axis. A visual example is shown in Fig. \ref{fig:zelterman_viz}.

%\begin{figure}[ht]
%\includegraphics[width=\textwidth]{ZeltermanPlot.pdf}
%\caption{
%A realization of $\mathbb{X}$ with $p=0.3$, $s=3$, and $t=10$. An arrow to the right indicates a coin flip of ``tail'' and
%an arrow in the vertical direction indicates ``head.''
%}
%\label{fig:zelterman_viz}
%\end{figure}

If we assume that $p$ has a Beta distribution, with constant prior parameters $\alpha$ and $\beta$ then we can calculate the closed form compound distribution of the SNB as
\begin{align}
\mathbb{P} \{Y = k | \mathbf{B}, \alpha, \beta \} &= {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} \{s \leq k \leq s+k-1\} + \nonumber \\
& {k-1 \choose k-t} \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} \{t \leq k \leq s+k-1\}
\end{align}
where $B$ is the Beta function and $\mathbf{B}$ is the sequence of Bernoulli data.

When $p$ is known, the probability mass function (PMF) of the number of steps until the process is stopped is given by (\ref{eqn:pmf}). However, many times this is not the case. In Phase II clinical trials for example patients are often enrolled sequentially until there are enough patients to guarantee the desired level of power or until a specified number of adverse events are recorded. In cases such as these the endpoints $s$ and $t$ are known and we would like to be able to estimate $p$ either during enrollment to provide odds of successful enrollment or after an endpoint has been reached to determine the amount of uncertainty in the endpoint. To do this, we will assume that $p$'s probability in the absence of available data is $\text{Beta}(1/2, 1/2)$, the Jeffrey's prior. This distribution was chosen for two reasons. First, it is the conjugate prior of the (Negative) Binomial distribution and updating the distribution as data are received is straightforward. Second, it is objective and limits the number of ``prior coin-flips'' to one. This means that data are more heavily influence the distributional characteristics of $p$ with more data providing more proportional influence.

\begin{prop}
The compound PMF of the Stopped Negative Binomial distribution with a Beta($\alpha$, $\beta$) prior is:
\begin{align}
f(k | s, t, \alpha, \beta) &= {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} \{s \leq k \leq s+k-1\} + \nonumber \\
& {k-1 \choose k-t} \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} \{t \leq k \leq s+k-1\}
\end{align}
where $B$ is the Beta function.
\end{prop}
\begin{proof}
For notational simplicity, assume that $s,t \leq k \leq s+k-1$. When this is not the case appropriate terms should be removed as dictated by the indicator functions.
\begin{align*}
f(k | s, t, \alpha, \beta) = \frac{1}{B(\alpha, \beta)} & \int_0^1 {k-1 \choose s-1} p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} + \\
 & {k-1 \choose k-t} p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp \\
= \frac{1}{B(\alpha, \beta)}  {k-1 \choose s-1} & \int_0^1  p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} + \\
 & \frac{1}{B(\alpha, \beta)} {k-1 \choose k-t} \int_0^1  p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp
\end{align*}
The result follows by applying the definition of the Beta function to the integral terms.
\end{proof}

\section{A Bayesian Exploration Simon Two-stage Optimal Design}

Returning to the breast cancer trial introduced in the first section, a response rate of 2/12 in the first stage corresponds to an individual probability of responding of $p=0.1667$. The standard deviation of this probability is $\sqrt{p(1-p)/12}=0.1076$. If we were to use the Central Limit Theorem then the 95\% confidence interval around the mean would be $[-0.0442, 0.1076]$. While this approach is problematic since the number of samples (12 in this case) is small it does indicate how much uncertainty is built into the first stage. If we consider samples from both stages then the standard deviation drops to 0.0630 and the 95\% confidence interval would be $[0.0432, 0.2901]$.

These calculations expose two difficulties with the Simon Two-stage Optimal Design. First, success in the first stage is not a good indicator of success in the second stage and at the same time failure in first stage in no way implies failure in the second stage, especially when $p$ is close to it's threshold value. The variance in the estimate so large that it is a poor estimator. The second, difficulty is that the design is not {\em enrollment efficient}. That is, if the first stage is successful then additional patients are enrolled. Enrollment, especially for certain types of cancer, can be very difficult and it can taken time to find eligible patients. Furthermore, if all 12 of the patients in the first stage respond positively, should we really need to enroll additional patients?

Properly quantifying uncertainty in the process implies quantifying uncertainty in $p$, which stems from two sources. First is the number of samples we receive. Any estimate of $p$ will be based on data and generally the more data we have, the better our estimate. If the data are small, as they often are in clinical trials, the less reliable our estimate is. The second source comes from heterogeneity in the data. While models may assume that the probability of success $p$ is a single value this is often not the case in real-world data. In clinical trials for example an individuals chance of having an adverse outcome may be influenced by any number of underlying factors, which may vary across individual. In these cases $p$ may not be a single numeric value and may be better conceptualized as a realization over an entire distribution. A frequentest approach will not capture this variation.

We would like to be able to estimate $p$ and capture both sources of uncertainty. For cases where the process has not yet reached an endpoint, we would like to be able to provide updated estimates of $p$, its uncertainty, and the probabilities of hitting each endpoint. If the process has reached an endpoint, we would again like a estimate of $p$ and its uncertainty along with the probability that another endpoint could have been reached, given the data.

\subsection{Quantifying Uncertainty in the Simon Two-stage Optimal Design}

Both of the problems in the Simon Two-stage Optimal Design stem from the fact that it fails to quantify the uncertainty in the estimate of the success probability. A framework based on Stopped Negative Binomial distribution addresses this failure and provides the added features of monitoring the trial as new results are received, providing estimates of how many new enrollees are needed to complete a trial, and quantifying the uncertainty in the success probability.

Let's start by recasting the breast cancer study in the context of the Stopped Negative Binomial framework, assuming sequential enrollment. The first stage will stop when either two positive responses are reached or 11 negative responses. This corresponds to a Stopped Negative Binomial Distribution with $s=2$ and $t=11$. Next, lets assume that, in the first stage, there were two successes. A second success could occur during the second, through a tenth enrollment. At each of these times we would like to know the estimate for $p$ along with the lower five percent quantile of the distribution.

\begin{figure}[ht]
\includegraphics[width=\textwidth]{uncertainty.pdf}
\caption{
A visualization of the mode and the 0.05\% quantile of the estimate of the distribution of $p$.
}
\label{fig:simon}
\end{figure}

To better understand how the distributional estimate of $p$ changes in the number of steps needed for two successes data were simulated for each potential stopping point from two to 11. This procedure essentially estimates the shape parameters of a beta distribution. Fig. \ref{fig:simon} shows the change in the mode of the distribution, which is also the Maximum Likelihood Estimator (MLE) as well as the five percent quantile of the distribution. The graph shows, for example, that if a second success is found with the third enrollment then the MLE estimate of $p$ is 0.5 and with 95\% confidence $p$ is at least 0.25. In this case investigators could have moved on to the next stage of the trial with only three enrollments as opposed to 12.

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|} \hline
{\bf Steps to 2 successes} & {\bf Mode of $p$} & {\bf 5\% quantile} \\ \hline
2 & 0.7500  & 0.4307 \\ \hline
3 & 0.5000  & 0.2355 \\ \hline
4 & 0.3000  & 0.1278 \\ \hline
5 & 0.1875  & 0.0763 \\ \hline
6 & 0.1250  & 0.0497 \\ \hline
7 & 0.0882  & 0.0347 \\ \hline
8 & 0.0652  & 0.0254 \\ \hline
9 & 0.0500  & 0.0194 \\ \hline
10 & 0.0398 & 0.0153 \\ \hline
11 & 0.0319 & 0.0123 \\ \hline
\end{tabular} 
\end{center}
\caption{
A table of the mode and the 0.05\% quantile of the estimate of the distribution of $p$.
}
\label{tab:simon}
\end{table}

Table \ref{tab:simon} shows the actual values used to generate Fig. \ref{fig:simon}. In the beginning of this section it was shown that, at the end of stage two, the breast cancer study actually tests whether the probability of success is greater than 0.0432 with 95\% confidence and requires 35 patients to do so. With the Stopped Negative Binomial model we only need six patients for the same level of confidence. This represents a tremendous amount of savings in time, effort and resources.

\subsection{Trial Monitoring}

Like most Bayesian Adaptive Clinical trials the SNB approach allows for monitoring of trial and can be updated as new results are received without losing power. Essentially this is because the object under investigation is not a point estimate of the success probability. It is the uncertainty in the distribution of the success probability. Point estimates of the success probability can be regarded as artifacts of our parameterization of the underlying distribution. Every point estimate calculated is accompanied by the quantification of the uncertainty around it. This estimate and it's uncertainty can be calculated at any time and may be useful in evaluating ongoing trials as well as providing a post-hoc evaluation of trials that have been completed.

\begin{figure}[ht]
\includegraphics[width=\textwidth]{hypo_traj.pdf}
\caption{
A visualization of the mode and the 0.05\% quantile of the estimate of the distribution of $p$ for a hypothetical sequence of patient enrollments.
}
\label{fig:hypo}
\end{figure}

Consider the trial described above. Suppose eight patients were enrolled and there was a single success on the third trial (visualized in Fig. \ref{fig:hypo}). Since the mode of the beta distribution can only be calculated when there is at least one success the maximum of the mean and the mode of the distribution is shown as the estimate. At the third time step there is an increase in the estimate and the 0.05 \% quantile corresponding to the success.

\begin{figure}[ht]
\includegraphics[width=\textwidth]{conditional_snb.pdf}
\caption{
The conditional Stopped Negative Binomial distribution after eight patients with one success.
}
\label{fig:conditional_snb}
\end{figure}

Fig. \ref{fig:conditional_snb} shows the conditional, compound SNB distribution. In each of the following enrollment a success would end the trial. If none of the following enrollments are success then the trial will stop at the eleventh patient. From Fig. \ref{fig:hypo} we know that the estimate of $p$ is about 0.19. The conditional compound distribution gives the probability of another success as 0.3883 and the probability of no more success (and the first stage failing) as 0.6117. Furthermore, we can see that the probability that the trial ends in the next step is approximately 20\% and the probability the trials end with the $11^{th}$ patient is about 65\%.

\subsection{Post-hoc Evaluation}

Suppose that in our example trial that a second successful outcome is reached on the $10^th$ patient. From the data collected we have shown how to create an estimate for the distribution of $p$. Based on that estimate we may like to see the probability of a different outcome. That is, what is the probability that we don't get two successful outcomes based on the data we have collected and what is the posterior distribution for the number of patients enrolled for the study?

In this case the success probability $p$ is distributed as $Beta(2.5, 8.8)$. Using the mode a point estimate yields $\hat{p} = 0.1667$. The standard deviation of the estimate is 0.1210. The probability of two successes during the trial is 0.6843 and the probability of failure is 0.3157. The distribution of the stopping times, along with contributions from two positive events and less than two positive events is shown below.

\begin{figure}[ht]
\includegraphics[width=\textwidth]{post_hoc.pdf}
\caption{
The post-hoc Stopped Negative Binomial distribution after the trial ends with 10 patient enrollments and two successes.
}
\label{fig:post_hoc}
\end{figure}

\section{Discussion and Conclusion}

We have presented a new discrete distribution by curtailed sampling rules common in early-stage clinical trials, which we refer to as the Stopped Negative Binomial distribution. The distribution models the stopping time of a sequential trial where the trial is stopped when a number of events are accumulated. The compound distribution was derived for the case when the event probability $p$ has a Beta distribution. Using a trial description from \url{clinicaltrials.gov} we showed how the SNB is an integral part of trial monitoring, and post-hoc analysis and can be used in a framework alternative to the Simon two-stage optimal design while providing better estimates of the event probability along with quantifying the uncertainty associated with the estimates. As a result, we were able to show fewer patient enrollments and achieve the same estimates, with added  in a hypothetical, but representative, clinical trial.

Current work focuses on the generalization of the distribution and the application in other areas of clinical trials. Adverse outcomes in particular are another area that could greatly benefit both from the presented distribution and its framework. For these trials monitoring may need to be performed not only on the outcome of an intervention but also on the safety of the trial. This especially in areas like late-stage cancer treatments where the treatment is harsh and patients may be forced to drop out as a result of the side-effects of the treatment; a matter independent of the outcome. Other application areas include providing design that allows clinicians to balance uncertainty and success probability with a minimum number of patient enrollees.

%   ------------------------    References     -----------------------

% -------------------------  References ---------------------------
\section*     {\bf References}
% -----------------------------------------------------------------

\begin{enumerate}

\item[\bibref{Kane 2015}]
Kane MJ. (2015). snb: The Stopped Negative Binomial. R package version 0.1. 2015. Available from \url{https://github.com/kaneplusplus/snb}

\item[\bibref{R Core Team 2015}]
R Core Team (2015). R: A language and environment for statistical
computing. R Foundation for Statistical Computing, Vienna, Austria. 2015. URL \url{http://www.R-project.org/}.

\item[\bibref{Sanofi 2013}]
Sanofi Pharmaceutical Company. A Phase 2, Single Arm Study of BSI-201 in Patients With BRCA-1 or BRCA-2 Associated Advanced Epithelial Ovarian, Fallopian Tube, or Primary Peritoneal Cancer. In: ClinicalTrials.gov [Internet]. Bethesda (MD): National Library of Medicine (US). 2000- [cited 2015 March 27]. Available from: \url{https://clinicaltrials.gov/ct2/show/study/NCT00677079} Identifier: NCT00677079.

\item[\bibref{Simon 1989}]
Simon R.  Optimal two-stage designs for Phase II clinical trials. {\it Controlled Clinical Trials\/} 1989; {\bf 10}: 1--10.

\item[\bibref{Uppuluri 1970}] 
Uppuluri VRR, Blot WJ. ``A probability distribution arising in a riff-shuffle.'' {\it Random Counts  in Scientific Work, 1: Random Counts in Models and Structures}, 1970. G.P. Patil (editor), University Park: Pennsylvania State University Press, pp  23--46.

\item[\bibref{Zhang 2000}]
Zhang Z, Burtness BA, Zelterman D.  The maximum negative binomial distribution. {\em Journal of Statistical Planning and Inference\/}  2000; {\bf 87}: 1--19.

\item[\bibref{Zelterman 2004}]
D Zelterman. {\it Discrete Distributions: Application in 
the Health Sciences}, New York: J. Wiley. 2004. xix + 277pp

\end{enumerate}

% ------------------------------------------------------------------

% ------------------------------------------------------------------
% --------------------   Tables and Figures   ----------------------
% ------------------------------------------------------------------

% -------------------------------------------------------------------

% -------------------------------------------------------------------
% ------------------------  FIGURES  --------------------------------
% -------------------------------------------------------------------

%\newpage 
% -------------------------------------------------------------------

%\newpage     %  Figure 1: functional forms of the mass function

% -------------------------------------------------------------------

\end{document}

% ------------------  end of this file --------------------------

### 2015-06-09 main.tex
Integrated Dan's newest comments.

@@
 %                 Doubly-Stopped Binomial Distribution
-
 %                        Kane and Zelterman
-
 \typeout{}\typeout{}\typeout{}\typeout{}
@@
 
-%\font\fiverm=cmr5                     % Font definition for PicTex
-%\input{prepic.tex}                    % Input three PicTeX files
-%\input{pictex.tex}
-%\input{postpic.tex}
-
 \newtheorem{prop}{Proposition}
@@
 % --------------- New command definitions ----------------------
-
-%  These commands allow exercises listed as \item[]'s to be listed in
-%  the .aux file and referenced in the text.
 
@@
 
-In the classic Simon 2-stage Optimal Design, Phase II clinical trial [\ref{Simon 1989}] a small number of patients are enrolled onto a therapeutic protocol for an initial stage and assess either success according to a clinically useful criteria or else as treatment failure.
-Failure might include adverse reaction to the drug or withdrawal due to toxicity.
-We use a large number of these early successes as evidence for whether to additional patients for a second stage.
-A statistically small number of successes lead us to terminate the trial.
+In the Simon 2-stage Phase II clinical trial [\ref{Simon 1989}], a small 
+number of patients are enrolled onto a therapeutic 
+protocol for an initial stage. Fore each patient, we assess either success 
+or failure according to a 
+clinically useful criteria. Success might include tumor response to treatment
+or perhaps that the response is maintained for a minimum specified duration.
+In addition to failing to achieve the primary endpoint, treatment failure 
+might include adverse reaction to the drug or withdrawal due to toxicity. We 
+use a large number of these early successes as evidence for whether to enroll
+additional patients for a second stage. A statistically small number of 
+successes in Stage I would lead us to terminate the trial without entering 
+Stage II.
 
-Consider the Phase 2 clinical trial that investigates the efficacy of iniparib (BSI-201/SAR240550) in patients with breast cancer gene-associated (BRCA) ovarian cancer [\ref{Sanofi 2013}].
-In the first stage 12 patients were treated with iniparib (BSI-201/SAR240550).
-If at least two of those patients respond, as defined by Response Evaluation Criteria in Solid Tumor (RECIST), then the trial proceeds to the next stage where additional patients were treated.
+Consider the Phase II clinical trial that investigates the efficacy of 
+iniparib in patients with breast cancer gene-associated (BRCA) ovarian 
+cancer [\ref{Sanofi 2013}]. In the first stage 12 patients were treated with 
+iniparib (BSI-201/SAR240550). The endpoint was defined by Response Evaluation 
+Criteria in Solid Tumor (RECIST). If at least two Stage I patients succeeded,
+then the trial would proceed to the next stage where additional patients were 
+treated. If fewer than two of the 12 Stage I patients is successful then 
+the trial would terminate.
 
-If, in the first stage, two or more are successful then we are lead to believe that the therapy is useful and will proceed to the second stage.
-Otherwise, the trial will terminate.
-Our goal is to know the distribution of the sample size of the design for planning purposes.
-If all 12 patients are enrolled at once, as in the classic design, then the sample size is 12.
-However, in most clinical trials, the patients are enrolled sequentially, often with one patient's outcome realized before the next one enters the trial.
-In the present example, observing two successful patients allows us to proceed to the next stage.
-Similarly 11 observed treatment failures also ends the stage.
-This sampling mechanism, in which the experiment ends as soon as any of the endpoints is reached, is call {\em curtailed sampling}.
-Under curtailed sampling the range of the sample size is between two and 11.
+Our goal is to know the distribution of the sample size of the design for 
+planning purposes.  If all 12 patients are enrolled at once, as in the classic 
+design, then the sample size is 12. However, in most clinical trials, the 
+patients are enrolled sequentially, often with one patient's outcome realized 
+before the next one enters the trial. In the present example, observing two 
+successful patients allows us to proceed to the next stage. Similarly 11 
+observed treatment failures also ends the stage. This sampling mechanism, in 
+which the experiment ends as soon as any of the endpoints is reached, is 
+call {\em curtailed sampling}. Under curtailed sampling the range of the 
+sample size is between two and 12.
 
@@
 
-Let $\,B_1, B_2, \ldots \,$ denote a sequence of independent, identically distributed, Bernoulli random variables with 
-$\mathbb{P}\{B_i=1\}=p$, for probability parameter $p \in \{0\leq p\leq 1\}$.
+Let $\,B_1, B_2, \ldots \,$ denote a sequence of independent, identically 
+distributed, Bernoulli random variables with $\mathbb{P}\{B_i=1\}=p$, for 
+probability parameter $0\leq p \leq 1$. In the clinical trial setting 
+$\,B_i = 0$ corresponds to a successful enrollment. 
 Let $s$ and $t$ be positive integers. 
-Define the Stopped Negative Binomial (SNB) random variable $Y$ as the smallest integer value such that $\,\{B_1, \ldots , B_Y\}\,$ contains either $\,s\,$ successes or else $\,t\,$ failures.
-The distribution of $\,Y\,$ has support on integer values in the range
+Define the stopped negative binomial (SNB) random variable $Y$ as the smallest 
+integer value such that $\,\{B_1, \ldots , B_Y\}\,$ contains {\em either} 
+$\,s\,$ successes or else $\,t\,$ failures. The distribution of $\,Y\,$ has 
+support on integer values in the range 
 \begin{equation*}                                     %   (3)
@@
 \begin{equation} \label{eqn:pmf}
-\mathbb{P}\{Y=k\} = S(k, p, s)  \{s \leq k \leq s+t-1\} + T(k, p, t) \{t \leq k \leq s+t-1 \} 
+\mathbb{P}\{Y=k\} = S(k, p, s)  \{s \leq k\} + T(k, p, t) \{t \leq k \} 
 \end{equation}
@@
 
-To show that (\ref{eqn:pmf}) is the distribution of $Y$, consider the process $\mathbf{X} = \{X(k) : k \in \mathbb{Z}_{\ge 0} \}$ with $X(0)=0$ and
+To illustrate that (\ref{eqn:pmf}) is the distribution of $Y$, consider the 
+process $\mathbf{X} = \{X(k) : k \in \mathbb{Z}_{\ge 0} \}$ with $X(0)=0$ and
 \begin{equation*} \label{eqn:proc}
-X_{k+1} = X_k + B_{k+1} \{ s \leq X_k \leq k-t \}
+X_{k+1} = X_k + B_{k+1} \{ k-t < X_k < s \}
 \end{equation*}
@@
 
-The probability that a given realization of $\mathbf{X}$ reaches $s$ at time $k$ is  the probability that, at time $k-1$ there are $s-1$ successful coin flips and $k-s$ unsuccessful coin flips times the probability of a successful coin flip at time $k$ (\ref{eqn:N}).
-The probability $X_k = k-t$ is the probability that, at time $k-1$ there are $k-t$ successful coin flips and $t-1$ unsuccessful coin flips times the probability of an unsuccessful coin flip at time $k$ (\ref{eqn:R}).
-
-Next, we need to show that the sum
+The probability that a given realization of $\mathbf{X}$ reaches $s$ at 
+time $k$ is  the probability that, at time $k-1$ there are $s-1$ successful 
+coin flips and $k-s$ unsuccessful coin flips times the probability of a 
+successful coin flip at time $k$ (\ref{eqn:N}). The probability $X_k = k-t$ 
+is the probability that, at time $k-1$ there are $k-t$ successful coin flips 
+and $t-1$ unsuccessful coin flips times the probability of an unsuccessful 
+coin flip at time $k$ (\ref{eqn:R}).  Next, we need to show that the sum
 \begin{equation} \label{eqn:sum_proof}
@@
 First, substitute $i=k-s$ in the first summation and
-$j=k-t$ in the second. Then (\ref{eqn:sum_proof}) becomes:
-\begin{equation*}
-R = \sum_{i=0}^{t-1} {i+s-1 \choose s-1} p^s (1-p)^i +
-\sum_{j=0}^{s-1} {j+t-1 \choose j} p^j (1-p)^t.
-\end{equation*}
-Write   
+$j=k-t$ in the second. 
 \begin{equation*}
@@
 \end{equation*}
-and so $R$ can be written
+Then %(\ref{eqn:sum_proof}) becomes:
+%\begin{equation*}
+%R = \sum_{i=0}^{t-1} {i+s-1 \choose s-1} p^s (1-p)^i +
+%\sum_{j=0}^{s-1} {j+t-1 \choose j} p^j (1-p)^t.
+%\end{equation*}
+$R$ can be written
 as the cumulative distribution function of two 
@@
 \begin{equation} \label{eqn:transformed_sum}
-R = \sum_{i=0}^{t-1} {i+s-1 \choose i} p^s (1-p)^i +
+R = \sum_{i=0}^{t-1} {i+s-1 \choose i} p^s (1-p)^i \; + \;
 \sum_{j=0}^{s-1} {j+t-1 \choose j} p^j (1-p)^t.
 \end{equation}
-Let $I_p(s, t)$ be the {\em regularized incomplete beta function}, which is also one minus the c.d.f. of the negative
-binomial distribution, and $I_p(s, t) = I_{1-p}(t, s)$.
-Then the result can be proven as:
+
+Let $I_p(s, t)$ be the {\em regularized incomplete beta function}, which is 
+also one minus the c.d.f. of the negative binomial distribution. This
+function satisfies $I_p(s, t) = I_{1-p}(t, s)$.  Then 
 \begin{align*}
@@
 \end{proof}
+Proving that \label{eqn:pmf} is a valid probability mass function.
+
 From (\ref{eqn:pmf}) one can see that the mass at any point on the support of the distribution may come from either $X_k = s$, captured in the function $S$, or $X_k = k-t$, captured in the function $T$.
@@
 These shapes are illustrated in Fig.~\ref{shapes.fig}.
-The SNB is related to the negative binomial distribution. Specifically, if $t$ is infinitely large then the $Y-s$ has a negative binomial distribution with
+The SNB is related to the negative binomial distribution. Specifically, if 
+$t$ is infinitely large then the $Y-s$ has a negative binomial distribution 
+with
 \begin{equation*}                                    %   (1)
@@
 \end{equation*}
-for $\,j=0, 1,\ldots\,$.
-Similarly, if $s$ is infinitely large then $\,Y - t\,$ has a negative binomial distribution with
-\begin{equation*}                                    %   (2)
-    \mathbb{P}\{Y=t+j \}        \label{nb2.eq} 
-               = {{t+j-1}\choose{t-1}} p^j(1-p)^t.
-\end{equation*}
+for $\,j=0, 1,\ldots\,$. A similar statement can be made when $s$ is large
+and $t$ is small.
 

### 2015-07-01 paper-2/main.tex
Moved stuff around and initial commit.

\documentclass{article}

% For visualization.
\usepackage{graphicx}

% For citations
\usepackage[authoryear,round]{natbib}

% For algorithms
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{authblk}

\hyphenation{general-iza-tions}
\hyphenation{distri-bution}
\hyphenation{Zel-ter-man}

\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{prop}{Proposition}

\begin{document}

\title{SNB Results Paper 2}

\author[1]{Michael J. Kane}
\author[1]{Daniel Zelterman}

\affil[1]{Department of Biostatistics, Yale University, New Haven CT, USA}

\maketitle

\begin{abstract}
This paper other properties of the snb...
\end{abstract}

\section{Introduce the Z-Plot Construction}

\begin{figure}[ht]
\includegraphics[width=\textwidth]{z-plot1.pdf}
\caption{
A hypothetical realization of the Bernoulli process.
}
\label{fig:z_plot}
\end{figure}

\section{Rotating the Z-Plot}

\bibliography{references}
\bibliographystyle{jss}

\end{document}

### 2015-07-01 paper-2/src/supp.r

@@ -1,0 +1,10 @@
+library(snb)
+
+s = 3
+t = 5
+p=0.5
+flips = c(0, 0, 1, 0, 0, 1, 0)
+pl = zplot(flips, s=s, t=t)
+
+ggsave("../z-plot1.pdf", pl, width=5, height=3)
+

### 2015-07-27 paper-1/main.tex
Almost done with editor minor comments.

--- 2015-06-09 main.tex
@@
 might include adverse reaction to the drug or withdrawal due to toxicity. We 
-use a large number of these early successes as evidence for whether to enroll
-additional patients for a second stage. A statistically small number of 
+use a large number of these early successes as evidence for whether to treat
+additional patients in a second stage. A statistically small number of 
 successes in Stage I would lead us to terminate the trial without entering 
@@
 also one minus the c.d.f. of the negative binomial distribution. This
-function satisfies $I_p(s, t) = I_{1-p}(t, s)$.  Then 
+function satisfies $I_p(s, t) = 1-I_{1-p}(t, s)$.  Then 
 \begin{align*}
@@
 
-From (\ref{eqn:pmf}) one can see that the mass at any point on the support of the distribution may come from either $X_k = s$, captured in the function $S$, or $X_k = k-t$, captured in the function $T$.
-The support of $S$ is defined by $s \leq S \leq s+t+1$ and the support of $T$ is defined by $t \leq T \leq s+t+1$.
+From (\ref{eqn:pmf}) one can see that the mass at any point on the support 
+of the distribution may come from either $X_k = s$, captured in the 
+function $S$, or $X_k = k-t$, captured in the function $T$. The support of 
+$S$ is defined by $s \leq S \leq s+t-1$ and the support of $T$ is defined 
+by $t \leq T \leq s+t-1$.
 
@@
 
-Returning to the breast cancer trial introduced in the first section, a response rate of 2/12 in the first stage corresponds to an individual probability of responding of $p=0.1667$.
-The standard deviation of this probability is $\sqrt{p(1-p)/12}=0.1076$.
-If we were to use the Central Limit Theorem then the 95\% confidence interval around the mean would be $[-0.0442, 0.1076]$.
-While this approach is problematic since the number of samples (12 in this case) is small it does indicate how much uncertainty is built into the first stage.
-If we consider samples from both stages then the standard deviation drops to 0.0630 and the 95\% confidence interval would be $[0.0432, 0.2901]$.
+Returning to the breast cancer trial introduced in the first section, a 
+response rate of 2/12 in the first stage corresponds to an individual 
+probability of responding of $p=0.1667$. The standard deviation of this 
+probability is $\sqrt{p(1-p)/12}=0.1076$. If we were to use the Central 
+Limit Theorem then the 95\% confidence interval around the mean would 
+be $[-0.0442, 0.3776]$. While this approach is problematic since the number 
+of samples (12 in this case) is small it does indicate how much uncertainty 
+is built into the first stage. If we consider samples from both stages then 
+the standard deviation drops to 0.0630 and the 95\% confidence interval 
+would be $[0.0432, 0.2901]$.
 
-These calculations expose two difficulties with the Simon Two-stage Optimal Design.
-First, success in the first stage is not a good indicator of success in the second stage and at the same time failure in first stage in no way implies failure in the second stage, especially when $p$ is close to it's threshold value.
-The variance in the estimate so large that it is a poor estimator.
-The second, difficulty is that the design is not {\em enrollment efficient}.
-That is, if the first stage is successful then additional patients are enrolled.
-Enrollment, especially for certain types of cancer, can be very difficult and it can taken time to find eligible patients.
-Furthermore, if all 12 of the patients in the first stage respond positively, should we really need to enroll additional patients?
+These calculations expose two difficulties with the Simon Two-stage Optimal 
+Design. First, success in the first stage is not a good indicator of success 
+in the second stage and at the same time failure in first stage in no way 
+implies failure in the second stage, especially when $p$ is close to it's 
+threshold value. The variance in the estimate is so large that it is a poor 
+estimator. The second difficulty is that the design is not {\em enrollment 
+efficient}. That is, if the first stage is successful then additional patients 
+are enrolled. Enrollment, especially for certain types of cancer, can be very 
+difficult and can take time to find eligible patients. Furthermore, if 
+all 12 of the patients in the first stage respond positively, should we 
+really need to enroll additional patients?
 
-Properly quantifying uncertainty in the process implies quantifying uncertainty in $p$, which stems from two sources.
-First is the number of samples we receive.
-Any estimate of $p$ will be based on data and generally the more data we have, the better our estimate.
-If the data are small, as they often are in clinical trials, the less reliable our estimate is.
-The second source comes from heterogeneity in the data.
-While models may assume that the probability of success $p$ is a single value this is often not the case in real-world data.
-In clinical trials for example an individuals chance of having an adverse outcome may be influenced by any number of underlying factors, which may vary across individual.
-In these cases $p$ may not be a single numeric value and may be better conceptualized as a realization over an entire distribution.
-A frequentest approach will not capture this variation.
+Properly quantifying uncertainty in the process implies quantifying 
+uncertainty in $p$, which stems from two sources. First is the number of 
+samples we receive. An estimate of $p$ will be based on data and generally 
+the more data we have, the better our estimate. If the data are small, as 
+they often are in clinical trials, the less reliable our estimate is. The 
+second source comes from heterogeneity in the data. While models may assume 
+that the probability of success $p$ is a single value this is often not the 
+case in real-world data. In clinical trials for example an individual's chance 
+of having an adverse outcome may be influenced by any number of underlying 
+factors, which may vary across individual. In these cases $p$ may not be a 
+single numeric value and may be better conceptualized as a realization over 
+an entire distribution. A frequentest approach will not capture this variation.
 
@@
 Both of the problems in the Simon Two-stage Optimal Design stem from the fact that it fails to quantify the uncertainty in the estimate of the success probability.
-A framework based on Stopped Negative Binomial distribution addresses this failure and provides the added features of monitoring the trial as new results are received, providing estimates of how many new enrollees are needed to complete a trial, and quantifying the uncertainty in the success probability.
+A framework based on the SNB distribution addresses this failure and
+provides the added features of monitoring the trial as new results are 
+received, providing estimates of how many new enrollees are needed to complete 
+a trial, and quantifying the uncertainty in the success probability.
 
-Let's start by recasting the breast cancer study in the context of the Stopped Negative Binomial framework, assuming sequential enrollment.
-The first stage will stop when either two positive responses are reached or 11 negative responses.
-This corresponds to a Stopped Negative Binomial Distribution with $s=2$ and $t=11$.
-Next, lets assume that, in the first stage, there were two successes.
-A second success could occur during the second, through a tenth enrollment.
-At each of these times we would like to know the estimate for $p$ along with the lower five percent quantile of the distribution.
+Let's start by recasting the breast cancer study in the context of the SNB 
+framework, assuming sequential enrollment. The first stage will stop when 
+either two positive responses are reached or 11 negative responses. This 
+corresponds to a SNB distribution with $s=2$ and $t=11$. Next, let's assume 
+that, in the first stage, there were two successes. A second success could 
+occur during the second, through a tenth enrollment. At each of these times we 
+would like to know the estimate for $p$ along with the lower five percent 
+quantile of the distribution.
 
@@
 \caption{
-A visualization of the mode and the 0.05\% quantile of the estimate of the distribution of $p$.
+A visualization of the mode and the 5\% quantile of the estimate of the distribution of $p$.
 }
@@
 \caption{
-A table of the mode and the 0.05\% quantile of the estimate of the distribution of $p$.
+A table of the mode and the 5\% quantile of the estimate of the distribution of $p$.
 }
@@
 \caption{
-A visualization of the mode and the 0.05\% quantile of the estimate of the distribution of $p$ for a hypothetical sequence of patient enrollments.
+A visualization of the mode and the 5\% quantile of the estimate of the distribution of $p$ for a hypothetical sequence of patient enrollments.
 }
@@
 
-Consider the trial described above.
-Suppose eight patients were enrolled and there was a single success on the third trial (visualized in Fig.
-\ref{fig:hypo}).
-Since the mode of the beta distribution can only be calculated when there is at least one success the maximum of the mean and the mode of the distribution is shown as the estimate.
-At the third time step there is an increase in the estimate and the 0.05 \% quantile corresponding to the success.
+Consider the trial described above. Suppose eight patients were enrolled and 
+there was a single success on the third trial (visualized in Fig. 
+\ref{fig:hypo}). The mode of the beta distribution can only be calculated when 
+there is at least one success. The mean of the distribution can be
+taken otherwise. At the third time step there is an increase in the estimate and the 5\% quantile corresponding to the success.
 
@@
 Fig.
-\ref{fig:conditional_snb} shows the conditional, compound SNB distribution.
+\ref{fig:conditional_snb} shows the compound SNB distribution.
 In each of the following enrollment a success would end the trial.
@@
 \ref{fig:hypo} we know that the estimate of $p$ is about 0.19.
-The conditional compound distribution gives the probability of another success as 0.3883 and the probability of no more success (and the first stage failing) as 0.6117.
+The compound distribution gives the probability of another success as 0.3883 and the probability of no more success (and the first stage failing) as 0.6117.
 Furthermore, we can see that the probability that the trial ends in the next step is approximately 20\% and the probability the trials end with the $11^{th}$ patient is about 65\%.
@@
 
-Suppose that in our example trial that a second successful outcome is reached on the $10^th$ patient.
-From the data collected we have shown how to create an estimate for the distribution of $p$.
-Based on that estimate we may like to see the probability of a different outcome.
-That is, what is the probability that we don't get two successful outcomes based on the data we have collected and what is the posterior distribution for the number of patients enrolled for the study?
+Suppose that, in our example trial, a second successful outcome was reached 
+on the $10^{th}$ patient. From the data collected we have shown how to create 
+an estimate for the distribution of $p$. Based on that estimate we may like to 
+see the probability of a different outcome. We may want to answer the
+question, "What is the probability that we don't get two successful outcomes?" 
+We may also ask what is the distribution for the number of patients 
+enrolled for a similar study?
 
-In this case the success probability $p$ is distributed as $Beta(2.5, 8.8)$.
-Using the mode a point estimate yields $\hat{p} = 0.1667$.
-The standard deviation of the estimate is 0.1210.
-The probability of two successes during the trial is 0.6843 and the probability of failure is 0.3157.
-The distribution of the stopping times, along with contributions from two positive events and less than two positive events is shown below.
+If there are two successful enrollments with the second one occurring 
+on the $10^{th}$ trial the success probability $p$ is distributed as 
+$Beta(2.5, 8.8)$. 
+Using the mode, a point estimate yields $\hat{p} = 0.1667$. The standard 
+deviation of the estimate is 0.1210. The probability of two successes during 
+the trial is 0.6843 and the probability of failure is 0.3157. The distribution 
+of the stopping times, along with contributions from two positive events and 
+less than two positive events is shown below.
 

### 2015-07-27 paper-2/main.tex
Almost done with editor minor comments.

@@
 \newtheorem{prop}{Proposition}
+\newtheorem{thm}{Theorem}
 
@@
 
-\author[1]{Michael J. Kane}
-\author[1]{Daniel Zelterman}
+\author{Michael J. Kane}
+\author{Daniel Zelterman}
 
-\affil[1]{Department of Biostatistics, Yale University, New Haven CT, USA}
+\affil{Department of Biostatistics, Yale University, New Haven CT, USA}
 
@@
 
-\begin{abstract}
-This paper other properties of the snb...
-\end{abstract}
+%\begin{abstract}
+%This paper other properties of the snb...
+%\end{abstract}
 
-\section{Introduce the Z-Plot Construction}
+\section{The Stopped Negative Binomial Distribution} \label{sect:intro}
+
+\begin{enumerate}
+\item Review of the SNB
+\item Show when the SNB is and isn't unimodel.
+\item Past efforts to show the asymptotic behavior using Sterling 
+approximations.
+\end{enumerate}
+
+\section{Main Proof}
+
+\begin{thm} \label{thm:main}
+The main theorem states that once there is a central limit effect 
+then the hitting times look the L\'{e}vy-Bachlier formula for the density 
+of the first hitting time of a sloped line.
+\end{thm}
+
+The proof of Theorem \ref{thm:main} proceeds as follows. In Section
+\ref{sect:process_construction} it is shown that the SNB can be 
+constructed as a binomial process on a grid. Under suitable 
+rotation and scaling the process is equivalent to a random walk 
+on the integers with stopping rules. Section \ref{sect:fclt} shows that
+after reparameterizing the process it converges to a Standard Brownian
+Motion with stopping rules. The density of the first hitting times for
+the two separate endpoints are each distributed according to the 
+L\'{e}vy-Bachlier formula. Section \ref{sect:stat} provides statistical 
+results, deriving the convergence when the rotation and scaling parameters
+are estimated from data. Section \ref{sect:discussion} ends the article
+with a discussion and proposes avenues for future work.
+
+\section{Process Construction} \label{sect:process_construction}
+
+As shown in Section \ref{sect:intro} the SNB can be conceptualized as
+a series of coin flips that stop when either $s$ heads or $t$ tail are
+reached. The time to reach either $s$ or $t$ is distributed as the SNB.
+The process can be visualized as in Figure \ref{z-plot1.pdf}. The process
+starts at the origin and at each step a coin is flipped if it is heads
+the process moves in the vertical direction one step, if tails it moves
+one step in the horizontal direction.
 
@@
 \caption{
-A hypothetical realization of the Bernoulli process.
+A hypothetical realization of the SNB coin-flipping process.
 }
@@
 
-\section{Rotating the Z-Plot}
+The expected slope of the process is determined by $p \in [0, 1]$, the 
+probability of a heads. If $s$ and $t$ are finite, then as $p$ increases 
+the probability a given realization of the process hits $s$ increases (and
+$t$ decreases). Likewise as the $p$ decreases the probability a realization 
+hits $s$ decreases (and $t$ increases).
+
+Letting $p = \alpha / \alpha + \beta$, the expected direction of the 
+process is $\theta = \text{atan}(\alpha / \beta)$. In Section \ref{sect:stat}
+it is shown that an estimate of $\theta$ converges at the same rate as
+an estimate of $p$. For now though, we would like to rotate the process
+clockwise by $\theta$ so that, in expectation, the rotate process angle
+is zero.
+
+\begin{lemma}
+After rotation, the line $y=s$ becomes
+\begin{equation}
+y' = -\tan(\theta) x' + s \left(\sin(\theta) \tan(\theta) + \cos(\theta) \right)
+\end{equation}
+where $y'$ and $x'$ are the vertical and horizontal axes in the rotation
+space.
+\end{lemma}
+
+\begin{enumerate}
+\item Show a picture of the z-plot before and after rotation/rescaling.
+\item Rotate the system clockwise by $\theta = p = \alpha/(\alpha+\beta)$ 
+where $\beta$ is the number of failures and $\alpha$ is the number of successes.
+\item Derive the rotated $t$- and $s$-lines.
+\item Calculate the scaling for the rotated $t$- and $s$- lines so that the
+reparameterized process goes right and either up on step or down one step...
+Like a random walk on the integers over time.
+\end{enumerate}
+
+\section{The FCLT for the process} \label{sect:fclt}
+
+\begin{enumerate}
+\item TODO: Find the appropriate scalings for $s$ an $t$ in terms of $n$
+\item Use the Heyde/Hall FCLT to show that the symmetric random walk
+converges to a Standard Brownian motion over the appropriate interval.
+\item Derive the probability of hitting $s$ or $t$ as a Girsinov process.
+\end{enumerate}
+
+Describe the properties of the Girsinov process and relate them back to 
+the multi-mode result from the beginning.
+
+\section{Estimating $\theta$ and the rotations from data} \label{sect:stat}
+
+\begin{enumerate}
+\item Show that if we need to estimate $\theta$ that it converges like 
+$\widehat{p}$. Include the graph of $\widehat{\theta}$ vs. $\alpha$.
+\item Show convergence results for for the $s$- and $t$-lines. 
+\end{enumerate}
+
+\section{Discussion and Future Work} \label{sect:discussion}
 

### 2015-08-19 paper-1/main.tex
Working through comments.

@@
 Our goal is to know the distribution of the sample size of the design for 
-planning purposes.  If all 12 patients are enrolled at once, as in the classic 
+planning purposes and to characterize the uncertainty in the estimate of the
+success probability.  If all 12 patients are enrolled at once, 
+as in the classic 
 design, then the sample size is 12. However, in most clinical trials, the 
@@
 
-The next section of this paper introduces our notation and basic results including the density of the SNB and compound SNB along with a description of it's relation to other distributions.
-Sections 2 and 3 derive the distribution based on a defined Bernoulli process and give some basic properties of the distribution.
-Section 4 derives the compound distribution using a Beta prior.
-Section 5 develops a post-hoc analysis of a completed trial.
-Section 6 is devoted to discussion and conclusions.
-The simulations, model fitting routines, and visualizations presented in this paper were generated using the snb package [\ref{Kane 2015}] for the R programming environment [\ref{R Core Team 2015}]
+The next section of this paper introduces our notation and basic results 
+including the density of the distribution along with a description of 
+it's relation to other distributions. Sections 2 and 3 derive the distribution 
+based on a defined Bernoulli process and give some basic properties of the 
+distribution. Section 4 derives the compound distribution using a Beta prior. 
+Section 5 develops a post-hoc analysis of a completed trial. Section 6 is 
+devoted to discussion and conclusions. The simulations, model fitting 
+routines, and visualizations presented in this paper were generated using the 
+snb package [\ref{Kane 2015}] for the R programming 
+environment [\ref{R Core Team 2015}]
 
@@
  & {k-1 \choose k-t} p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp \\
-= \frac{1}{B(\alpha, \beta)}  {k-1 \choose s-1} & \int_0^1  p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} + \\
+= \frac{1}{B(\alpha, \beta)}  {k-1 \choose s-1} & \int_0^1  p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} dp + \\
  & \frac{1}{B(\alpha, \beta)} {k-1 \choose k-t} \int_0^1  p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp
@@
 
-Like most Bayesian Adaptive Clinical trials the SNB approach allows for monitoring of trial and can be updated as new results are received without losing power.
-Essentially this is because the object under investigation is not a point estimate of the success probability.
-It is the uncertainty in the distribution of the success probability.
-Point estimates of the success probability can be regarded as artifacts of our parameterization of the underlying distribution.
-Every point estimate calculated is accompanied by the quantification of the uncertainty around it.
-This estimate and it's uncertainty can be calculated at any time and may be useful in evaluating ongoing trials as well as providing a post-hoc evaluation of trials that have been completed.
+Like most Bayesian Adaptive Clinical trials the SNB approach allows for monitoring of trial and can be updated as new results are received without losing 
+power. This is because the object under investigation is not a point estimate 
+of the success probability; it is the uncertainty in the distribution of the 
+success probability. Point estimates of the success probability can be 
+regarded as artifacts of our parameterization of the underlying distribution. 
+Every point estimate calculated is accompanied by the quantification of the 
+uncertainty around it. This estimate and its uncertainty can be calculated at 
+any time and may be useful in evaluating ongoing trials as well as providing a 
+post-hoc evaluation of trials that have been completed.
 
@@
 
-Fig.
-\ref{fig:conditional_snb} shows the compound SNB distribution.
-In each of the following enrollment a success would end the trial.
-If none of the following enrollments are success then the trial will stop at the eleventh patient.
-From Fig.
-\ref{fig:hypo} we know that the estimate of $p$ is about 0.19.
-The compound distribution gives the probability of another success as 0.3883 and the probability of no more success (and the first stage failing) as 0.6117.
-Furthermore, we can see that the probability that the trial ends in the next step is approximately 20\% and the probability the trials end with the $11^{th}$ patient is about 65\%.
+Fig \ref{fig:hypo} shows that at the eight enrollment the estimate of $p$ is 
+about 0.19 and, from the lower 5\% quantile, we can see that there is
+a 95\% probability that $p$ is greater than 2.5\%. 
+Fig. \ref{fig:conditional_snb} shows the compound SNB distribution for this 
+hypothetical trial at the eighth step.  Successes may occur 
+at steps nine through 11 with probability 26.7\%. The remaining probability
+mass corresponds to reaching the $11^{th}$ step without a second success.
 
@@
 see the probability of a different outcome. We may want to answer the
-question, "What is the probability that we don't get two successful outcomes?" 
+question, ``What is the probability that we don't get two successful outcomes?''
 We may also ask what is the distribution for the number of patients 

### 2015-08-19 paper-1/script.r

@@ -1,0 +1,75 @@
+library(ggplot2)
+
+stairs <- function(p, xstart, xend) {
+  x <- c(xstart, rep((xstart+1):xend, each=2))
+  y <- rep(0:(xend-xstart), each=2)
+  y <- y[-length(y)]
+  for (i in 1:(length(x)-1)) {
+    p <- p + geom_segment(x=x[i], y=y[i], xend=x[i+1], yend=y[i+1], 
+      color="green")
+  }
+  p
+}
+
+s <- 3
+t <- 10
+flips <- rbinom(2*t, 1, 0.3)
+x <- c(0, cumsum( flips))
+X <- data.frame(list(X=x, k=0:(length(x)-1)))
+X1 <- X[X$X <= s,]
+
+p1 <- qplot(k, X, data=X1, geom="line") +
+  scale_x_continuous(breaks=0:t, limits=c(0, t)) +
+  scale_y_continuous(breaks=0:s, limits=c(0, s)) +
+  geom_segment(x=0, y=s, xend=t, yend=s, color="red") + 
+  geom_segment(x=t, y=0, xend=t, yend=(s-1), color="green")
+
+ggsave("BernoulliProcess.pdf", p1, width=7, height=5)
+
+X$head <- c(0, cumsum(flips))
+X$tail<- c(0, cumsum(!(flips)))
+X$headEnd <- c(X$head[-1], NA)
+X$tailEnd <- c(X$tail[-1], NA)
+
+X2 <- X[1:(min(which(X$headEnd >= s), which(X$tailEnd >= t))),]
+
+p2 <- ggplot(data=X2) + #right, up, data=X) + #, geom="segment") +
+  scale_x_continuous(breaks=0:t, limits=c(0, t)) +
+  scale_y_continuous(breaks=0:s, limits=c(0, s)) +
+  geom_segment(mapping=aes(x=tail, y=head, xend=tailEnd,
+    yend=headEnd), arrow=arrow()) +
+  geom_segment(x=0, y=s, xend=t-1, yend=s, color="red") +
+  geom_segment(x=t, y=0, xend=t, yend=s-1, color="green")
+
+ggsave("ZeltermanPlot.pdf", p2, width=7, height=5)
+
+X3 <- X[1:(min(which(X$X >= s), which(X$tail >= t))),]
+p3 <- qplot(k, X, data=X3, geom="line") +
+  scale_x_continuous(breaks=0:(t+s), limits=c(0, t+s)) +
[…]

### 2015-08-21 paper-1/script.r
Updated script for paper.

@@ -56,20 +56,7 @@ ggsave("KanePlot.pdf", p3, width=7, height=5)
 s=2
 t=11
 trial = c(0, 0, 1, 0, 0, 0, 0, 0)
-fit = fit_flips(trial, s=s, t=t)
-# Use the mode for the 
-phat = (fit['shape1'] - 1) / (fit['shape1'] + fit['shape2'] - 2)
-total_range = s+t-1
-conditional_range = (length(trial):total_range)
-conditional_s = s - sum(trial)
-conditional_t = t - sum(1-trial)
-# Use the mode unless we don't have enough data.
-p_hat = if(fit[1] < 1 || fit[2] < 1) {
-          fit[1] / (fit[1] + fit[2]) 
-        } else {
-          (fit[1] - 1) / (fit[1] + fit[2] - 2)
-        }
+p4 = cdsnb_stack_plot(trial, s, t)
+ggsave("conditional_snb.pdf", p4, width=7, height=5)
 
-step_offset = length(trial)
-p4 = dsnb_stack_plot(p_hat, conditional_s, conditional_t, offset=step_offset)
-ggsave("conditional_snb.pdf", width=7, height=5)
+cdsnb(trial, s, t)

### 2015-09-13 paper-1/main.tex
Most of Michelle's comments are addressed.

@@
 \typeout{}\typeout{}
-\typeout{ Michael Kane \& Dan Zelterman }
+\typeout{ Michelle DeVeaux \& Michael Kane \& Dan Zelterman }
 \typeout{}\typeout{}\typeout{}\typeout{}
@@
 \usepackage{url}
+\DeclareMathOperator*{\argmin}{argmin}
+\newcommand*{\argminl}{\argmin\limits}
 \begin{document}
@@
 \begin{tabular}{l} 
+ {\bf Michelle DeVeaux} \\
  {\bf Michael Kane${}^*$} \\
@@
 number of patients are enrolled onto a therapeutic 
-protocol for an initial stage. Fore each patient, we assess either success 
+protocol for an initial stage. For each patient, we assess either success 
 or failure according to a 
@@
 % ---------------------------  2. Notation ----------------------
-\section   {Notation and summary of distributional results}
+\section   {Notation and Summary of Distributional Results}
 \label{notation.section}
@@
 \end{equation}
-is the negative binomial probability of $t$ failures.
+is the negative binomial probability of $t$ failures. It should be noted that
+this result expressed using the de Finetti where indicator functions and
+sets are equivalent. A more complete explanation of this equivalence can be 
+found in [\ref{Pollard 2002}].
 
@@
 \begin{equation*} \label{eqn:proc}
-X_{k+1} = X_k + B_{k+1} \{ k-t < X_k < s \}
+X_{k+1} = X_k + B_{k+1} \left[ k-t < X_k < s \right]
 \end{equation*}
@@
 \begin{prop}
-The distribution of the stopping time $\{k: s \leq X_k \leq k-t \}$ is equal to the SNB and is given in (\ref{eqn:pmf}). 
+The distribution of the stopping time $\argminl_k [X_k \geq s, X_k \leq k-t]$ 
+is equal to the SNB and is given in (\ref{eqn:pmf}). 
 \end{prop}
@@
 \end{proof}
-Proving that \label{eqn:pmf} is a valid probability mass function.
+Proving that \ref{eqn:pmf} is a valid probability mass function.
 
@@
 \begin{center}
-\includegraphics[scale=.9]{shapes.pdf}
+\includegraphics[width=\textwidth]{shapes.pdf}
 \end{center}
-\caption{Different shapes of the SNB distribution with parameters ($s$, $t$, $p$), as given. \label{shapes.fig}}
+\caption{Different shapes of the SNB distribution with parameters ($s$, $t$, $p$), as given. Red indicates mass contributed by hitting $s$, teal indicates
+mass contributed by hitting $t$. \label{shapes.fig}}
 \end{figure}
@@
 
-For the special case of $\,s=t,$ the distribution of $\,Y\,$ is the riff-shuffle, or minimum negative binomial distribution~[\ref{Uppuluri 1970}].
-Similar derivations of the closely-related maximum negative binomial discrete distributions also appear in~[\ref{Zhang 2000}, \ref{Zelterman 2004}].
+For the special case of $\,s=t,$ the distribution of $\,Y\,$ is the 
+riff-shuffle, or minimum negative binomial distribution~[\ref{Uppuluri 1970}].
+Similar derivations of the closely-related maximum negative binomial discrete 
+distributions also appear in~[\ref{Zhang 2000}, \ref{Zelterman 2004}]. However,
+where the maximum negative binomial is the smallest number of coin-flips 
+to observe at least $c$ successes {\em and} $c$ failures, the snb give
+the number of coin flips to observe {\em either} $s$ heads or $t$ tails.
 
@@
 
-\section{The Compound Distribution}
+\section{The Posterior Distribution}
 
@@
 
-If we assume that $p$ has a Beta distribution, with constant prior parameters $\alpha$ and $\beta$ then we can calculate the closed form compound distribution of the SNB as
+If we assume that $p$ has a Beta distribution, with constant prior parameters $\alpha$ and $\beta$ then we can calculate the closed form posterior 
+distribution of the SNB as
 \begin{align}
@@
 
-When $p$ is known, the probability mass function (PMF) of the number of steps until the process is stopped is given by (\ref{eqn:pmf}).
-However, many times this is not the case.
-In Phase II clinical trials for example patients are often enrolled sequentially until there are enough patients to guarantee the desired level of power or until a specified number of adverse events are recorded.
-In cases such as these the endpoints $s$ and $t$ are known and we would like to be able to estimate $p$ either during enrollment to provide odds of successful enrollment or after an endpoint has been reached to determine the amount of uncertainty in the endpoint.
-To do this, we will assume that $p$'s probability in the absence of available data is $\text{Beta}(1/2, 1/2)$, the Jeffrey's prior.
-This distribution was chosen for two reasons.
-First, it is the conjugate prior of the (Negative) Binomial distribution and updating the distribution as data are received is straightforward.
-Second, it is objective and limits the number of ``prior coin-flips'' to one.
-This means that data are more heavily influence the distributional characteristics of $p$ with more data providing more proportional influence.
+When $p$ is known, the probability mass function (PMF) of the number of steps 
+until the process is stopped is given by (\ref{eqn:pmf}). However, many times 
+this is not the case. In Phase II clinical trials for example patients are 
+often enrolled sequentially until there are enough patients to guarantee the 
+desired level of power or until a specified number of adverse events are 
+recorded. In cases such as these the endpoints $s$ and $t$ are known and we 
+would like to be able to estimate $p$ either during enrollment to provide odds 
+of successful enrollment or after an endpoint has been reached to determine 
+the amount of uncertainty in the endpoint. 
+%To do this, we will assume that $p$'s probability in the absence of 
+%available data is $\text{Beta}(1/2, 1/2)$, the Jeffrey's prior. This 
+%distribution was chosen for two reasons. First, it is the conjugate prior 
+%of the (Negative) Binomial distribution and updating the distribution as data 
+%are received is straightforward. Second, it is objective and limits the number 
+%of ``prior coin-flips'' to one. This means that data are more heavily 
+%influence the distributional characteristics of $p$ with more data providing 
+%more proportional influence.
 
 \begin{prop}
-The compound PMF of the Stopped Negative Binomial distribution with a Beta($\alpha$, $\beta$) prior is:
+The posterior PMF of the Stopped Negative Binomial distribution with a Beta($\alpha$, $\beta$) prior is:
 \begin{align}
@@
 
-Returning to the breast cancer trial introduced in the first section, a 
-response rate of 2/12 in the first stage corresponds to an individual 
+Returning to the breast cancer trial introduced in the first section, consider
+the first stage, having a sampel size of 12. A
+response rate of 2/12 stage corresponds to an individual 
 probability of responding of $p=0.1667$. The standard deviation of this 
@@
 single numeric value and may be better conceptualized as a realization over 
-an entire distribution. A frequentest approach will not capture this variation.
+an entire distribution. A frequentist approach will not capture this variation.
 
@@
 
-To better understand how the distributional estimate of $p$ changes in the number of steps needed for two successes data were simulated for each potential stopping point from two to 11.
-This procedure essentially estimates the shape parameters of a beta distribution.
-Fig.
-\ref{fig:simon} shows the change in the mode of the distribution, which is also the Maximum Likelihood Estimator (MLE) as well as the five percent quantile of the distribution.
-The graph shows, for example, that if a second success is found with the third enrollment then the MLE estimate of $p$ is 0.5 and with 95\% confidence $p$ is at least 0.25.
-In this case investigators could have moved on to the next stage of the trial with only three enrollments as opposed to 12.
+To better understand how the distributional estimate of $p$ changes in the 
+number of steps needed for two successes data were simulated for each 
+potential stopping point from two to 11. That is, we will fit the
+Beta distribution with the Jeffrey's prior to the sequence of coin flips 
+$\{[1,1], [1, 0, 1], [1, 0, 0, 1], ... \}$ in the support of the SNB with
+$s=2$ and $t=11$.  Fig. \ref{fig:simon} shows the 
+change in the mode of the beta distribution, which is also the Maximum 
+Likelihood Estimator (MLE) as well as the five percent quantile of the 
+distribution.  The graph shows, for example, that if a second success is 
+found with the third enrollment then the MLE estimate of $p$ is 0.5 and with 
+95\% confidence $p$ is at least 0.25. In this case investigators could have 
+moved on to the next stage of the trial with only three enrollments as 
+opposed to 12.
 
@@
 \ref{fig:simon}. 
-In the beginning of this section it was shown that, at the end of stage two, the breast cancer study actually tests whether the probability of success is greater than 0.0432 with 95\% confidence and requires 35 patients to do so.
-With the Stopped Negative Binomial model we only need six patients for the same level of confidence.
-This represents a tremendous amount of savings in time, effort and resources.
+%In the beginning of this section it was shown that, at the 
+%end of stage two, the breast cancer study actually tests whether the 
+%probability of success is greater than 0.0432 with 95\% confidence and 
+%requires 35 patients to do so. With the Stopped Negative Binomial model we 
+%only need six patients for the same level of confidence. This represents a 
+%tremendous amount of savings in time, effort and resources.
 
@@
 
-Fig \ref{fig:hypo} shows that at the eight enrollment the estimate of $p$ is 
+Fig \ref{fig:hypo} shows that at the eighth enrollment the estimate of $p$ is 
 about 0.19 and, from the lower 5\% quantile, we can see that there is
 a 95\% probability that $p$ is greater than 2.5\%. 
-Fig. \ref{fig:conditional_snb} shows the compound SNB distribution for this 
+Fig. \ref{fig:conditional_snb} shows the posterior SNB distribution for this 
 hypothetical trial at the eighth step.  Successes may occur 
@@
 on the $10^{th}$ trial the success probability $p$ is distributed as 
-$Beta(2.5, 8.8)$. 
+$Beta(2.5, 8.5)$ (assuming the Jeffrey's prior). 
 Using the mode, a point estimate yields $\hat{p} = 0.1667$. The standard 
@@
 The distribution models the stopping time of a sequential trial where the trial is stopped when a number of events are accumulated.
-The compound distribution was derived for the case when the event probability $p$ has a Beta distribution.
-Using a trial description from \url{clinicaltrials.gov} we showed how the SNB is an integral part of trial monitoring, and post-hoc analysis and can be used in a framework alternative to the Simon two-stage optimal design while providing better estimates of the event probability along with quantifying the uncertainty associated with the estimates.
-As a result, we were able to show fewer patient enrollments and achieve the same estimates, with added  in a hypothetical, but representative, clinical trial.
+The
+posterior distribution was derived for the case when the event probability $p$ 
+has a Beta distribution. Using a trial description from 
+\url{clinicaltrials.gov} we showed how the SNB is an integral part of trial 
+monitoring, and post-hoc analysis and can be used in a framework alternative 
+to the Simon two-stage optimal design while providing better estimates of the 
+event probability along with quantifying the uncertainty associated with the 
+estimates. As a result, we were able to show fewer patient enrollments and 
+achieve the same estimates, when added in a hypothetical, but representative, 
+clinical trial.
 
@@
 For these trials monitoring may need to be performed not only on the outcome of an intervention but also on the safety of the trial.
-This especially in areas like late-stage cancer treatments where the treatment is harsh and patients may be forced to drop out as a result of the side-effects of the treatment; a matter independent of the outcome.
-Other application areas include providing design that allows clinicians to balance uncertainty and success probability with a minimum number of patient enrollees.
+This especially
+true in areas like late-stage cancer treatments where the treatment is harsh 
+and patients may be forced to drop out as a result of the side-effects of the 
+treatment; a matter independent of the outcome. Other application areas 
+include providing designs that allows clinicians to balance uncertainty and 
+success probability with a minimum number of patient enrollees.
 
@@
 Kane MJ. (2015). snb: The Stopped Negative Binomial. R package version 0.1. 2015. Available from \url{https://github.com/kaneplusplus/snb}
+
+\item[\bibref{Pollard 2002}]
+Pollard D. {\it A User's Guide to Measure Theoretic Probability}, 
+Cambridge University Press, 2002.
 

### 2015-09-13 paper-1/script.r

@@ -60,3 +60,42 @@ p4 = cdsnb_stack_plot(trial, s, t)
 ggsave("conditional_snb.pdf", p4, width=7, height=5)
 
 cdsnb(trial, s, t)
+
+
+approx_plots_start = list(
+  dsnb_stack_plot(0.2, 15, 75) + scale_x_discrete(breaks=seq(15, 95, by=10)) + 
+    labs(x="", y="", title="bimodal example (15, 75, .2)"),
+  dsnb_stack_plot(0.35, 50, 50) + scale_x_discrete(breaks=seq(50, 105, by=10))+
+    labs(x="", y="", title="approximate normal (50, 50, .35)"),
+  dsnb_stack_plot(0.06, 10, 10) + scale_x_discrete(breaks=seq(10, 20, by=2))+
+    labs(x="", y="", title="geometric (10, 10, .06)"),
+  dsnb_stack_plot(0.06, 3, 175) + scale_x_discrete(breaks=seq(3, 180, by=30))+ 
+    labs(x="", y="", title="gamma approximation (3, 175, .06)"),
+  dsnb_stack_plot(0.5, 25, 25) + scale_x_discrete(breaks=seq(25, 50, by=5))+
+    labs(x="", y="", title="lower half normal (25, 25, .5)"),
+  dsnb_stack_plot(0.98, 175, 2) + scale_x_discrete(breaks=seq(2, 180, by=20))+
+    labs(x="", y="", title="local mode at top of range (175, 2, .98)"))
+
+library(foreach)
+library(grid)
+
+approx_plots = foreach(p=approx_plots_start) %do% {
+  p + theme(legend.position="none") 
+}
+
+vplayout <- function(x, y) viewport(layout.pos.row = x, layout.pos.col = y)
+
+pdf("shapes.pdf", width=10, height=11)
+grid.newpage()
+p = pushViewport(viewport(layout = grid.layout(3, 2)))
+
+k=1
+for (i in 1:3) {
+  for (j in 1:2) {
+    print(approx_plots[[k]], vp=vplayout(i, j))
+    k = k+1
+  }
+}
+dev.off()
+
+

### 2015-09-13 paper-1/main.tex
Yeah.

@@
 
-This paper introduces a new discrete distribution suggested by curtailed sampling rules common in early-stage clinical trials.
-We derive the distribution of the smallest number of independent Bernoulli($p$) trials needed in order to observe either $s$ successes or $t$ failures.
-The closed form expression for the distribution as well as the compound distribution are derived.
-Properties of the distribution are shown and discussed.
-A case study is presented showing how the distribution can be used to monitor sequential enrollment of clinical trials with binary outcomes as well as providing post-hoc analysis of completed trials.
+This paper introduces a new discrete distribution suggested by curtailed 
+sampling rules common in early-stage clinical trials. We derive the 
+distribution of the smallest number of independent Bernoulli($p$) trials 
+needed in order to observe either $s$ successes or $t$ failures. The closed 
+form expression for the distribution as well as the compound distribution are 
+derived. Properties of the distribution are shown and discussed. A case study 
+is presented showing how the distribution can be used to monitor sequential 
+enrollment of clinical trials with binary outcomes as well as providing 
+post-hoc analysis of completed trials.
 
@@
 
-Consider the Phase II clinical trial that investigates the efficacy of 
+Consider the Phase II clinical trial that investigated the efficacy of 
 iniparib in patients with breast cancer gene-associated (BRCA) ovarian 
@@
 Criteria in Solid Tumor (RECIST). If at least two Stage I patients succeeded,
-then the trial would proceed to the next stage where additional patients were 
-treated. If fewer than two of the 12 Stage I patients is successful then 
-the trial would terminate.
+then the trial would have proceed to the next stage where additional patients 
+were treated. If fewer than two of the 12 Stage I patients is successful then 
+the trial would have terminated.
 
-Our goal is to know the distribution of the sample size of the design for 
-planning purposes and to characterize the uncertainty in the estimate of the
+Our goal is to describe the distribution of the sample size of the design for 
+planning purposes and to use this distribution to estimate the
 success probability.  If all 12 patients are enrolled at once, 
@@
 
-We will assume each of the patient enrollments can be modeled as independent, identically distributed Bernoulli($p$) random variables.
-The realization of the trial is a sequence of these random variables that stops when either a specified number of success or failures is reached.
-As an example, suppose in the previously described trial, two successes were reached after enrolling 10 patients (one in the third step and one at the $10^{th}$).
-The trial is visualized in Fig \ref{fig:kane_viz}.
-The vertical axis denotes the number of successful outcomes.
-The horizontal axis denotes the number of patients that have been enrolled.
-The horizontal and vertical line segments represent endpoints for the trial.
+We assume each of the patient outcomes can be modeled as independent, 
+identically distributed Bernoulli($p$) random variables. The realization of 
+the trial is a sequence of these random variables that stops when either a 
+specified number of success or failures is reached. As an example, suppose in 
+the previously described trial, two successes were reached after enrolling 10 
+patients (one in the third step and one at the $10^{th}$). The trial is 
+visualized in Fig \ref{fig:kane_viz}. The vertical axis denotes the number of 
+successful outcomes. The horizontal axis denotes the number of patients that 
+have been enrolled. The horizontal and vertical line segments represent 
+endpoints for the trial.
 
@@
 
-Let $\,B_1, B_2, \ldots \,$ denote a sequence of independent, identically 
-distributed, Bernoulli random variables with $\mathbb{P}\{B_i=1\}=p$, for 
+Let $\,b_1, b_2, \ldots \,$ denote a sequence of independent, identically 
+distributed, Bernoulli random variables with $\mathbb{P}\{b_i=1\}=p$, for 
 probability parameter $0\leq p \leq 1$. In the clinical trial setting 
-$\,B_i = 0$ corresponds to a successful enrollment. 
-Let $s$ and $t$ be positive integers. 
-Define the stopped negative binomial (SNB) random variable $Y$ as the smallest 
-integer value such that $\,\{B_1, \ldots , B_Y\}\,$ contains {\em either} 
-$\,s\,$ successes or else $\,t\,$ failures. The distribution of $\,Y\,$ has 
+$\,b_i = 0$ corresponds to an unsuccessful outcome.  Let $s$ and $t$ be 
+positive integers.  Define the stopped negative binomial (SNB) random 
+variable $Y$ as the smallest 
+integer value such that $\,\{b_1, \ldots , b_Y\}\,$ contains {\em either} 
+$\,s\,$ successes {\em or} $\,t\,$ failures. The distribution of $\,Y\,$ has 
 support on integer values in the range 
@@
 \begin{equation} \label{eqn:pmf}
-\mathbb{P}\{Y=k\} = S(k, p, s)  \{s \leq k\} + T(k, p, t) \{t \leq k \} 
+\mathbb{P}[Y=k] = S(k, p, s)  [s \leq k] + T(k, p, t) [t \leq k ] 
 \end{equation}
@@
 \end{equation}
-is the negative binomial probability of $s$ successes and
+is the negative binomial probability of $s$ successful outcomes and
 \begin{equation} \label{eqn:R}
@@
 \end{equation}
-is the negative binomial probability of $t$ failures. It should be noted that
-this result expressed using the de Finetti where indicator functions and
-sets are equivalent. A more complete explanation of this equivalence can be 
-found in [\ref{Pollard 2002}].
+is the negative binomial probability of $t$ unsuccessful outcomes. 
+From (\ref{eqn:pmf}) one can see that the mass at any point on the support 
+of the distribution may come from either $X_k = s$, captured in the 
+function $S$, or $X_k = k-t$, captured in the function $T$. The support of 
+$S$ is defined by $s \leq S \leq s+t-1$ and the support of $T$ is defined 
+by $t \leq T \leq s+t-1$.
+
+%It should be noted that
+%this result expressed using the de Finetti where indicator functions and
+%sets are equivalent. A more complete explanation of this equivalence can be 
+%found in [\ref{Pollard 2002}].
 
@@
 
-To illustrate that (\ref{eqn:pmf}) is the distribution of $Y$, consider the 
+To show the result in Equation (\ref{eqn:pmf}) consider the 
 process $\mathbf{X} = \{X(k) : k \in \mathbb{Z}_{\ge 0} \}$ with $X(0)=0$ and
 \begin{equation*} \label{eqn:proc}
-X_{k+1} = X_k + B_{k+1} \left[ k-t < X_k < s \right]
+X_{k+1} = X_k + b_{k+1} \left[ k-t < X_k < s \right].
 \end{equation*}
-where $B_{k+1}$ is once again distributed as Bernoulli($p$) for all $k$ in the domain of $\mathbf{X}$.
-The process can be conceptualized as a series of coin flips that stops when either $s$ heads or $t$ tails are reached.
-At each step a coin is flipped.
-If it is heads, the process advances one diagonally in the positive horizontal and vertical direction.
-Otherwise, it advances in the positive horizontal direction only.
-The process stops, $X_k$'s value does not change, after either $X_k = s$ or $X_k = k-t$.
-One example is given in Fig.
-\ref{fig:kane_viz}.
+As before the $b_{k+1}$'s are distributed as Bernoulli($p$) for all $k$ in the 
+domain of $\mathbf{X}$. The process can be conceptualized as a series of coin 
+flips that stops when either $s$ heads or $t$ tails are reached. At each step 
+a coin is flipped. If it is heads, the process advances one diagonally in the 
+positive horizontal and vertical direction. Otherwise, it advances in the 
+positive horizontal direction only. When the process stops, $X_k$'s and 
+remains either $X_k = s$ or $X_k = k-t$. One example is given in 
+Fig. \ref{fig:kane_viz}.
 
@@
 \begin{proof}
-The proof will proceed in two parts.
-First, a combinatorial justification will be given for the probability mass value on each element of the support.
-Second, it will be shown that the sum of the masses over the support sums to one.
+%The proof will proceed in two parts. First, a combinatorial justification 
+%will be given for the probability mass value on each element of the support. 
+%Second, it will be shown that the sum of the masses over the support sums to 
+%one.
 
 The probability that a given realization of $\mathbf{X}$ reaches $s$ at 
-time $k$ is  the probability that, at time $k-1$ there are $s-1$ successful 
-coin flips and $k-s$ unsuccessful coin flips times the probability of a 
-successful coin flip at time $k$ (\ref{eqn:N}). The probability $X_k = k-t$ 
-is the probability that, at time $k-1$ there are $k-t$ successful coin flips 
-and $t-1$ unsuccessful coin flips times the probability of an unsuccessful 
-coin flip at time $k$ (\ref{eqn:R}).  Next, we need to show that the sum
+the $k$th outcome is the probability that, at time $k-1$ there are $s-1$ 
+successful outcomes and $k-s$ unsuccessful outcomes multiplied by 
+the probability of a 
+success at time $k$ (\ref{eqn:N}). The probability $X_k = k-t$ 
+is the probability that, at outcome $k-1$ there are $k-t$ successful outcomes 
+and $t-1$ unsuccessful outcomes multiplied by the probability of an 
+unsuccessful outcome at time $k$ (\ref{eqn:R}).  Finally, we need to show that 
+the sum
 \begin{equation} \label{eqn:sum_proof}
@@
 also one minus the c.d.f. of the negative binomial distribution. This
-function satisfies $I_p(s, t) = 1-I_{1-p}(t, s)$.  Then 
+function satisfies $I_p(s, t) = 1-I_{1-p}(t, s)$.  Then it can be proven
+that \ref{eqn:pmf} is a valid probability mass by showing that the
+probability mass function sums to one.
 \begin{align*}
@@
 \end{proof}
-Proving that \ref{eqn:pmf} is a valid probability mass function.
 
-From (\ref{eqn:pmf}) one can see that the mass at any point on the support 
-of the distribution may come from either $X_k = s$, captured in the 
-function $S$, or $X_k = k-t$, captured in the function $T$. The support of 
-$S$ is defined by $s \leq S \leq s+t-1$ and the support of $T$ is defined 
-by $t \leq T \leq s+t-1$.
+\begin{prop} Let $S$ be distributed SNB with parameters $p$, $s$, and $t$.
+Then the moment generating function (MGF) of $S$ is
+\begin{equation} \label{eqn:mgf}
+\mathbb{E}~e^{xS} = \frac{p e^x}{1 - qe^x} I_{qe^x} (s, t) + \frac{qe^x}{1-pe^x} I_{pe^x}(s, t)
+\end{equation}
+for $q = 1-p$ when $x \leq -\log(q)$ and $x \leq -\log(p)$.
+\end{prop}
+\begin{proof}
+By definition, the MFG of the SNB is:
+\begin{equation*}
+\mathbb{E}~e^{xS} = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s q^{k-s} e^{kx} 
+  + \sum_{k=t}^{s+t-1} {k-1 \choose k-t} p^{k-t} q^t e^{kx}
+\end{equation*}
+and this can be rewritten as:
+\begin{equation} \label{eqn:first_sum}
+\mathbb{E}~e^{xS} = \sum_{k=s}^{s+t-1}{k-1 \choose s-1} (pe)^{sx} (qe^x)^{k-s} 
+  + \sum_{k=t}^{s+t-1}{k-1 \choose k-t} (qe^x)^t (pe^x)^{k-t}.
+\end{equation}
+Taking the first summation in Equation \ref{eqn:first_sum}:
+\begin{align*}
+\sum_{k=s}^{s+t-1}{k-1 \choose s-1} (pe)^{sx} (qe^x)^{k-s} &= 
+  \left(\frac{pe^x}{1 - qe^x}\right)^s \ \ \sum_{k=s}^{s+t-1} {k-1 \choose s-1} 
+    (qe^x)^{k-s} (1-qe^x)^s) \\
+  &= \left(\frac{pe^x}{1 - qe^x}\right)^s I_{qe^x}(s, t).
+\end{align*}
+Since the incomplete beta function has support on zero to one 
+$qe^x \leq 1$. This implies that $x \leq -\log(q)$.
+
+A similar expression can be derived using the same calculation with 
+the constraint that $x \leq -\log(p)$. The result
+follows from the afore mentioned property of the regularized incomplete
+beta function.
+\end{proof}
 
@@
 The SNB is related to the negative binomial distribution. Specifically, if 
-$t$ is infinitely large then the $Y-s$ has a negative binomial distribution 
+$t$ is large then the $Y-s$ has a negative binomial distribution 
 with
@@
 Similar derivations of the closely-related maximum negative binomial discrete 
-distributions also appear in~[\ref{Zhang 2000}, \ref{Zelterman 2004}]. However,
-where the maximum negative binomial is the smallest number of coin-flips 
-to observe at least $c$ successes {\em and} $c$ failures, the snb give
+distributions also appear in~[\ref{Zhang 2000}, \ref{Zelterman 2004}]. 
+The maximum negative binomial is the smallest number of outcomes necessary
+to observe at least $c$ successes {\em and} $c$ failures, but the SNB give
 the number of coin flips to observe {\em either} $s$ heads or $t$ tails.
@@
 
-\section{The Posterior Distribution}
+\section{A Bayesian Analysis}
 
@@
 
-If we assume that $p$ has a Beta distribution, with constant prior parameters $\alpha$ and $\beta$ then we can calculate the closed form posterior 
-distribution of the SNB as
+Let us assume that $p$ has a Beta distribution, with constant prior 
+parameters $\alpha$ and $\beta$. Then the closed form 
+posterior distribution of the SNB is
 \begin{align}
-\mathbb{P} \{Y = k | \mathbf{B}, \alpha, \beta \} &= {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} \{s \leq k \leq s+k-1\} + \nonumber \\
-& {k-1 \choose k-t} \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} \{t \leq k \leq s+k-1\}
+\mathbb{P} \left[Y = k | \mathbf{b}, \alpha, \beta \right] &= 
+  {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} 
+  [s \leq k \leq s+k-1] + \nonumber \\
+  & {k-1 \choose k-t} 
+  \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} 
+  [t \leq k \leq s+k-1]
 \end{align}
-where $B$ is the Beta function and $\mathbf{B}$ is the sequence of Bernoulli data.
+where $B$ is the Beta function and $\mathbf{b}$ is the sequence of Bernoulli data.
 
-When $p$ is known, the probability mass function (PMF) of the number of steps 
-until the process is stopped is given by (\ref{eqn:pmf}). However, many times 
-this is not the case. In Phase II clinical trials for example patients are 
-often enrolled sequentially until there are enough patients to guarantee the 
-desired level of power or until a specified number of adverse events are 
-recorded. In cases such as these the endpoints $s$ and $t$ are known and we 
-would like to be able to estimate $p$ either during enrollment to provide odds 
-of successful enrollment or after an endpoint has been reached to determine 
-the amount of uncertainty in the endpoint. 
+%When $p$ is known, the probability mass function (PMF) of the number of steps 
+%until the process is stopped is given by (\ref{eqn:pmf}). However, many times 
+%this is not the case. 
 %To do this, we will assume that $p$'s probability in the absence of 
@@
 The posterior PMF of the Stopped Negative Binomial distribution with a Beta($\alpha$, $\beta$) prior is:
-\begin{align}
-f(k | s, t, \alpha, \beta) &= {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} \{s \leq k \leq s+k-1\} + \nonumber \\
-& {k-1 \choose k-t} \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} \{t \leq k \leq s+k-1\}
+\begin{align} \label{eqn:posterior}
+f(k | s, t, \alpha, \beta) &= 
+  {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)}     [s \leq k \leq s+k-1] + \nonumber \\
+  & {k-1 \choose k-t} 
+    \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} 
+    [t \leq k \leq s+k-1]
 \end{align}
@@
 \begin{proof}
-For notational simplicity, assume that $s,t \leq k \leq s+k-1$. When this is not the case appropriate terms should be removed as dictated by the indicator functions.
+For notational simplicity, assume that $s,t \leq k \leq s+t-1$. When this is not the case appropriate terms should be removed as dictated by the indicator functions.
 \begin{align*}
@@
 
-\section{A Bayesian Exploration Simon Two-stage Optimal Design}
+Equation \ref{eqn:posterior} suggests an alternative construction for 
+Phase II clinical trials.  Patients could be 
+enrolled sequentially until there are enough patients to guarantee the 
+desired level of power or until a specified number of adverse events are 
+recorded. In cases such as these the endpoints $s$ and $t$ are known and 
+new outcomes can be incorporated into the estimate of
+$p$ to provide updated odds of successful enrollment. Later it will be shown 
+that estimates of the distribution of $p$ can even be used after an endpoint 
+has been reached to determine the amount of uncertainty in the endpoint.
 
-Returning to the breast cancer trial introduced in the first section, consider
-the first stage, having a sampel size of 12. A
-response rate of 2/12 stage corresponds to an individual 
-probability of responding of $p=0.1667$. The standard deviation of this 
-probability is $\sqrt{p(1-p)/12}=0.1076$. If we were to use the Central 
-Limit Theorem then the 95\% confidence interval around the mean would 
-be $[-0.0442, 0.3776]$. While this approach is problematic since the number 
-of samples (12 in this case) is small it does indicate how much uncertainty 
-is built into the first stage. If we consider samples from both stages then 
-the standard deviation drops to 0.0630 and the 95\% confidence interval 
-would be $[0.0432, 0.2901]$.
+\section{A Bayesian Design for Clinical Trials}
 
-These calculations expose two difficulties with the Simon Two-stage Optimal 
-Design. First, success in the first stage is not a good indicator of success 
+%Returning to the breast cancer trial introduced in the first section, consider
+%the first stage, having a sample size of 12. A
+%response rate of 2/12 stage corresponds to an individual 
+%probability of responding of $p=0.1667$. The standard deviation of this 
+%probability is $\sqrt{p(1-p)/12}=0.1076$. If we were to use the Central 
+%Limit Theorem then the 95\% confidence interval around the mean would 
+%be $[-0.0442, 0.3776]$. While this approach is problematic since the number 
+%of samples (12 in this case) is small it does indicate how much uncertainty 
+%is built into the first stage. If we consider samples from both stages then 
+%the standard deviation drops to 0.0630 and the 95\% confidence interval 
+%would be $[0.0432, 0.2901]$.
+
+The Simon Two-stage Optimal Design faces two problems. 
+First, success in the first stage is not a good indicator of success 
 in the second stage and at the same time failure in first stage in no way 
@@
 
-\item[\bibref{Pollard 2002}]
-Pollard D. {\it A User's Guide to Measure Theoretic Probability}, 
-Cambridge University Press, 2002.
+%\item[\bibref{Pollard 2002}]
+%Pollard D. {\it A User's Guide to Measure Theoretic Probability}, 
+%Cambridge University Press, 2002.
 

### 2015-10-07 paper-1/main.tex
Changes made based on Dan's and Michelle's suggestions.

@@
 \noindent${}^*$Email address for author: {\tt
-michael.kane@yale.edu}.  This research was supported by grants 
-	R01CA131301, R01CA157749, R01CA148996, R01CA168733 awarded by the National Cancer Institute, and support from the Yale Comprehensive Cancer Center.
+michael.kane@yale.edu}.  
+This research was supported by grants R01CA131301, R01CA157749, R01CA148996, 
+R01CA168733, and PC50CA196530 awarded by the National Cancer Institute, and 
+support from the Yale Comprehensive Cancer Center.
 
@@
 
-In the Simon 2-stage Phase II clinical trial [\ref{Simon 1989}], a small 
-number of patients are enrolled onto a therapeutic 
-protocol for an initial stage. For each patient, we assess either success 
-or failure according to a 
-clinically useful criteria. Success might include tumor response to treatment
-or perhaps that the response is maintained for a minimum specified duration.
-In addition to failing to achieve the primary endpoint, treatment failure 
-might include adverse reaction to the drug or withdrawal due to toxicity. We 
-use a large number of these early successes as evidence for whether to treat
-additional patients in a second stage. A statistically small number of 
-successes in Stage I would lead us to terminate the trial without entering 
-Stage II.
+%In the Simon 2-stage Phase II clinical trial [\ref{Simon 1989}], a small 
+%number of patients are enrolled onto a therapeutic 
+%protocol for an initial stage. For each patient, we assess either success 
+%or failure according to a 
+%clinically useful criteria. Success might include tumor response to treatment
+%or perhaps that the response is maintained for a minimum specified duration.
+%In addition to failing to achieve the primary endpoint, treatment failure 
+%might include adverse reaction to the drug or withdrawal due to toxicity. We 
+%use a large number of these early successes as evidence for whether to treat
+%additional patients in a second stage. A statistically small number of 
+%successes in Stage I would lead us to terminate the trial without entering 
+%Stage II.
 
-Consider the Phase II clinical trial that investigated the efficacy of 
-iniparib in patients with breast cancer gene-associated (BRCA) ovarian 
-cancer [\ref{Sanofi 2013}]. In the first stage 12 patients were treated with 
-iniparib (BSI-201/SAR240550). The endpoint was defined by Response Evaluation 
-Criteria in Solid Tumor (RECIST). If at least two Stage I patients succeeded,
-then the trial would have proceed to the next stage where additional patients 
-were treated. If fewer than two of the 12 Stage I patients is successful then 
-the trial would have terminated.
+Consider the design for a Phase II clinical trial to investigated the 
+efficacy of iniparib in patients with breast cancer gene-associated (BRCA) 
+ovarian cancer [\ref{Sanofi 2013}]. In the first stage, 12 patients were 
+treated with iniparib (BSI-201/SAR240550). The endpoint was defined by 
+Response Evaluation Criteria in Solid Tumor (RECIST). If at least two of
+these patients respond favorably, then the trial will proceed to the next 
+stage where additional patients are able to be treated. If fewer than two 
+respond then the trial would be terminated.
 
-Our goal is to describe the distribution of the sample size of the design for 
+The maxiumu sample size is 12 but the number of patients necessary to 
+reach and endpoint can be less. Our goal is to describe the distribution of 
+the enrollment size of the design for 
 planning purposes and to use this distribution to estimate the
@@
 the trial is a sequence of these random variables that stops when either a 
-specified number of success or failures is reached. As an example, suppose in 
-the previously described trial, two successes were reached after enrolling 10 
+specified number of success or failures is reached. As in the 
+previous example, two successes were reached after enrolling 10 
 patients (one in the third step and one at the $10^{th}$). The trial is 
-visualized in Fig \ref{fig:kane_viz}. The vertical axis denotes the number of 
+visualized in Fig.~\ref{fig:kane_viz}. The vertical axis denotes the number of 
 successful outcomes. The horizontal axis denotes the number of patients that 
 have been enrolled. The horizontal and vertical line segments represent 
-endpoints for the trial.
+boundayr endpoints for the trial.
 
@@
 \begin{equation} \label{eqn:pmf}
-\mathbb{P}[Y=k] = S(k, p, s)  [s \leq k] + T(k, p, t) [t \leq k ] 
+\mathbb{P}[Y=k] = S(k, p, s)\left[s \leq k\right] + 
+  S(k, 1-p, t) \left[ t \leq k \right]
 \end{equation}
@@
 \end{equation}
-is the negative binomial probability of $s$ successful outcomes and
-\begin{equation} \label{eqn:R}
-T(k, p, t) = {k-1 \choose k-t} p^{k-t} (1-p)^t
-\end{equation}
-is the negative binomial probability of $t$ unsuccessful outcomes. 
-From (\ref{eqn:pmf}) one can see that the mass at any point on the support 
-of the distribution may come from either $X_k = s$, captured in the 
-function $S$, or $X_k = k-t$, captured in the function $T$. The support of 
+$S$ is the negative binomial probability truncated to a subset of the support 
+- in this case, $s$ and $t$ respectively.
+From Eqn.~\ref{eqn:pmf} one can see that the mass at 
+any point on the support of the distribution may come from either 
+$X_k = s$, captured in the function $S$, or $X_k = k-t$, captured in the 
+function $T$. The support of 
 $S$ is defined by $s \leq S \leq s+t-1$ and the support of $T$ is defined 
@@
 
-To show the result in Equation (\ref{eqn:pmf}) consider the 
+To show the result in Eqn.~\ref{eqn:pmf} consider the 
 process $\mathbf{X} = \{X(k) : k \in \mathbb{Z}_{\ge 0} \}$ with $X(0)=0$ and
@@
 positive horizontal and vertical direction. Otherwise, it advances in the 
-positive horizontal direction only. When the process stops, $X_k$'s and 
-remains either $X_k = s$ or $X_k = k-t$. One example is given in 
-Fig. \ref{fig:kane_viz}.
+positive horizontal direction only. When the process stops
+either $X_k = s$ or $X_k = k-t$. One example is given in 
+Fig.~\ref{fig:kane_viz}.
 
@@
 \begin{proof}
-By definition, the MFG of the SNB is:
+By definition, the MGF of the SNB is:
 \begin{equation*}
@@
 
-Equation \ref{eqn:posterior} suggests an alternative construction for 
+Eqn.~\ref{eqn:posterior} suggests an alternative construction for 
 Phase II clinical trials.  Patients could be 

### 2015-10-07 paper-1/main.tex
Updated the kplot based on Dan's feedback.

@@
 
-The maxiumu sample size is 12 but the number of patients necessary to 
+The maximum sample size is 12 but the number of patients necessary to 
 reach and endpoint can be less. Our goal is to describe the distribution of 
@@
 have been enrolled. The horizontal and vertical line segments represent 
-boundayr endpoints for the trial.
+boundary endpoints for the trial.
 
@@
 To show the result in Eqn.~\ref{eqn:pmf} consider the 
-process $\mathbf{X} = \{X(k) : k \in \mathbb{Z}_{\ge 0} \}$ with $X(0)=0$ and
+process $\mathbf{X} = \left[X(k) : k \in \mathbb{Z}_{\ge 0} \right]$ 
+with $X(0)=0$ and
 \begin{equation*} \label{eqn:proc}
@@
 and $t-1$ unsuccessful outcomes multiplied by the probability of an 
-unsuccessful outcome at time $k$ (\ref{eqn:R}).  Finally, we need to show that 
+unsuccessful outcome at time $k$.  Finally, we need to show that 
 the sum

### 2015-10-07 paper-1/script.r

@@ -24,7 +24,60 @@ p1 <- qplot(k, X, data=X1, geom="line") +
   geom_segment(x=0, y=s, xend=t, yend=s, color="red") + 
   geom_segment(x=t, y=0, xend=t, yend=(s-1), color="green")
 
-ggsave("BernoulliProcess.pdf", p1, width=7, height=5)
+stairs = function (p, xstart, xend) {
+    x = c(xstart, rep((xstart + 1):xend, each = 2))
+    y = rep(0:(xend - xstart), each = 2)
+    y = y[-length(y)]
+    for (i in 1:(length(x) - 1)) {
+        p = p + geom_segment(x = x[i], y = y[i], xend = x[i + 
+            1], yend = y[i + 1], color = "green", linetype=1)
+    }
+    p
+}
+
+kplot = function(flips, s, t) 
+{
+  if (!is.list(flips)) {
+    d = flips_to_kplot_df(flips)
+    p = qplot(k, path, data = d, geom = "line") + 
+      scale_x_continuous(breaks = 0:(t + s), limits = c(0, t + s)) + 
+      scale_y_continuous(breaks = 0:s, 
+      limits = c(0, s)) + 
+      geom_segment(x = s, y = s, xend = (t + s - 1), yend = s, color = "red", 
+                   linetype=1)
+    p = stairs(p, t, s + t - 1)
+  } else {
+    flip_set = lapply(flips, flips_to_kplot_df)
+    for (i in 1:length(flip_set)) {
+      flip_set[[i]]$num = as.factor(i)
+      flip_set[[i]]$k = jitter(flip_set[[i]]$k)
+      flip_set[[i]]$k[flip_set[[i]]$k < 0] = 0
+    }
+    d = Reduce(rbind, flip_set)[, -(4:5)]
+    p = qplot(k, path, data = d, geom = "path", group = num) + 
+        scale_x_continuous(breaks = 0:(t + s), limits = c(0, 
+            t + s)) + scale_y_continuous(breaks = 0:s, limits = c(0, 
+        s)) + geom_segment(x = s, y = s, xend = (t + s - 
+        1), yend = s, color = "red")
+    p = stairs(p, t, s + t - 1)
+    p
+  }
+  p
+}
+
+flips_to_kplot_df = function (flips) {
[…]

### 2015-10-12 paper-1/main.tex
Changes based on Dan's feedback.

@@
 sampling rules common in early-stage clinical trials. We derive the 
-distribution of the smallest number of independent Bernoulli($p$) trials 
+distribution of the smallest number of independent and identically 
+distributed Bernoulli trials 
 needed in order to observe either $s$ successes or $t$ failures. The closed 
@@
 
-Consider the design for a Phase II clinical trial to investigated the 
+Consider the design for a Phase II clinical trial that investigates the 
 efficacy of iniparib in patients with breast cancer gene-associated (BRCA) 
 ovarian cancer [\ref{Sanofi 2013}]. In the first stage, 12 patients were 
-treated with iniparib (BSI-201/SAR240550). The endpoint was defined by 
-Response Evaluation Criteria in Solid Tumor (RECIST). If at least two of
+treated with iniparib (BSI-201/SAR240550). The outcome was defined by 
+Response Evaluation Criteria in Solid Tumor (RECIST). There were two 
+endpoints: terminate or enroll additional patients at the end of this 
+stage. If at least two of
 these patients respond favorably, then the trial will proceed to the next 
-stage where additional patients are able to be treated. If fewer than two 
-respond then the trial would be terminated.
+stage where additional patients were able to be treated. If fewer than two 
+respond then the trial would have been terminated.
 
-The maximum sample size is 12 but the number of patients necessary to 
-reach and endpoint can be less. Our goal is to describe the distribution of 
+The maximum sample size was 12 but the number of patients necessary to 
+reach any endpoint could be less. Our goal is to describe the distribution of 
 the enrollment size of the design for 
@@
 before the next one enters the trial. In the present example, observing two 
-successful patients allows us to proceed to the next stage. Similarly 11 
+successful patients allows us reach one endpoint so the sample required 
+could be as small as two. So, 12 might not be necessary. Similarly 11 
 observed treatment failures also ends the stage. This sampling mechanism, in 
@@
 the trial is a sequence of these random variables that stops when either a 
-specified number of success or failures is reached. As in the 
-previous example, two successes were reached after enrolling 10 
-patients (one in the third step and one at the $10^{th}$). The trial is 
-visualized in Fig.~\ref{fig:kane_viz}. The vertical axis denotes the number of 
-successful outcomes. The horizontal axis denotes the number of patients that 
-have been enrolled. The horizontal and vertical line segments represent 
-boundary endpoints for the trial.
+specified number of success or failures is reached. In the 
+previous example suppose two successes were reached after enrolling 10 
+patients (one in the third step and one at the $10^{th}$). The sample
+path is illustrated 
+in Fig.~\ref{fig:kane_viz}. The vertical axis denotes the number of 
+successful outcomes. The horizontal axis counts the number of patients that 
+have been enrolled. The horizontal and vertical boundaries represent 
+endpoints for the trial.
 
@@
 Let $\,b_1, b_2, \ldots \,$ denote a sequence of independent, identically 
-distributed, Bernoulli random variables with $\mathbb{P}\{b_i=1\}=p$, for 
+distributed, Bernoulli random variables with $\mathbb{E}I_{\{b_i=1\}}=p$ and
+$\mathbb{E}I_{\{b_i = 0\}} = 1-p$, for 
 probability parameter $0\leq p \leq 1$. In the clinical trial setting 
-$\,b_i = 0$ corresponds to an unsuccessful outcome.  Let $s$ and $t$ be 
+$\,b_i = 1$ corresponds to a successful outcome.  Let $s$ and $t$ be 
 positive integers.  Define the stopped negative binomial (SNB) random 
@@
 integer value such that $\,\{b_1, \ldots , b_Y\}\,$ contains {\em either} 
-$\,s\,$ successes {\em or} $\,t\,$ failures. The distribution of $\,Y\,$ has 
-support on integer values in the range 
-\begin{equation*}                                     %   (3)
+$\,s\,$ successes {\em or} $\,t\,$ failures. That is, either 
+$\sum_i^Y b_i s$ or else $\sum_i^Y 1-b_i = t$.
+
+The distribution of $\,Y\,$ has support on integer values in the range 
+\begin{equation*}               
      \min(s,t) \leq \; Y \;\leq s+t-1  \label{range.y.eq}
@@
 \begin{equation} \label{eqn:pmf}
-\mathbb{P}[Y=k] = S(k, p, s)\left[s \leq k\right] + 
-  S(k, 1-p, t) \left[ t \leq k \right]
+\mathbb{E}\  I_{\{Y=k\}} = S(k, p, s) I_{\{s \leq k\}} + 
+  S(k, 1-p, t) I_{\{ t \leq k \}}
 \end{equation}
-where 
+where $I_{\{f\}}$ is one if $f$ is true and zero otherwise and
 \begin{equation} \label{eqn:N}
@@
 \end{equation}
-$S$ is the negative binomial probability truncated to a subset of the support 
-- in this case, $s$ and $t$ respectively.
-From Eqn.~\ref{eqn:pmf} one can see that the mass at 
-any point on the support of the distribution may come from either 
-$X_k = s$, captured in the function $S$, or $X_k = k-t$, captured in the 
-function $T$. The support of 
-$S$ is defined by $s \leq S \leq s+t-1$ and the support of $T$ is defined 
-by $t \leq T \leq s+t-1$.
+$S$ is the negative binomial probability 
+In~\ref{eqn:pmf} we can see that the mass at 
+any point on the support of the distribution may come from $S(p)$; $S(1-p)$;
+or both of their probabilities.
 
@@
 To show the result in Eqn.~\ref{eqn:pmf} consider the 
-process $\mathbf{X} = \left[X(k) : k \in \mathbb{Z}_{\ge 0} \right]$ 
+process $\mathbf{X} = \left\{X(k) : k = 0,1,... \right\}$ 
 with $X(0)=0$ and
 \begin{equation*} \label{eqn:proc}
-X_{k+1} = X_k + b_{k+1} \left[ k-t < X_k < s \right].
+X_{k+1} = X_k + b_{k+1} I_{\{ k-t < X_k < s\}}.
 \end{equation*}
-As before the $b_{k+1}$'s are distributed as Bernoulli($p$) for all $k$ in the 
-domain of $\mathbf{X}$. The process can be conceptualized as a series of coin 
-flips that stops when either $s$ heads or $t$ tails are reached. At each step 
-a coin is flipped. If it is heads, the process advances one diagonally in the 
-positive horizontal and vertical direction. Otherwise, it advances in the 
-positive horizontal direction only. When the process stops
-either $X_k = s$ or $X_k = k-t$. One example is given in 
-Fig.~\ref{fig:kane_viz}.
+The process can be conceptualized as a series of coin 
+flips that stops when either $s$ successes or $t$ failures are reached. 
+At each step a coin is flipped. If it is heads, the process advances 
+one diagonally in the 
+positive horizontal and vertical direction in Fig.~\ref{fig:kane_viz}. 
+Otherwise, it advances in the positive horizontal direction only. When the 
+process stops either $X_k = s$ or $X_k = k-t$.
 
 \begin{prop}
-The distribution of the stopping time $\argminl_k [X_k \geq s, X_k \leq k-t]$ 
+The distribution of the stopping time 
+$\argminl_k \left\{ X_k \geq s \cup X_k \leq k-t \right\}$ 
 is equal to the SNB and is given in (\ref{eqn:pmf}). 
@@
 the sum
-\begin{equation} \label{eqn:sum_proof}
-R = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s} + \sum_{k=t}^{s+t-1} {k-1 \choose k-t} p^{k-t} (1-p)^t
-\end{equation}
+\begin{align} \label{eqn:sum_proof}
+R &= \sum_{k=s}^{s+t-1} S(k, p, s) + \sum_{k=t}^{s+t-1} S(k, 1-p, t) \\
+  &= \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s} + \sum_{k=t}^{s+t-1} {k-1 \choose k-t} p^{k-t} (1-p)^t
+\end{align}
 is equal to one.
-First, substitute $i=k-s$ in the first summation and
+
+Substitute $i=k-s$ in the first summation and
 $j=k-t$ in the second. 
-\begin{equation*}
-{j+s-1 \choose s-1} = {j+s-1 \choose j}
-\end{equation*}
-Then %(\ref{eqn:sum_proof}) becomes:
+%\begin{equation*}
+%{j+s-1 \choose s-1} = {j+s-1 \choose j}
+%\end{equation*}
+%Then %(\ref{eqn:sum_proof}) becomes:
 %\begin{equation*}
@@
 %\end{equation*}
-$R$ can be written
+Then $R$ can be written
 as the cumulative distribution function of two 
@@
 
-Let $I_p(s, t)$ be the {\em regularized incomplete beta function}, which is 
+Let $\mathcal{I}_p(s, t)$ be the {\em regularized incomplete beta function}, which is 
 also one minus the c.d.f. of the negative binomial distribution. This
-function satisfies $I_p(s, t) = 1-I_{1-p}(t, s)$.  Then it can be proven
+function satisfies $\mathcal{I}_p(s, t) = 1-\mathcal{I}_{1-p}(t, s)$.  
+Then it can be proven
 that \ref{eqn:pmf} is a valid probability mass by showing that the
@@
 \sum_{j=0}^{s-1}  {j+t-1 \choose j} p^j  (1-p)^t \\
-   &= 1-I_p(s, t) + 1 - I_{1-p}(t, s) \\
+   &= 1-\mathcal{I}_p(s, t) + 1 - \mathcal{I}_{1-p}(t, s) \\
    &= 1. 
@@
 \begin{equation} \label{eqn:mgf}
-\mathbb{E}~e^{xS} = \frac{p e^x}{1 - qe^x} I_{qe^x} (s, t) + \frac{qe^x}{1-pe^x} I_{pe^x}(s, t)
+\mathbb{E}~e^{xS} = \frac{p e^x}{1 - qe^x} \mathcal{I}_{qe^x} (s, t) + 
+  \frac{qe^x}{1-pe^x} \mathcal{I}_{pe^x}(s, t)
 \end{equation}
@@
     (qe^x)^{k-s} (1-qe^x)^s) \\
-  &= \left(\frac{pe^x}{1 - qe^x}\right)^s I_{qe^x}(s, t).
+  &= \left(\frac{pe^x}{1 - qe^x}\right)^s \mathcal{I}_{qe^x}(s, t).
 \end{align*}
@@
 \begin{align}
-\mathbb{P} \left[Y = k | \mathbf{b}, \alpha, \beta \right] &= 
+f \left(k | \mathbf{b}, \alpha, \beta \right) &= 
   {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} 
-  [s \leq k \leq s+k-1] + \nonumber \\
+  I_{\{s \leq k \leq s+k-1\}} + \nonumber \\
   & {k-1 \choose k-t} 
   \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} 
-  [t \leq k \leq s+k-1]
+  I_{\{t \leq k \leq s+k-1\}}
 \end{align}
-where $B$ is the Beta function and $\mathbf{b}$ is the sequence of Bernoulli data.
+where $B$ is the Beta function and $\mathbf{b}$ is the realization 
+of Bernoulli data.
 
@@
 f(k | s, t, \alpha, \beta) &= 
-  {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)}     [s \leq k \leq s+k-1] + \nonumber \\
+  {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} 
+    \ I_{\{s \leq k \leq s+k-1\}} \ + \nonumber \\
   & {k-1 \choose k-t} 
     \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} 
-    [t \leq k \leq s+k-1]
+    \ I_{\{t \leq k \leq s+k-1\}}
 \end{align}

### 2015-10-12 paper-1/script.r

@@ -35,17 +35,14 @@ stairs = function (p, xstart, xend) {
     p
 }
 
-kplot = function(flips, s, t) 
-{
+kplot = function(flips, s, t) {
   if (!is.list(flips)) {
     d = flips_to_kplot_df(flips)
     p = qplot(k, path, data = d, geom = "line") + 
       scale_x_continuous(breaks = 0:(t + s), limits = c(0, t + s)) + 
-      scale_y_continuous(breaks = 0:s, 
-      limits = c(0, s)) + 
-      geom_segment(x = s, y = s, xend = (t + s - 1), yend = s, color = "red", 
-                   linetype=1)
-    p = stairs(p, t, s + t - 1)
+      scale_y_continuous(breaks = 0:s, limits=c(0, s+0.15)) + 
+      geom_segment(x=s, y=s, xend=(t+s-1), yend=s, color="green", linetype=1) +
+      geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, col="red")
   } else {
     flip_set = lapply(flips, flips_to_kplot_df)
     for (i in 1:length(flip_set)) {
@@ -55,11 +52,10 @@ kplot = function(flips, s, t)
     }
     d = Reduce(rbind, flip_set)[, -(4:5)]
     p = qplot(k, path, data = d, geom = "path", group = num) + 
-        scale_x_continuous(breaks = 0:(t + s), limits = c(0, 
-            t + s)) + scale_y_continuous(breaks = 0:s, limits = c(0, 
-        s)) + geom_segment(x = s, y = s, xend = (t + s - 
-        1), yend = s, color = "red")
-    p = stairs(p, t, s + t - 1)
+        scale_x_continuous(breaks=0:(t+s), limits = c(0, t+s)) + 
+        geom_segment(x = s, y = s, xend = (t + s - 1), yend = s, 
+                     color = "green") + 
+        geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, col="red")
     p
   }
   p
@@ -88,7 +84,7 @@ X2 <- X[1:(min(which(X$headEnd >= s), which(X$tailEnd >= t))),]
 
 p2 <- ggplot(data=X2) + #right, up, data=X) + #, geom="segment") +
   scale_x_continuous(breaks=0:t, limits=c(0, t)) +
-  scale_y_continuous(breaks=0:s, limits=c(0, s)) +
[…]

### 2016-02-29 paper-1/main.tex
Stat and Prob Letter draft started.

@@
 
-\begin{figure}[ht]
+\begin{figure}[t!]
 \includegraphics[width=\textwidth]{KanePlot.pdf}
@@
 Let $\,b_1, b_2, \ldots \,$ denote a sequence of independent, identically 
-distributed, Bernoulli random variables with $\mathbb{E}I_{\{b_i=1\}}=p$ and
-$\mathbb{E}I_{\{b_i = 0\}} = 1-p$, for 
+distributed, Bernoulli random variables with $\mathbb{P}[b_i=1]=p$ and
+$\mathbb{P}[b_i = 0] = 1-p$, for 
 probability parameter $0\leq p \leq 1$. In the clinical trial setting 
@@
 $\,s\,$ successes {\em or} $\,t\,$ failures. That is, either 
-$\sum_i^Y b_i s$ or else $\sum_i^Y 1-b_i = t$.
+$\sum_i^Y b_i = s$ or $\sum_i^Y 1-b_i = t$.
 
@@
 \begin{equation} \label{eqn:pmf}
-\mathbb{E}\  I_{\{Y=k\}} = S(k, p, s) I_{\{s \leq k\}} + 
-  S(k, 1-p, t) I_{\{ t \leq k \}}
+\mathbb{E}\  I_{\{Y=k\}} = S(k, p, s) \ I_{\{s \leq k\}} + 
+  S(k, 1-p, t) \ I_{\{ t \leq k \}}
 \end{equation}
@@
 \begin{equation*} \label{eqn:proc}
-X_{k+1} = X_k + b_{k+1} I_{\{ k-t < X_k < s\}}.
+X_{k+1} = X_k + b_{k+1} \ I_{\{ k-t < X_k < s\}}.
 \end{equation*}
@@
 also one minus the c.d.f. of the negative binomial distribution. This
-function satisfies $\mathcal{I}_p(s, t) = 1-\mathcal{I}_{1-p}(t, s)$.  
+function satisfies $\mathcal{I}_p(s, t) = 1-\mathcal{I}_{1-p}(t, s)$
+[\ref{abramowitz1964}].
 Then it can be proven
@@
   {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} 
-  I_{\{s \leq k \leq s+k-1\}} + \nonumber \\
+  \ I_{\{s \leq k \leq s+k-1\}} + \nonumber \\
   & {k-1 \choose k-t} 
   \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} 
-  I_{\{t \leq k \leq s+k-1\}}
+  \ I_{\{t \leq k \leq s+k-1\}}
 \end{align}
@@
 \item[\bibref{Kane 2015}]
-Kane MJ. (2015). snb: The Stopped Negative Binomial. R package version 0.1. 2015. Available from \url{https://github.com/kaneplusplus/snb}
+MJ Kane. (2015). snb: The Stopped Negative Binomial. R package version 0.1. 2015. Available from \url{https://github.com/kaneplusplus/snb}
 
@@
 computing. R Foundation for Statistical Computing, Vienna, Austria. 2015. URL \url{http://www.R-project.org/}.
+
+\item[\bibref{abramowitz1964}]
+M Abramowitz, IA Stegun. {\it Handbook of mathematical functions: with 
+formulas, graphs, and mathematical tables}. Courier Corporation. 1964;
+{\bf 55}: 263.
 
@@
 \item[\bibref{Simon 1989}]
-Simon R.  Optimal two-stage designs for Phase II clinical trials. {\it Controlled Clinical Trials\/} 1989; {\bf 10}: 1--10.
+R Simon.  Optimal two-stage designs for Phase II clinical trials. {\it Controlled Clinical Trials\/} 1989; {\bf 10}: 1--10.
 
 \item[\bibref{Uppuluri 1970}] 
-Uppuluri VRR, Blot WJ.
+VRR Uppuluri, WJ Blot.
 ``A probability distribution arising in a riff-shuffle.'' {\it Random Counts  in Scientific Work, 1: Random Counts in Models and Structures}, 1970.
@@
 \item[\bibref{Zhang 2000}]
-Zhang Z, Burtness BA, Zelterman D.  The maximum negative binomial distribution. {\em Journal of Statistical Planning and Inference\/}  2000; {\bf 87}: 1--19.
+Z Zhang, BA Burtness, D Zelterman.  The maximum negative binomial distribution. {\em Journal of Statistical Planning and Inference\/}  2000; {\bf 87}: 1--19.
 

### 2016-02-29 stat-and-prob-letters/snb-paper.tex
Stat and Prob Letter draft started.

--- 2016-02-29 paper-1/main.tex
@@
-%                 Doubly-Stopped Binomial Distribution
-%                        Kane and Zelterman
-\typeout{}\typeout{}\typeout{}\typeout{}
-\typeout{ Two-Stage, Phase II Designs }
-\typeout{}\typeout{}
-\typeout{ Michelle DeVeaux \& Michael Kane \& Dan Zelterman }
-\typeout{}\typeout{}\typeout{}\typeout{}
+\documentclass[review]{elsarticle}
 
-\documentclass[12pt]{article}         %LaTeX preamble
-\parskip   .4em
-\jot  2ex
-\parindent .3in
-\textwidth      6.5in
-\oddsidemargin    0in
-\textheight 8.75in
-\topmargin      -.5in
-\baselineskip    26pt
-\renewcommand{\baselinestretch}{1.5}
-\renewcommand{\arraystretch}{.75}     % vertical spacing in tables
-\usepackage{graphicx}                 % Needed to import pdf files
-\usepackage{adjustbox}
+\usepackage{lineno,hyperref}
+
 \usepackage{amsmath}
@@
 \newcommand*{\argminl}{\argmin\limits}
+\newtheorem{prop}{Proposition}
+
+\modulolinenumbers[5]
+
+\journal{Statistics and Probability Letters}
+
+%%%%%%%%%%%%%%%%%%%%%%%
+%% Elsevier bibliography styles
+%%%%%%%%%%%%%%%%%%%%%%%
+%% To change the style, put a % in front of the second line of the current style and
+%% remove the % from the second line of the style you would like to use.
+%%%%%%%%%%%%%%%%%%%%%%%
+
+%% Numbered
+%\bibliographystyle{model1-num-names}
+
+%% Numbered without titles
+%\bibliographystyle{model1a-num-names}
+
+%% Harvard
+%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
+
+%% Vancouver numbered
+%\usepackage{numcompress}\bibliographystyle{model3-num-names}
+
+%% Vancouver name/year
+%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
+
+%% APA style
+%\bibliographystyle{model5-names}\biboptions{authoryear}
+
+%% AMA style
+%\usepackage{numcompress}\bibliographystyle{model6-num-names}
+
+%% `Elsevier LaTeX' style
+\bibliographystyle{elsarticle-num}
+%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{document}
 
-\newtheorem{prop}{Proposition}
+\begin{frontmatter}
 
-% --------------- New command definitions ----------------------
+\title{A Stopped Negative Binomial Distribution}
+\tnotetext[mytitlenote]{This research was supported by grants R01CA131301, 
+R01CA157749, R01CA148996, R01CA168733, and PC50CA196530 awarded by the 
+National Cancer Institute, and support from the Yale Comprehensive Cancer 
+Center.}
 
-\newcounter{bibno}%            Bibliographic counter
+%% Group authors per affiliation:
+%\author{Elsevier\fnref{myfootnote}}
+%\address{Radarweg 29, Amsterdam}
+%\fntext[myfootnote]{Since 1880.}
 
-\def\waux{1}%                  Write number for the .AUX file
+%% or include affiliations in footnotes:
+\author[mymainaddress]{Michelle DeVeaux}
+\ead{michelle.deveaux@yale.edu}
 
-\newcommand\writelabel[2]%     Write label[#1] and value[#2] to .aux file
-   {%                          General command, called by \Aexer and \Texer
-     \immediate\write\waux%         Write this information to the .aux file:
-          {\noexpand\newlabel{#1}%  Arguments #1, #2, and the page number
-             {{#2}{\thepage}}}%     Warning: the page number will be incorrect
-   }%                               because I use the \immediate command
+\author[mymainaddress]{Michael J. Kane\corref{mycorrespondingauthor}}
+\cortext[mycorrespondingauthor]{Corresponding author}
+\ead{michael.kane@yale.edu}
 
-\newcommand\bibref[1]%              Reference name and my label (#1)
-   {%
-    \addtocounter{bibno}{1}%        Increment the counter
-    \def\biblab{\thebibno}%         name is same as counter with dot
-    \writelabel{#1}{\biblab}%       Write the label to .AUX file 
-    \noindent\biblab.%              Write this exercise name for \item[ ]
-   }%
+\author[mymainaddress]{Daniel Zelterman}
+\ead{daniel.zelterman@yale.edu}
 
-% -------------------  global definitions  ------------------------
+\address[mymainaddress]{Department of Biostatistics\\ School of Epidemiology and Public Health\\ Yale University, New Haven, CT}
 
-\newcommand{\0}{\mbox{\boldmath $0$}}         %  bold zero in math mode
-\newcommand{\bbeta}{\mbox{\boldmath $\beta$}} %  bold beta in math mode
-\newcommand{\bd}{\mbox{\boldmath $d$}}        %  bold d in math mode
-\newcommand{\bg}{\mbox{\boldmath $\gamma$}}   %  bold gamma in math mode
-\newcommand{\blam}{\mbox{\boldmath $\lambda$}}%  bold lambda in math mode
-\newcommand{\bpi}{\mbox{\boldmath $\pi$}}     %  bold pi in math mode
-\newcommand{\bsig}{\mbox{\boldmath $\sigma$}} %  bold sigma in math mode
-\newcommand{\bth}{\mbox{\boldmath$\theta$}}   %  bold theta in math mode
-\newcommand{\dt}{\widetilde{d}}               %  d tilde in math mode
-\newcommand{\m}{\mbox{\boldmath $m$}}         %  bold m in math mode
-\newcommand{\M}{\mbox{\boldmath $M$}}         %  bold M in math mode
-\newcommand{\mi}{{$-$}}                       %  large math minus sign
-\newcommand{\n}{\mbox{\boldmath $n$}}         %  bold n in math mode
-\newcommand{\ndot}{n{\rm\bf .}}               %  n sub-dot
-\newcommand{\Nh}{\widehat{N}}                 %  N-hat
-\newcommand{\nsp}{\hspace*{-.3in}}            %  large negative space
-\newcommand{\p}{\phantom{5}}                  %  space for one digit
-\newcommand{\pih}{\widehat{\pi}}              %  pi-hat
-\newcommand{\pmi}{\phantom{$-$}}              %  space for wide minus
-\newcommand{\ps}{\phantom{$*$}}               %  space for * footnote
-\newcommand{\pc}{{\scriptsize$\bullet$}}      %  plotting dot character
-\newcommand{\q}{\phantom{55}}                 %  space for two digits
-\newcommand{\qs}{\hspace*{.1in}}              %  wide space
-\newcommand{\re}{{\rm e}}                     %  Roman e (...to the power)
-\newcommand{\bt}{\mbox{\boldmath $t$}}        %  bold t in math mode
-\newcommand{\bT}{\mbox{\boldmath $T$}}        %  bold T in math mode
-\newcommand{\U}{\mbox{\boldmath $u$}}         %  bold u in math mode
-\newcommand{\X}{\mbox{\boldmath $X$}}         %  bold X in math mode
-\newcommand{\x}{\mbox{\boldmath $x$}}         %  bold x in math mode
-\newcommand{\Y}{\mbox{\boldmath $Y$}}         %  bold Y in math mode
-\newcommand{\y}{\mbox{\boldmath $y$}}         %  bold y in math mode
-\newcommand{\z}{\mbox{\boldmath $z$}}         %  bold z in math mode
-
-% ------------------- my hyphenation rules ---------------------
-
-\hyphenation{general-iza-tions}
-\hyphenation{distri-bution}
-\hyphenation{Zel-ter-man}
-
-% ---------------  cover page  --------------------------------
-
-\thispagestyle{empty}
-\begin{center}
-\vspace*{\fill}
-
-{\Large\bf A Stopped Negative Binomial Distribution} \\ [1ex]
-
-\vspace*{1.25in}
-\begin{tabular}{l} 
- {\bf Michelle DeVeaux} \\
- {\bf Michael Kane${}^*$} \\
- {\bf Daniel Zelterman}  \\[.1in]
-
- Department of Biostatistics \\
- School of Epidemiology  \\
- \hspace*{.15in} and Public Health  \\
- Yale University       \\
- New Haven, CT  06520  \\[.2in]
-
-\today
-\end{tabular}
-\end{center}
-
-\vspace*{\fill}
-
-\noindent${}^*$Email address for author: {\tt
-michael.kane@yale.edu}.  
-This research was supported by grants R01CA131301, R01CA157749, R01CA148996, 
-R01CA168733, and PC50CA196530 awarded by the National Cancer Institute, and 
-support from the Yale Comprehensive Cancer Center.
-
-% ------------------------ Abstract ----------------------------
-
-\newpage
-\thispagestyle{empty}
-
-\section*    {\bf   Abstract}
-
+\begin{abstract}
 This paper introduces a new discrete distribution suggested by curtailed
@@
 distribution of the smallest number of independent and identically
-distributed Bernoulli trials 
-needed in order to observe either $s$ successes or $t$ failures. The closed 
-form expression for the distribution as well as the compound distribution are 
-derived. Properties of the distribution are shown and discussed. A case study 
-is presented showing how the distribution can be used to monitor sequential 
-enrollment of clinical trials with binary outcomes as well as providing 
-post-hoc analysis of completed trials.
+distributed Bernoulli trials needed in order to observe either $s$ successes 
+or $t$ failures. The closed form expression for the distribution as well as 
+the compound distribution are derived. Properties of the distribution are 
+shown and discussed. A case study is presented showing how the distribution 
+can be used to monitor sequential enrollment of clinical trials with binary 
+outcomes as well as providing post-hoc analysis of completed trials.
+\end{abstract}
 
-\bigskip
+\begin{keyword}
+discrete distribution\sep curtailed sampling
+\end{keyword}
 
-\noindent{\bf Keywords:}  discrete distribution; curtailed sampling
+\end{frontmatter}
 
-% ---------------------------------------------------------------
+\linenumbers
 
-\thispagestyle{empty}
-\setcounter{page}{1}
-
-% --------------------  1. Introduction   ----------------------
-\section            {Introduction and Motivation}
-% --------------------------------------------------------------
-
-%In the Simon 2-stage Phase II clinical trial [\ref{Simon 1989}], a small 
-%number of patients are enrolled onto a therapeutic 
-%protocol for an initial stage. For each patient, we assess either success 
-%or failure according to a 
-%clinically useful criteria. Success might include tumor response to treatment
-%or perhaps that the response is maintained for a minimum specified duration.
-%In addition to failing to achieve the primary endpoint, treatment failure 
-%might include adverse reaction to the drug or withdrawal due to toxicity. We 
-%use a large number of these early successes as evidence for whether to treat
-%additional patients in a second stage. A statistically small number of 
-%successes in Stage I would lead us to terminate the trial without entering 
-%Stage II.
+\section{Introdution and Motivation}
 
@@
 
-%Our goals for this paper are three-fold.
-First, we derive the distribution for the number patients that need to be enrolled in the described sequential trial, which we refer to as the {\em Stopped Negative Binomial} (SNB) distribution.
-Using a sequential trial it is often possible to reduce the number of patients needed when compared to traditional design.
-Understanding of this distribution allows to create better, more {\em enrollment efficient} designs.
-Second, assuming a Beta distribution on the probability of success, we will derive the compound distribution of the SNB.
-Both the success probability and the SNB compound distribution can be updated as the individual outcome of patients are known and, as a result, better estimates of the distribution of patient-enrollment size can be derived as new data are received.
-Third, we will use the derived results to propose an enrollment design framework alternative to the Simon Two-stage Optimal trial design.
-We will show a hypothetical but representative case where the number of enrollments is less than those proposed by the Simon Two-stage and the uncertainty associated with both the individual success outcome as well as the trial is fully characterized.
-
 The next section of this paper introduces our notation and basic results
@@
 
-%Their individual outcomes are summarized as independent Bernoulli, %with either success according to a clinically useful criteria or else %as treatment failure.
-Failure might include adverse reaction to the %drug or withdrawal due to toxicity.
-Successes occur with probability %$\,p\; (0\leq p\leq 1)$.
-
-%More generally, suppose we have a sequence of independent Bernoulli$(p)\,$random variables $\,Z_1, Z_2, \ldots$.
-For positive integers $\,s\,$ and $\,t,$  we study the distribution of the smallest number $\,Y\,$ such that $\,Z_1, Z_2, \ldots, Z_Y\,$ %contains either $\,s\,$ successes of $\,t\,$ failures.
-
-% ---------------------------  2. Notation ----------------------
-\section   {Notation and Summary of Distributional Results}
+\section{Notation and Summary of Distributional Results}
 \label{notation.section}
-% --------------------------------------------------------------
 
@@
 or both of their probabilities.
-
-%It should be noted that
-%this result expressed using the de Finetti where indicator functions and
-%sets are equivalent. A more complete explanation of this equivalence can be 
-%found in [\ref{Pollard 2002}].
-
-% ------------   Section 3: Random Walk  --------------
-%\section {Deriving the SNB with a Random Walk}
-%\label{random.walk.section}
-% --------------------------------------------------------------
 
@@
 $j=k-t$ in the second.
-%\begin{equation*}
-%{j+s-1 \choose s-1} = {j+s-1 \choose j}
-%\end{equation*}
-%Then %(\ref{eqn:sum_proof}) becomes:
-%\begin{equation*}
-%R = \sum_{i=0}^{t-1} {i+s-1 \choose s-1} p^s (1-p)^i +
-%\sum_{j=0}^{s-1} {j+t-1 \choose j} p^j (1-p)^t.
-%\end{equation*}
 Then $R$ can be written
@@
 function satisfies $\mathcal{I}_p(s, t) = 1-\mathcal{I}_{1-p}(t, s)$
-[\ref{abramowitz1964}].
+\citep{Abramowitz1964}.
 Then it can be proven
@@
 
-The probability mass function of $\,Y\,$ has a variety of shapes for different choices of the parameters $(s,\, t,\,p)$.
-These shapes are illustrated in Fig.~\ref{shapes.fig}.
-The SNB is related to the negative binomial distribution. Specifically, if 
-$t$ is large then the $Y-s$ has a negative binomial distribution 
-with
+The probability mass function of $\,Y\,$ has a variety of shapes for different choices of the parameters $(s,\, t,\,p)$.  These shapes are illustrated in 
+Fig.~\ref{shapes.fig}.The SNB is related to the negative binomial 
+distribution. Specifically, if $t$ is large then the $Y-s$ has a negative 
+binomial distribution with
 \begin{equation*}                                    %   (1)   
@@
 For the special case of $\,s=t,$ the distribution of $\,Y\,$ is the
-riff-shuffle, or minimum negative binomial distribution~[\ref{Uppuluri 1970}].
-Similar derivations of the closely-related maximum negative binomial discrete 
-distributions also appear in~[\ref{Zhang 2000}, \ref{Zelterman 2004}]. 
-The maximum negative binomial is the smallest number of outcomes necessary
-to observe at least $c$ successes {\em and} $c$ failures, but the SNB give
-the number of coin flips to observe {\em either} $s$ heads or $t$ tails.
+riff-shuffle, or minimum negative binomial distribution~\citep{Uppuluri1970}.
+Similar derivations of the closely-related maximum negative binomial 
+discretedistributions also appear in~\cite{Zhang 2000}
+and \cite{Zelterman2004}.
+The maximum negative binomial is the smallest number of outcomes necessary to 
+observe at least $c$ successes {\em and} $c$ failures, but the SNB givethe 
+number of coin flips to observe {\em either} $s$ heads or $t$ tails.
 
-%\subsection{Updating the SNB distribution}
+\section{Connection to the Binomial Tail Probability}
 
-%Under the presented random walk construction if the process is observed before reaching one of its endpoints the conditional distribution is a reparametrized SNB distribution whose parameters are determined by $X_k$ and $k$, the process value and time step.
-For example, in the breast cancer trial the distribution of the number of patients is parameterized as $SNB(2, 11, p)$.
-At the third enrollment there is one successful outcome.
-Conditioned on this information, the number of additional patient enrollments is $SNB(1, 9, p)$.
-As new outcomes are collected the conditional distribution can be re-derived to get better estimates of the total patient enrollment.
-Furthermore, if our task is to estimate $p$ along with the patient enrollment distribution, this parameter can also be updated as new data are received.
-
-\section{A Bayesian Analysis}
-
-%The previous section derives a new distribution for a sequential Bernoulli process where the process is stopped when either a total number of successes is reached or a total number of failure is reached.
-The process can be alternatively conceptualize as a series of coin flips that stops when either $s$ heads or $t$ tails are reached.
-At each step a coin is flipped.
-If it is heads then one step is taken in the positive direction on the vertical axis.
-Otherwise, a step is taken in the positive direction on the horizontal axis.
-The walk stops when either $s$ is reached on the vertical axis or $t$ is reached on the horizontal axis.
-A visual example is shown in Fig.
-\ref{fig:zelterman_viz}.
-
-%\begin{figure}[ht]
-%\includegraphics[width=\textwidth]{ZeltermanPlot.pdf}
-%\caption{
-%A realization of $\mathbb{X}$ with $p=0.3$, $s=3$, and $t=10$. An arrow to the right indicates a coin flip of ``tail'' and
-%an arrow in the vertical direction indicates ``head.''
-%}
-%\label{fig:zelterman_viz}
-%\end{figure}
+\section{The Posterior Distribution}
 
@@
 of Bernoulli data.
-
-%When $p$ is known, the probability mass function (PMF) of the number of steps 
-%until the process is stopped is given by (\ref{eqn:pmf}). However, many times 
-%this is not the case. 
-%To do this, we will assume that $p$'s probability in the absence of 
-%available data is $\text{Beta}(1/2, 1/2)$, the Jeffrey's prior. This 
-%distribution was chosen for two reasons. First, it is the conjugate prior 
-%of the (Negative) Binomial distribution and updating the distribution as data 
-%are received is straightforward. Second, it is objective and limits the number 
-%of ``prior coin-flips'' to one. This means that data are more heavily 
-%influence the distributional characteristics of $p$ with more data providing 
-%more proportional influence.
 
@@
 
-\section{A Bayesian Design for Clinical Trials}
-
-%Returning to the breast cancer trial introduced in the first section, consider
-%the first stage, having a sample size of 12. A
-%response rate of 2/12 stage corresponds to an individual 
-%probability of responding of $p=0.1667$. The standard deviation of this 
-%probability is $\sqrt{p(1-p)/12}=0.1076$. If we were to use the Central 
-%Limit Theorem then the 95\% confidence interval around the mean would 
-%be $[-0.0442, 0.3776]$. While this approach is problematic since the number 
-%of samples (12 in this case) is small it does indicate how much uncertainty 
-%is built into the first stage. If we consider samples from both stages then 
-%the standard deviation drops to 0.0630 and the 95\% confidence interval 
-%would be $[0.0432, 0.2901]$.
-
-The Simon Two-stage Optimal Design faces two problems. 
-First, success in the first stage is not a good indicator of success 
-in the second stage and at the same time failure in first stage in no way 
-implies failure in the second stage, especially when $p$ is close to it's 
-threshold value. The variance in the estimate is so large that it is a poor 
-estimator. The second difficulty is that the design is not {\em enrollment 
-efficient}. That is, if the first stage is successful then additional patients 
-are enrolled. Enrollment, especially for certain types of cancer, can be very 
-difficult and can take time to find eligible patients. Furthermore, if 
-all 12 of the patients in the first stage respond positively, should we 
-really need to enroll additional patients?
-
-Properly quantifying uncertainty in the process implies quantifying 
-uncertainty in $p$, which stems from two sources. First is the number of 
-samples we receive. An estimate of $p$ will be based on data and generally 
-the more data we have, the better our estimate. If the data are small, as 
-they often are in clinical trials, the less reliable our estimate is. The 
-second source comes from heterogeneity in the data. While models may assume 
-that the probability of success $p$ is a single value this is often not the 
-case in real-world data. In clinical trials for example an individual's chance 
-of having an adverse outcome may be influenced by any number of underlying 
-factors, which may vary across individual. In these cases $p$ may not be a 
-single numeric value and may be better conceptualized as a realization over 
-an entire distribution. A frequentist approach will not capture this variation.
-
-We would like to be able to estimate $p$ and capture both sources of uncertainty.
-For cases where the process has not yet reached an endpoint, we would like to be able to provide updated estimates of $p$, its uncertainty, and the probabilities of hitting each endpoint.
-If the process has reached an endpoint, we would again like a estimate of $p$ and its uncertainty along with the probability that another endpoint could have been reached, given the data.
-
-\subsection{Quantifying Uncertainty in the Simon Two-stage Optimal Design}
-
-Both of the problems in the Simon Two-stage Optimal Design stem from the fact that it fails to quantify the uncertainty in the estimate of the success probability.
-A framework based on the SNB distribution addresses this failure and
-provides the added features of monitoring the trial as new results are 
-received, providing estimates of how many new enrollees are needed to complete 
-a trial, and quantifying the uncertainty in the success probability.
-
-Let's start by recasting the breast cancer study in the context of the SNB 
-framework, assuming sequential enrollment. The first stage will stop when 
-either two positive responses are reached or 11 negative responses. This 
-corresponds to a SNB distribution with $s=2$ and $t=11$. Next, let's assume 
-that, in the first stage, there were two successes. A second success could 
-occur during the second, through a tenth enrollment. At each of these times we 
-would like to know the estimate for $p$ along with the lower five percent 
-quantile of the distribution.
-
-\begin{figure}[ht]
-\includegraphics[width=\textwidth]{uncertainty.pdf}
-\caption{
-A visualization of the mode and the 5\% quantile of the estimate of the distribution of $p$.
-}
-\label{fig:simon}
-\end{figure}
-
-To better understand how the distributional estimate of $p$ changes in the 
-number of steps needed for two successes data were simulated for each 
-potential stopping point from two to 11. That is, we will fit the
-Beta distribution with the Jeffrey's prior to the sequence of coin flips 
-$\{[1,1], [1, 0, 1], [1, 0, 0, 1], ... \}$ in the support of the SNB with
-$s=2$ and $t=11$.  Fig. \ref{fig:simon} shows the 
-change in the mode of the beta distribution, which is also the Maximum 
-Likelihood Estimator (MLE) as well as the five percent quantile of the 
-distribution.  The graph shows, for example, that if a second success is 
-found with the third enrollment then the MLE estimate of $p$ is 0.5 and with 
-95\% confidence $p$ is at least 0.25. In this case investigators could have 
-moved on to the next stage of the trial with only three enrollments as 
-opposed to 12.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{|c|c|c|} \hline
-{\bf Steps to 2 successes} & {\bf Mode of $p$} & {\bf 5\% quantile} \\ \hline
-2 & 0.7500  & 0.4307 \\ \hline
-3 & 0.5000  & 0.2355 \\ \hline
-4 & 0.3000  & 0.1278 \\ \hline
-5 & 0.1875  & 0.0763 \\ \hline
-6 & 0.1250  & 0.0497 \\ \hline
-7 & 0.0882  & 0.0347 \\ \hline
-8 & 0.0652  & 0.0254 \\ \hline
-9 & 0.0500  & 0.0194 \\ \hline
-10 & 0.0398 & 0.0153 \\ \hline
-11 & 0.0319 & 0.0123 \\ \hline
-\end{tabular} 
-\end{center}
-\caption{
-A table of the mode and the 5\% quantile of the estimate of the distribution of $p$.
-}
-\label{tab:simon}
-\end{table}
-
-Table \ref{tab:simon} shows the actual values used to generate Fig. 
-\ref{fig:simon}. 
-%In the beginning of this section it was shown that, at the 
-%end of stage two, the breast cancer study actually tests whether the 
-%probability of success is greater than 0.0432 with 95\% confidence and 
-%requires 35 patients to do so. With the Stopped Negative Binomial model we 
-%only need six patients for the same level of confidence. This represents a 
-%tremendous amount of savings in time, effort and resources.
-
-\subsection{Trial Monitoring}
-
-Like most Bayesian Adaptive Clinical trials the SNB approach allows for monitoring of trial and can be updated as new results are received without losing 
-power. This is because the object under investigation is not a point estimate 
-of the success probability; it is the uncertainty in the distribution of the 
-success probability. Point estimates of the success probability can be 
-regarded as artifacts of our parameterization of the underlying distribution. 
-Every point estimate calculated is accompanied by the quantification of the 
-uncertainty around it. This estimate and its uncertainty can be calculated at 
-any time and may be useful in evaluating ongoing trials as well as providing a 
-post-hoc evaluation of trials that have been completed.
-
-\begin{figure}[ht]
-\includegraphics[width=\textwidth]{hypo_traj.pdf}
-\caption{
-A visualization of the mode and the 5\% quantile of the estimate of the distribution of $p$ for a hypothetical sequence of patient enrollments.
-}
-\label{fig:hypo}
-\end{figure}
-
-Consider the trial described above. Suppose eight patients were enrolled and 
-there was a single success on the third trial (visualized in Fig. 
-\ref{fig:hypo}). The mode of the beta distribution can only be calculated when 
-there is at least one success. The mean of the distribution can be
-taken otherwise. At the third time step there is an increase in the estimate and the 5\% quantile corresponding to the success.
-
-\begin{figure}[ht]
-\includegraphics[width=\textwidth]{conditional_snb.pdf}
-\caption{
-The conditional Stopped Negative Binomial distribution after eight patients with one success.
-}
-\label{fig:conditional_snb}
-\end{figure}
-
-Fig \ref{fig:hypo} shows that at the eighth enrollment the estimate of $p$ is 
-about 0.19 and, from the lower 5\% quantile, we can see that there is
-a 95\% probability that $p$ is greater than 2.5\%. 
-Fig. \ref{fig:conditional_snb} shows the posterior SNB distribution for this 
-hypothetical trial at the eighth step.  Successes may occur 
-at steps nine through 11 with probability 26.7\%. The remaining probability
-mass corresponds to reaching the $11^{th}$ step without a second success.
-
-\subsection{Post-hoc Evaluation}
-
-Suppose that, in our example trial, a second successful outcome was reached 
-on the $10^{th}$ patient. From the data collected we have shown how to create 
-an estimate for the distribution of $p$. Based on that estimate we may like to 
-see the probability of a different outcome. We may want to answer the
-question, ``What is the probability that we don't get two successful outcomes?''
-We may also ask what is the distribution for the number of patients 
-enrolled for a similar study?
-
-If there are two successful enrollments with the second one occurring 
-on the $10^{th}$ trial the success probability $p$ is distributed as 
-$Beta(2.5, 8.5)$ (assuming the Jeffrey's prior). 
-Using the mode, a point estimate yields $\hat{p} = 0.1667$. The standard 
-deviation of the estimate is 0.1210. The probability of two successes during 
-the trial is 0.6843 and the probability of failure is 0.3157. The distribution 
-of the stopping times, along with contributions from two positive events and 
-less than two positive events is shown below.
-
-\begin{figure}[ht]
-\includegraphics[width=\textwidth]{post_hoc.pdf}
-\caption{
-The post-hoc Stopped Negative Binomial distribution after the trial ends with 10 patient enrollments and two successes.
-}
-\label{fig:post_hoc}
-\end{figure}
-
 \section{Discussion and Conclusion}
@@
 
-%   ------------------------    References     -----------------------
+\section*{References}
 
-% -------------------------  References ---------------------------
-\section*     {\bf References}
-% -----------------------------------------------------------------
-
-\begin{enumerate}
-
-\item[\bibref{Kane 2015}]
-MJ Kane. (2015). snb: The Stopped Negative Binomial. R package version 0.1. 2015. Available from \url{https://github.com/kaneplusplus/snb}
-
-%\item[\bibref{Pollard 2002}]
-%Pollard D. {\it A User's Guide to Measure Theoretic Probability}, 
-%Cambridge University Press, 2002.
-
-\item[\bibref{R Core Team 2015}]
-R Core Team (2015). R: A language and environment for statistical
-computing. R Foundation for Statistical Computing, Vienna, Austria. 2015. URL \url{http://www.R-project.org/}.
-
-\item[\bibref{abramowitz1964}]
-M Abramowitz, IA Stegun. {\it Handbook of mathematical functions: with 
-formulas, graphs, and mathematical tables}. Courier Corporation. 1964;
-{\bf 55}: 263.
-
-\item[\bibref{Sanofi 2013}]
-Sanofi Pharmaceutical Company.
-A Phase 2, Single Arm Study of BSI-201 in Patients With BRCA-1 or BRCA-2 Associated Advanced Epithelial Ovarian, Fallopian Tube, or Primary Peritoneal Cancer.
-In: ClinicalTrials.gov [Internet].
-Bethesda (MD): National Library of Medicine (US).
-2000- [cited 2015 March 27].
-Available from: \url{https://clinicaltrials.gov/ct2/show/study/NCT00677079} Identifier: NCT00677079.
-
-\item[\bibref{Simon 1989}]
-R Simon.  Optimal two-stage designs for Phase II clinical trials. {\it Controlled Clinical Trials\/} 1989; {\bf 10}: 1--10.
-
-\item[\bibref{Uppuluri 1970}] 
-VRR Uppuluri, WJ Blot.
-``A probability distribution arising in a riff-shuffle.'' {\it Random Counts  in Scientific Work, 1: Random Counts in Models and Structures}, 1970.
-G.P.
-Patil (editor), University Park: Pennsylvania State University Press, pp  23--46.
-
-\item[\bibref{Zhang 2000}]
-Z Zhang, BA Burtness, D Zelterman.  The maximum negative binomial distribution. {\em Journal of Statistical Planning and Inference\/}  2000; {\bf 87}: 1--19.
-
-\item[\bibref{Zelterman 2004}]
-D Zelterman. {\it Discrete Distributions: Application in 
-the Health Sciences}, New York: J. Wiley. 2004. xix + 277pp
-
-\end{enumerate}
-
-% ------------------------------------------------------------------
-
-% ------------------------------------------------------------------
-% --------------------   Tables and Figures   ----------------------
-% ------------------------------------------------------------------
-
-% -------------------------------------------------------------------
-
-% -------------------------------------------------------------------
-% ------------------------  FIGURES  --------------------------------
-% -------------------------------------------------------------------
-
-%\newpage 
-% -------------------------------------------------------------------
-
-%\newpage     %  Figure 1: functional forms of the mass function
-
-% -------------------------------------------------------------------
+\bibliography{mybibfile}
 
 \end{document}
-
-% ------------------  end of this file --------------------------

### 2016-02-29 paper-1/script.r

@@ -72,9 +72,6 @@ flips_to_kplot_df = function (flips) {
     d
 }
 
-
-
-
 X$head <- c(0, cumsum(flips))
 X$tail<- c(0, cumsum(!(flips)))
 X$headEnd <- c(X$head[-1], NA)
@@ -82,26 +79,101 @@ X$tailEnd <- c(X$tail[-1], NA)
 
 X2 <- X[1:(min(which(X$headEnd >= s), which(X$tailEnd >= t))),]
 
-p2 <- ggplot(data=X2) + #right, up, data=X) + #, geom="segment") +
-  scale_x_continuous(breaks=0:t, limits=c(0, t)) +
-  scale_y_continuous(breaks=0:s, limits=c(0, s), labels=as.character(0:s)) +
-  geom_segment(mapping=aes(x=tail, y=head, xend=tailEnd,
-    yend=headEnd), arrow=arrow()) +
-  geom_segment(x=0, y=s, xend=t-1, yend=s, color="red") +
-  geom_segment(x=t, y=0, xend=t, yend=s-1, color="green")
+#p2 <- ggplot(data=X2) + #right, up, data=X) + #, geom="segment") +
+#  scale_x_continuous(breaks=0:t, limits=c(0, t)) +
+#  scale_y_continuous(breaks=0:s, limits=c(0, s), labels=as.character(0:s)) +
+#  geom_segment(mapping=aes(x=tail, y=head, xend=tailEnd,
+#    yend=headEnd), arrow=arrow()) +
+#  geom_segment(x=0, y=s, xend=t-1, yend=s, color="red") +
+#  geom_segment(x=t, y=0, xend=t, yend=s-1, color="green")
 
-ggsave("ZeltermanPlot.pdf", p2, width=7, height=5)
-
+#ggsave("ZeltermanPlot.pdf", p2, width=7, height=5)
 
 p3 = kplot( c(rep(0, 3), 1, rep(0, 6), 1), s=2, t=11) + 
   xlab("Number of Patients Enrolled") + ylab("Number of Responders") +
   geom_text(data=NULL, x=7, y=2.1, label="Success Boundary") +
   geom_text(data=NULL, x=12, y=0.5, label="Failure Boundary", 
-            angle=80)
-
-
+            angle=(45+90)/2+9)
 
 ggsave("KanePlot.pdf", p3, width=7, height=5)
+
+# Regions s < t
+s = 3
+t = 5
+flips = rep(0., t)
[…]

### 2016-02-29 stat-and-prob-letters/snb-paper.tex
Initial commit.

@@
 \begin{abstract}
-This paper introduces a new discrete distribution suggested by curtailed
+This paper introduces a novel discrete distribution suggested by curtailed
 sampling rules common in early-stage clinical trials. We derive the
@@
 distributed Bernoulli trials needed in order to observe either $s$ successes 
-or $t$ failures. The closed form expression for the distribution as well as 
-the compound distribution are derived. Properties of the distribution are 
-shown and discussed. A case study is presented showing how the distribution 
-can be used to monitor sequential enrollment of clinical trials with binary 
-outcomes as well as providing post-hoc analysis of completed trials.
+or $t$ failures. The closed-form expression for the distribution as well as 
+its characteristics and properties are explored.
 \end{abstract}
@@
 
-Consider the design for a Phase II clinical trial that investigates the
-efficacy of iniparib in patients with breast cancer gene-associated (BRCA)
-ovarian cancer [\ref{Sanofi 2013}]. In the first stage, 12 patients were
-treated with iniparib (BSI-201/SAR240550). The outcome was defined by
-Response Evaluation Criteria in Solid Tumor (RECIST). There were two
-endpoints: terminate or enroll additional patients at the end of this
-stage. If at least two of
-these patients respond favorably, then the trial will proceed to the next
-stage where additional patients were able to be treated. If fewer than two
-respond then the trial would have been terminated.
+Consider a prototypical Phase II single-arm clinical trial. 12 patients
+are enrolled and treated. If at least 2 of those patients respond
+favorably, then the trial will proceed to the next stage. If fewer than 
+two respond then the trial would have been terminated.
 
-The maximum sample size was 12 but the number of patients necessary to
+The maximum sample size is 12 but the number of patients necessary to
 reach any endpoint could be less. Our goal is to describe the distribution of
-the enrollment size of the design for
-planning purposes and to use this distribution to estimate the
-success probability.  If all 12 patients are enrolled at once,
-as in the classic
+the enrollment size. This distribution will be used for planning purposes. 
+If all 12 patients are enrolled at once, as in the classic
 design, then the sample size is 12. However, in most clinical trials, the
@@
 successful patients allows us reach one endpoint so the sample required
-could be as small as two. So, 12 might not be necessary. Similarly 11
+could be as small as two - 12 might not be necessary. Similarly 11
 observed treatment failures also ends the stage. This sampling mechanism, in
@@
 
-We assume each of the patient outcomes can be modeled as independent,
-identically distributed Bernoulli($p$) random variables. The realization of
-the trial is a sequence of these random variables that stops when either a
+Assume each of patient outcome can be modeled as an independent,
+identically distributed Bernoulli($p$) random variable. The trial is realized
+as a sequence of these random variables that stops when either a
 specified number of success or failures is reached. In the
@@
 \caption{
-A hypothetical realization of the breast cancer trial.
+A hypothetical realization of a trial.
 }
@@
 
-The next section of this paper introduces our notation and basic results
+The rest of this paper derives the distribution for the number of
+patients needed to end a trial and explores the characteristics and
+properties of this distribution.
+The next section introduces our notation and basic results
 including the density of the distribution along with a description of
-it's relation to other distributions. Sections 2 and 3 derive the distribution
-based on a defined Bernoulli process and give some basic properties of the
-distribution. Section 4 derives the compound distribution using a Beta prior.
-Section 5 develops a post-hoc analysis of a completed trial. Section 6 is
-devoted to discussion and conclusions. The simulations, model fitting
-routines, and visualizations presented in this paper were generated using the
-snb package [\ref{Kane 2015}] for the R programming
-environment [\ref{R Core Team 2015}]
+it's relation to other distributions. Sections 2 derives the distribution
+based on a defined Bernoulli process and give some basic properties.
+Section 3 provides a connection to the Binomail tail probability.
+Section 4 derives the posterior distribution using a Beta prior.
+Section 5 provides a brief discussion on the use of the distribution
+in clinical trials along with future avenues for generalization.
 
-\section{Notation and Summary of Distributional Results}
+\section{Distributional Result}
 \label{notation.section}
@@
 $S$ is the negative binomial probability
-In~\ref{eqn:pmf} we can see that the mass at
-any point on the support of the distribution may come from $S(p)$; $S(1-p)$;
-or both of their probabilities.
 
-To show the result in Eqn.~\ref{eqn:pmf} consider the
+To prove the result in Eqn.~\ref{eqn:pmf} consider the
 process $\mathbf{X} = \left\{X(k) : k = 0,1,... \right\}$
@@
 one diagonally in the
-positive horizontal and vertical direction in Fig.~\ref{fig:kane_viz}.
-Otherwise, it advances in the positive horizontal direction only. When the
-process stops either $X_k = s$ or $X_k = k-t$.
+positive horizontal and vertical direction.
+Otherwise, it advances in the positive horizontal direction only. The
+process stops when either $X_k = s$ or $X_k = k-t$.
 
@@
 $\argminl_k \left\{ X_k \geq s \cup X_k \leq k-t \right\}$
-is equal to the SNB and is given in (\ref{eqn:pmf}).
+is equal to Eqn.~\ref{eqn:pmf}.
 \end{prop}
@@
 
-The probability that a given realization of $\mathbf{X}$ reaches $s$ at
+The probability a given realization of $\mathbf{X}$ reaches $s$ at
 the $k$th outcome is the probability that, at time $k-1$ there are $s-1$
@@
 the probability of a
-success at time $k$ (\ref{eqn:N}). The probability $X_k = k-t$
+success at time $k$. 
+\begin{equation} \label{stop_s}
+S(k, p, s) = {k-1 \choose s-1} p^s (1-p)^{k-s}
+\end{equation}
+The probability a given realization reaches $k-t$
 is the probability that, at outcome $k-1$ there are $k-t$ successful outcomes
 and $t-1$ unsuccessful outcomes multiplied by the probability of an
-unsuccessful outcome at time $k$.  Finally, we need to show that
-the sum
+unsuccessful outcome at time $k$.  
+\begin{equation} \label{stop_t}
+S'(k, p, t) = {k-1 \choose k-t} p^{k-t} (1-p)^t
+\end{equation}
+It can be shown that $S(k, p, s) = S'(k, 1-p, s)$ by realizing that 
+${k-1 \choose k-s} = {k-1 \choose s-1}$.
+
+To show that sum of Eqn. \ref{stop_s} and \ref{stop_t} sum to one
+over their support 
 \begin{align} \label{eqn:sum_proof}
@@
 \end{align}
-is equal to one.
-
-Substitute $i=k-s$ in the first summation and
-$j=k-t$ in the second.
+substitute $i=k-s$ in the first summation and $j=k-t$ in the second.
 Then $R$ can be written
@@
 
-Let $\mathcal{I}_p(s, t)$ be the {\em regularized incomplete beta function}, which is
-also one minus the c.d.f. of the negative binomial distribution. This
-function satisfies $\mathcal{I}_p(s, t) = 1-\mathcal{I}_{1-p}(t, s)$
-\citep{Abramowitz1964}.
-Then it can be proven
-that \ref{eqn:pmf} is a valid probability mass by showing that the
-probability mass function sums to one.
+Let $\mathcal{I}_p(s, t)$ be the {\em regularized incomplete beta function}
+and recall this function satisfies 
+$\mathcal{I}_p(s, t) = 1-\mathcal{I}_{1-p}(t, s)$ \citep{Abramowitz1964}.
 \begin{align*}
@@
 \end{proof}
+
+\section{Shape and Basic Properties}
+
+\begin{figure}[p!]
+\begin{center}
+\includegraphics[width=\textwidth]{shapes.pdf}
+\end{center}
+\caption{Different shapes of the SNB distribution with parameters ($s$, $t$, $p$), as given. Red indicates mass contributed by hitting $s$, teal indicates
+mass contributed by hitting $t$. \label{shapes.fig}}
+\end{figure}
+
+The probability mass function of $\,Y\,$ has a variety of shapes for different 
+choices of the parameters $(s,\, t,\,p)$.  These shapes are illustrated in 
+Fig.~\ref{shapes.fig}. The SNB is related to the negative binomial 
+distribution. Specifically, if $t$ is large then the $Y-s$ has a negative 
+binomial distribution with
+\begin{equation*}                                    %   (1)   
+\mathbb{P}[Y=s+j]        \label{nb1.eq}          
+  = {{s+j-1}\choose{s-1}} p^s (1-p)^j
+\end{equation*}
+for $\,j=0, 1,\ldots\,$. A similar statement can be made when $s$ is large
+and $t$ is small.
+
+For the special case of $\,s=t,$ the distribution of $\,Y\,$ is the
+riff-shuffle, or minimum negative binomial distribution~\citep{Uppuluri1970}.
+Similar derivations of the closely-related maximum negative binomial 
+discrete distributions also appear in~\cite{Zhang2000}
+and \cite{Zelterman2005}.
+The maximum negative binomial is the smallest number of outcomes necessary to 
+observe at least $c$ successes {\em and} $c$ failures, but the SNB givethe 
+number of coin flips to observe {\em either} $s$ heads or $t$ tails.
+
+\section{Connection to the Binomial Tail Probability}
+
+\begin{prop} Let $Y$ be distributed as SNB($p$, $s$, $t$) and let 
+$B$ be distributed Binomial with size $n=s+t-1$ and success probability
+$p$. Then
+\begin{equation}
+\mathbb{P}[B \geq s] = \mathbb{P} [Y \leq n | \text{\#success} = s].
+\end{equation}
+That is, the probability that the number of successes is at least $s$
+in the Binomial model is the same that the trial stops with $s$ 
+successes in the SNB model.
+\end{prop}
+\begin{proof}
+The Binomial tail probability is
+\begin{align*}
+\mathbb{P}[B \geq s] &= \sum_{k=s}^{s+t-1} {n \choose k} p^k (1-p)^{n-k} \\
+  &= 1 - \sum_{k=0}^{s-1} p^k (1-p)^{n-k} \\
+  & 1 - \mathcal{I}_{1-p}.
+\end{align*}
+Next, consider the SNB probability
+\begin{equation*}
+\mathbb{P} [Y \leq n | \text{\#success} = s] = \sum_{k=s}^{s+t-1}
+  {k-1 \choose s-1} p^s (1-p)^k-s.
+\end{equation*}
+Let $i=k-s$ then using the fact that ${i+s-1 \choose s-1} = {i+s-1 \choose i}$
+the summation can be rewritten as
+\begin{align}
+\mathbb{P} [Y \leq n | \text{\#success} = s] &= \sum_{i=0}^{t-1} 
+  {i+s-1 \choose i} p^s (1-p)^i\\
+  &= 1 - \mathcal{I}_{1-p}(t, s).
+\end{align}
+\end{proof}
+
+\section{The Moment Generating Function}
 
@@
 \end{equation*}
-and this can be rewritten as:
+and can be rewritten as:
 \begin{equation} \label{eqn:first_sum}
@@
 \end{proof}
-
-\begin{figure}[p!]
-\begin{center}
-\includegraphics[width=\textwidth]{shapes.pdf}
-\end{center}
-\caption{Different shapes of the SNB distribution with parameters ($s$, $t$, $p$), as given. Red indicates mass contributed by hitting $s$, teal indicates
-mass contributed by hitting $t$. \label{shapes.fig}}
-\end{figure}
-
-The probability mass function of $\,Y\,$ has a variety of shapes for different choices of the parameters $(s,\, t,\,p)$.  These shapes are illustrated in 
-Fig.~\ref{shapes.fig}.The SNB is related to the negative binomial 
-distribution. Specifically, if $t$ is large then the $Y-s$ has a negative 
-binomial distribution with
-\begin{equation*}                                    %   (1)   
-\mathbb{P}\{Y=s+j \}        \label{nb1.eq}          
-  = {{s+j-1}\choose{s-1}} p^s (1-p)^j
-\end{equation*}
-for $\,j=0, 1,\ldots\,$. A similar statement can be made when $s$ is large
-and $t$ is small.
-
-For the special case of $\,s=t,$ the distribution of $\,Y\,$ is the
-riff-shuffle, or minimum negative binomial distribution~\citep{Uppuluri1970}.
-Similar derivations of the closely-related maximum negative binomial 
-discretedistributions also appear in~\cite{Zhang 2000}
-and \cite{Zelterman2004}.
-The maximum negative binomial is the smallest number of outcomes necessary to 
-observe at least $c$ successes {\em and} $c$ failures, but the SNB givethe 
-number of coin flips to observe {\em either} $s$ heads or $t$ tails.
-
-\section{Connection to the Binomial Tail Probability}
 
@@
 
-Eqn.~\ref{eqn:posterior} suggests an alternative construction for
-Phase II clinical trials.  Patients could be
-enrolled sequentially until there are enough patients to guarantee the
-desired level of power or until a specified number of adverse events are
-recorded. In cases such as these the endpoints $s$ and $t$ are known and
-new outcomes can be incorporated into the estimate of
-$p$ to provide updated odds of successful enrollment. Later it will be shown
-that estimates of the distribution of $p$ can even be used after an endpoint
-has been reached to determine the amount of uncertainty in the endpoint.
-
 \section{Discussion and Conclusion}

### 2016-02-29 stat-and-prob-letters/s.r

@@ -1,0 +1,21 @@
+library(snb)
+
+p = 0.2
+s = 2
+t = 12
+n = s+t-1
+
+dsnb_stack_plot(p, s, t) + ylab("SNB(0.2, 2, 12)")
+ggsave("snb_density.pdf")
+
+s1 = dsnb_stacked(min(s,t):n, p, s, t)
+
+b1 = dbinom(0:n, n, p)
+
+d = data.frame(list(x=0:n, y=b1))
+d$Outcome = factor(c(rep("t", s), rep("s", n-s+1)))
+d$Outcome = relevel(d$Outcome, "s")
+
+ggplot(d, aes(x=x, y=y, fill=Outcome))+geom_bar(stat="identity") + xlab("k")+
+  ylab("Bin(0.2, 13)")
+ggsave("bin_density.pdf")

### 2016-02-29 stat-and-prob-letters/snb-paper.tex
yeah.

@@
 \usepackage{url}
+\usepackage{subfig}
 \DeclareMathOperator*{\argmin}{argmin}
@@
 
-The probability mass function of $\,Y\,$ has a variety of shapes for different 
-choices of the parameters $(s,\, t,\,p)$.  These shapes are illustrated in 
-Fig.~\ref{shapes.fig}. The SNB is related to the negative binomial 
-distribution. Specifically, if $t$ is large then the $Y-s$ has a negative 
-binomial distribution with
+The SNB is related to the negative 
+binomial distribution. Specifically, if $t$ is large then the $Y-s$ has a 
+negative binomial distribution with
 \begin{equation*}                                    %   (1)   
@@
 for $\,j=0, 1,\ldots\,$. A similar statement can be made when $s$ is large
-and $t$ is small.
+and $t$ is small. As a result, with proper parameter choice, the SNB
+can mimic other probability distributions in a manner similar to 
+that described in \cite{Peizer1968} and \cite{Best1974}. Examples are
+shown in in Fig.~\ref{shapes.fig}.
 
@@
 The maximum negative binomial is the smallest number of outcomes necessary to 
-observe at least $c$ successes {\em and} $c$ failures, but the SNB givethe 
+observe at least $c$ successes {\em and} $c$ failures. The SNB is the 
 number of coin flips to observe {\em either} $s$ heads or $t$ tails.
@@
 \section{Connection to the Binomial Tail Probability}
+
+\begin{figure}[t!]
+\centering
+\subfloat[The SNB distribution.]{\includegraphics[width=0.5\textwidth]{snb_density.pdf}}
+\hfill
+\subfloat[The Binomial distribution.]{\includegraphics[width=0.5\textwidth]{bin_density.pdf}}
+\caption{
+SNB(0.2, 2, 12) with mass contributed from 
+$s$ successes (red) or $t$ failures (teal) along with Bin(0.2, 13) with
+at least 2 successes (red) or fewer (teal).
+}
+\label{fig:kane_viz}
+\end{figure}
 

### 2016-03-01 stat-and-prob-letters/snb-paper.tex
Draft for my review.

@@
 
-The rest of this paper derives the distribution for the number of
-patients needed to end a trial and explores the characteristics and
-properties of this distribution.
+\begin{table}[t!]
+\caption{Stopped Negative Binomial Distribution Characteristics}
+\label{tab:snb}
+\begin{center}
+\begin{tabular}{|c|l|} \hline
+Notation & SNB($p$, $s$, $t$) \\ \hline
+Parameters & $p$ the success probability ($q = 1-p$) \\
+           & $s$ the number of successes before stopping \\
+           & $t$ the number of failures before stopping \\ \hline
+Support & min$(s,t) \leq k \leq s+t-1$  \\ \hline
+PMF & ${k-1 \choose s-1} p^s (1-p)^{k-s} + {k-1 \choose t-1} (1-p)^s p^{k-t}$\\ \hline
+CDF & $2 - \mathcal{I}_{1-p}(k+1, s) - \mathcal{I}_{p}(k+1, t)$\\ 
+    & where $\mathcal{I}$ is the regularized incomplete beta function.\\ \hline
+Mean & $\frac{s}{p} \mathcal{I}_p(s,t) + \frac{q^{t-1} p^{s-1}}{B(s,t)} +
+  \frac{t}{q} \mathcal{I}_q(t,s) + \frac{q^{s-1}p^{t-1}}{B(s,t)}$\\ 
+  & where $B$ is the beta function \\ \hline
+MGF & $\left(\frac{p e^x}{1 - qe^x}\right)^s 
+  \mathcal{I}_{1-qe^x} (s, t) + \left(\frac{qe^x}{1-pe^x}\right)^t 
+  \mathcal{I}_{1-pe^x}(t, s) $\\ \hline
+\end{tabular}
+\end{center}
+\end{table}
+
+Tab.~\ref{tab:snb} summarizes the distribution and 
+some of its characteristics.
+The rest of this paper derives these results
+and explores properties of the distribution.
 The next section introduces our notation and basic results
@@
 based on a defined Bernoulli process and give some basic properties.
-Section 3 provides a connection to the Binomail tail probability.
+Section 3 provides a connection to the Binomial tail probability.
 Section 4 derives the posterior distribution using a Beta prior.
@@
 The distribution of the stopping time
-$\argminl_k \left\{ X_k \geq s \cup X_k \leq k-t \right\}$
+$\argminl_k \left[X_k \geq s \cup X_k \leq k-t \right]$
 is equal to Eqn.~\ref{eqn:pmf}.
@@
 
-\section{Connection to the Binomial Tail Probability}
+\section{Connection Between $S$ and the Binomial Tail Probability}
+
+\begin{prop} \label{binomial_tail}
+Let $Y$ be distributed as SNB($p$, $s$, $t$) and let 
+$B$ be distributed Binomial with size $n=s+t-1$ and success probability
+$p$. Then
+\begin{equation}
+\mathbb{P}[B \geq s] = \mathbb{P} [Y \leq n\ |\ \text{\#success} = s].
+\end{equation}
+That is, the probability that the number of successes is at least $s$
+in the Binomial model is the same that the trial stops with $s$ 
+successes in the SNB model.
+\end{prop}
+\begin{proof}
+The Binomial tail probability is
+\begin{align*}
+\mathbb{P}[B \geq s] &= \sum_{k=s}^{s+t-1} {n \choose k} p^k (1-p)^{n-k} \\
+  &= 1 - \sum_{k=0}^{s-1} {n \choose k} p^k (1-p)^{n-k} \\
+  & 1 - \mathcal{I}_{1-p}.
+\end{align*}
+Next, consider the SNB probability
+\begin{align*}
+\mathbb{P} [Y \leq n\ |\ \text{\#success}=s] &= \sum_{k=s}^{s+t-1} S(k, p, s)\\
+ &= {k-1 \choose s-1} p^s (1-p)^{k-s}.
+\end{align*}
+Let $i=k-s$ then using the fact that ${i+s-1 \choose s-1} = {i+s-1 \choose i}$
+the summation can be rewritten as
+\begin{align}
+\mathbb{P} [Y \leq n\ |\ \text{\#success} = s] &= \sum_{i=0}^{t-1} 
+  {i+s-1 \choose i} p^s (1-p)^i\\
+  &= 1 - \mathcal{I}_{1-p}(t, s).
+\end{align}
+\end{proof}
+
+This result is visualized in Fig.~\ref{fig:snb_bin_compare} for the case
+where $s=2$, $t=12$, and $p=0.2$. The probability masses represented in
+red are equal as are the masses in teal. The probability that $s$
+successes are reached in the SNB process is the same as the binomial 
+probability of at least two successes. Likewise, the probability that $t$ 
+failures are reached in the SNB process is the same as the binomial
+probability of zero or one successes.
 
@@
 }
-\label{fig:kane_viz}
+\label{fig:snb_bin_compare}
 \end{figure}
-
-\begin{prop} Let $Y$ be distributed as SNB($p$, $s$, $t$) and let 
-$B$ be distributed Binomial with size $n=s+t-1$ and success probability
-$p$. Then
-\begin{equation}
-\mathbb{P}[B \geq s] = \mathbb{P} [Y \leq n | \text{\#success} = s].
-\end{equation}
-That is, the probability that the number of successes is at least $s$
-in the Binomial model is the same that the trial stops with $s$ 
-successes in the SNB model.
-\end{prop}
-\begin{proof}
-The Binomial tail probability is
-\begin{align*}
-\mathbb{P}[B \geq s] &= \sum_{k=s}^{s+t-1} {n \choose k} p^k (1-p)^{n-k} \\
-  &= 1 - \sum_{k=0}^{s-1} p^k (1-p)^{n-k} \\
-  & 1 - \mathcal{I}_{1-p}.
-\end{align*}
-Next, consider the SNB probability
-\begin{equation*}
-\mathbb{P} [Y \leq n | \text{\#success} = s] = \sum_{k=s}^{s+t-1}
-  {k-1 \choose s-1} p^s (1-p)^k-s.
-\end{equation*}
-Let $i=k-s$ then using the fact that ${i+s-1 \choose s-1} = {i+s-1 \choose i}$
-the summation can be rewritten as
-\begin{align}
-\mathbb{P} [Y \leq n | \text{\#success} = s] &= \sum_{i=0}^{t-1} 
-  {i+s-1 \choose i} p^s (1-p)^i\\
-  &= 1 - \mathcal{I}_{1-p}(t, s).
-\end{align}
-\end{proof}
 
@@
 
-\begin{prop} Let $S$ be distributed SNB with parameters $p$, $s$, and $t$.
-Then the moment generating function (MGF) of $S$ is
+\begin{prop} Let $Y$ be distributed SNB with parameters $p$, $s$, and $t$.
+Then the moment generating function (MGF) of $Y$ is
 \begin{equation} \label{eqn:mgf}
-\mathbb{E}~e^{xS} = \frac{p e^x}{1 - qe^x} \mathcal{I}_{qe^x} (s, t) + 
-  \frac{qe^x}{1-pe^x} \mathcal{I}_{pe^x}(s, t)
+\mathbb{E}~e^{xY} = \left(\frac{p e^x}{1 - qe^x}\right)^s 
+  \mathcal{I}_{1-qe^x} (s, t) + \left(\frac{qe^x}{1-pe^x}\right)^t 
+  \mathcal{I}_{1-pe^x}(t, s)
 \end{equation}
@@
 \begin{equation*}
-\mathbb{E}~e^{xS} = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s q^{k-s} e^{kx} 
+\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1} {k-1 \choose k-s} p^s q^{k-s} e^{kx} 
   + \sum_{k=t}^{s+t-1} {k-1 \choose k-t} p^{k-t} q^t e^{kx}
@@
 \begin{equation} \label{eqn:first_sum}
-\mathbb{E}~e^{xS} = \sum_{k=s}^{s+t-1}{k-1 \choose s-1} (pe)^{sx} (qe^x)^{k-s} 
+\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1}{k-1 \choose k-s} (pe)^{sx} (qe^x)^{k-s} 
   + \sum_{k=t}^{s+t-1}{k-1 \choose k-t} (qe^x)^t (pe^x)^{k-t}.
@@
 \begin{align*}
-\sum_{k=s}^{s+t-1}{k-1 \choose s-1} (pe)^{sx} (qe^x)^{k-s} &= 
-  \left(\frac{pe^x}{1 - qe^x}\right)^s \ \ \sum_{k=s}^{s+t-1} {k-1 \choose s-1} 
-    (qe^x)^{k-s} (1-qe^x)^s) \\
+\sum_{k=s}^{s+t-1}{k-1 \choose k-s} (pe)^{sx} (qe^x)^{k-s} &= 
+  \left(\frac{pe^x}{1 - qe^x}\right)^s \ \ \sum_{k=s}^{s+t-1} {k-1 \choose k-s} 
+    (qe^x)^{k-s} (1-qe^x)^s \\
   &= \left(\frac{pe^x}{1 - qe^x}\right)^s \mathcal{I}_{qe^x}(s, t).
@@
 \section{The Posterior Distribution}
-
-Let us assume that $p$ has a Beta distribution, with constant prior
-parameters $\alpha$ and $\beta$. Then the closed form
-posterior distribution of the SNB is
-\begin{align}
-f \left(k | \mathbf{b}, \alpha, \beta \right) &= 
-  {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} 
-  \ I_{\{s \leq k \leq s+k-1\}} + \nonumber \\
-  & {k-1 \choose k-t} 
-  \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} 
-  \ I_{\{t \leq k \leq s+k-1\}}
-\end{align}
-where $B$ is the Beta function and $\mathbf{b}$ is the realization
-of Bernoulli data.
 

### 2016-03-01 paper-1/script.r

@@ -165,8 +165,6 @@ p6 = ggplot(x, aes(x=x, y=y, group=id)) +
 ggsave("Region-s-eq-t.pdf", p6, width=7, height=5)
 
 
-stop("here")
-
 d = snb:::flips_to_kplot_df(flips)
 qplot(k, path, data = d, geom = "blank") +
   scale_x_continuous(breaks = 0:(t + s), limits = c(0, t + s)) +
@@ -186,17 +184,17 @@ cdsnb(trial, s, t)
 
 approx_plots_start = list(
   dsnb_stack_plot(0.2, 15, 75) + scale_x_discrete(breaks=seq(15, 95, by=10)) + 
-    labs(x="", y="", title="bimodal example (15, 75, .2)"),
+    labs(x="", y="", title="bimodal example (0.2, 15, 75)"),
   dsnb_stack_plot(0.35, 50, 50) + scale_x_discrete(breaks=seq(50, 105, by=10))+
-    labs(x="", y="", title="approximate normal (50, 50, .35)"),
+    labs(x="", y="", title="approximate normal (0.35, 50, 50)"),
   dsnb_stack_plot(0.06, 10, 10) + scale_x_discrete(breaks=seq(10, 20, by=2))+
-    labs(x="", y="", title="geometric (10, 10, .06)"),
+    labs(x="", y="", title="geometric (0.06, 10, 10)"),
   dsnb_stack_plot(0.06, 3, 175) + scale_x_discrete(breaks=seq(3, 180, by=30))+ 
-    labs(x="", y="", title="gamma approximation (3, 175, .06)"),
+    labs(x="", y="", title="gamma approximation (0.06, 3, 175)"),
   dsnb_stack_plot(0.5, 25, 25) + scale_x_discrete(breaks=seq(25, 50, by=5))+
-    labs(x="", y="", title="lower half normal (25, 25, .5)"),
+    labs(x="", y="", title="lower half normal (0.5, 25, 25)"),
   dsnb_stack_plot(0.98, 175, 2) + scale_x_discrete(breaks=seq(2, 180, by=20))+
-    labs(x="", y="", title="local mode at top of range (175, 2, .98)"))
+    labs(x="", y="", title="local mode at top of range (0.98, 175, 2)"))
 
 library(foreach)
 library(grid)

### 2016-03-10 stat-and-prob-letters/snb-paper.tex
Made the easy revisions suggested by Dan.

@@
 R01CA157749, R01CA148996, R01CA168733, and PC50CA196530 awarded by the 
-National Cancer Institute, and support from the Yale Comprehensive Cancer 
-Center.}
+National Cancer Institute along with support from the Yale Comprehensive Cancer 
+Center and the Yale Center for Outcomes Research. We would also like 
+to thank Rick Landin at Ignyta Inc. for his suggestions.}
 
@@
 \begin{abstract}
-This paper introduces a novel discrete distribution suggested by curtailed
+We introduce a discrete distribution suggested by curtailed
 sampling rules common in early-stage clinical trials. We derive the
 distribution of the smallest number of independent and identically
-distributed Bernoulli trials needed in order to observe either $s$ successes 
+distributed Bernoulli trials needed to observe either $s$ successes 
 or $t$ failures. The closed-form expression for the distribution as well as 
-its characteristics and properties are explored.
+its characteristics are derived and properties of the distribution are explored.
 \end{abstract}
@@
 
-Consider a prototypical Phase II single-arm clinical trial. 12 patients
-are enrolled and treated. If at least 2 of those patients respond
-favorably, then the trial will proceed to the next stage. If fewer than 
-two respond then the trial would have been terminated.
+Consider a prototypical Phase II single-arm clinical trial in which 12 patients
+are enrolled and treated. If two or more respond then we reject the null
+hypothesis that the treatment is not effective and the trial proceeds to the 
+next stage. If fewer that two respond then the null is not rejected and the 
+trial is terminated.
 
-The maximum sample size is 12 but the number of patients necessary to
-reach any endpoint could be less. Our goal is to describe the distribution of
-the enrollment size. This distribution will be used for planning purposes. 
 If all 12 patients are enrolled at once, as in the classic
-design, then the sample size is 12. However, in most clinical trials, the
+design, then the sample size is 12. However, as in most clinical trials, the
 patients are enrolled sequentially, often with one patient's outcome realized
@@
 successful patients allows us reach one endpoint so the sample required
-could be as small as two - 12 might not be necessary. Similarly 11
+could be as small as two. Similarly 11
 observed treatment failures also ends the stage. This sampling mechanism, in
@@
 
-Tab.~\ref{tab:snb} summarizes the distribution and 
-some of its characteristics.
+In this work we derive the distribution of the number of enrollees needed
+for either $s$ successes or $t$ failures. We refer to this distribution
+as the Stopped Negative Binomial (SNB) and some of its characteristics are
+summarized in Tab.~\ref{tab:snb}.
 The rest of this paper derives these results
@@
 including the density of the distribution along with a description of
-it's relation to other distributions. Sections 2 derives the distribution
-based on a defined Bernoulli process and give some basic properties.
+it's relation to other distributions. Section 2 derives the distribution
+based on a defined Bernoulli process and gives some basic properties.
 Section 3 provides a connection to the Binomial tail probability.
@@
 
-\section{Distributional Result}
+\section{Probability Mass Function}
 \label{notation.section}
@@
 $\,b_i = 1$ corresponds to a successful outcome.  Let $s$ and $t$ be
-positive integers.  Define the stopped negative binomial (SNB) random
+positive integers.  Define the SNB random
 variable $Y$ as the smallest
 integer value such that $\,\{b_1, \ldots , b_Y\}\,$ contains {\em either}
-$\,s\,$ successes {\em or} $\,t\,$ failures. That is, either
+$\,s\,$ successes {\em or} $\,t\,$ failures. That is, the SNB distribution
+of $Y$ is the smallest integer such that either
 $\sum_i^Y b_i = s$ or $\sum_i^Y 1-b_i = t$.
@@
 \end{equation}
-$S$ is the negative binomial probability
+is the negative binomial probability
 
-To prove the result in Eqn.~\ref{eqn:pmf} consider the
+To prove (\ref{eqn:pmf}), consider the
 process $\mathbf{X} = \left\{X(k) : k = 0,1,... \right\}$
@@
 \end{equation*}
-The process can be conceptualized as a series of coin
-flips that stops when either $s$ successes or $t$ failures are reached.
-At each step a coin is flipped. If it is heads, the process advances
-one diagonally in the
+At each step a patient's outcome is measured. If it is success, the process 
+advances one diagonally in the
 positive horizontal and vertical direction.
 Otherwise, it advances in the positive horizontal direction only. The
-process stops when either $X_k = s$ or $X_k = k-t$.
+process continues until either $X_k = s$ or $X_k = k-t$.
 
@@
 $\argminl_k \left[X_k \geq s \cup X_k \leq k-t \right]$
-is equal to Eqn.~\ref{eqn:pmf}.
+is given at (\ref{eqn:pmf}).
 \end{prop}
@@
 successful outcomes and $k-s$ unsuccessful outcomes multiplied by
-the probability of a
-success at time $k$. 
-\begin{equation} \label{stop_s}
-S(k, p, s) = {k-1 \choose s-1} p^s (1-p)^{k-s}
-\end{equation}
-The probability a given realization reaches $k-t$
+the probability of a success at time $k$. This expression is given
+in (\ref{eqn:N}). 
+Similarly, probability a given realization reaches $k-t$
 is the probability that, at outcome $k-1$ there are $k-t$ successful outcomes
@@
 unsuccessful outcome at time $k$.
+
+Next, define
 \begin{equation} \label{stop_t}
@@
 \end{equation}
-It can be shown that $S(k, p, s) = S'(k, 1-p, s)$ by realizing that 
+and notice that $S(k, p, s) = S'(k, 1-p, s)$ by writing
 ${k-1 \choose k-s} = {k-1 \choose s-1}$.
 
-To show that sum of Eqn. \ref{stop_s} and \ref{stop_t} sum to one
-over their support 
+To show that (\ref{eqn:N}) and (\ref{stop_t}) sum to one
+over their support, let
 \begin{align} \label{eqn:sum_proof}
@@
 \end{align}
-substitute $i=k-s$ in the first summation and $j=k-t$ in the second.
-Then $R$ can be written
-as the cumulative distribution function of two
-negative binomial distributions
+where we substitute $i=k-s$ in the first summation and $j=k-t$ in the second.
+
+Then $R$ can be written as the cumulative distribution function of two
+negative binomial distributions:
 \begin{equation} \label{eqn:transformed_sum}
@@
 Let $\mathcal{I}_p(s, t)$ be the {\em regularized incomplete beta function} 
-and recall this function satisfies 
+\citep{Olver2010} and recall this function satisfies 
 $\mathcal{I}_p(s, t) = 1-\mathcal{I}_{1-p}(t, s)$ \citep{Abramowitz1964}.
@@
 \end{align*}
+This completes the proof that (\ref{eqn:pmf}) is a valid probability mass
+function.
 \end{proof}
@@
 
-The SNB is related to the negative 
-binomial distribution. Specifically, if $t$ is large then the $Y-s$ has a 
+The SNB is a generalization of the negative 
+binomial distribution. If $t$ is large then the $Y-s$ has a 
 negative binomial distribution with
@@
 can mimic other probability distributions in a manner similar to 
-that described in \cite{Peizer1968} and \cite{Best1974}. Examples are
+those described in \cite{Peizer1968} and \cite{Best1974}. Examples are
 shown in in Fig.~\ref{shapes.fig}.
@@
 The maximum negative binomial is the smallest number of outcomes necessary to 
-observe at least $c$ successes {\em and} $c$ failures. The SNB is the 
+observe at least $s$ successes {\em and} $s$ failures. The SNB is the 
 number of coin flips to observe {\em either} $s$ heads or $t$ tails.
 
-\section{Connection Between $S$ and the Binomial Tail Probability}
+\section{Connection Between the SNB and the Binomial Tail Probability}
 
@@
 \end{align*}
-Next, consider the SNB probability
+The corresponding SNB probability is
 \begin{align*}
@@
 \end{align*}
-Let $i=k-s$ then using the fact that ${i+s-1 \choose s-1} = {i+s-1 \choose i}$
-the summation can be rewritten as
+Let $i=k-s$. Use ${i+s-1 \choose s-1} = {i+s-1 \choose i}$
+so the last summation can be rewritten as
 \begin{align}
@@
   {i+s-1 \choose i} p^s (1-p)^i\\
-  &= 1 - \mathcal{I}_{1-p}(t, s).
+  &= 1 - \mathcal{I}_{1-p}(t, s)
 \end{align}
+completing the proof.
 \end{proof}
 
-This result is visualized in Fig.~\ref{fig:snb_bin_compare} for the case
+Fig.~\ref{fig:snb_bin_compare} shows the case
 where $s=2$, $t=12$, and $p=0.2$. The probability masses represented in
@@
 \end{equation}
-for $q = 1-p$ when $x \leq -\log(q)$ and $x \leq -\log(p)$.
+for $q = 1-p$ when $x \leq \min \left\{\log(1/p), \log(1/q) \right\}$.
 \end{prop}
 \begin{proof}
-By definition, the MGF of the SNB is:
+The MGF of the SNB is:
 \begin{equation*}
@@
 \end{align*}
-Since the incomplete beta function has support on zero to one
-$qe^x \leq 1$. This implies that $x \leq -\log(q)$.
+Since the incomplete beta function has support on zero to one, we have
+$qe^x \leq 1$. This also shows $x \leq -\log(q)$.
 
@@
 
+Let us consider a Bayesian analysis where there is a Beta prior distribution
+on $p$.
 \begin{prop}
@@
     \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} 
-    \ I_{\{t \leq k \leq s+k-1\}}
+    \ I_{\{t \leq k \leq s+k-1\}}.
 \end{align}
-where $B$ is the Beta function.
 \end{prop}
@@
 
-We have presented a new discrete distribution by curtailed sampling rules common in early-stage clinical trials, which we refer to as the Stopped Negative Binomial distribution.
-The distribution models the stopping time of a sequential trial where the trial is stopped when a number of events are accumulated.
-The
-posterior distribution was derived for the case when the event probability $p$
-has a Beta distribution. Using a trial description from
-\url{clinicaltrials.gov} we showed how the SNB is an integral part of trial
-monitoring, and post-hoc analysis and can be used in a framework alternative
-to the Simon two-stage optimal design while providing better estimates of the
-event probability along with quantifying the uncertainty associated with the
-estimates. As a result, we were able to show fewer patient enrollments and
-achieve the same estimates, when added in a hypothetical, but representative,
-clinical trial.
+We have presented a new discrete distribution by curtailed sampling rules 
+common in early-stage clinical trials, which we refer to as the Stopped 
+Negative Binomial distribution. The distribution models the stopping time of 
+a sequential trial where the trial is stopped when a number of events are 
+accumulated. A connection between the Binomial tail probability and the SNB
+distribution was shown; It's MGF was derived; and
+the posterior distribution was derived for the case when the 
+event probability $p$ has a Beta distribution.
 
-Current work focuses on the generalization of the distribution and the application in other areas of clinical trials.
-Adverse outcomes in particular are another area that could greatly benefit both from the presented distribution and its framework.
-For these trials monitoring may need to be performed not only on the outcome of an intervention but also on the safety of the trial.
-This especially
-true in areas like late-stage cancer treatments where the treatment is harsh
-and patients may be forced to drop out as a result of the side-effects of the
-treatment; a matter independent of the outcome. Other application areas
-include providing designs that allows clinicians to balance uncertainty and
-success probability with a minimum number of patient enrollees.
+Current work focuses on two different areas. First, the distribution
+is being applied to clinical trial design
+where the number of enrollees is smaller than traditional trials.
+This scenario is increasingly common in targeted cancer
+therapy where treatment may only be appropriate for small subpopulations.
+Second, generalizations and extensions of the distribution are being pursued
+for applications power analysis. The posterior construction allows us
+to quantify the uncertainty both in the response probability as well as the
+time until a trial is complete. This suggests new approaches to sample
+size calculations as well as trial monitoring.
 

### 2016-03-10 stat-and-prob-letters/s.r

@@ -1,5 +1,96 @@
 library(snb)
+library(ggplot2)
 
+# kplot
+
+kplot = function(flips, s, t) {
+  if (!is.list(flips)) {
+    d = flips_to_kplot_df(flips)
+    p = qplot(k, path, data = d, geom = "line") +
+      scale_x_continuous(breaks = 0:(t + s), limits = c(0, t + s)) +
+      scale_y_continuous(breaks = 0:s, limits=c(0, s+0.15)) +
+      geom_segment(x=s, y=s, xend=(t+s-1), yend=s, color="green", linetype=1) +
+      geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, col="red")
+  } else {
+    flip_set = lapply(flips, flips_to_kplot_df)
+    for (i in 1:length(flip_set)) {
+      flip_set[[i]]$num = as.factor(i)
+      flip_set[[i]]$k = jitter(flip_set[[i]]$k)
+      flip_set[[i]]$k[flip_set[[i]]$k < 0] = 0
+    }
+    d = Reduce(rbind, flip_set)[, -(4:5)]
+    p = qplot(k, path, data = d, geom = "path", group = num) +
+        scale_x_continuous(breaks=0:(t+s), limits = c(0, t+s)) +
+        geom_segment(x = s, y = s, xend = (t + s - 1), yend = s,
+                     color = "green") +
+        geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, col="red")
+    p
+  }
+  p
+}
+
+flips_to_kplot_df = function (flips) {
+    d = data.frame(k = 0:length(flips))
+    d$head = c(0, cumsum(flips))
+    d$tail = c(0, cumsum(1 - flips))
+    d$headEnd = c(d$head[-1], NA)
+    d$tailEnd = c(d$tail[-1], NA)
+    d$path = c(0, cumsum(flips))
+    d$k = 0:(nrow(d) - 1)
+    d
+}
+
+p = kplot( c(rep(0, 3), 1, rep(0, 6), 1), s=2, t=11) +
+  xlab("Number of Patients Enrolled") + ylab("Number of Responders") +
+  geom_text(data=NULL, x=7, y=2.1, label="Success Boundary") +
+  geom_text(data=NULL, x=12, y=0.5, label="Failure Boundary",
+            angle=(45+90)/2+9) + 
[…]

### 2016-03-22 stat-and-prob-letters/snb-paper.tex
Yeah.

@@
 \newtheorem{prop}{Proposition}
+\newtheorem{corollary}{Corollary}
 
@@
 are enrolled and treated. If two or more respond then we reject the null
-hypothesis that the treatment is not effective and the trial proceeds to the 
-next stage. If fewer that two respond then the null is not rejected and the 
+hypothesis that the treatment is not effective and the trial is deemed
+successful. If fewer that two respond then the null is not rejected and the 
 trial is terminated.
@@
 
-Assume each of patient outcome can be modeled as an independent,
+Assume each patient outcome can be modeled as an independent,
 identically distributed Bernoulli($p$) random variable. The trial is realized
@@
 \end{equation*}
-At each step a patient's outcome is measured. If it is success, the process 
+At each step a patient's outcome is measured. If it is a success, the process 
 advances one diagonally in the
@@
 and notice that $S(k, p, s) = S'(k, 1-p, s)$ by writing
-${k-1 \choose k-s} = {k-1 \choose s-1}$.
+\begin{equation*}
+{k-1 \choose k-s} = {k-1 \choose s-1}.
+\end{equation*}
 
@@
 
+\begin{corollary} \label{conditional_distribution}
+Assume a trial that stops on the $s$th successs or $t$failure as before. 
+Furthermore, assume we observer $s'$ successes and $t'$ failures
+where $s' < s$ and $t' < t$.
+That is, we see the number of success and failures at a time before
+reaching an endpoint. Then the number of subsequent enrollements needed 
+to reach an endpoint is SNB($p$, $s-s'$, $t-t'$).
+\end{corollary}
+\begin{proof}
+The result follows directly by realizing that, conditioned on $s'$
+success and $t'$ failures, there are $s-s'$ successes needed to reach the
+success endpoint and $t-t'$ failures needed to reach the failure endpoint.
+\end{proof}
+
 \section{Shape and Basic Properties}
@@
 \end{center}
-\caption{Different shapes of the SNB distribution with parameters ($s$, $t$, $p$), as given. Red indicates mass contributed by hitting $s$, teal indicates
+\caption{Different shapes of the SNB distribution with parameters ($s$, $t$, $p$), as given. Black indicates mass contributed by hitting $s$, grey indicates
 mass contributed by hitting $t$. \label{shapes.fig}}
@@
 The SNB is a generalization of the negative 
-binomial distribution. If $t$ is large then the $Y-s$ has a 
+binomial distribution. If $t$ is large then $Y-s$ has a 
 negative binomial distribution with
@@
 observe at least $s$ successes {\em and} $s$ failures. The SNB is the 
-number of coin flips to observe {\em either} $s$ heads or $t$ tails.
+number of outcomes to observe {\em either} $s$ successes or $t$ failures.
 
@@
 
+The probability of reaching the success endpoint in an 
+SNB($p$, $s$, $t$) random variable is 
+equal to the probability of at least $s$ successes in a Binomial distributed 
+random variable with size $s+t-1$ and success probability $p$. 
+Likewise, the probability of reaching the failure endpoint is equal
+to at most $s-1$ successes in a Binomial distribution.
 \begin{prop} \label{binomial_tail}
@@
 where $s=2$, $t=12$, and $p=0.2$. The probability masses represented in
-red are equal as are the masses in teal. The probability that $s$
+black are equal as are the masses in grey. The probability that $s$
 successes are reached in the SNB process is the same as the binomial 
@@
 SNB(0.2, 2, 12) with mass contributed from 
-$s$ successes (red) or $t$ failures (teal) along with Bin(0.2, 13) with
-at least 2 successes (red) or fewer (teal).
+$s$ successes (black) or $t$ failures (grey) along with Bin(0.2, 13) with
+at least 2 successes (black) or fewer (grey).
 }
@@
 \section{The Moment Generating Function}
+
+The moment generating function is calculated in a manner similar to 
+that of two negative binomial distributions. However, in the negative
+binomial derivation the
+term involving the regularized incomplete beta function integrates to
+one since it is integrated over it's entire support. This is not the
+case in the SNB and it appears in both terms.
 
@@
 on $p$.
-\begin{prop}
+\begin{prop} \label{prop:bayesian}
 The posterior PMF of the Stopped Negative Binomial distribution with a Beta($\alpha$, $\beta$) prior is:
@@
 
-\section{Discussion and Conclusion}
+\section{Discussion}
 
-We have presented a new discrete distribution by curtailed sampling rules 
-common in early-stage clinical trials, which we refer to as the Stopped 
-Negative Binomial distribution. The distribution models the stopping time of 
-a sequential trial where the trial is stopped when a number of events are 
-accumulated. A connection between the Binomial tail probability and the SNB
-distribution was shown; It's MGF was derived; and
-the posterior distribution was derived for the case when the 
-event probability $p$ has a Beta distribution.
+The SNB can be used to extend clinical trial methodology
+in at least two areas. First, it can be used to construct nested criteria
+for early stopping and efficacy. These trials consist of two-stages.
+If the failure endpoint is reached in the first stage then the 
+trial stops due to futility. Otherwise, the trial enters the second 
+stage, which consists of responders from the first stage along with
+new enrollees in the second. Thus the design provides
+both this provide early testing for futility along with the ability
+to make statistically sound decisions sooner than traditional trials.
 
-Current work focuses on two different areas. First, the distribution
-is being applied to clinical trial design
-where the number of enrollees is smaller than traditional trials.
-This scenario is increasingly common in targeted cancer
-therapy where treatment may only be appropriate for small subpopulations.
-Second, generalizations and extensions of the distribution are being pursued
-for applications power analysis. The posterior construction allows us
-to quantify the uncertainty both in the response probability as well as the
-time until a trial is complete. This suggests new approaches to sample
-size calculations as well as trial monitoring.
+A second avenue for methodological innovation comes directly from
+Corollary \ref{conditional_distribution} and Proposition \ref{prop:bayesian}.
+While a trial is being conducted the response probabilities can be fitted
+and updated as new results are received. The number of individual successes
+and failure are used to learn the shape characteristics of the corresponding
+Beta distribution and resulting posterior
+distribution. An estimate of the probability of reaching the success (and
+failure) endpoints can be provided at each step of the trial.
+
+There are also several avenues for extending the SNB itself. First is the 
+development of a stopped negative multinomial distribution allowing
+for more than two endpoints. This extension would have clinical trial
+applications including dose escalation where the outcomes are response
+to a therapy, failure to respond, and an adverse outcome endpoint where
+and enrollee is removed from the trial with neither response or non-response
+due to side effects.
+
+A second, more theoretical extension is to the asymptotic setting. Under
+proper reparameterization, the process from which the SNB is derived 
+can be shown to converge to a Brownian motion. In this case, we conjecture
+that the SNB can be defined in terms of the L\'{e}vy-Bachelier 
+\citep{Bachelier1900} formula
+thereby relating the SNB to the classic result in stochastic calculus.
 

### 2016-03-22 stat-and-prob-letters/s.r

@@ -9,8 +9,10 @@ kplot = function(flips, s, t) {
     p = qplot(k, path, data = d, geom = "line") +
       scale_x_continuous(breaks = 0:(t + s), limits = c(0, t + s)) +
       scale_y_continuous(breaks = 0:s, limits=c(0, s+0.15)) +
-      geom_segment(x=s, y=s, xend=(t+s-1), yend=s, color="green", linetype=1) +
-      geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, col="red")
+      geom_segment(x=s, y=s, xend=(t+s-1), yend=s, linetype=2) +
+      geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, linetype=2)
+#      geom_segment(x=s, y=s, xend=(t+s-1), yend=s, color="green", linetype=1) +
+#      geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, col="red")
   } else {
     flip_set = lapply(flips, flips_to_kplot_df)
     for (i in 1:length(flip_set)) {
@@ -22,22 +24,25 @@ kplot = function(flips, s, t) {
     p = qplot(k, path, data = d, geom = "path", group = num) +
         scale_x_continuous(breaks=0:(t+s), limits = c(0, t+s)) +
         geom_segment(x = s, y = s, xend = (t + s - 1), yend = s,
-                     color = "green") +
-        geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, col="red")
+                     linetype=2) +
+        geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, linetype=2)
+#        geom_segment(x = s, y = s, xend = (t + s - 1), yend = s,
+#                     color = "green") +
+#        geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, col="red")
     p
   }
   p
 }
 
 flips_to_kplot_df = function (flips) {
-    d = data.frame(k = 0:length(flips))
-    d$head = c(0, cumsum(flips))
-    d$tail = c(0, cumsum(1 - flips))
-    d$headEnd = c(d$head[-1], NA)
-    d$tailEnd = c(d$tail[-1], NA)
-    d$path = c(0, cumsum(flips))
-    d$k = 0:(nrow(d) - 1)
-    d
[…]

### 2016-03-23 stat-and-prob-letters/snb-paper.tex
Integrated Michelle's suggestions.

@@
 hypothesis that the treatment is not effective and the trial is deemed
-successful. If fewer that two respond then the null is not rejected and the 
+successful. If fewer than two respond then the null is not rejected and the 
 trial is terminated.
@@
 before the next one enters the trial. In the present example, observing two
-successful patients allows us reach one endpoint so the sample required
+successful patients allows us to reach one endpoint so the sample required
 could be as small as two. Similarly 11
-observed treatment failures also ends the stage. This sampling mechanism, in
+observed treatment failures also ends the trial. This sampling mechanism, in
 which the experiment ends as soon as any of the endpoints is reached, is
@@
 as a sequence of these random variables that stops when either a
-specified number of success or failures is reached. In the
+specified number of successes or failures is reached. In the
 previous example suppose two successes were reached after enrolling 10
@@
 including the density of the distribution along with a description of
-it's relation to other distributions. Section 2 derives the distribution
+its relation to other distributions. Section 2 derives the distribution
 based on a defined Bernoulli process and gives some basic properties.
@@
 \begin{corollary} \label{conditional_distribution}
-Assume a trial that stops on the $s$th successs or $t$failure as before. 
+Assume a trial that stops on the $s^{th}$ success or $t^{th}$ failure as 
+before. 
 Furthermore, assume we observer $s'$ successes and $t'$ failures
@@
 where $s=2$, $t=12$, and $p=0.2$.  The probability masses represented in 
-black are equal as are the masses in grey. The probability that $s$
+black are equal in a) and b) as are the masses in grey.
+The probability that $s$
 successes are reached in the SNB process is the same as the binomial 
@@
 the constraint that $x \leq -\log(p)$. The result
-follows from the afore mentioned property of the regularized incomplete
+follows from the aforementioned property of the regularized incomplete
 beta function.
@@
 new enrollees in the second. Thus the design provides
-both this provide early testing for futility along with the ability
+both early testing for futility along with the ability
 to make statistically sound decisions sooner than traditional trials.
@@
 to a therapy, failure to respond, and an adverse outcome endpoint where
-and enrollee is removed from the trial with neither response or non-response
+an enrollee is removed from the trial with neither response or non-response
 due to side effects.

### 2016-03-30 stat-and-prob-letters/snb-paper.tex
Integrating Dan's changes.

@@
 distributed Bernoulli trials needed to observe either $s$ successes 
-or $t$ failures. The closed-form expression for the distribution as well as 
-its characteristics are derived and properties of the distribution are explored.
+or $t$ failures. This report provides a closed-form expression for the 
+mass function and illustrates limiting approximations.
 \end{abstract}
@@
 Consider a prototypical Phase II single-arm clinical trial in which 12 patients
-are enrolled and treated. If two or more respond then we reject the null
-hypothesis that the treatment is not effective and the trial is deemed
-successful. If fewer than two respond then the null is not rejected and the 
-trial is terminated.
+are enrolled and treated. The binomial probability of success is $p=0.045$ for
+each patient under the null hypothesis that treatment is no effective.
+If two or more patients respond to the treatment then we reject this null
+hypothesis, and the trial is deemed successful at significance level of $0.1$. 
+If fewer than two respond then the null is not rejected.
 
@@
 call {\em curtailed sampling}. Under curtailed sampling the range of the
-sample size is between two and 12.
+sample size for this trial is between two and 12.
 
-Assume each patient outcome can be modeled as an independent,
+Let us assume each patient outcome can be modeled as an independent,
 identically distributed Bernoulli($p$) random variable. The trial is realized
-as a sequence of these random variables that stops when either a
-specified number of successes or failures is reached. In the
-previous example suppose two successes were reached after enrolling 10
-patients (one in the third step and one at the $10^{th}$). The sample
-path is illustrated
-in Fig.~\ref{fig:kane_viz}. The vertical axis denotes the number of
+as a sequence of these random variables stopping when either a
+specified number of successes or failures has been reached.
+
+A hypothetical sample path is illustrated in Fig.~\ref{fig:kane_viz}.
+The vertical axis denotes the number of
 successful outcomes. The horizontal axis counts the number of patients that
 have been enrolled. The horizontal and vertical boundaries represent
-endpoints for the trial.
+endpoints for the trial. In this case two successes were reached after
+enrolling 10 patients: one in the third outcome and on in the tenth.
 
@@
 \begin{table}[t!]
-\caption{Stopped Negative Binomial Distribution Characteristics}
+\caption{Characteristics of the Stopped Negative Binomial Distribution}
 \label{tab:snb}
@@
 \begin{tabular}{|c|l|} \hline
-Notation & SNB($p$, $s$, $t$) \\ \hline
-Parameters & $p$ the success probability ($q = 1-p$) \\
-           & $s$ the number of successes before stopping \\
-           & $t$ the number of failures before stopping \\ \hline
+Parameters & $p$ the success probability ($0\leq p \leq 1$; $q = 1-p$) \\
+           & $s$ the number of successes before stopping ($s=1, 2, ...$)\\
+           & $t$ the number of failures before stopping ($t=1, 2, ...$)\\ \hline
 Support & min$(s,t) \leq k \leq s+t-1$  \\ \hline
-PMF & ${k-1 \choose s-1} p^s (1-p)^{k-s} + {k-1 \choose t-1} (1-p)^s p^{k-t}$\\ \hline
-CDF & $2 - \mathcal{I}_{1-p}(k+1, s) - \mathcal{I}_{p}(k+1, t)$\\ 
+$\mathbb{P}[Y=k]$ & ${k-1 \choose s-1} p^s (1-p)^{k-s} + {k-1 \choose t-1} (1-p)^s p^{k-t}$\\ \hline
+$\mathbb{P}[Y \leq k]$ & $2 - \mathcal{I}_{q}(k+1, s) - \mathcal{I}_{p}(k+1, t)$\\ 
     & where $\mathcal{I}$ is the regularized incomplete beta function.\\ \hline
-Mean & $\frac{s}{p} \mathcal{I}_p(s,t) + \frac{q^{t-1} p^{s-1}}{B(s,t)} +
-  \frac{t}{q} \mathcal{I}_q(t,s) + \frac{q^{s-1}p^{t-1}}{B(s,t)}$\\ 
+Mean & $\frac{s}{p} \mathcal{I}_p(s,t) + \frac{p^{s-1} q^{t-1}}{B(s,t)} +
+  \frac{t}{q} \mathcal{I}_q(t,s) + \frac{p^{t-1} q^{s-1}}{B(s,t)}$\\ 
   & where $B$ is the beta function \\ \hline
@@
 Section 3 provides a connection to the Binomial tail probability.
-Section 4 derives the posterior distribution using a Beta prior.
-Section 5 provides a brief discussion on the use of the distribution
+Section 4 derives the posterior distribution using a Beta prior distribution
+for $p$. Section 5 provides a brief discussion on the use of the distribution
 in clinical trials along with future avenues for generalization.
@@
 probability parameter $0\leq p \leq 1$. In the clinical trial setting
-$\,b_i = 1$ corresponds to a successful outcome.  Let $s$ and $t$ be
-positive integers.  Define the SNB random
+$\,b_i = 1$ corresponds to a successful patient outcome following treatment.  
+Let $s$ and $t$ be positive integers.  Define the SNB random
 variable $Y$ as the smallest
@@
 \begin{equation*}               
-     \min(s,t) \leq \; Y \;\leq s+t-1  \label{range.y.eq}
+     \min(s,t) \leq \; Y \;\leq s+t-1  \label{range.y.eq}.
 \end{equation*}
-and it is distributed as
+The probability mass function is
 \begin{equation} \label{eqn:pmf}
-\mathbb{E}\  I_{\{Y=k\}} = S(k, p, s) \ I_{\{s \leq k\}} + 
+\mathbb{P} [Y=k] = S(k, p, s) \ I_{\{s \leq k\}} + 
   S(k, 1-p, t) \ I_{\{ t \leq k \}}
 \end{equation}
-where $I_{\{f\}}$ is one if $f$ is true and zero otherwise and
+where $I_{\{f\}}$ is one if $f$ is true and zero otherwise, and
 \begin{equation} \label{eqn:N}
@@
 \end{equation}
-is the negative binomial probability
+is the negative binomial probability mass.
 

### 2016-03-30 stat-and-prob-letters/s.r

@@ -7,7 +7,7 @@ kplot = function(flips, s, t) {
   if (!is.list(flips)) {
     d = flips_to_kplot_df(flips)
     p = qplot(k, path, data = d, geom = "line") +
-      scale_x_continuous(breaks = 0:(t + s), limits = c(0, t + s)) +
+      scale_x_continuous(breaks = 0:(t + s), limits = c(0, t + s - 1)) +
       scale_y_continuous(breaks = 0:s, limits=c(0, s+0.15)) +
       geom_segment(x=s, y=s, xend=(t+s-1), yend=s, linetype=2) +
       geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, linetype=2)
@@ -46,14 +46,15 @@ flips_to_kplot_df = function (flips) {
 }
 
 p = kplot( c(rep(0, 3), 1, rep(0, 6), 1), s=2, t=11) +
-  xlab("Number of Patients Enrolled") + ylab("Number of Responders") +
+  xlab("Number of Patients Enrolled") + 
+  ylab("Number of Responders to Treatment") +
   geom_text(data=NULL, x=7, y=2.1, label="Success Boundary") +
-  geom_text(data=NULL, x=12, y=0.5, label="Failure Boundary",
-            angle=(45+90)/2+9) + 
+  geom_text(data=NULL, x=11.75, y=0.45, label="Failure Boundary",
+            angle=(45+90)/2-5) + 
   geom_text(data=NULL, x=7, y=1.1, label="Sample Path") +
   theme_bw() + scale_fill_grey()
 
-ggsave("KanePlot.pdf", p, width=7, height=5)
+ggsave("KanePlot.pdf", p, width=10, height=3.5)
 
 
 # SNB distribution shapes.

### 2016-04-07 stat-and-prob-letters/snb-paper.tex
Updated based on Dan's comments.

@@
 \end{equation*}
-At each step a patient's outcome is measured. If it is a success, the process 
+At each step a patient's outcome is measured. In Fig.~\ref{fig:kane_viz} 
+we consider a graphical illustration of the plot $X_k$ against
+$k$. If the outcome of the $k$th patient is a success then the process 
 advances one diagonally in the
-positive horizontal and vertical direction.
-Otherwise, it advances in the positive horizontal direction only. The
+positive horizontal and vertical direction. If the $k$th patient fails
+then the sample path advances in the positive horizontal direction only. The
 process continues until either $X_k = s$ or $X_k = k-t$.
@@
 
-Next, define
-\begin{equation} \label{stop_t}
-S'(k, p, t) = {k-1 \choose k-t} p^{k-t} (1-p)^t
-\end{equation}
-and notice that $S(k, p, s) = S'(k, 1-p, s)$ by writing
-\begin{equation*}
-{k-1 \choose k-s} = {k-1 \choose s-1}.
-\end{equation*}
+%Next, define
+%\begin{equation} \label{stop_t}
+%S'(k, p, t) = {k-1 \choose k-t} p^{k-t} (1-p)^t
+%\end{equation}
+%and notice that $S(k, p, s) = S'(k, 1-p, s)$ by writing
+%\begin{equation*}
+%{k-1 \choose k-s} = {k-1 \choose s-1}.
+%\end{equation*}
 
-To show that (\ref{eqn:N}) and (\ref{stop_t}) sum to one
-over their support, let
+To show that Equation \ref{eqn:pmf} sums to one, define
 \begin{align} \label{eqn:sum_proof}
@@
 \end{align*}
-This completes the proof that (\ref{eqn:pmf}) is a valid probability mass
+This completes the proof (\ref{eqn:pmf}) is is the distribution of the
+stopping time and a valid probability mass
 function.
@@
 
+Next, we consider an interim analysis of a clinical trial after $s'$ 
+successes and $t'$ failures have been observed for $s' < s$ and $t' < t$.
 \begin{corollary} \label{conditional_distribution}
-Assume a trial that stops on the $s^{th}$ success or $t^{th}$ failure as 
-before. 
-Furthermore, assume we observer $s'$ successes and $t'$ failures
-where $s' < s$ and $t' < t$.
-That is, we see the number of success and failures at a time before
-reaching an endpoint. Then the number of subsequent enrollements needed 
+The number of subsequent enrollements needed 
 to reach an endpoint is SNB($p$, $s-s'$, $t-t'$).
 \end{corollary}
-\begin{proof}
-The result follows directly by realizing that, conditioned on $s'$
-success and $t'$ failures, there are $s-s'$ successes needed to reach the
+Having observed $s'$
+successes and $t'$ failures, there are $s-s'$ successes needed to reach the
 success endpoint and $t-t'$ failures needed to reach the failure endpoint.
-\end{proof}
 
-\section{Shape and Basic Properties}
+\section{Connections and Approximations by Other Distributions}
 
@@
 
-\section{Connection Between the SNB and the Binomial Tail Probability}
+%\section{Connection Between the SNB and the Binomial Tail Probability}
 
+There is a close connection between the tail areas of the SNB and the 
+binomial distributions.
 The probability of reaching the success endpoint in an 
@@
 Likewise, the probability of reaching the failure endpoint is equal
-to at most $s-1$ successes in a Binomial distribution.
+to at most $s-1$ successes in a Binomial distribution. That is, the
+probability that the number of successes is at least $s$ in the Binomial
+model is the same that the trial succeeds (reaches $s$) in the SNB model.
 \begin{prop} \label{binomial_tail}
@@
 \begin{equation}
-\mathbb{P}[B \geq s] = \mathbb{P} [Y \leq n\ |\ \text{\#success} = s].
+\mathbb{P}[B \geq s] = \mathbb{P} [Y \leq n\ |\ \text{trial succeeds}].
 \end{equation}
-That is, the probability that the number of successes is at least $s$
-in the Binomial model is the same that the trial stops with $s$ 
-successes in the SNB model.
 \end{prop}
@@
   &= 1 - \sum_{k=0}^{s-1} {n \choose k} p^k (1-p)^{n-k} \\
-  & 1 - \mathcal{I}_{1-p}.
+  &= 1 - \mathcal{I}_{1-p}.
 \end{align*}
 The corresponding SNB probability is
-\begin{align*}
-\mathbb{P} [Y \leq n\ |\ \text{\#success}=s] &= \sum_{k=s}^{s+t-1} S(k, p, s)\\
- &= {k-1 \choose s-1} p^s (1-p)^{k-s}.
-\end{align*}
-Let $i=k-s$. Use ${i+s-1 \choose s-1} = {i+s-1 \choose i}$
-so the last summation can be rewritten as
+\begin{equation*}
+\mathbb{P} [Y \leq n\ |\ \text{trial succeeds}=s] 
+  = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s}.
+\end{equation*}
+Let $i=k-s$. Using the fact that
+\begin{equation*}
+{i+s-1 \choose s-1} = {i+s-1 \choose i}
+\end{equation*}
+the last summation can be rewritten as
 \begin{align}
@@
 
-Fig.~\ref{fig:snb_bin_compare} shows the case
+To illustrate this result, let us return to our initial example, shown
+in Fig.~\ref{fig:snb_bin_compare}, 
 where $s=2$, $t=12$, and $p=0.2$.  The probability masses represented in 
@@
 
-The moment generating function is calculated in a manner similar to 
-that of two negative binomial distributions. However, in the negative
-binomial derivation the
-term involving the regularized incomplete beta function integrates to
-one since it is integrated over it's entire support. This is not the
-case in the SNB and it appears in both terms.
-
+The moment generating function for the SNB is calculated in a manner similar to 
+that of two negative binomial distributions. 
 \begin{prop} Let $Y$ be distributed SNB with parameters $p$, $s$, and $t$.
@@
 \end{equation}
-Taking the first summation in Equation \ref{eqn:first_sum}:
+The first summation in Equation \ref{eqn:first_sum} satisfies
 \begin{align*}
@@
 \end{align*}
-Since the incomplete beta function has support on zero to one, we have
-$qe^x \leq 1$. This also shows $x \leq -\log(q)$.
+Since the incomplete beta function's subscript parameter has support on zero 
+to one, we have $qe^x \leq 1$. This also shows $x \leq -\log(q)$.
 
-A similar expression can be derived using the same calculation with
-the constraint that $x \leq -\log(p)$. The result
-follows from the aforementioned property of the regularized incomplete
-beta function.
+A similar expression can be derived for the second summation in 
+Equation \ref{eqn:first_sum} and results in
+the constraint $x \leq -\log(p)$.
 \end{proof}
@@
 
-Let us consider a Bayesian analysis where there is a Beta prior distribution
-on $p$.
+Let us consider a Bayesian analysis with a Beta prior distribution on $p$.
 \begin{prop} \label{prop:bayesian}
-The posterior PMF of the Stopped Negative Binomial distribution with a Beta($\alpha$, $\beta$) prior is:
+The posterior PMF of the Stopped Negative Binomial distribution with a 
+Beta($\alpha$, $\beta$) prior is
 \begin{align} \label{eqn:posterior}
@@
 \begin{proof}
-For notational simplicity, assume that $s,t \leq k \leq s+t-1$. When this is not the case appropriate terms should be removed as dictated by the indicator functions.
+For notational simplicity, assume that $\min(s,t) \leq k \leq s+t-1$. When 
+this is not the case appropriate terms should be removed as dictated by the indicator functions.
 \begin{align*}
@@
 = \frac{1}{B(\alpha, \beta)}  {k-1 \choose s-1} & \int_0^1  p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} dp + \\
- & \frac{1}{B(\alpha, \beta)} {k-1 \choose k-t} \int_0^1  p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp
+ & \frac{1}{B(\alpha, \beta)} {k-1 \choose k-t} \int_0^1  p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp.
 \end{align*}
@@
 in at least two areas. First, it can be used to construct nested criteria
-for early stopping and efficacy. These trials consist of two-stages.
-If the failure endpoint is reached in the first stage then the 
-trial stops due to futility. Otherwise, the trial enters the second 
-stage, which consists of responders from the first stage along with
-new enrollees in the second. Thus the design provides
+for early stopping and efficacy. Current approaches are often inspired by
+the Simon 2-stage Phase II clinical trial \cite{Simon1989} where a
+a small number of patients are enrolled for an initial stage. For patients,
+success or failure is assessed. If at least a pre-specified number of
+successes are observed then the trial enters a second stage with more
+patients. If a sufficient number of successes are not observed then the
+trial fails.
+
+The SNB is being used to generalize the Simon design by providing separate
+criteria for futility and efficacy.
+The Simon design tests for a single response probability and time to event
+parameter for both the first
+and second stage. The SNB is being used to generalize this design by 
+allowing these parameters to vary for the two stages, and even allow
+information about patients in the first stage to be used in the second
+stage. Early results show the design provides
 both early testing for futility along with the ability
@@
 Corollary \ref{conditional_distribution} and Proposition \ref{prop:bayesian}.
-While a trial is being conducted the response probabilities can be fitted
-and updated as new results are received. The number of individual successes
+Our knowledge of the response probability $p$ comes from published reports
+yielding a prior distribution.
+While a trial is being conducted the response probabilities are 
+updated as new results are received. The number of individual successes
 and failure are used to learn the shape characteristics of the corresponding
@@
 
-There are also several avenues for extending the SNB itself. First is the 
-development of a stopped negative multinomial distribution allowing
-for more than two endpoints. This extension would have clinical trial
-applications including dose escalation where the outcomes are response
-to a therapy, failure to respond, and an adverse outcome endpoint where
-an enrollee is removed from the trial with neither response or non-response
-due to side effects.
+%There are also several avenues for extending the SNB itself. First is the 
+%development of a stopped negative multinomial distribution allowing
+%for more than two endpoints. This extension would have clinical trial
+%applications including dose escalation where the outcomes are response
+%to a therapy, failure to respond, and an adverse outcome endpoint where
+%an enrollee is removed from the trial with neither response or non-response
+%due to side effects.
 
-A second, more theoretical extension is to the asymptotic setting. Under
-proper reparameterization, the process from which the SNB is derived 
-can be shown to converge to a Brownian motion. In this case, we conjecture
-that the SNB can be defined in terms of the L\'{e}vy-Bachelier 
-\citep{Bachelier1900} formula
-thereby relating the SNB to the classic result in stochastic calculus.
+%A second, more theoretical extension is to the asymptotic setting. Under
+%proper reparameterization, the process from which the SNB is derived 
+%can be shown to converge to a Brownian motion. In this case, we conjecture
+%that the SNB can be defined in terms of the L\'{e}vy-Bachelier 
+%\citep{Bachelier1900} formula
+%thereby relating the SNB to the classic result in stochastic calculus.
 

### 2016-04-07 stat-and-prob-letters/s.r

@@ -1,38 +1,38 @@
 library(snb)
 library(ggplot2)
 
-# kplot
-
-kplot = function(flips, s, t) {
-  if (!is.list(flips)) {
-    d = flips_to_kplot_df(flips)
-    p = qplot(k, path, data = d, geom = "line") +
-      scale_x_continuous(breaks = 0:(t + s), limits = c(0, t + s - 1)) +
-      scale_y_continuous(breaks = 0:s, limits=c(0, s+0.15)) +
-      geom_segment(x=s, y=s, xend=(t+s-1), yend=s, linetype=2) +
-      geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, linetype=2)
-#      geom_segment(x=s, y=s, xend=(t+s-1), yend=s, color="green", linetype=1) +
-#      geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, col="red")
-  } else {
-    flip_set = lapply(flips, flips_to_kplot_df)
-    for (i in 1:length(flip_set)) {
-      flip_set[[i]]$num = as.factor(i)
-      flip_set[[i]]$k = jitter(flip_set[[i]]$k)
-      flip_set[[i]]$k[flip_set[[i]]$k < 0] = 0
-    }
-    d = Reduce(rbind, flip_set)[, -(4:5)]
-    p = qplot(k, path, data = d, geom = "path", group = num) +
-        scale_x_continuous(breaks=0:(t+s), limits = c(0, t+s)) +
-        geom_segment(x = s, y = s, xend = (t + s - 1), yend = s,
-                     linetype=2) +
-        geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, linetype=2)
+## kplot
+#
+#kplot = function(flips, s, t) {
+#  if (!is.list(flips)) {
+#    d = flips_to_kplot_df(flips)
+#    p = qplot(k, path, data = d, geom = "line") +
+#      scale_x_continuous(breaks = 0:(t + s), limits = c(0, t + s - 1)) +
+#      scale_y_continuous(breaks = 0:s, limits=c(0, s+0.15)) +
+#      geom_segment(x=s, y=s, xend=(t+s-1), yend=s, linetype=2) +
+#      geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, linetype=2)
+##      geom_segment(x=s, y=s, xend=(t+s-1), yend=s, color="green", linetype=1) +
[…]

### 2016-05-30 stat-and-prob-letters/snb-paper.tex
More revisions.

@@
 
-\usepackage{lineno,hyperref}
+\usepackage{lineno}
 
@@
 \usepackage{amsthm}
-\usepackage{url}
+%\usepackage{url}
 \usepackage{subfig}
@@
 are enrolled and treated. The binomial probability of success is $p=0.045$ for
-each patient under the null hypothesis that treatment is no effective.
+each patient under the null hypothesis that treatment is not effective.
 If two or more patients respond to the treatment then we reject this null
-hypothesis, and the trial is deemed successful at significance level of $0.1$. 
+hypothesis and the trial is deemed successful at significance level of $0.1$. 
 If fewer than two respond then the null is not rejected.
@@
 If all 12 patients are enrolled at once, as in the classic
-design, then the sample size is 12. However, as in most clinical trials, the
+design, then the sample size is 12. However, in most clinical trials the
 patients are enrolled sequentially, often with one patient's outcome realized
@@
 observed treatment failures also ends the trial. This sampling mechanism, in
-which the experiment ends as soon as any of the endpoints is reached, is
+which the experiment ends as soon as a predefined endpoints is reached, is
 call {\em curtailed sampling}. Under curtailed sampling the range of the
@@
 endpoints for the trial. In this case two successes were reached after
-enrolling 10 patients: one in the third outcome and on in the tenth.
+enrolling 10 patients: one in the fourth outcome and on in the eleventh.
 
@@
 and explores properties of the distribution.
-The next section introduces our notation and basic results
-including the density of the distribution along with a description of
-its relation to other distributions. Section 2 derives the distribution
+%The next section introduces our notation and basic results
+%including the density of the distribution along with a description of
+%its relation to other distributions. 
+Section 2 derives the distribution
 based on a defined Bernoulli process and gives some basic properties.
-Section 3 provides a connection to the Binomial tail probability.
-Section 4 derives the posterior distribution using a Beta prior distribution
-for $p$. Section 5 provides a brief discussion on the use of the distribution
+Section 3 shows how the distribution is related to other standard
+distributions and connects the SNB tail probability to the Binomial tail 
+probability.
+Section 4 derives the moment generating function.
+Section 5 derives the posterior distribution using a Beta prior distribution
+for $p$. Section 6 provides a brief discussion on the use of the distribution
 in clinical trials along with future avenues for generalization.
@@
 \end{equation}
-where $I_{\{f\}}$ is one if $f$ is true and zero otherwise, and
+where $I_{\{f\}}$ is the {\emph indicator function}, taking a value 
+of one if $f$ is true and zero otherwise, and
 \begin{equation} \label{eqn:N}
@@
 $k$. If the outcome of the $k$th patient is a success then the process 
-advances one diagonally in the
-positive horizontal and vertical direction. If the $k$th patient fails
+advances diagonally in the positive horizontal and vertical direction. 
+If the $k$th patient fails
 then the sample path advances in the positive horizontal direction only. The
@@
 where we substitute $i=k-s$ in the first summation and $j=k-t$ in the second.
-
 Then $R$ can be written as the cumulative distribution function of two
@@
 \citep{Olver2010} and recall this function satisfies 
-$\mathcal{I}_p(s, t) = 1-\mathcal{I}_{1-p}(t, s)$ \citep{Abramowitz1964}.
+$\mathcal{I}_p(s, t) = 1-\mathcal{I}_{1-p}(t, s)$ \citep{Uppuluri1970}.
 \begin{align*}
@@
 
+The SNB also generalizes both the minimum (riff-shuffle) and maximum negative
+binomial distributions up to a translation of the support.
 For the special case of $\,s=t,$ the distribution of $\,Y\,$ is the
@@
 
-There is a close connection between the tail areas of the SNB and the 
+There is a close connection between the tail probabilities of the SNB and the 
 binomial distributions.
@@
 \begin{equation*}
-\mathbb{P} [Y \leq n\ |\ \text{trial succeeds}=s] 
+\mathbb{P} [Y \leq n\ |\ \text{trial succeeds}] 
   = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s}.
@@
 \begin{align}
-\mathbb{P} [Y \leq n\ |\ \text{\#success} = s] &= \sum_{i=0}^{t-1} 
+\mathbb{P} [Y \leq n\ |\ \text{trial succeeds}] &= \sum_{i=0}^{t-1} 
   {i+s-1 \choose i} p^s (1-p)^i\\
@@
 in Fig.~\ref{fig:snb_bin_compare}, 
-where $s=2$, $t=12$, and $p=0.2$.  The probability masses represented in 
+where $s=2$, $t=12$, and $p=0.045$.  The probability masses represented in 
 black are equal in a) and b) as are the masses in grey.
@@
 \caption{
-SNB(0.2, 2, 12) with mass contributed from 
-$s$ successes (black) or $t$ failures (grey) along with Bin(0.2, 13) with
+SNB(0.045, 2, 12) with mass contributed from 
+$s$ successes (black) or $t$ failures (grey) along with Bin(0.045, 13) with
 at least 2 successes (black) or fewer (grey).
@@
 to one, we have $qe^x \leq 1$. This also shows $x \leq -\log(q)$.
-
 A similar expression can be derived for the second summation in 
@@
 \begin{proof}
-For notational simplicity, assume that $\min(s,t) \leq k \leq s+t-1$. When 
+For notational simplicity, assume that $k$ is in the range
+$\min(s,t) \leq k \leq s+t-1$. When 
 this is not the case appropriate terms should be removed as dictated by the indicator functions.
@@
 
-The SNB is being used to generalize the Simon design by providing separate
-criteria for futility and efficacy.
-The Simon design tests for a single response probability and time to event
-parameter for both the first
-and second stage. The SNB is being used to generalize this design by 
-allowing these parameters to vary for the two stages, and even allow
-information about patients in the first stage to be used in the second
-stage. Early results show the design provides
-both early testing for futility along with the ability
-to make statistically sound decisions sooner than traditional trials.
+%The SNB is being used to generalize the Simon design by providing separate
+%criteria for futility and efficacy.
+%The Simon design tests for a single response probability and time-to-event
+%parameter for both the first
+%and second stage. The SNB is being used to generalize this design by 
+%allowing these parameters to vary for the two stages, and even allow
+%information about patients in the first stage to be used in the second
+%stage. Early results show the design provides
+%both early testing for futility along with the ability
+%to make statistically sound decisions sooner than traditional trials.
 

### 2016-05-30 stat-and-prob-letters/s.r

@@ -2,37 +2,37 @@ library(snb)
 library(ggplot2)
 
 ## kplot
-#
-#kplot = function(flips, s, t) {
-#  if (!is.list(flips)) {
-#    d = flips_to_kplot_df(flips)
-#    p = qplot(k, path, data = d, geom = "line") +
-#      scale_x_continuous(breaks = 0:(t + s), limits = c(0, t + s - 1)) +
-#      scale_y_continuous(breaks = 0:s, limits=c(0, s+0.15)) +
-#      geom_segment(x=s, y=s, xend=(t+s-1), yend=s, linetype=2) +
-#      geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, linetype=2)
-##      geom_segment(x=s, y=s, xend=(t+s-1), yend=s, color="green", linetype=1) +
-##      geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, col="red")
-#  } else {
-#    flip_set = lapply(flips, flips_to_kplot_df)
-#    for (i in 1:length(flip_set)) {
-#      flip_set[[i]]$num = as.factor(i)
-#      flip_set[[i]]$k = jitter(flip_set[[i]]$k)
-#      flip_set[[i]]$k[flip_set[[i]]$k < 0] = 0
-#    }
-#    d = Reduce(rbind, flip_set)[, -(4:5)]
-#    p = qplot(k, path, data = d, geom = "path", group = num) +
-#        scale_x_continuous(breaks=0:(t+s), limits = c(0, t+s)) +
+
+kplot = function(flips, s, t) {
+  if (!is.list(flips)) {
+    d = flips_to_kplot_df(flips)
+    p = qplot(k, path, data = d, geom = "line") +
+      scale_x_continuous(breaks = 0:(t + s), limits = c(0, t + s - 1)) +
+      scale_y_continuous(breaks = 0:s, limits=c(0, s+0.15)) +
+      geom_segment(x=s, y=s, xend=(t+s-1), yend=s, linetype=2) +
+      geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, linetype=2)
+#      geom_segment(x=s, y=s, xend=(t+s-1), yend=s, color="green", linetype=1) +
+#      geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, col="red")
+  } else {
+    flip_set = lapply(flips, flips_to_kplot_df)
[…]

### 2016-06-03 stat-and-prob-letters/snb-paper.tex
More paper revisions.

@@
 
-Consider a prototypical Phase II single-arm clinical trial in which 12 patients
+Consider a prototypical early phase, single-arm clinical trial in which 
+12 patients
 are enrolled and treated. The binomial probability of success is $p=0.045$ for
 each patient under the null hypothesis that treatment is not effective.
-If two or more patients respond to the treatment then we reject this null
-hypothesis and the trial is deemed successful at significance level of $0.1$. 
-If fewer than two respond then the null is not rejected.
+If two or more patients out of these 12 respond to the treatment then we 
+reject this null
+hypothesis and the treatment is deemed successful at significance 
+level of $0.1$.  If fewer than two respond then the null is not rejected
+and the treatment is judged as ineffective.
 
@@
 observed treatment failures also ends the trial. This sampling mechanism, in
-which the experiment ends as soon as a predefined endpoints is reached, is
+which the experiment ends as soon as a predefined endpoint is reached, is
 call {\em curtailed sampling}. Under curtailed sampling the range of the
@@
 The vertical axis denotes the number of
-successful outcomes. The horizontal axis counts the number of patients that
-have been enrolled. The horizontal and vertical boundaries represent
+successful outcomes. The horizontal axis counts the number of patients 
+enrolled. The horizontal and vertical boundaries represent
 endpoints for the trial. In this case two successes were reached after
@@
 \caption{
-A hypothetical realization of a trial.
+A hypothetical realization of the trial.
 }
@@
   \mathcal{I}_{1-pe^x}(t, s) $\\ \hline
+Posterior & 
+${k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)}+
+{k-1 \choose t-1} \frac{B\left(\alpha + k-t, t+\beta\right)}{B(\alpha,\beta)}$\\
+& for Beta($\alpha$, $\beta$) prior. \\ \hline
 \end{tabular}
@@
 The distribution of the stopping time
-$\argminl_k \left[X_k \geq s \cup X_k \leq k-t \right]$
+\begin{equation*}
+Y = \argminl_k \left[X_k \geq s \cup X_k \leq k-t \right]
+\end{equation*}
 is given at (\ref{eqn:pmf}).
@@
 
-To show that Equation \ref{eqn:pmf} sums to one, define
+To show that (\ref{eqn:pmf}) sums to one, define
 \begin{align} \label{eqn:sum_proof}
@@
 Let $\mathcal{I}_p(s, t)$ be the {\em regularized incomplete beta function} 
-\citep{Olver2010} and recall this function satisfies 
-$\mathcal{I}_p(s, t) = 1-\mathcal{I}_{1-p}(t, s)$ \citep{Uppuluri1970}.
+\citep{Olver2010}. This function satisfies 
+$\mathcal{I}_p(s, t) = 1-\mathcal{I}_{1-p}(t, s)$ \citep{Uppuluri1970}. Then
 \begin{align*}
@@
 \end{align*}
-This completes the proof (\ref{eqn:pmf}) is is the distribution of the
-stopping time and a valid probability mass
-function.
+This completes the proof (\ref{eqn:pmf}) is the distribution of the
+stopping time and is a valid probability mass function.
 \end{proof}
@@
 \end{center}
-\caption{Different shapes of the SNB distribution with parameters ($s$, $t$, $p$), as given. Black indicates mass contributed by hitting $s$, grey indicates
-mass contributed by hitting $t$. \label{shapes.fig}}
+\caption{Different shapes of the SNB distribution with parameters 
+($p$, $s$, $t$), as given. Black indicates mass contributed by reaching
+$s$ successes before $t$ failures, grey indicates
+mass contributed by reaching $t$ failures first. \label{shapes.fig}}
 \end{figure}
@@
 Likewise, the probability of reaching the failure endpoint is equal
-to at most $s-1$ successes in a Binomial distribution. That is, the
-probability that the number of successes is at least $s$ in the Binomial
-model is the same that the trial succeeds (reaches $s$) in the SNB model.
+to the probability of at most $s-1$ successes in a Binomial distribution. 
+That is, the probability that the number of successes is at least $s$ in the 
+Binomial model is the same that the trial succeeds (reaches $s$) in the SNB 
+model.
 \begin{prop} \label{binomial_tail}
@@
 \begin{equation}
-\mathbb{P}[B \geq s] = \mathbb{P} [Y \leq n\ |\ \text{trial succeeds}].
+\mathbb{P}[B \geq s] = \mathbb{P} [\text{trial succeeds}].
 \end{equation}
@@
 \begin{equation*}
-\mathbb{P} [Y \leq n\ |\ \text{trial succeeds}] 
+\mathbb{P} [\text{trial succeeds}] 
   = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s}.
@@
 \begin{align}
-\mathbb{P} [Y \leq n\ |\ \text{trial succeeds}] &= \sum_{i=0}^{t-1} 
+\mathbb{P} [\text{trial succeeds}] &= \sum_{i=0}^{t-1} 
   {i+s-1 \choose i} p^s (1-p)^i\\
@@
 
-To illustrate this result, let us return to our initial example, shown
-in Fig.~\ref{fig:snb_bin_compare}, 
-where $s=2$, $t=12$, and $p=0.045$.  The probability masses represented in 
-black are equal in a) and b) as are the masses in grey.
+To illustrate this result, let us return to our initial example
+where $s=2$, $t=12$, and $p=0.045$.  The probability masses in
+Fig.~\ref{fig:snb_bin_compare} represented in 
+black are equal in panels a) and b) as are the masses in grey.
 The probability that $s$
@@
 SNB(0.045, 2, 12) with mass contributed from 
-$s$ successes (black) or $t$ failures (grey) along with Bin(0.045, 13) with
+$s$ successes (black) or $t$ failures (grey) along with Bin(0.045, 12) with
 at least 2 successes (black) or fewer (grey).
@@
 \end{equation}
-The first summation in Equation \ref{eqn:first_sum} satisfies
+The first summation in (\ref{eqn:first_sum}) satisfies
 \begin{align*}
@@
 Since the incomplete beta function's subscript parameter has support on zero 
-to one, we have $qe^x \leq 1$. This also shows $x \leq -\log(q)$.
+to one, we have $qe^x \leq 1$. This also shows we must restrict
+$x \leq -\log(q)$.
 A similar expression can be derived for the second summation in 
@@
 \begin{prop} \label{prop:bayesian}
-The posterior PMF of the Stopped Negative Binomial distribution with a 
+The posterior probability mass function of the Stopped Negative Binomial 
+distribution with a 
 Beta($\alpha$, $\beta$) prior is
 \begin{align} \label{eqn:posterior}
-f(k | s, t, \alpha, \beta) &= 
+\mathbb{P} [Y = k | s, t, \alpha, \beta ] &= 
   {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} 
     \ I_{\{s \leq k \leq s+k-1\}} \ + \nonumber \\
-  & {k-1 \choose k-t} 
+  & {k-1 \choose t-1} 
     \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} 
@@
 f(k | s, t, \alpha, \beta) = \frac{1}{B(\alpha, \beta)} & \int_0^1 {k-1 \choose s-1} p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} + \\
- & {k-1 \choose k-t} p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp \\
+ & {k-1 \choose t-1} p^{k-t+\alpha-1}\left(1-p\right)^{t+\beta-1} dp \\
 = \frac{1}{B(\alpha, \beta)}  {k-1 \choose s-1} & \int_0^1  p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} dp + \\
- & \frac{1}{B(\alpha, \beta)} {k-1 \choose k-t} \int_0^1  p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp.
+ & \frac{1}{B(\alpha, \beta)} {k-1 \choose t-1} \int_0^1  p^{k-t+\alpha-1}\left(1-p\right)^{t+\beta-1} dp.
 \end{align*}
-The result follows by applying the definition of the Beta function to the integral terms.
+The result in (\ref{eqn:posterior}) follows by the definition of the 
+Beta function.
 \end{proof}
+
+To better understand the posterior distribution, let us once again return to 
+our example where $s=2$, $t=12$. However, we assume $p$ is a Beta distribution
+with shape parameters $\alpha = 0.045 c$ and 
+$\beta = 0.955 c$ where $c$ will take values between one and 10. 
+This parameterization allows us to examine the relationship between 
+uncertainty in $p$ and the SNB.
+
+\begin{figure}[t!]
+\centering
+\subfloat[The expected sample size.]{\includegraphics[width=0.5\textwidth]{bayesian-sample-expectation.pdf}}
+\hfill
+\subfloat[The variance of the sample size.]{\includegraphics[width=0.5\textwidth]{bayesian-sample-variance.pdf}}
+\caption{
+The expected size and variance of the posterior SNB varying the uncertainty 
+in the prior distribution of $p$. A larger value of $c$
+corresponds to a Beta prior with smaller variance.
+}
+\label{fig:bayesian-sample-size}
+\end{figure}
+
+Fig.~\ref{fig:bayesian-sample-size} shows how the expected sample size
+and variance of the posterior SNB. As the certainty increases the 
+posterior values converge to corresponding expected values with 
+$p = 0.045$. It may be noted that convergence is relatively slow. This
+is because when $c$ is small, the prior is heavily skewed. As $c$ increases
+this effect is minimized.
+
+Slow convergence in the mean will occur when $p$ takes values close to 
+zero or one. In these cases it may be preferable to select shape paramters
+based on the median or mode of the beta distribution. These values
+will vary less in the uncertainty of the distribution.
 

### 2016-06-03 stat-and-prob-letters/s.r

@@ -123,3 +123,54 @@ d$Outcome = relevel(d$Outcome, "s")
 ggplot(d, aes(x=x, y=y, fill=Outcome))+geom_bar(stat="identity") + xlab("k")+
   ylab("Bin(0.045, 13)") + theme_bw() + scale_fill_grey()
 ggsave("bin_density.pdf")
+
+# Sample size and variance for posterior distribution.
+s = 2
+t = 12
+p = 0.045
+x1 = cdsnb_stacked(min(s, t):(s+t-1), s, t, c(0.045, 0.955))
+stacked_plot(x1$k, x1$s, x1$t)
+dev.new()
+x2 = cdsnb_stacked(min(s, t):(s+t-1), s, t, c(0.045*5, 0.955*5))
+stacked_plot(x2$k, x2$s, x2$t)
+dev.new()
+plot(dbeta(seq(0, 1, by=0.01), 0.045*5, 0.955*5), type="l", col="green")
+lines(dbeta(seq(0, 1, by=0.01), 0.045, 0.955), type="l", col="red")
+#lines(dbeta(seq(0, 1, by=0.01), 0.045*10, 0.955*10), type="l", col="blue")
+
+ecsnb(c(0.5, 0.5), s, t)
+
+x2 = cdsnb_stacked(min(s, t):(s+t-1), s, t, c(0.1*10, 0.9*10))
+stacked_plot(x2$k, x2$s, x2$t)
+
+# Probality of success in Bayesian model.
+# Variance of success in Bayesian model.
+
+shape = c(0.045, 0.955)
+is = seq(1, 100, length.out=100)
+x = foreach (i=is, .combine=rbind) %do% {
+  cbind(i, ecsnb(shape*i, s, t), vcsnb(shape*i, s, t))
+}
+x = as.data.frame(x)
+names(x) = c("n", "e", "v")
+
+exp_size = esnb(0.045, s, t)
+ggplot(x, aes(x=n, y=e)) + geom_line() + 
+  geom_segment(y=exp_size,x=min(is),yend=exp_size, xend=max(is), color="grey") +
+  ylim(range(c(x$e, exp_size, 12))) +
+  geom_text(data=NULL, x=50, y=11.98, label="Deterministic Limit (11.96)") + 
+    theme_bw() + scale_fill_grey() + ylab("Expected Posterior") + xlab("c")
+ggsave("bayesian-sample-expectation.pdf")
+
+exp_var = vsnb(0.045, s, t)
[…]

### 2016-06-07 stat-and-prob-letters/snb-paper.tex
More of Dan's comments integrated.

@@
 enrolling 10 patients: one in the fourth outcome and on in the eleventh.
+Since the success boundary is reached, we say that the trial succeeds,
+or more correctly, the treatment succeeds.
 
@@
 In this work we derive the distribution of the number of enrollees needed
-for either $s$ successes or $t$ failures. We refer to this distribution
-as the Stopped Negative Binomial (SNB) and some of its characteristics are
+to observe either $s$ successes or $t$ failures. We refer to this distribution
+as the Stopped Negative Binomial (SNB). Some of its characteristics are
 summarized in Tab.~\ref{tab:snb}.
@@
 %its relation to other distributions. 
-Section 2 derives the distribution
+Section 2 derives the distribution function
 based on a defined Bernoulli process and gives some basic properties.
@@
 \end{equation}
-where $I_{\{f\}}$ is the {\emph indicator function}, taking a value 
+where $I_{\{f\}}$ is the {\emph indicator function}, taking the value 
 of one if $f$ is true and zero otherwise, and
@@
 The probability a given realization of $\mathbf{X}$ reaches $s$ at
-the $k$th outcome is the probability that, at time $k-1$ there are $s-1$
+the $k$th outcome is the probability that, at time $k-1$, there are $s-1$
 successful outcomes and $k-s$ unsuccessful outcomes multiplied by
-the probability of a success at time $k$. This expression is given
+the probability of the final success at time $k$. This expression is given
 in (\ref{eqn:N}). 
-Similarly, probability a given realization reaches $k-t$
+Similarly, the probability a given realization reaches $k-t$
 is the probability that, at outcome $k-1$ there are $k-t$ successful outcomes
-and $t-1$ unsuccessful outcomes multiplied by the probability of an
+and $t-1$ unsuccessful outcomes multiplied by the probability of a final
 unsuccessful outcome at time $k$.
@@
 To show that (\ref{eqn:pmf}) sums to one, define
-\begin{align} \label{eqn:sum_proof}
-R &= \sum_{k=s}^{s+t-1} S(k, p, s) + \sum_{k=t}^{s+t-1} S(k, 1-p, t) \\
-  &= \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s} + \sum_{k=t}^{s+t-1} {k-1 \choose k-t} p^{k-t} (1-p)^t
-\end{align}
-where we substitute $i=k-s$ in the first summation and $j=k-t$ in the second.
-Then $R$ can be written as the cumulative distribution function of two
+%\begin{align} \label{eqn:sum_proof}
+%R &= \sum_{k=s}^{s+t-1} S(k, p, s) + \sum_{k=t}^{s+t-1} S(k, 1-p, t) \\
+%  &= \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s} + \sum_{k=t}^{s+t-1} {k-1 \choose k-t} p^{k-t} (1-p)^t
+%\end{align}
+\begin{equation*} 
+R = \sum_{k=s}^{s+t-1} S(k, p, s) + \sum_{k=t}^{s+t-1} S(k, 1-p, t).
+\end{equation*}
+If we substitute $i=k-s$ in the first summation and $j=k-t$ in the second then
+$R$ can be written as the cumulative distribution function of two
 negative binomial distributions:
@@
 \begin{corollary} \label{conditional_distribution}
-The number of subsequent enrollements needed 
-to reach an endpoint is SNB($p$, $s-s'$, $t-t'$).
+The number of subsequent enrollments needed 
+to reach either endpoint behaves as SNB($p$, $s-s'$, $t-t'$).
 \end{corollary}
@@
 ($p$, $s$, $t$), as given. Black indicates mass contributed by reaching
-$s$ successes before $t$ failures, grey indicates
+$s$ successes before $t$ failures. Grey indicates
 mass contributed by reaching $t$ failures first. \label{shapes.fig}}
@@
 The Binomial tail probability is
-\begin{align*}
-\mathbb{P}[B \geq s] &= \sum_{k=s}^{s+t-1} {n \choose k} p^k (1-p)^{n-k} \\
-  &= 1 - \sum_{k=0}^{s-1} {n \choose k} p^k (1-p)^{n-k} \\
-  &= 1 - \mathcal{I}_{1-p}.
-\end{align*}
+\begin{equation*}
+\mathbb{P}[B \geq s] = 1 - \mathcal{I}_{1-p}.
+\end{equation*}
+%\begin{align*}
+%\mathbb{P}[B \geq s] &= \sum_{k=s}^{s+t-1} {n \choose k} p^k (1-p)^{n-k} \\
+%  &= 1 - \sum_{k=0}^{s-1} {n \choose k} p^k (1-p)^{n-k} \\
+%  &= 1 - \mathcal{I}_{1-p}.
+%\end{align*}
 The corresponding SNB probability is
@@
 \begin{equation*}
-\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1} {k-1 \choose k-s} p^s q^{k-s} e^{kx} 
-  + \sum_{k=t}^{s+t-1} {k-1 \choose k-t} p^{k-t} q^t e^{kx}
+\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s q^{k-s} e^{kx} 
+  + \sum_{k=t}^{s+t-1} {k-1 \choose t-1} p^{k-t} q^t e^{kx}
 \end{equation*}
@@
 \begin{equation} \label{eqn:first_sum}
-\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1}{k-1 \choose k-s} (pe)^{sx} (qe^x)^{k-s} 
-  + \sum_{k=t}^{s+t-1}{k-1 \choose k-t} (qe^x)^t (pe^x)^{k-t}.
+\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1}{k-1 \choose s-1} (pe)^{sx} (qe^x)^{k-s} 
+  + \sum_{k=t}^{s+t-1}{k-1 \choose t-1} (qe^x)^t (pe^x)^{k-t}.
 \end{equation}
@@
 \begin{align*}
-\sum_{k=s}^{s+t-1}{k-1 \choose k-s} (pe)^{sx} (qe^x)^{k-s} &= 
-  \left(\frac{pe^x}{1 - qe^x}\right)^s \ \ \sum_{k=s}^{s+t-1} {k-1 \choose k-s} 
+\sum_{k=s}^{s+t-1}{k-1 \choose s-1} (pe)^{sx} (qe^x)^{k-s} &= 
+  \left(\frac{pe^x}{1 - qe^x}\right)^s \ \ \sum_{k=s}^{s+t-1} {k-1 \choose s-1} 
     (qe^x)^{k-s} (1-qe^x)^s \\
@@
 
-\section{The Posterior Distribution}
+The SNB's ability to approximate the geometric, normal, and gamma 
+distributions follow from it generalizing the negative binomial. To 
+recover the MGF of the negative binomial consider the case where
+$t \rightarrow \infty$ in (\ref{eqn:mgf}). The regularized incomplete
+beta function in the first term goes to one and zero in the second term
+and we are left with the MGF of the negative binomial distribution. 
+When $s=1$ this is the same as the geometric distribution. The negative
+binomial can therefore be seen as a sum of i.i.d. geometric distributions.
+For an appropriately large number of samples the central limit theorem
+hold yielding a normal approximation.
+
+Drawing a connection to the gamma distribution is more complicated.
+However it is a well-studied problem in the literature (see 
+\cite{Ord1968,Guenther1972,Best1974} for examples). Methods for these
+approximation generally work by equating cumulants and then showing 
+that errors are not too far off.
+
+The lower-half normal distribution can be approximated by setting $s=t$
+for appropriately large $s$ and $t$ and $p = 0.5$. In this
+case the SNB can be viewed as identical, negative binomials
+approximating a normal and truncated at the median.
+
+\section{The Compound Distribution}
 
@@
 \begin{prop} \label{prop:bayesian}
-The posterior probability mass function of the Stopped Negative Binomial 
-distribution with a 
-Beta($\alpha$, $\beta$) prior is
+The compound distribution of the Stopped Negative Binomial 
+distribution with a Beta($\alpha$, $\beta$) prior is
 \begin{align} \label{eqn:posterior}
@@
 
-To better understand the posterior distribution, let us once again return to 
+To better understand the compound distribution, let us once again return to 
 our example where $s=2$, $t=12$. However, we assume $p$ is a Beta distribution
 with shape parameters $\alpha = 0.045 c$ and 
-$\beta = 0.955 c$ where $c$ will take values between one and 10. 
+$\beta = 0.955 c$ for any $c > 0$.
 This parameterization allows us to examine the relationship between 
@@
 \caption{
-The expected size and variance of the posterior SNB varying the uncertainty 
+The expected size and variance of the compound SNB varying the uncertainty 
 in the prior distribution of $p$. A larger value of $c$
@@
 Fig.~\ref{fig:bayesian-sample-size} shows how the expected sample size
-and variance of the posterior SNB. As the certainty increases the 
-posterior values converge to corresponding expected values with 
+and variance of the compound SNB. As the certainty increases the 
+compound distribution sample values converge to corresponding expected 
+values with 
 $p = 0.045$. It may be noted that convergence is relatively slow. This
@@
 Slow convergence in the mean will occur when $p$ takes values close to 
-zero or one. In these cases it may be preferable to select shape paramters
+zero or one. In these cases it may be preferable to select shape parameters
 based on the median or mode of the beta distribution. These values
@@
 
-\section{Discussion}
-
-The SNB can be used to extend clinical trial methodology
-in at least two areas. First, it can be used to construct nested criteria
-for early stopping and efficacy. Current approaches are often inspired by
-the Simon 2-stage Phase II clinical trial \cite{Simon1989} where a
-a small number of patients are enrolled for an initial stage. For patients,
-success or failure is assessed. If at least a pre-specified number of
-successes are observed then the trial enters a second stage with more
-patients. If a sufficient number of successes are not observed then the
-trial fails.
+%\section{Discussion}
+%
+%The SNB can be used to extend clinical trial methodology
+%in at least two areas. First, it can be used to construct nested criteria
+%for early stopping and efficacy. Current approaches are often inspired by
+%the Simon 2-stage Phase II clinical trial \cite{Simon1989} where a
+%a small number of patients are enrolled for an initial stage. For patients,
+%success or failure is assessed. If at least a pre-specified number of
+%successes are observed then the trial enters a second stage with more
+%patients. If a sufficient number of successes are not observed then the
+%trial fails.
 
@@
 
-A second avenue for methodological innovation comes directly from
-Corollary \ref{conditional_distribution} and Proposition \ref{prop:bayesian}.
-Our knowledge of the response probability $p$ comes from published reports
-yielding a prior distribution.
-While a trial is being conducted the response probabilities are 
-updated as new results are received. The number of individual successes
-and failure are used to learn the shape characteristics of the corresponding
-Beta distribution and resulting posterior
-distribution. An estimate of the probability of reaching the success (and
-failure) endpoints can be provided at each step of the trial.
+%A second avenue for methodological innovation comes directly from
+%Corollary \ref{conditional_distribution} and Proposition \ref{prop:bayesian}.
+%Our knowledge of the response probability $p$ comes from published reports
+%yielding a prior distribution.
+%While a trial is being conducted the response probabilities are 
+%updated as new results are received. The number of individual successes
+%and failure are used to learn the shape characteristics of the corresponding
+%Beta distribution and resulting posterior
+%distribution. An estimate of the probability of reaching the success (and
+%failure) endpoints can be provided at each step of the trial.
 

### 2016-06-15 stat-and-prob-letters/snb-paper.tex
More changes based on Dan’s comments.

@@
   \mathcal{I}_{1-pe^x}(t, s) $\\ \hline
-Posterior & 
+Compound Distribution & 
 ${k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)}+
 {k-1 \choose t-1} \frac{B\left(\alpha + k-t, t+\beta\right)}{B(\alpha,\beta)}$\\
-& for Beta($\alpha$, $\beta$) prior. \\ \hline
+& for Beta($\alpha$, $\beta$). \\ \hline
 \end{tabular}
@@
 Section 4 derives the moment generating function.
-Section 5 derives the posterior distribution using a Beta prior distribution
-for $p$. Section 6 provides a brief discussion on the use of the distribution
-in clinical trials along with future avenues for generalization.
+Section 5 derives the compound distribution using a Beta distribution
+for $p$.
 
@@
 That is, the probability that the number of successes is at least $s$ in the 
-Binomial model is the same that the trial succeeds (reaches $s$) in the SNB 
+Binomial model is the same that the treatment succeeds (reaches $s$) in the SNB 
 model.
@@
 \begin{equation}
-\mathbb{P}[B \geq s] = \mathbb{P} [\text{trial succeeds}].
+\mathbb{P}[B \geq s] = \mathbb{P} [\text{treatment succeeds}].
 \end{equation}
@@
 \begin{equation*}
-\mathbb{P} [\text{trial succeeds}] 
+\mathbb{P} [\text{treatment succeeds}] 
   = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s}.
@@
 \begin{align}
-\mathbb{P} [\text{trial succeeds}] &= \sum_{i=0}^{t-1} 
+\mathbb{P} [\text{treatment succeeds}] &= \sum_{i=0}^{t-1} 
   {i+s-1 \choose i} p^s (1-p)^i\\
@@
 
-\section{The Compound Distribution}
+\section{The Compound Probability Distribution}
 
-Let us consider a Bayesian analysis with a Beta prior distribution on $p$.
+Let us consider a Bayesian analysis with a Beta distribution on $p$.
 \begin{prop} \label{prop:bayesian}
 The compound distribution of the Stopped Negative Binomial 
-distribution with a Beta($\alpha$, $\beta$) prior is
+distribution when $p$ is distributed Beta($\alpha$, $\beta$) is
 \begin{align} \label{eqn:posterior}
@@
 \centering
-\subfloat[The expected sample size.]{\includegraphics[width=0.5\textwidth]{bayesian-sample-expectation.pdf}}
+\subfloat[The compound distribution's expected sample size.]{\includegraphics[width=0.5\textwidth]{bayesian-sample-expectation.pdf}}
 \hfill
-\subfloat[The variance of the sample size.]{\includegraphics[width=0.5\textwidth]{bayesian-sample-variance.pdf}}
+\subfloat[The compound distribution's variance.]{\includegraphics[width=0.5\textwidth]{bayesian-sample-variance.pdf}}
 \caption{
-The expected size and variance of the compound SNB varying the uncertainty 
-in the prior distribution of $p$. A larger value of $c$
-corresponds to a Beta prior with smaller variance.
+The expected size and variance of the compound SNB with $s=2$, $t=12$
+with $\mathbb{E}\ p = 0.045$ varying the uncertainty 
+in the distribution of $p$. A larger value of $c$
+corresponds to a Beta with smaller variance.
 }
@@
 values with 
-$p = 0.045$. It may be noted that convergence is relatively slow. This
-is because when $c$ is small, the prior is heavily skewed. As $c$ increases
-this effect is minimized.
+$p = 0.045$. It may be noted that the compound distribution's mean
+approaches the deterministic limit slowly. This
+is because when $c$ is small, the distribution of $p$ is heavily skewed. 
+As $c$ increases this effect is minimized.
 

### 2016-06-15 stat-and-prob-letters/s.r

@@ -3,49 +3,50 @@ library(ggplot2)
 
 ## kplot
 
-kplot = function(flips, s, t) {
-  if (!is.list(flips)) {
-    d = flips_to_kplot_df(flips)
-    p = qplot(k, path, data = d, geom = "line") +
-      scale_x_continuous(breaks = 0:(t + s), limits = c(0, t + s - 1)) +
-      scale_y_continuous(breaks = 0:s, limits=c(0, s+0.15)) +
-      geom_segment(x=s, y=s, xend=(t+s-1), yend=s, linetype=2) +
-      geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, linetype=2)
-#      geom_segment(x=s, y=s, xend=(t+s-1), yend=s, color="green", linetype=1) +
-#      geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, col="red")
-  } else {
-    flip_set = lapply(flips, flips_to_kplot_df)
-    for (i in 1:length(flip_set)) {
-      flip_set[[i]]$num = as.factor(i)
-      flip_set[[i]]$k = jitter(flip_set[[i]]$k)
-      flip_set[[i]]$k[flip_set[[i]]$k < 0] = 0
-    }
-    d = Reduce(rbind, flip_set)[, -(4:5)]
-    p = qplot(k, path, data = d, geom = "path", group = num) +
-        scale_x_continuous(breaks=0:(t+s), limits = c(0, t+s)) +
-        geom_segment(x = s, y = s, xend = (t + s - 1), yend = s,
-                     linetype=2) +
-        geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, linetype=2)
+#kplot = function(flips, s, t) {
+#  if (!is.list(flips)) {
+#    d = flips_to_kplot_df(flips)
+#    p = qplot(k, path, data = d, geom = "line") +
+#      scale_x_continuous(breaks = 0:(t + s), limits = c(0, t + s - 1)) +
+#      scale_y_continuous(breaks = 0:s, limits=c(0, s+0.15)) +
+#      geom_segment(x=s, y=s, xend=(t+s-1), yend=s, linetype=2) +
+#      geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, linetype=2)
[…]

### 2016-06-15 stat-and-prob-letters/snb-paper.tex
Yeah.

@@
 
+\begin{figure}[bp!]
+\includegraphics[width=\textwidth]{KanePlot.pdf}
+\caption{
+A hypothetical realization of the trial.
+}
+\label{fig:kane_viz}
+\end{figure}
+
 A hypothetical sample path is illustrated in Fig.~\ref{fig:kane_viz}.
@@
 
-\begin{figure}[t!]
-\includegraphics[width=\textwidth]{KanePlot.pdf}
+\begin{figure}[bp!]
+\centering
+\includegraphics[width=\textwidth]{snb-first-plot.pdf}
 \caption{
-A hypothetical realization of the trial.
+The probability mass function for a trial where $p$ 
+is 0.045 and where the trial stops
+after two patients respond or 11 patients do not respond.
 }
-\label{fig:kane_viz}
+\label{fig:first-plot}
 \end{figure}
+
+More generally the distribution for number of trial enrollments is shown in 
+Fig.~\ref{fig:first-plot}. The probability mass is concentrated at the 
+$11^{\text{th}}$ and $12^{\text{th}}$ step since $p$ 
+is small and for most realizations reach 11 
+non-responders before reaching 2 responders.
+
+\begin{figure}[bp!]
+\centering
+\subfloat[Expected value]{\includegraphics[width=0.5\textwidth]{dist-mean.pdf}}
+\hfill
+\subfloat[Variance]{\includegraphics[width=0.5\textwidth]{dist-var.pdf}}
+\caption{
+The mean and variance for a trial that stops after two patients respond
+or 11 patients do not respond when $p$ varies between zero and one.
+}
+\label{fig:exp-and-var}
+\end{figure}
+
+Fig.~\ref{fig:exp-and-var} shows the expected value and variance for the
+number of trial enrollments varying $p$ between zero an 1. When $p$ is
+small the probability of an individual response is small, the treatment
+is more likely to fail, and the number of enrollees is large. When $p$
+is large the treatment is more likely to succeed and the number of 
+enrolees is small. Since $p=1$ and $p=0$ represent the deterministic
+the corresponding variance is zero. The variance is right-skewed because
+the support of the distribution corresponding to successful treatment
+is larger than that of failure.
 
@@
 
-In this work we derive the distribution of the number of enrollees needed
+In the rest of this work we derive the distribution of the number of 
+enrollees needed
 to observe either $s$ successes or $t$ failures. We refer to this distribution
@@
 summarized in Tab.~\ref{tab:snb}.
-The rest of this paper derives these results
+This paper derives these results
 and explores properties of the distribution.
@@
 
-\begin{figure}[t!]
+\begin{figure}[bp!]
 \centering

### 2016-06-15 stat-and-prob-letters/s.r

@@ -183,7 +183,7 @@ ggsave("bayesian-sample-variance.pdf")
 # a) b)
 
 dsnb_plot(p, s, t) + ylab("Probability")
-ggsave("snb-first-plot.pdf")
+ggsave("snb-first-plot.pdf", width=10, height=4)
 
 ps = seq(0, 1, by=0.01)
 moments = foreach(p=ps, .combine=rbind) %do% {
@@ -194,9 +194,9 @@ m = as.data.frame(cbind(ps, moments))
 names(m) = c("p", "exp", "var")
 rownames(m) = NULL
 
-ggplot(m, aes(x=p, y=exp)) + geom_line() + ylab("Distribution Mean")
-ggsave("dist-mean.pdf")
+p = ggplot(m, aes(x=p, y=exp)) + geom_line() + ylab("Distribution Mean")
+ggsave("dist-mean.pdf", p)
 
-ggplot(m, aes(x=p, y=var)) + geom_line() + ylab("Distribution Variance")
-ggsave("dist-var.pdf")
+p = ggplot(m, aes(x=p, y=var)) + geom_line() + ylab("Distribution Variance")
+ggsave("dist-var.pdf", p)
 

### 2016-06-16 stat-and-prob-letters/snb-paper.tex
Release candidate?

@@
 %% or include affiliations in footnotes:
-\author[mymainaddress]{Michelle DeVeaux}
+\author{Michelle DeVeaux}
 \ead{michelle.deveaux@yale.edu}
 
-\author[mymainaddress]{Michael J. Kane\corref{mycorrespondingauthor}}
-\cortext[mycorrespondingauthor]{Corresponding author}
+\author{Michael J. Kane$^*$\footnote{* Corresponding author}}
 \ead{michael.kane@yale.edu}
 
-\author[mymainaddress]{Daniel Zelterman}
+\author{Daniel Zelterman}
 \ead{daniel.zelterman@yale.edu}
 
-\address[mymainaddress]{Department of Biostatistics\\ School of Epidemiology and Public Health\\ Yale University, New Haven, CT}
+\address{Department of Biostatistics\\ School of Epidemiology and Public Health\\ Yale University, New Haven, CT}
 
@@
 
-\section{Introdution and Motivation}
+\section{Introduction and Motivation}
 
@@
 which the experiment ends as soon as a predefined endpoint is reached, is
-call {\em curtailed sampling}. Under curtailed sampling the range of the
+called {\em curtailed sampling}. Under curtailed sampling the range of the
 sample size for this trial is between two and 12.
@@
 Similarly, the probability a given realization reaches $k-t$
-is the probability that, at outcome $k-1$ there are $k-t$ successful outcomes
+is the probability that, at outcome $k-1$, there are $k-t$ successful outcomes
 and $t-1$ unsuccessful outcomes multiplied by the probability of a final
@@
 \end{align*}
-This completes the proof (\ref{eqn:pmf}) is the distribution of the
+This completes the proof that (\ref{eqn:pmf}) is the distribution of the
 stopping time and is a valid probability mass function.
@@
 observe at least $s$ successes {\em and} $s$ failures. The SNB is the 
-number of outcomes to observe {\em either} $s$ successes or $t$ failures.
+number of outcomes to observe {\em either} $s$ successes {\em or} $t$ failures.
 
@@
 To illustrate this result, let us return to our initial example
-where $s=2$, $t=12$, and $p=0.045$.  The probability masses in
+where $s=2$, $t=11$, and $p=0.045$.  The probability masses in
 Fig.~\ref{fig:snb_bin_compare} represented in 
@@
 \caption{
-SNB(0.045, 2, 12) with mass contributed from 
+SNB(0.045, 2, 11) with mass contributed from 
 $s$ successes (black) or $t$ failures (grey) along with Bin(0.045, 12) with
@@
 
-The SNB's ability to approximate the geometric, normal, and gamma 
+The SNB's ability to approximate the geometric, normal, gamma, and poisson
 distributions follow from it generalizing the negative binomial. To 
@@
 $t \rightarrow \infty$ in (\ref{eqn:mgf}). The regularized incomplete
-beta function in the first term goes to one and zero in the second term
-and we are left with the MGF of the negative binomial distribution. 
-When $s=1$ this is the same as the geometric distribution. The negative
+beta function in the first term goes to one and zero in the second term.
+We are left with the MGF of the negative binomial distribution.
+
+When $s=1$ this is the geometric distribution. The negative
 binomial can therefore be seen as a sum of i.i.d. geometric distributions.
 For an appropriately large number of samples the central limit theorem
-hold yielding a normal approximation.
+yields a normal approximation.
 
-Drawing a connection to the gamma distribution is more complicated.
-However it is a well-studied problem in the literature (see 
-\cite{Ord1968,Guenther1972,Best1974} for examples). Methods for these
-approximation generally work by equating cumulants and then showing 
-that errors are not too far off.
+Drawing connections to the gamma and poisson distributions are more complicated.
+However a connection to the gamma distribution well-studied problem in the 
+literature (see \cite{Ord1968,Guenther1972,Best1974} for examples). 
+A connection to the poisson appears in \cite{Anscombe1950} where
+it is shown that if the mean of a poisson is proportional
+to a $\chi^2_{2k}$ distribution then the negative binomial is obtained. 
+All of these results work by equating cumulants and then showing 
+that differences between betwen the cumulant generating
+functions converge to zero.
 
@@
 To better understand the compound distribution, let us once again return to 
-our example where $s=2$, $t=12$. However, we assume $p$ is a Beta distribution
+our example where $s=2$, $t=11$. However, we assume $p$ is a Beta distribution
 with shape parameters $\alpha = 0.045 c$ and 
@@
 
-\begin{figure}[bp!]
+\begin{figure}[h!]
 \centering
@@
 \caption{
-The expected size and variance of the compound SNB with $s=2$, $t=12$
+The expected size and variance of the compound SNB with $s=2$, $t=11$
 with $\mathbb{E}\ p = 0.045$ varying the uncertainty 

### 2016-06-20 stat-and-prob-letters/snb-paper.tex
Plots and captions are based on SNB(.2, 7, 11).

@@
 
+%\begin{figure}[bp!]
+%\centering
+%\includegraphics[width=\textwidth]{snb-first-plot.pdf}
+%\caption{
+%The probability mass function for a trial where $p$ 
+%is 0.045 and where the trial stops
+%after two patients respond or 11 patients do not respond.
+%}
+%\label{fig:first-plot}
+%\end{figure}
+
 \begin{figure}[bp!]
 \centering
-\includegraphics[width=\textwidth]{snb-first-plot.pdf}
+\subfloat[enrollment distribution]{\includegraphics[width=0.5\textwidth]{snb-first-plot.pdf}}
+\hfill
+\subfloat[effect of varying $p$]{\includegraphics[width=0.5\textwidth]{mean-and-variance.pdf}}
 \caption{
-The probability mass function for a trial where $p$ 
-is 0.045 and where the trial stops
-after two patients respond or 11 patients do not respond.
+The distribution of a trial that stops after seven patients respond
+or 11 patients do not. (a) shows the distribution when
+$p = 0.2$. (b) shows the result of varying $p$ between 
+zero and one.
 }
-\label{fig:first-plot}
+\label{fig:exp-and-var}
 \end{figure}
@@
 More generally the distribution for number of trial enrollments is shown in 
-Fig.~\ref{fig:first-plot}. The probability mass is concentrated at the 
+Fig.~\ref{fig:first-plot} (a). The probability mass is concentrated at the 
 $11^{\text{th}}$ and $12^{\text{th}}$ step since $p$ 
@@
 non-responders before reaching 2 responders.
-
-\begin{figure}[bp!]
-\centering
-\subfloat[Expected value]{\includegraphics[width=0.5\textwidth]{dist-mean.pdf}}
-\hfill
-\subfloat[Variance]{\includegraphics[width=0.5\textwidth]{dist-var.pdf}}
-\caption{
-The mean and variance for a trial that stops after two patients respond
-or 11 patients do not respond when $p$ varies between zero and one.
-}
-\label{fig:exp-and-var}
-\end{figure}
 
@@
 \centering
-\subfloat[The SNB distribution.]{\includegraphics[width=0.5\textwidth]{snb_density.pdf}}
+\subfloat[SNB distribution]{\includegraphics[width=0.5\textwidth]{snb_density.pdf}}
 \hfill
-\subfloat[The Binomial distribution.]{\includegraphics[width=0.5\textwidth]{bin_density.pdf}}
+\subfloat[Binomial distribution.]{\includegraphics[width=0.5\textwidth]{bin_density.pdf}}
 \caption{
@@
 \centering
-\subfloat[The compound distribution's expected sample size.]{\includegraphics[width=0.5\textwidth]{bayesian-sample-expectation.pdf}}
+\subfloat[mean]{\includegraphics[width=0.5\textwidth]{bayesian-sample-expectation.pdf}}
 \hfill
-\subfloat[The compound distribution's variance.]{\includegraphics[width=0.5\textwidth]{bayesian-sample-variance.pdf}}
+\subfloat[variance]{\includegraphics[width=0.5\textwidth]{bayesian-sample-variance.pdf}}
 \caption{

### 2016-06-20 stat-and-prob-letters/s.r

@@ -45,18 +45,19 @@ library(ggplot2)
 #  d
 #}
 
+outcomes = c(rep(0, 2), 1, 0, 1, 0, 0, 1, 0, 1, 1, 0, 1, 0, 1)
 
-p = kplot( c(rep(0, 3), 1, rep(0, 6), 1), s=2, t=11, bw=TRUE) +
+p = kplot( outcomes, s=7, t=11, bw=TRUE) +
   xlab("Number of Patients Enrolled") + 
   ylab("Number of Responders to Treatment") +
-  geom_text(data=NULL, x=7, y=2.1, label="Success Boundary") +
-  geom_text(data=NULL, x=11.75, y=0.45, label="Failure Boundary",
-            angle=(45+90)/2-5) + 
-  geom_text(data=NULL, x=7, y=1.1, label="Sample Path") +
+  geom_text(data=NULL, x=11.5, y=6.5, label="Success Boundary") +
+  geom_text(data=NULL, x=14.5, y=2.75, label="Failure Boundary",
+            angle=(45+90)/2-25) + 
+  geom_text(data=NULL, x=6, y=2.5, label="Sample Path") +
   theme_bw() + scale_fill_grey()
+p
 
 ggsave("KanePlot.pdf", p, width=10, height=3.5)
-
 
 # SNB distribution shapes.
 
@@ -74,10 +75,11 @@ approx_plots_start = list(
     labs(x="", y="", title="gamma approximation (0.06, 3, 175)") + theme_bw()+
     scale_fill_grey(),
   dsnb_stack_plot(0.5, 25, 25) + scale_x_discrete(breaks=seq(25, 50, by=5))+
-    labs(x="", y="", title="lower half normal (0.5, 25, 25)") + 
+    labs(x="", y="", title="lower half normal / riff-shuffle  (0.5, 25, 25)") + 
     theme_bw() + scale_fill_grey(),
-  dsnb_stack_plot(0.98, 175, 2) + scale_x_discrete(breaks=seq(2, 180, by=20))+
-    labs(x="", y="", title="local mode at top of range (0.98, 175, 2)") + 
+  dsnb_stack_plot(p=0.45, 25, 25) + 
+    scale_x_discrete(breaks=seq(25, 49, by=3))+
+    labs(x="", y="", title="riff-shuffle (0.45, 25, 25)") + 
     theme_bw() + scale_fill_grey())
 
 library(foreach)
[…]

### 2016-06-22 stat-and-prob-letters/snb-paper.tex
Another day another version.

@@
 Consider a prototypical early phase, single-arm clinical trial in which 
-12 patients
-are enrolled and treated. The binomial probability of success is $p=0.045$ for
-each patient under the null hypothesis that treatment is not effective.
-If two or more patients out of these 12 respond to the treatment then we 
-reject this null
-hypothesis and the treatment is deemed successful at significance 
-level of $0.1$.  If fewer than two respond then the null is not rejected
-and the treatment is judged as ineffective.
+17 patients
+are enrolled and treated. The binomial probability a patient 
+responds to treatment is $p=0.2$ 
+under the null hypothesis that the treatment is not effective.
+If seven or more patients out of these 17 respond to the treatment then we 
+reject the hypothesis and the treatment is deemed successful at 
+significance level of $0.1$.  If fewer than seven respond then the null is not 
+rejected and the treatment is deemed ineffective.
 
-If all 12 patients are enrolled at once, as in the classic
-design, then the sample size is 12. However, in most clinical trials the
-patients are enrolled sequentially, often with one patient's outcome realized
-before the next one enters the trial. In the present example, observing two
-successful patients allows us to reach one endpoint so the sample required
-could be as small as two. Similarly 11
+If all 17 patients are enrolled at once, as in the classic
+design, then the sample size is 17. However, in most clinical trials the
+patients are enrolled sequentially with one patient's outcome realized
+before the next one enters the trial. In the present example, observing seven
+successful patients is one endpoint and the number of enrollees required
+could be as small as seven. Similarly 11
 observed treatment failures also ends the trial. This sampling mechanism, in
-which the experiment ends as soon as a predefined endpoint is reached, is
+which the experiment ends as soon as any predefined endpoint is reached, is
 called {\em curtailed sampling}. Under curtailed sampling the range of the
-sample size for this trial is between two and 12.
+sample size for this trial is between seven and 17.
 
@@
 identically distributed Bernoulli($p$) random variable. The trial is realized
-as a sequence of these random variables stopping when either a
-specified number of successes or failures has been reached.
+as a sequence of Bernoulli samples that stops when either a
+specified number of responders or non-responders have been reached.
 
@@
 enrolled. The horizontal and vertical boundaries represent
-endpoints for the trial. In this case two successes were reached after
-enrolling 10 patients: one in the fourth outcome and on in the eleventh.
-Since the success boundary is reached, we say that the trial succeeds,
-or more correctly, the treatment succeeds.
+endpoints for the trial. In this case a seventh response was reached on
+the $15^{\text{th}}$ enrollment.
+Since the success boundary is reached, we say the treatment succeeds.
 
@@
 \caption{
-The distribution of a trial that stops after seven patients respond
+The distribution of a trial that stops after seven patients respond to treatment
 or 11 patients do not. (a) shows the distribution when
-$p = 0.2$. (b) shows the result of varying $p$ between 
-zero and one.
+$p = 0.2$. (b) shows the mean and variance of the distribution varying $p$ 
+between zero and one.
 }
@@
 
-More generally the distribution for number of trial enrollments is shown in 
-Fig.~\ref{fig:first-plot} (a). The probability mass is concentrated at the 
-$11^{\text{th}}$ and $12^{\text{th}}$ step since $p$ 
-is small and for most realizations reach 11 
-non-responders before reaching 2 responders.
-
-Fig.~\ref{fig:exp-and-var} shows the expected value and variance for the
-number of trial enrollments varying $p$ between zero an 1. When $p$ is
+More generally the distribution of the number of trial enrollments is shown in 
+Fig.~\ref{fig:exp-and-var} (a). There is relatively little probability mass
+for values of seven through 10 since $p$ is small and the treatment succeeding
+quickly is unlikely.
+%The probability mass is concentrated at the 
+%$11^{\text{th}}$ and $12^{\text{th}}$ step since $p$ 
+%is small and for most realizations reach 11 
+%non-responders before reaching 2 responders.
+Fig.~\ref{fig:exp-and-var} (b) shows the expected value and variance for the
+number of trial enrollments varying $p$ between zero and one. When $p$ is
 small the probability of an individual response is small, the treatment
-is more likely to fail, and the number of enrollees is large. When $p$
-is large the treatment is more likely to succeed and the number of 
-enrolees is small. Since $p=1$ and $p=0$ represent the deterministic
-the corresponding variance is zero. The variance is right-skewed because
-the support of the distribution corresponding to successful treatment
-is larger than that of failure.
+is more likely to fail shortly after the $11^{\text{th}}$ enrollment.
+When $p$ is large the treatment is more likely to succeed and the number of 
+enrollees approaches seven from above. 
+Since $p=1$ and $p=0$ represent the deterministic
+case the corresponding variance is zero. The ``bump'' around $p=0.25$ results 
+from inequality in the endpoints.
 
@@
 Section 3 shows how the distribution is related to other standard
-distributions and connects the SNB tail probability to the Binomial tail 
+distributions and connects the SNB tail probability to the binomial tail 
 probability.
 Section 4 derives the moment generating function.
-Section 5 derives the compound distribution using a Beta distribution
-for $p$.
+Section 5 derives the compound distribution when $p$ has a beta distribution.
 
@@
 probability parameter $0\leq p \leq 1$. In the clinical trial setting
-$\,b_i = 1$ corresponds to a successful patient outcome following treatment.  
+$\,b_i = 1$ corresponds to a patient responding to treatment.  
 Let $s$ and $t$ be positive integers.  Define the SNB random
@@
 integer value such that $\,\{b_1, \ldots , b_Y\}\,$ contains {\em either}
-$\,s\,$ successes {\em or} $\,t\,$ failures. That is, the SNB distribution
-of $Y$ is the smallest integer such that either
+$\,s\,$ responders {\em or} $\,t\,$ non-responders. That is, the SNB 
+distribution of $Y$ is the smallest integer such that either
 $\sum_i^Y b_i = s$ or $\sum_i^Y 1-b_i = t$.
@@
 \end{equation}
-where $I_{\{f\}}$ is the {\emph indicator function}, taking the value 
+where $I_{\{f\}}$ is the {\em indicator function}, taking the value 
 of one if $f$ is true and zero otherwise, and
@@
 we consider a graphical illustration of the plot $X_k$ against
-$k$. If the outcome of the $k$th patient is a success then the process 
+$k$. If the outcome of the $k$th patient responds to treatment then the process 
 advances diagonally in the positive horizontal and vertical direction. 
-If the $k$th patient fails
+If the $k$th patient does not respond
 then the sample path advances in the positive horizontal direction only. The
@@
 successful outcomes and $k-s$ unsuccessful outcomes multiplied by
-the probability of the final success at time $k$. This expression is given
+the probability of a final success at time $k$. This expression is given
 in (\ref{eqn:N}). 
@@
 Next, we consider an interim analysis of a clinical trial after $s'$ 
-successes and $t'$ failures have been observed for $s' < s$ and $t' < t$.
+patients respond to treatment 
+and $t'$ do not for $s' < s$ and $t' < t$.
 \begin{corollary} \label{conditional_distribution}
@@
 \end{corollary}
-Having observed $s'$
-successes and $t'$ failures, there are $s-s'$ successes needed to reach the
-success endpoint and $t-t'$ failures needed to reach the failure endpoint.
+Having observed $s'$ responders and $t'$ non-responders, there are $s-s'$ 
+more responders needed to reach the success endpoint and $t-t'$ more 
+non-responders needed to reach the failure endpoint.
 
-\section{Connections and Approximations by Other Distributions}
+\section{Connections and Approximations to Other Distributions}
 
@@
 ($p$, $s$, $t$), as given. Black indicates mass contributed by reaching
-$s$ successes before $t$ failures. Grey indicates
-mass contributed by reaching $t$ failures first. \label{shapes.fig}}
+$s$ responders before $t$ non-responders. Grey indicates
+mass contributed by reaching $t$ non-responders first. \label{shapes.fig}}
 \end{figure}
@@
 The SNB is a generalization of the negative 
-binomial distribution. If $t$ is large then $Y-s$ has a 
+binomial distribution. When $t$ is large $Y-s$ has a 
 negative binomial distribution with
 \begin{equation*}                                    %   (1)   
-\mathbb{P}[Y=s+j]        \label{nb1.eq}          
-  = {{s+j-1}\choose{s-1}} p^s (1-p)^j
+\mathbb{P}[Y=s+j\ |\ t \text{ is large\ }]        \label{nb1.eq}          
+  \rightarrow = {{s+j-1}\choose{s-1}} p^s (1-p)^j
 \end{equation*}
@@
 
-The SNB also generalizes both the minimum (riff-shuffle) and maximum negative
+The SNB is the 
+number of enrollment outcomes to observe {\em either} $s$ responders {\em or} 
+$t$ non-responders.
+The SNB generalizes both the minimum (riff-shuffle) and maximum negative
 binomial distributions up to a translation of the support.
-For the special case of $\,s=t,$ the distribution of $\,Y\,$ is the
+For the special case of $\,s=t,$ the distribution of $\,Y\,$ is exactly the
 riff-shuffle, or minimum negative binomial distribution~\citep{Uppuluri1970}.
-Similar derivations of the closely-related maximum negative binomial 
-discrete distributions also appear in~\cite{Zhang2000}
-and \cite{Zelterman2005}.
-The maximum negative binomial is the smallest number of outcomes necessary to 
-observe at least $s$ successes {\em and} $s$ failures. The SNB is the 
-number of outcomes to observe {\em either} $s$ successes {\em or} $t$ failures.
+The maximum negative binomial \cite{Zhang2000,Zelterman2005} is the smallest 
+number of outcomes necessary to 
+observe at least $s$ responders {\em and} $s$ non-responders and is equivalent
+to a translated version of the riff-shuffle.
 
@@
 
-There is a close connection between the tail probabilities of the SNB and the 
-binomial distributions.
-The probability of reaching the success endpoint in an 
-SNB($p$, $s$, $t$) random variable is 
-equal to the probability of at least $s$ successes in a Binomial distributed 
-random variable with size $s+t-1$ and success probability $p$. 
-Likewise, the probability of reaching the failure endpoint is equal
-to the probability of at most $s-1$ successes in a Binomial distribution. 
-That is, the probability that the number of successes is at least $s$ in the 
-Binomial model is the same that the treatment succeeds (reaches $s$) in the SNB 
+%There is a close connection between the tail probabilities of the SNB and the 
+%binomial distributions.
+%The probability of reaching the success endpoint in an 
+%SNB($p$, $s$, $t$) random variable is 
+%equal to the probability of at least $s$ successes in a binomial distributed 
+%random variable with size $s+t-1$ and success probability $p$. 
+%Likewise, the probability of reaching the failure endpoint is equal
+%to the probability of at most $s-1$ successes in a binomial distribution. 
+There is an equivalence between the probability of reaching an endpoint 
+in the SNB model and the tail probability of the binomial distribution.
+That is, the probability that the number of responders is at least $s$ in the 
+binomial model is the same that the treatment succeeds (reaches $s$) in the SNB 
 model.
@@
 Let $Y$ be distributed as SNB($p$, $s$, $t$) and let 
-$B$ be distributed Binomial with size $n=s+t-1$ and success probability
+$B$ be distributed binomial with size $n=s+t-1$ and response probability
 $p$. Then
@@
 \begin{proof}
-The Binomial tail probability is
+The binomial tail probability is
 \begin{equation*}
@@
 To illustrate this result, let us return to our initial example
-where $s=2$, $t=11$, and $p=0.045$.  The probability masses in
+where $s=7$, $t=11$, and $p=0.2$.  The probability masses in
 Fig.~\ref{fig:snb_bin_compare} represented in 
-black are equal in panels a) and b) as are the masses in grey.
+black are equal in panels (a) and (b) as are the masses in grey.
 The probability that $s$
-successes are reached in the SNB process is the same as the binomial 
-probability of at least two successes. Likewise, the probability that $t$ 
-failures are reached in the SNB process is the same as the binomial
-probability of zero or one successes.
+responders are reached in the SNB process is the same as the binomial 
+probability of at least seven responders. Likewise, the probability that $t$ 
+non-responders are reached in the SNB process is the same as the binomial
+probability of zero through six responders.
 
@@
 \hfill
-\subfloat[Binomial distribution.]{\includegraphics[width=0.5\textwidth]{bin_density.pdf}}
+\subfloat[binomial distribution]{\includegraphics[width=0.5\textwidth]{bin_density.pdf}}
 \caption{
-SNB(0.045, 2, 11) with mass contributed from 
-$s$ successes (black) or $t$ failures (grey) along with Bin(0.045, 12) with
-at least 2 successes (black) or fewer (grey).
+SNB(0.2, 7, 11) with mass contributed from 
+$s$ responders (black) or $t$ non-responders (grey) along with 
+Binomial(0.2, 17) with at least 2 responders (black) or fewer (grey).
 }
@@
 The SNB's ability to approximate the geometric, normal, gamma, and poisson
-distributions follow from it generalizing the negative binomial. To 
-recover the MGF of the negative binomial consider the case where
+distributions follow from it generalizing the negative binomial distribution. 
+To recover the MGF of the negative binomial consider the case where
 $t \rightarrow \infty$ in (\ref{eqn:mgf}). The regularized incomplete
@@
 
-When $s=1$ this is the geometric distribution. The negative
+When $s=1$ (and $t$ is still large) the SNB's MGF is the same as that of the
+geometric distribution. The negative
 binomial can therefore be seen as a sum of i.i.d. geometric distributions.
@@
 to a $\chi^2_{2k}$ distribution then the negative binomial is obtained. 
-All of these results work by equating cumulants and then showing 
+Both of these approximations work by equating cumulants and then showing 
 that differences between betwen the cumulant generating
@@
 
-Let us consider a Bayesian analysis with a Beta distribution on $p$.
+Let us consider the SNB distribution where $p$ has a beta distribution.
 \begin{prop} \label{prop:bayesian}
-The compound distribution of the Stopped Negative Binomial 
+The compound distribution of the SNB
 distribution when $p$ is distributed Beta($\alpha$, $\beta$) is
@@
 The result in (\ref{eqn:posterior}) follows by the definition of the 
-Beta function.
+beta function.
 \end{proof}
@@
 To better understand the compound distribution, let us once again return to 
-our example where $s=2$, $t=11$. However, we assume $p$ is a Beta distribution
-with shape parameters $\alpha = 0.045 c$ and 
-$\beta = 0.955 c$ for any $c > 0$.
+our example where $s=7$, $t=11$. However, we assume $p$ is a beta distribution
+with shape parameters $\alpha = 0.2 c$ and 
+$\beta = 0.8 c$ for any $c > 0$.
 This parameterization allows us to examine the relationship between 
@@
 \caption{
-The expected size and variance of the compound SNB with $s=2$, $t=11$
-with $\mathbb{E}\ p = 0.045$ varying the uncertainty 
+The mean and variance of the compound SNB with $s=7$, $t=11$
+with $\mathbb{E}\ p = 0.2$ varying the uncertainty 
 in the distribution of $p$. A larger value of $c$
-corresponds to a Beta with smaller variance.
+corresponds to a beta with smaller variance.
 }
@@
 
-Fig.~\ref{fig:bayesian-sample-size} shows how the expected sample size
-and variance of the compound SNB. As the certainty increases the 
-compound distribution sample values converge to corresponding expected 
-values with 
-$p = 0.045$. It may be noted that the compound distribution's mean
-approaches the deterministic limit slowly. This
-is because when $c$ is small, the distribution of $p$ is heavily skewed. 
-As $c$ increases this effect is minimized.
-
-Slow convergence in the mean will occur when $p$ takes values close to 
-zero or one. In these cases it may be preferable to select shape parameters
-based on the median or mode of the beta distribution. These values
-will vary less in the uncertainty of the distribution.
+Fig.~\ref{fig:bayesian-sample-size} shows how the mean
+and variance of the compound distribution is related to the SNB's 
+mean and variance. As $c$ increases the 
+compound distribution sample statistics converge to corresponding 
+statistics for the SNB with $p = 0.2$. 
+%It may be noted that the compound 
+%distribution's mean
+%approaches the deterministic limit slowly. This
+%is because when $c$ is small, the distribution of $p$ is heavily skewed. 
+%As $c$ increases this effect is minimized.
+When the distribution of $p$ is centered close to zero or one the mean 
+estimate approach the deterministic limit relatively slowly in the uncertainty
+parameter $c$.  Because of this, we note that for estimation it may be 
+preferable to select shape parameters based on the median or mode of the 
+beta distribution since they are less sensitive to distributional uncertainty.
 
@@
 %and failure are used to learn the shape characteristics of the corresponding
-%Beta distribution and resulting posterior
+%beta distribution and resulting posterior
 %distribution. An estimate of the probability of reaching the success (and

### 2016-06-22 stat-and-prob-letters/s.r

@@ -1,5 +1,6 @@
 library(snb)
 library(ggplot2)
+library(tidyr)
 
 ## kplot
 
@@ -66,11 +67,11 @@ approx_plots_start = list(
     labs(x="", y="", title="bimodal (0.2, 15, 75)") + theme_bw() +
     scale_fill_grey(),
   dsnb_stack_plot(0.35, 50, 50) + scale_x_discrete(breaks=seq(50, 105, by=10))+
-    labs(x="", y="", title="approximate normal (0.35, 50, 50)") + theme_bw() +
+    labs(x="", y="", title="normal approximation (0.35, 50, 50)") + theme_bw() +
     scale_fill_grey(),
   dsnb_stack_plot(0.06, 10, 10) + scale_x_discrete(breaks=seq(10, 20, by=2))+
-    labs(x="", y="", title="geometric (0.06, 10, 10)") + theme_bw() +
-    scale_fill_grey(),
+    labs(x="", y="", title="geometric approximation (0.06, 10, 10)") + 
+    theme_bw() + scale_fill_grey(),
   dsnb_stack_plot(0.06, 3, 175) + scale_x_discrete(breaks=seq(3, 180, by=30))+
     labs(x="", y="", title="gamma approximation (0.06, 3, 175)") + theme_bw()+
     scale_fill_grey(),
@@ -113,7 +114,8 @@ n = s+t-1
 
 plot.new()
 dsnb_stack_plot(p, s, t) + title("SNB(0.2, 7, 11)") + theme_bw() + 
-  scale_fill_grey() + ylab("Probability") + ylim(0, 0.25) 
+  scale_fill_grey() + ylab("Probability") + ylim(0, 0.25) +
+  theme(text = element_text(size=15))
 ggsave("snb_density.pdf")
 
 s1 = dsnb_stacked(min(s,t):n, p, s, t)
@@ -127,7 +129,7 @@ d$Outcome = relevel(d$Outcome, "s")
 ggplot(d, aes(x=factor(x), y=y, fill=Outcome)) +
   geom_bar(stat="identity") + xlab("k")+
   title("Bin(0.2, 17)") + theme_bw() + scale_fill_grey() +
-  ylab("") + ylim(0, 0.25) 
+  ylab("") + ylim(0, 0.25) + theme(text = element_text(size=15))
 ggsave("bin_density.pdf")
 
 # Sample size and variance for posterior distribution.

### 2016-07-15 stat-and-prob-letters/snb-paper.tex
Addressing comments and fixed a typo.

@@
 If seven or more patients out of these 17 respond to the treatment then we 
-reject the hypothesis and the treatment is deemed successful at 
+reject this hypothesis and the treatment is deemed successful at 
 significance level of $0.1$.  If fewer than seven respond then the null is not 
@@
 identically distributed Bernoulli($p$) random variable. The trial is realized
-as a sequence of Bernoulli samples that stops when either a
-specified number of responders or non-responders have been reached.
+as a sequence of Bernoulli samples stopping when either a
+specified number of responders or non-responders has been reached.
 
@@
 \hfill
-\subfloat[effect of varying $p$]{\includegraphics[width=0.5\textwidth]{mean-and-variance.pdf}}
+\subfloat[mean and variance varying $p$]{\includegraphics[width=0.5\textwidth]{mean-and-variance.pdf}}
 \caption{
 The distribution of a trial that stops after seven patients respond to treatment
-or 11 patients do not. (a) shows the distribution when
-$p = 0.2$. (b) shows the mean and variance of the distribution varying $p$ 
-between zero and one.
+or 11 patients do not. Panel (a) shows the distribution when
+$p = 0.2$. The mean and variance of the distribution varying $p$ 
+between zero and one are shown in Panel (b).
 }
@@
 Fig.~\ref{fig:exp-and-var} (a). There is relatively little probability mass
-for values of seven through 10 since $p$ is small and the treatment succeeding
-quickly is unlikely.
+for values of seven through 10 since $p$ is small and it is unlikely 
+the treatment will succeed quickly.
 %The probability mass is concentrated at the 
@@
 number of trial enrollments varying $p$ between zero and one. When $p$ is
-small the probability of an individual response is small, the treatment
+small the probability of an individual response is small and the treatment
 is more likely to fail shortly after the $11^{\text{th}}$ enrollment.
@@
 enrollees approaches seven from above.
+
 Since $p=1$ and $p=0$ represent the deterministic
-case the corresponding variance is zero. The ``bump'' around $p=0.25$ results 
-from inequality in the endpoints.
+case the corresponding variances are zero. The variance is
+determined by the size of the support of the two endpoints and $p$, which 
+weights the contribution of an endpoint based on how likely it is reached.
+When $p$ is small it is more likely the trial will reach the failure endpoint
+which has a support size of seven. As $p$ increases it becomes more likely the
+trial reaches the success endpoint with a support size of 11.
+The ``bump'' around $p=0.25$ results from inequality in the size of the
+support of the two endpoints.
 
@@
 Support & min$(s,t) \leq k \leq s+t-1$  \\ \hline
-$\mathbb{P}[Y=k]$ & ${k-1 \choose s-1} p^s (1-p)^{k-s} + {k-1 \choose t-1} (1-p)^s p^{k-t}$\\ \hline
+$\mathbb{P}[Y=k]$ & ${k-1 \choose s-1} p^s (1-p)^{k-s} I_{\{s \leq k \leq s+t-1\}\}}+ {k-1 \choose t-1} (1-p)^s p^{k-t} I_{\{t \leq k \leq s+t-1\}}$\\ \hline
 $\mathbb{P}[Y \leq k]$ & $2 - \mathcal{I}_{q}(k+1, s) - \mathcal{I}_{p}(k+1, t)$\\ 
@@
   \mathcal{I}_{1-pe^x}(t, s) $\\ \hline
-Compound Distribution & 
+Predictive Distribution & 
 ${k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)}+
 {k-1 \choose t-1} \frac{B\left(\alpha + k-t, t+\beta\right)}{B(\alpha,\beta)}$\\
-& for Beta($\alpha$, $\beta$). \\ \hline
+& for $p$ distributed as Beta($\alpha$, $\beta$). \\ \hline
 \end{tabular}
@@
 Section 4 derives the moment generating function.
-Section 5 derives the compound distribution when $p$ has a beta distribution.
+Section 5 derives the predictive distribution when $p$ has a beta distribution.
 
@@
 \begin{equation} \label{eqn:pmf}
-\mathbb{P} [Y=k] = S(k, p, s) \ I_{\{s \leq k\}} + 
-  S(k, 1-p, t) \ I_{\{ t \leq k \}}
+\mathbb{P} [Y=k] = S(k, p, s) \ I_{\{s \leq k \leq s+t-1\}} + 
+  S(k, 1-p, t) \ I_{\{ t \leq k \leq s+t-1 \}}
 \end{equation}
@@
 \mathbb{P}[Y=s+j\ |\ t \text{ is large\ }]        \label{nb1.eq}          
-  \rightarrow = {{s+j-1}\choose{s-1}} p^s (1-p)^j
+  = {{s+j-1}\choose{s-1}} p^s (1-p)^j
 \end{equation*}
@@
 binomial distributions up to a translation of the support.
-For the special case of $\,s=t,$ the distribution of $\,Y\,$ is exactly the
+For the special case of $\,s=t,$ the distribution of $\,Y\,$ is the
 riff-shuffle, or minimum negative binomial distribution~\citep{Uppuluri1970}.
@@
 That is, the probability that the number of responders is at least $s$ in the 
-binomial model is the same that the treatment succeeds (reaches $s$) in the SNB 
+binomial model is the same as the probability the treatment succeeds 
+(reaches $s$) in the SNB 
 model.
 \begin{prop} \label{binomial_tail}
-Let $Y$ be distributed as SNB($p$, $s$, $t$) and let 
+Let $Y$ be distributed as SNB($p$, $s$, $t$) and recall that $X_Y$ corresponds
+to the number of responders in a trial. Let 
 $B$ be distributed binomial with size $n=s+t-1$ and response probability
@@
 \begin{equation}
-\mathbb{P}[B \geq s] = \mathbb{P} [\text{treatment succeeds}].
+\mathbb{P}[B \geq s] = \mathbb{P} [X_Y = s].
 \end{equation}
@@
 %\end{align*}
-The corresponding SNB probability is
+The corresponding success probability is
 \begin{equation*}
-\mathbb{P} [\text{treatment succeeds}] 
+\mathbb{P} [X_Y = s] 
   = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s}.
 \end{equation*}
-Let $i=k-s$. Using the fact that
+Let $i=k-s$. Since
 \begin{equation*}
-{i+s-1 \choose s-1} = {i+s-1 \choose i}
+{i+s-1 \choose s-1} = {i+s-1 \choose i},
 \end{equation*}
@@
 \begin{align}
-\mathbb{P} [\text{treatment succeeds}] &= \sum_{i=0}^{t-1} 
-  {i+s-1 \choose i} p^s (1-p)^i\\
+\mathbb{P} [X_Y = s] &= \sum_{i=0}^{t-1} {i+s-1 \choose i} p^s (1-p)^i\\
   &= 1 - \mathcal{I}_{1-p}(t, s)
@@
 SNB(0.2, 7, 11) with mass contributed from 
-$s$ responders (black) or $t$ non-responders (grey) along with 
+7 responders (black) or 11 non-responders (grey) along with 
 Binomial(0.2, 17) with at least 2 responders (black) or fewer (grey).
@@
 \end{equation}
-for $q = 1-p$ when $x \leq \min \left\{\log(1/p), \log(1/q) \right\}$.
+for $q = 1-p$ defined for $x \leq \min \left\{\log(1/p), \log(1/q) \right\}$.
 \end{prop}
@@
 \begin{equation} \label{eqn:first_sum}
-\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1}{k-1 \choose s-1} (pe)^{sx} (qe^x)^{k-s} 
+\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1}{k-1 \choose s-1} (pe^x)^{s} (qe^x)^{k-s} 
   + \sum_{k=t}^{s+t-1}{k-1 \choose t-1} (qe^x)^t (pe^x)^{k-t}.
@@
     (qe^x)^{k-s} (1-qe^x)^s \\
-  &= \left(\frac{pe^x}{1 - qe^x}\right)^s \mathcal{I}_{qe^x}(s, t).
+  &= \left(\frac{pe^x}{1 - qe^x}\right)^s \mathcal{I}_{1-qe^x}(s, t).
 \end{align*}
@@
 A similar expression can be derived for the second summation in 
-Equation \ref{eqn:first_sum} and results in
+(\ref{eqn:first_sum}) and results in
 the constraint $x \leq -\log(p)$.
@@
 
-\section{The Compound Probability Distribution}
+\section{The Predictive Probability Distribution}
 
@@
 \begin{prop} \label{prop:bayesian}
-The compound distribution of the SNB
+The predictive distribution of the SNB
 distribution when $p$ is distributed Beta($\alpha$, $\beta$) is
@@
 
-To better understand the compound distribution, let us once again return to 
+To better understand the predictive distribution, let us once again return to 
 our example where $s=7$, $t=11$. However, we assume $p$ is a beta distribution
@@
 \caption{
-The mean and variance of the compound SNB with $s=7$, $t=11$
+The mean and variance of the predictive SNB with $s=7$, $t=11$
 with $\mathbb{E}\ p = 0.2$ varying the uncertainty 
@@
 Fig.~\ref{fig:bayesian-sample-size} shows how the mean
-and variance of the compound distribution is related to the SNB's 
+and variance of the predictive distribution is related to the SNB's 
 mean and variance. As $c$ increases the 
-compound distribution sample statistics converge to corresponding 
+predictive distribution sample statistics converge to corresponding 
 statistics for the SNB with $p = 0.2$. 

### 2016-07-15 stat-and-prob-letters/s.r

@@ -70,7 +70,7 @@ approx_plots_start = list(
     labs(x="", y="", title="normal approximation (0.35, 50, 50)") + theme_bw() +
     scale_fill_grey(),
   dsnb_stack_plot(0.06, 10, 10) + scale_x_discrete(breaks=seq(10, 20, by=2))+
-    labs(x="", y="", title="geometric approximation (0.06, 10, 10)") + 
+    labs(x="", y="", title="poisson approximation (0.06, 10, 10)") + 
     theme_bw() + scale_fill_grey(),
   dsnb_stack_plot(0.06, 3, 175) + scale_x_discrete(breaks=seq(3, 180, by=30))+
     labs(x="", y="", title="gamma approximation (0.06, 3, 175)") + theme_bw()+
@@ -113,9 +113,19 @@ t = 11
 n = s+t-1
 
 plot.new()
-dsnb_stack_plot(p, s, t) + title("SNB(0.2, 7, 11)") + theme_bw() + 
-  scale_fill_grey() + ylab("Probability") + ylim(0, 0.25) +
-  theme(text = element_text(size=15))
+d = as.data.frame(dsnb_stacked(min(s,t):(t+s-1), p = p, s = s, t = t))
+names(d)[2:3] = c("success", "failure")
+d = gather(d, x)
+names(d)[2] = "Outcome"
+d$Outcome = factor(d$Outcome)
+d$Outcome = relevel(d$Outcome, "success")
+ggplot(data = d, aes(x = factor(x), y = value, fill = Outcome)) + 
+    geom_bar(position = "stack", stat = "identity") + xlab("k") + 
+    title("SNB(0.2, 7, 11)") + 
+    ylab("Probability") + ylim(0, 0.25) +
+    theme(text = element_text(size=15)) + 
+    theme_bw() + scale_fill_grey()
+
 ggsave("snb_density.pdf")
 
 s1 = dsnb_stacked(min(s,t):n, p, s, t)
@@ -123,8 +133,8 @@ s1 = dsnb_stacked(min(s,t):n, p, s, t)
 b1 = dbinom(0:n, n, p)
 
 d = data.frame(list(x=0:n, y=b1))
-d$Outcome = factor(c(rep("t", s), rep("s", n-s+1)))
-d$Outcome = relevel(d$Outcome, "s")
+d$Outcome = factor(c(rep("failure", s), rep("success", n-s+1)))
[…]

### 2016-07-18 stat-and-prob-letters/snb-paper.tex
Another day another draft.

@@
 design, then the sample size is 17. However, in most clinical trials the
-patients are enrolled sequentially with one patient's outcome realized
-before the next one enters the trial. In the present example, observing seven
+patients are enrolled sequentially over time.
+In the present example, observing seven
 successful patients is one endpoint and the number of enrollees required
@@
 called {\em curtailed sampling}. Under curtailed sampling the range of the
-sample size for this trial is between seven and 17.
+sample size for this trial is seven to 17.
 
@@
 $p = 0.2$. The mean and variance of the distribution varying $p$ 
-between zero and one are shown in Panel (b).
+between zero and one are shown in panel (b).
 }
@@
 Section 4 derives the moment generating function.
-Section 5 derives the predictive distribution when $p$ has a beta distribution.
+Section 5 derives the predictive and posterior distributions when $p$ has a 
+beta distribution.
 
@@
 \end{equation}
-for $q = 1-p$ defined for $x \leq \min \left\{\log(1/p), \log(1/q) \right\}$.
+for $q = 1-p$ defined for $x < \min \left\{\log(1/p), \log(1/q) \right\}$.
 \end{prop}
@@
 Since the incomplete beta function's subscript parameter has support on zero 
-to one, we have $qe^x \leq 1$. This also shows we must restrict
-$x \leq -\log(q)$.
+to one, we have $0 \leq qe^x < 1$. This also shows we must restrict
+$0 \leq x < -\log(q)$.
 A similar expression can be derived for the second summation in 
 (\ref{eqn:first_sum}) and results in
-the constraint $x \leq -\log(p)$.
+the constraint $0 \leq x < -\log(p)$.
 \end{proof}
@@
 distributions follow from it generalizing the negative binomial distribution. 
-To recover the MGF of the negative binomial consider the case where
-$t \rightarrow \infty$ in (\ref{eqn:mgf}). The regularized incomplete
-beta function in the first term goes to one and zero in the second term.
-We are left with the MGF of the negative binomial distribution.
+\begin{prop}
+The MGF of the SNB converges to that of the negative binomial when either
+$s$ or $t$ gets large. That is
+\begin{equation*}
+\mathbb{E} e^{xY} \rightarrow \left( \frac{pe^x}{1-qe^x} \right)^s
+\end{equation*}
+as $t \rightarrow \infty$. The analogous result holds when 
+$s \rightarrow \infty$.
+\end{prop}
+\begin{proof}
+The second regularized incomplete beta term in (\ref{eqn:mgf}) can be written
+in terms of a cumulative binomial distribution
+\begin{equation*}
+\mathcal{I}_{1-pe^x}(t, s) = \mathbb{P}\left[ B \leq s-1 \right]
+\end{equation*}
+where $B$ is distributed as
+Binomial($t-k$, $pe^x$). By Hoeffding's inequality \cite{Hoeffding1963}
+it follows that
+\begin{equation} \label{eqn:hoeffding}
+\mathbb{P}\left[ B \leq s-1 \right] \leq 
+  \exp \left(-2\frac{((t-k)pe^x - s-1)^2}{t-k} \right).
+\end{equation}
+As $t$ gets large $\mathcal{I}_{1-pe^x}(t, s)$ goes to zero
+Likewise, $\mathcal{I}_{1-qe^x}(s, t)$ goes
+to one. The proof follows by realizing 
+\begin{equation*}
+0 < \frac{qe^x}{1-pe^x} < 1
+\end{equation*}
+over the support of $x$.
+\end{proof}
 
@@
 
-\section{The Predictive Probability Distribution}
+\section{The Posterior and Predictive Probability Distribution}
 
-Let us consider the SNB distribution where $p$ has a beta distribution.
-\begin{prop} \label{prop:bayesian}
+Let us consider the joint SNB distribution when $p$ has a beta distribution.
+\begin{prop} \label{prop:posterior}
+When $p$ is distributed Beta($\alpha$, $\beta$), the 
+posterior density under the SNB model is
+\begin{align*}
+f(p | s, t, \alpha, \beta) &= 
+  \frac{ {k-1 \choose s-1} }{B(\alpha, \beta)} p^{\alpha +s -1} 
+    (1-p)^{k+\beta-s-1} + \\
+  & \frac{ {k-1 \choose t-1} }{B(\alpha, \beta)} p^{k+\alpha -t -1} 
+    (1-p)^{\beta+t-1}.
+\end{align*}
+\end{prop}
+\begin{proof}
+This result can be found directly by multiplying the probability mass function
+in (\ref{eqn:pmf}) by the density of the Beta distribution.
+\end{proof}
+
+\begin{corollary} \label{prop:bayesian}
 The predictive distribution of the SNB
@@
 \end{align}
-\end{prop}
+\end{corollary}
 \begin{proof}
-For notational simplicity, assume that $k$ is in the range
-$\min(s,t) \leq k \leq s+t-1$. When 
-this is not the case appropriate terms should be removed as dictated by the indicator functions.
-\begin{align*}
-f(k | s, t, \alpha, \beta) = \frac{1}{B(\alpha, \beta)} & \int_0^1 {k-1 \choose s-1} p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} + \\
- & {k-1 \choose t-1} p^{k-t+\alpha-1}\left(1-p\right)^{t+\beta-1} dp \\
-= \frac{1}{B(\alpha, \beta)}  {k-1 \choose s-1} & \int_0^1  p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} dp + \\
- & \frac{1}{B(\alpha, \beta)} {k-1 \choose t-1} \int_0^1  p^{k-t+\alpha-1}\left(1-p\right)^{t+\beta-1} dp.
-\end{align*}
-The result in (\ref{eqn:posterior}) follows by the definition of the 
-beta function.
+This result can be found by integrating $p$ over the interval zero to one and
+applying the definition of the beta function.
 \end{proof}
+%\begin{proof}
+%For notational simplicity, assume that $k$ is in the range
+%$\min(s,t) \leq k \leq s+t-1$. When 
+%this is not the case appropriate terms should be removed as dictated by the indicator functions.
+%\begin{align*}
+%f(k | s, t, \alpha, \beta) = \frac{1}{B(\alpha, \beta)} & \int_0^1 {k-1 \choose s-1} p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} + \\
+% & {k-1 \choose t-1} p^{k-t+\alpha-1}\left(1-p\right)^{t+\beta-1} dp \\
+%= \frac{1}{B(\alpha, \beta)}  {k-1 \choose s-1} & \int_0^1  p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} dp + \\
+% & \frac{1}{B(\alpha, \beta)} {k-1 \choose t-1} \int_0^1  p^{k-t+\alpha-1}\left(1-p\right)^{t+\beta-1} dp.
+%\end{align*}
+%The result in (\ref{eqn:posterior}) follows by the definition of the 
+%beta function.
+%\end{proof}
 
 To better understand the predictive distribution, let us once again return to 
-our example where $s=7$, $t=11$. However, we assume $p$ is a beta distribution
-with shape parameters $\alpha = 0.2 c$ and 
-$\beta = 0.8 c$ for any $c > 0$.
-This parameterization allows us to examine the relationship between 
-uncertainty in $p$ and the SNB.
+our example where $s=7$, $t=11$. However, we assume $p$ has a beta distribution
+with shape parameters $\alpha = 0.2 c$ and $\beta = 0.8 c$ for any $c > 0$.
+This parameterization has mean 0.2 for all $c > 0$ and variance 
+$0.16 / (c+1)$.
 
@@
 in the distribution of $p$. A larger value of $c$
-corresponds to a beta with smaller variance.
+corresponds to a beta variance.
 }
@@
 
-Fig.~\ref{fig:bayesian-sample-size} shows how the mean
-and variance of the predictive distribution is related to the SNB's 
-mean and variance. As $c$ increases the 
+As $c$ increases the 
 predictive distribution sample statistics converge to corresponding 
 statistics for the SNB with $p = 0.2$. 
-%It may be noted that the compound 
-%distribution's mean
-%approaches the deterministic limit slowly. This
-%is because when $c$ is small, the distribution of $p$ is heavily skewed. 
-%As $c$ increases this effect is minimized.
-When the distribution of $p$ is centered close to zero or one the mean 
-estimate approach the deterministic limit relatively slowly in the uncertainty
-parameter $c$.  Because of this, we note that for estimation it may be 
+For estimation it may be 
 preferable to select shape parameters based on the median or mode of the 
@@
 
+%%It may be noted that the compound 
+%%distribution's mean
+%%approaches the deterministic limit slowly. This
+%%is because when $c$ is small, the distribution of $p$ is heavily skewed. 
+%%As $c$ increases this effect is minimized.
+%When the distribution of $p$ is centered close to zero or one the mean 
+%estimate approach the deterministic limit relatively slowly in the uncertainty
+%parameter $c$.  
 %\section{Discussion}

### 2016-07-21 stat-and-prob-letters/snb-paper.tex
Revisions and visualization for posterior underway.

@@
 %% remove the % from the second line of the style you would like to use.
-%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Numbered
-%\bibliographystyle{model1-num-names}
-
-%% Numbered without titles
-%\bibliographystyle{model1a-num-names}
-
+%%%%%%%%%%%%%%%%%%%%%%% %% Numbered %\bibliographystyle{model1-num-names} %% Numbered without titles %\bibliographystyle{model1a-num-names} 
 %% Harvard
@@
 17 patients
-are enrolled and treated. The binomial probability a patient 
-responds to treatment is $p=0.2$ 
+are enrolled and treated. The binomial probability of a patient 
+responding to treatment is $p=0.2$ 
 under the null hypothesis that the treatment is not effective.
@@
 reject this hypothesis and the treatment is deemed successful at 
-significance level of $0.1$.  If fewer than seven respond then the null is not 
+significance level of $0.1$.  If fewer than seven respond then the null 
+hypothesis is not 
 rejected and the treatment is deemed ineffective.
@@
 called {\em curtailed sampling}. Under curtailed sampling the range of the
-sample size for this trial is seven to 17.
+sample size for this trial is seven through 17.
 
@@
 Fig.~\ref{fig:exp-and-var} (a). There is relatively little probability mass
-for values of seven through 10 since $p$ is small and it is unlikely 
-the treatment will succeed quickly.
+for values of sample size equal to seven through 10 since $p$ is small and it 
+is unlikely the treatment will succeed quickly.
 %The probability mass is concentrated at the 
@@
 number of trial enrollments varying $p$ between zero and one. When $p$ is
-small the probability of an individual response is small and the treatment
+small then the treatment
 is more likely to fail shortly after the $11^{\text{th}}$ enrollment.
-When $p$ is large the treatment is more likely to succeed and the number of 
-enrollees approaches seven from above.
+When $p$ is large then the treatment is more likely to succeed and the 
+number of enrollees approaches seven from above.
 
-Since $p=1$ and $p=0$ represent the deterministic
-case the corresponding variances are zero. The variance is
-determined by the size of the support of the two endpoints and $p$, which 
-weights the contribution of an endpoint based on how likely it is reached.
-When $p$ is small it is more likely the trial will reach the failure endpoint
-which has a support size of seven. As $p$ increases it becomes more likely the
-trial reaches the success endpoint with a support size of 11.
-The ``bump'' around $p=0.25$ results from inequality in the size of the
-support of the two endpoints.
+When $p=0$ or $1$ the processes are deterministic and variance is zero.
+Value of $p$ between zero and one change the mixing proportions of 
+the two endpoints. The saddle around $p=0.25$ results from inequality 
+in the size of the support of the two endpoints.
 
-\begin{table}[t!]
-\caption{Characteristics of the Stopped Negative Binomial Distribution}
-\label{tab:snb}
-\begin{center}
-\begin{tabular}{|c|l|} \hline
-Parameters & $p$ the success probability ($0\leq p \leq 1$; $q = 1-p$) \\
-           & $s$ the number of successes before stopping ($s=1, 2, ...$)\\
-           & $t$ the number of failures before stopping ($t=1, 2, ...$)\\ \hline
-Support & min$(s,t) \leq k \leq s+t-1$  \\ \hline
-$\mathbb{P}[Y=k]$ & ${k-1 \choose s-1} p^s (1-p)^{k-s} I_{\{s \leq k \leq s+t-1\}\}}+ {k-1 \choose t-1} (1-p)^s p^{k-t} I_{\{t \leq k \leq s+t-1\}}$\\ \hline
-$\mathbb{P}[Y \leq k]$ & $2 - \mathcal{I}_{q}(k+1, s) - \mathcal{I}_{p}(k+1, t)$\\ 
-    & where $\mathcal{I}$ is the regularized incomplete beta function.\\ \hline
-Mean & $\frac{s}{p} \mathcal{I}_p(s,t) + \frac{p^{s-1} q^{t-1}}{B(s,t)} +
-  \frac{t}{q} \mathcal{I}_q(t,s) + \frac{p^{t-1} q^{s-1}}{B(s,t)}$\\ 
-  & where $B$ is the beta function \\ \hline
-MGF & $\left(\frac{p e^x}{1 - qe^x}\right)^s 
-  \mathcal{I}_{1-qe^x} (s, t) + \left(\frac{qe^x}{1-pe^x}\right)^t 
-  \mathcal{I}_{1-pe^x}(t, s) $\\ \hline
-Predictive Distribution & 
-${k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)}+
-{k-1 \choose t-1} \frac{B\left(\alpha + k-t, t+\beta\right)}{B(\alpha,\beta)}$\\
-& for $p$ distributed as Beta($\alpha$, $\beta$). \\ \hline
-\end{tabular}
-\end{center}
-\end{table}
+%\begin{table}[t!]
+%\caption{Characteristics of the Stopped Negative Binomial Distribution}
+%\label{tab:snb}
+%\begin{center}
+%\begin{tabular}{|c|l|} \hline
+%Parameters & $p$ the success probability ($0\leq p \leq 1$; $q = 1-p$) \\
+%           & $s$ the number of successes before stopping ($s=1, 2, ...$)\\
+%           & $t$ the number of failures before stopping ($t=1, 2, ...$)\\ \hline
+%Support & min$(s,t) \leq k \leq s+t-1$  \\ \hline
+%$\mathbb{P}[Y=k]$ & ${k-1 \choose s-1} p^s (1-p)^{k-s} I_{\{s \leq k \leq s+t-1\}\}}+ {k-1 \choose t-1} (1-p)^s p^{k-t} I_{\{t \leq k \leq s+t-1\}}$\\ \hline
+%$\mathbb{P}[Y \leq k]$ & $2 - \mathcal{I}_{q}(k+1, s) - \mathcal{I}_{p}(k+1, t)$\\ 
+%    & where $\mathcal{I}$ is the regularized incomplete beta function.\\ \hline
+%Mean & $\frac{s}{p} \mathcal{I}_p(s,t) + \frac{p^{s-1} q^{t-1}}{B(s,t)} +
+%  \frac{t}{q} \mathcal{I}_q(t,s) + \frac{p^{t-1} q^{s-1}}{B(s,t)}$\\ 
+%  & where $B$ is the beta function \\ \hline
+%MGF & $\left(\frac{p e^x}{1 - qe^x}\right)^s 
+%  \mathcal{I}_{1-qe^x} (s, t) + \left(\frac{qe^x}{1-pe^x}\right)^t 
+%  \mathcal{I}_{1-pe^x}(t, s) $\\ \hline
+%Predictive Distribution & 
+%${k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)}+
+%{k-1 \choose t-1} \frac{B\left(\alpha + k-t, t+\beta\right)}{B(\alpha,\beta)}$\\
+%& for $p$ distributed as Beta($\alpha$, $\beta$). \\ \hline
+%\end{tabular}
+%\end{center}
+%\end{table}
 
@@
 to observe either $s$ successes or $t$ failures. We refer to this distribution
-as the Stopped Negative Binomial (SNB). Some of its characteristics are
-summarized in Tab.~\ref{tab:snb}.
-This paper derives these results
-and explores properties of the distribution.
+as the Stopped Negative Binomial (SNB). 
+%Some of its characteristics are summarized in Tab.~\ref{tab:snb}.
+This paper derives this distribution and explores its properties.
 %The next section introduces our notation and basic results
@@
 Section 5 derives the predictive and posterior distributions when $p$ has a 
-beta distribution.
+beta prior distribution.
 
@@
 then the sample path advances in the positive horizontal direction only. The
-process continues until either $X_k = s$ or $X_k = k-t$.
+process continues until either $X_k = s$ or $X_k = k-t$ corresponding to the
+success and failure boundaries in Fig. \ref{fig:kane_viz}, respectively.
 
@@
 \citep{Olver2010}. This function satisfies 
-$\mathcal{I}_p(s, t) = 1-\mathcal{I}_{1-p}(t, s)$ \citep{Uppuluri1970}. Then
+$\mathcal{I}_p(s, t) = 1-\mathcal{I}_{1-p}(t, s)$ \citep{Uppuluri1967}. Then
 \begin{align*}
@@
 patients respond to treatment 
-and $t'$ do not for $s' < s$ and $t' < t$.
+and $t'$ fail to respond for $s' < s$ and $t' < t$.
 \begin{corollary} \label{conditional_distribution}
 The number of subsequent enrollments needed 
-to reach either endpoint behaves as SNB($p$, $s-s'$, $t-t'$).
+to reach either $s$ or $t$ endpoints behaves as SNB($p$, $s-s'$, $t-t'$).
 \end{corollary}
@@
 The SNB is a generalization of the negative 
-binomial distribution. When $t$ is large $Y-s$ has a 
+binomial distribution. When $t$ is large then $Y-s$ has a 
 negative binomial distribution with
@@
 for $\,j=0, 1,\ldots\,$. A similar statement can be made when $s$ is large
-and $t$ is small. As a result, with proper parameter choice, the SNB
+and $t$ is moderate. As a result, with proper parameter choice, the SNB
 can mimic other probability distributions in a manner similar to 
@@
 
-The SNB is the 
-number of enrollment outcomes to observe {\em either} $s$ responders {\em or} 
-$t$ non-responders.
 The SNB generalizes both the minimum (riff-shuffle) and maximum negative
@@
 For the special case of $\,s=t,$ the distribution of $\,Y\,$ is the
-riff-shuffle, or minimum negative binomial distribution~\citep{Uppuluri1970}.
+riff-shuffle, or minimum negative binomial distribution~\citep{Uppuluri1967}.
 The maximum negative binomial \cite{Zhang2000,Zelterman2005} is the smallest 
@@
 \begin{prop} \label{binomial_tail}
-Let $Y$ be distributed as SNB($p$, $s$, $t$) and recall that $X_Y$ corresponds
-to the number of responders in a trial. Let 
+Let $Y$ be distributed as SNB($p$, $s$, $t$) and let $X_Y$ correspond
+to the number of responders at the end of the trial. Let 
 $B$ be distributed binomial with size $n=s+t-1$ and response probability
@@
 \begin{equation*}
-\mathbb{P}[B \geq s] = 1 - \mathcal{I}_{1-p}.
+\mathbb{P}[B \geq s] = 1 - \mathcal{I}_{1-p}(s, t)
 \end{equation*}
@@
 %\end{align*}
-The corresponding success probability is
-\begin{equation*}
+where $\mathcal{I}_{1-p}(s, t)$ is the {\em regularized incomplete beta} 
+function. The corresponding success probability is
+\begin{equation} \label{eqn:success}
 \mathbb{P} [X_Y = s] 
   = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s}.
-\end{equation*}
+\end{equation}
 Let $i=k-s$. Since
@@
 \end{equation*}
-the last summation can be rewritten as
-\begin{align}
+the summation in (\ref{eqn:success}) can be rewritten as
+\begin{align*}
 \mathbb{P} [X_Y = s] &= \sum_{i=0}^{t-1} {i+s-1 \choose i} p^s (1-p)^i\\
   &= 1 - \mathcal{I}_{1-p}(t, s)
-\end{align}
+\end{align*}
 completing the proof.
@@
 \end{equation}
-for $q = 1-p$ defined for $x < \min \left\{\log(1/p), \log(1/q) \right\}$.
+for $q = 1-p$ and is defined for 
+$x < \min \left\{\log(1/p), \log(1/q) \right\}$.
 \end{prop}
@@
 \end{align*}
-Since the incomplete beta function's subscript parameter has support on zero 
-to one, we have $0 \leq qe^x < 1$. This also shows we must restrict
-$0 \leq x < -\log(q)$.
+Since the regularized incomplete beta function's subscript parameter 
+has support on zero 
+to one, we have $0 \leq qe^x < 1$. This implies
+$x < -\log(q)$.
 A similar expression can be derived for the second summation in 
 (\ref{eqn:first_sum}) and results in
-the constraint $0 \leq x < -\log(p)$.
+the constraint $x < -\log(p)$.
 \end{proof}
 
-The SNB's ability to approximate the geometric, normal, gamma, and poisson
+The SNB's ability to approximate the geometric, normal, gamma, and Poisson
 distributions follow from it generalizing the negative binomial distribution. 
@@
 
-When $s=1$ (and $t$ is still large) the SNB's MGF is the same as that of the
+When $s=1$ (and $t$ is still large) the SNB's MGF is approximately 
+the same as that of the
 geometric distribution. The negative
 binomial can therefore be seen as a sum of i.i.d. geometric distributions.
-For an appropriately large number of samples the central limit theorem
+For an appropriately large number of samples, the central limit theorem
 yields a normal approximation.
 
-Drawing connections to the gamma and poisson distributions are more complicated.
+Drawing connections to the gamma and Poisson distributions are more complicated.
 However a connection to the gamma distribution well-studied problem in the 
 literature (see \cite{Ord1968,Guenther1972,Best1974} for examples). 
-A connection to the poisson appears in \cite{Anscombe1950} where
-it is shown that if the mean of a poisson is proportional
-to a $\chi^2_{2k}$ distribution then the negative binomial is obtained. 
+A connection to the Poisson appears in \cite{Anscombe1950} where
+it is shown that if the mean of a Poisson is proportional
+to a chi-square distribution with $2k$ degrees of freedom
+then the negative binomial is obtained. 
 Both of these approximations work by equating cumulants and then showing 
@@
 
-Let us consider the joint SNB distribution when $p$ has a beta distribution.
-\begin{prop} \label{prop:posterior}
-When $p$ is distributed Beta($\alpha$, $\beta$), the 
-posterior density under the SNB model is
-\begin{align*}
-f(p | s, t, \alpha, \beta) &= 
-  \frac{ {k-1 \choose s-1} }{B(\alpha, \beta)} p^{\alpha +s -1} 
-    (1-p)^{k+\beta-s-1} + \\
+Let us consider the posterior distribution of $p$ with an SNB likelihood 
+and Beta($\alpha$, $\beta$) prior distribution.
+The prior times the likelihood is given by the function
+\begin{align} \label{eqn:ptl}
+f_p(x | \alpha, \beta ) f_{Y|p}(x | k, s, t) &= 
+  \frac{ {k-1 \choose s-1} }{B(\alpha, \beta)} x^{\alpha +s -1} 
+    (1-x)^{k+\beta-s-1} + \\
   & \frac{ {k-1 \choose t-1} }{B(\alpha, \beta)} p^{k+\alpha -t -1} 
-    (1-p)^{\beta+t-1}.
-\end{align*}
-\end{prop}
-\begin{proof}
+    (1-x)^{\beta+t-1} \nonumber
+\end{align}
+where $0 \leq x \leq 1$, $s \leq k \leq s+k-1$, and $t \leq k \leq s+k-1$.
 This result can be found directly by multiplying the probability mass function
 in (\ref{eqn:pmf}) by the density of the Beta distribution.
-\end{proof}
 
-\begin{corollary} \label{prop:bayesian}
-The predictive distribution of the SNB
-distribution when $p$ is distributed Beta($\alpha$, $\beta$) is
-\begin{align} \label{eqn:posterior}
-\mathbb{P} [Y = k | s, t, \alpha, \beta ] &= 
+The predictive distribution of the SNB can be found by integrating
+$x$ over the interval zero to one and applying the definition of the 
+beta function.
+\begin{align} \label{eqn:predictive}
+\int_0^1 f_p(x | \alpha, \beta )  f_{Y|p}(x | k, s, t) & dx = 
   {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} 
-    \ I_{\{s \leq k \leq s+k-1\}} \ + \nonumber \\
+    + \nonumber \\
   & {k-1 \choose t-1} 
-    \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} 
-    \ I_{\{t \leq k \leq s+k-1\}}.
+    \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)}.
 \end{align}
-\end{corollary}
-\begin{proof}
-This result can be found by integrating $p$ over the interval zero to one and
-applying the definition of the beta function.
-\end{proof}
+
+The ratio of (\ref{eqn:ptl}) divided by (\ref{eqn:predictive}) gives the 
+posterior distribution of $p$. It is a mixture of beta distributions. The
+mixing parameter depend on the endpoints ($s$ and $t$), the number of enrollees needed to reach an endpoint ($k$), and the prior parameters ($\alpha$ and
+$\beta$).
+
 %\begin{proof}

### 2016-07-21 stat-and-prob-letters/s.r

@@ -70,7 +70,7 @@ approx_plots_start = list(
     labs(x="", y="", title="normal approximation (0.35, 50, 50)") + theme_bw() +
     scale_fill_grey(),
   dsnb_stack_plot(0.06, 10, 10) + scale_x_discrete(breaks=seq(10, 20, by=2))+
-    labs(x="", y="", title="poisson approximation (0.06, 10, 10)") + 
+    labs(x="", y="", title="Poisson approximation (0.06, 10, 10)") + 
     theme_bw() + scale_fill_grey(),
   dsnb_stack_plot(0.06, 3, 175) + scale_x_discrete(breaks=seq(3, 180, by=30))+
     labs(x="", y="", title="gamma approximation (0.06, 3, 175)") + theme_bw()+
@@ -226,4 +226,19 @@ p = ggplot(y, aes(x=p, y=value)) + geom_line() +
   facet_grid(type ~ ., scales="free") + ylab("") +
   theme(text = element_text(size=15))
 ggsave("mean-and-variance.pdf")
+
+# Assume the Haldane prior.
+pp = function(x, k, s, t) {
+  denom = choose(k-1, s-1) * beta(s, k-s) + choose(k-1, t-1) * beta(k-t, t)
+  ret = foreach(p = x, .combine=rbind) %do% {
+    c(choose(k-1, s-1) * p^(s-1) * (1-p)^(k-s-1), 
+      choose(k-1, t-1) * p^(k-t-1) * (1-p)^(t-1))
+  }
+  ret = cbind(ret, x)
+  colnames(ret) = c("s", "t", "p")
+  rownames(ret) = NULL
+  ret
+}
+
+x= pp(seq(0, 1, by=0.01), 15, s, t)
 

### 2016-08-16 stat-and-prob-letters/snb-paper.tex
Improved hypergeometric distribution interpretation.

@@
 reject this hypothesis and the treatment is deemed successful at 
-significance level of $0.1$.  If fewer than seven respond then the null 
+a significance level of $0.1$.  If fewer than seven respond then the null 
 hypothesis is not 
@@
 which the experiment ends as soon as any predefined endpoint is reached, is
-called {\em curtailed sampling}. Under curtailed sampling the range of the
+called {\em curtailed sampling}. Under curtailed sampling, the range of the
 sample size for this trial is seven through 17.
@@
 enrolled. The horizontal and vertical boundaries represent
-endpoints for the trial. In this case a seventh response was reached on
+endpoints for the trial. In this case, a seventh response was reached on
 the $15^{\text{th}}$ enrollment.
@@
 When $p=0$ or $1$ the processes are deterministic and variance is zero.
-Value of $p$ between zero and one change the mixing proportions of 
+Values between zero and one change the mixing proportions of 
 the two endpoints. The saddle around $p=0.25$ results from inequality 
@@
 
-In the rest of this work we derive the distribution of the number of 
+In the rest of this work, we derive the distribution of the number of 
 enrollees needed
@@
 For the special case of $\,s=t,$ the distribution of $\,Y\,$ is the
-riff-shuffle, or minimum negative binomial distribution~\citep{Uppuluri1967}.
-The maximum negative binomial \cite{Zhang2000,Zelterman2005} is the smallest 
-number of outcomes necessary to 
+riff-shuffle, or minimum negative binomial 
+distribution~\citep{Uppuluri1967,Johnson2005}.
+The maximum negative binomial \cite{Zhang2000,Zelterman2005,Johnson2005} is 
+the smallest number of outcomes necessary to 
 observe at least $s$ responders {\em and} $s$ non-responders and is equivalent
@@
 where $B$ is distributed as
-Binomial($t-k$, $pe^x$). By Hoeffding's inequality \cite{Hoeffding1963}
+Binomial($t-k$, $pe^x$). By Chebychev's inequality %\cite{Hoeffding1963}
 it follows that
@@
 \mathbb{P}\left[ B \leq s-1 \right] \leq 
-  \exp \left(-2\frac{((t-k)pe^x - s-1)^2}{t-k} \right).
+  \frac{ (t-k) pe^x (1-pe^x) }{ \left(s - (t-k)pe^x\right)^2 }
 \end{equation}
@@
 Both of these approximations work by equating cumulants and then showing 
-that differences between betwen the cumulant generating
+that differences between the cumulant generating
 functions converge to zero.
@@
 for appropriately large $s$ and $t$ and $p = 0.5$. In this
-case the SNB can be viewed as identical, negative binomials
+case, the SNB can be viewed as identical, negative binomials
 approximating a normal and truncated at the median.
@@
 
-Let us consider the posterior distribution of $p$ with an SNB likelihood 
-and Beta($\alpha$, $\beta$) prior distribution.
+Let us consider the case where the rate parameter is distributed 
+as Beta($\alpha$, $\beta$) and denoted $P$.
 The prior times the likelihood is given by the function
 \begin{align} \label{eqn:ptl}
-f_p(x | \alpha, \beta ) f_{Y|p}(x | k, s, t) &= 
-  \frac{ {k-1 \choose s-1} }{B(\alpha, \beta)} x^{\alpha +s -1} 
-    (1-x)^{k+\beta-s-1} + \\
+f_P(p, \alpha, \beta ) f_{Y|P}(p, k, s, t) &= 
+  \frac{ {k-1 \choose s-1} }{B(\alpha, \beta)} p^{\alpha +s -1} 
+    (1-p)^{k+\beta-s-1} + \\
   & \frac{ {k-1 \choose t-1} }{B(\alpha, \beta)} p^{k+\alpha -t -1} 
-    (1-x)^{\beta+t-1} \nonumber
+    (1-p)^{\beta+t-1} \nonumber
 \end{align}
-where $0 \leq x \leq 1$, $s \leq k \leq s+k-1$, and $t \leq k \leq s+k-1$.
+where $0 \leq p \leq 1$, $s \leq k \leq s+k-1$, and $t \leq k \leq s+k-1$.
 This result can be found directly by multiplying the probability mass function
@@
 The predictive distribution of the SNB can be found by integrating
-$x$ over the interval zero to one and applying the definition of the 
+$p$ over the interval zero to one and applying the definition of the 
 beta function.
 \begin{align} \label{eqn:predictive}
-\int_0^1 f_p(x | \alpha, \beta )  f_{Y|p}(x | k, s, t) & dx = 
-  {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} 
-    + \nonumber \\
-  & {k-1 \choose t-1} 
+f_{Y}(k, s, &t, \alpha, \beta) = 
+  \int_0^1 f_P(p | \alpha, \beta )  f_{Y|P}(p, k, s, t) dp \nonumber \\ 
+ & = {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} 
+    + {k-1 \choose t-1} 
     \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)}.
 \end{align}
+If both $\alpha$ and $\beta$ are non-negative integers then the predictive
+distribution is a mixture of hypergeometric distributions. 
+\begin{align*} \label{eqn:hypergeo}
+f_{Y}(k, s, t, \alpha, \beta) = 
+  \frac{ {k - 1 \choose s - 1 } 
+         {\alpha + \beta \choose \alpha} }{
+         {\alpha + \beta + k - 1 \choose \alpha + s}}
+  \frac{\alpha}{\alpha + \beta}
+  \frac{ \beta }{ k-s+\beta } +
+  \frac{ {k - 1 \choose t - 1} 
+         {\alpha + \beta \choose \beta} 
+         }{
+         {\alpha + \beta + k -1 \choose t + \beta}
+       }
+  \frac{ \beta }{ \alpha + \beta}
+  \frac{ \alpha }{ k-t + \alpha}
+\end{align*}
+The ratio of combinations in the first term can be interpreted as the
+probability of $s-1$ responders from $k-1$ patients in $\alpha + s$ draws
+from a population size of $\alpha + \beta + k - 1$. This value is multiplied
+by $\alpha / (\alpha + \beta)$, the expected response rate of the prior. 
+The final term in the product weigths the prior based on the number
+of non-responders ($k-s$). Terms in the second summand are interpreted similarly
+for non-responders.
 
 The ratio of (\ref{eqn:ptl}) divided by (\ref{eqn:predictive}) gives the 
-posterior distribution of $p$. It is a mixture of beta distributions. The
+posterior distribution of $P$. It is a mixture of beta distributions. The
 mixing parameter depend on the endpoints ($s$ and $t$), the number of enrollees needed to reach an endpoint ($k$), and the prior parameters ($\alpha$ and
 $\beta$).
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=\textwidth]{beta-mixture.pdf}
+\caption{
+The posterior distribution of the response probability with 
+the Jeffreys prior ($\alpha=\beta=1/2$) for the trial where
+$s=7$, $t=11$, $k=15$.
+}
+\label{fig:beta_mixture}
+\end{figure}
+
+The posterior result above is for the case where the parameters are known
+and the endpoint is not. That is, we do not which boundary was reached.
+Fig.~\ref{fig:beta_mixture} shows this distribution based on the 
+parameters in the hypothetical trial assuming the
+Jeffreys prior \citep{Jeffreys1946}.  If we include the fact that the
+trajectory reaches the endpoint then the second terms in (\ref{eqn:ptl}) and
+(\ref{eqn:predictive}) are both zero and the posterior distribution is
+Beta(7.5, 8.5). This is proportional to the area labeled ``success'' in 
+Fig.~\ref{fig:beta_mixture}.
 
@@
 
-To better understand the predictive distribution, let us once again return to 
-our example where $s=7$, $t=11$. However, we assume $p$ has a beta distribution
-with shape parameters $\alpha = 0.2 c$ and $\beta = 0.8 c$ for any $c > 0$.
-This parameterization has mean 0.2 for all $c > 0$ and variance 
-$0.16 / (c+1)$.
+%To better understand the predictive distribution, let us once again return to 
+%our example where $s=7$, $t=11$. However, we assume $p$ has a beta distribution
+%with shape parameters $\alpha = 0.2 c$ and $\beta = 0.8 c$ for any $c > 0$.
+%This parameterization has mean 0.2 for all $c > 0$ and variance 
+%$0.16 / (c+1)$.
+%
+%
+%\begin{figure}[h!]
+%\centering
+%\subfloat[mean]{\includegraphics[width=0.5\textwidth]{bayesian-sample-expectation.pdf}}
+%\hfill
+%\subfloat[variance]{\includegraphics[width=0.5\textwidth]{bayesian-sample-variance.pdf}}
+%\caption{
+%The mean and variance of the predictive SNB with $s=7$, $t=11$
+%with $\mathbb{E}\ p = 0.2$ varying the uncertainty 
+%in the distribution of $p$. A larger value of $c$
+%corresponds to a beta variance.
+%}
+%\label{fig:bayesian-sample-size}
+%\end{figure}
 
-\begin{figure}[h!]
-\centering
-\subfloat[mean]{\includegraphics[width=0.5\textwidth]{bayesian-sample-expectation.pdf}}
-\hfill
-\subfloat[variance]{\includegraphics[width=0.5\textwidth]{bayesian-sample-variance.pdf}}
-\caption{
-The mean and variance of the predictive SNB with $s=7$, $t=11$
-with $\mathbb{E}\ p = 0.2$ varying the uncertainty 
-in the distribution of $p$. A larger value of $c$
-corresponds to a beta variance.
-}
-\label{fig:bayesian-sample-size}
-\end{figure}
-
-As $c$ increases the 
-predictive distribution sample statistics converge to corresponding 
-statistics for the SNB with $p = 0.2$. 
-For estimation it may be 
-preferable to select shape parameters based on the median or mode of the 
-beta distribution since they are less sensitive to distributional uncertainty.
+%As $c$ increases the 
+%predictive distribution sample statistics converge to corresponding 
+%statistics for the SNB with $p = 0.2$. 
+%For estimation it may be 
+%preferable to select shape parameters based on the median or mode of the 
+%beta distribution since they are less sensitive to distributional uncertainty.
 

### 2016-08-16 stat-and-prob-letters/s.r

@@ -1,6 +1,7 @@
 library(snb)
 library(ggplot2)
 library(tidyr)
+library(latex2exp)
 
 ## kplot
 
@@ -118,13 +119,13 @@ names(d)[2:3] = c("success", "failure")
 d = gather(d, x)
 names(d)[2] = "Outcome"
 d$Outcome = factor(d$Outcome)
-d$Outcome = relevel(d$Outcome, "success")
+d$Outcome = relevel(d$Outcome, "failure")
 ggplot(data = d, aes(x = factor(x), y = value, fill = Outcome)) + 
     geom_bar(position = "stack", stat = "identity") + xlab("k") + 
     title("SNB(0.2, 7, 11)") + 
     ylab("Probability") + ylim(0, 0.25) +
     theme(text = element_text(size=15)) + 
-    theme_bw() + scale_fill_grey()
+    theme_bw() + scale_fill_grey(start=0.8, end=0.2)
 
 ggsave("snb_density.pdf")
 
@@ -134,11 +135,11 @@ b1 = dbinom(0:n, n, p)
 
 d = data.frame(list(x=0:n, y=b1))
 d$Outcome = factor(c(rep("failure", s), rep("success", n-s+1)))
-d$Outcome = relevel(d$Outcome, "success")
+d$Outcome = relevel(d$Outcome, "failure")
 
 ggplot(d, aes(x=factor(x), y=y, fill=Outcome)) +
   geom_bar(stat="identity") + xlab("k")+
-  title("Bin(0.2, 17)") + theme_bw() + scale_fill_grey() +
+  title("Bin(0.2, 17)") + theme_bw() + scale_fill_grey(start=0.8, end=0.2) +
   ylab("") + ylim(0, 0.25) + theme(text = element_text(size=15))
 ggsave("bin_density.pdf")
 
@@ -198,7 +199,8 @@ ggsave("bayesian-sample-variance.pdf")
 # Variance of CSNB size
 # a) b)
 
-dsnb_plot(p, s, t) + ylab("Probability") + theme(text = element_text(size=15))
+dsnb_plot(p, s, t) + ylab("Probability") + theme(text = element_text(size=15)) +
+  theme_bw()
 ggsave("snb-first-plot.pdf")
 
 ps = seq(0, 1, by=0.01)
@@ -224,21 +226,29 @@ y$type[y$type=='exp'] = "Mean"
[…]

### 2016-08-22 stat-and-prob-letters/snb-paper.tex
Paper submitted.

@@
 
-\modulolinenumbers[5]
+%\modulolinenumbers[5]
 
@@
 
-\linenumbers
+%\linenumbers
 
@@
 In the present example, observing seven
-successful patients is one endpoint and the number of enrollees required
+successful patients ends the trial and so the number of enrollees required
 could be as small as seven. Similarly 11
@@
 \begin{figure}[bp!]
-\includegraphics[width=\textwidth]{KanePlot.pdf}
+\centering
+\includegraphics[width=0.9\textwidth]{KanePlot.pdf}
 \caption{
@@
 \centering
-\subfloat[enrollment distribution]{\includegraphics[width=0.5\textwidth]{snb-first-plot.pdf}}
+\subfloat[enrollment distribution]{\includegraphics[width=0.45\textwidth]{snb-first-plot.pdf}}
 \hfill
-\subfloat[mean and variance varying $p$]{\includegraphics[width=0.5\textwidth]{mean-and-variance.pdf}}
+\subfloat[mean and variance varying $p$]{\includegraphics[width=0.45\textwidth]{mean-and-variance.pdf}}
 \caption{
-The distribution of a trial that stops after seven patients respond to treatment
+The distribution of the sample size in a trial that stops after seven 
+patients respond to treatment
 or 11 patients do not. Panel (a) shows the distribution when
@@
 If we substitute $i=k-s$ in the first summation and $j=k-t$ in the second then
-$R$ can be written as the cumulative distribution function of two
+$R$ can be written as the cumulative distribution function (CDF) of two
 negative binomial distributions:
@@
 
-Let $\mathcal{I}_p(s, t)$ be the {\em regularized incomplete beta function} 
-\citep{Olver2010}. This function satisfies 
+The CDF of the negative binomial is commonly expressed in terms of the
+{\em regularized incomplete beta function} \citep{Olver2010}. Using standard
+notation, the first summation in (\ref{eqn:transformed_sum})
+is written as $\mathcal{I}_{1-p}(t, s)$.
+This function satisfies 
 $\mathcal{I}_p(s, t) = 1-\mathcal{I}_{1-p}(t, s)$ \citep{Uppuluri1967}. Then
@@
 \begin{center}
-\includegraphics[width=\textwidth]{shapes.pdf}
+\includegraphics[width=0.8\textwidth]{shapes.pdf}
 \end{center}
@@
 distribution~\citep{Uppuluri1967,Johnson2005}.
-The maximum negative binomial \cite{Zhang2000,Zelterman2005,Johnson2005} is 
+The maximum negative binomial \cite{Johnson2005,Zhang2000,Zelterman2005} is 
 the smallest number of outcomes necessary to 
@@
 However a connection to the gamma distribution well-studied problem in the 
-literature (see \cite{Ord1968,Guenther1972,Best1974} for examples). 
+literature (see \cite{Best1974,Ord1968,Guenther1972} for examples). 
 A connection to the Poisson appears in \cite{Anscombe1950} where

### 2017-02-27 stat-and-prob-letters/snb-paper.tex
yeah.

@@
-\documentclass[review]{elsarticle}
+%\documentclass[review]{elsarticle}
+\documentclass[12pt]{article}
 
@@
 \usepackage{amsthm}
+\usepackage{graphicx}
+\usepackage{natbib}
+\usepackage{setspace}
+\doublespacing
 %\usepackage{url}
@@
 
-\journal{Statistics and Probability Letters}
+%\journal{Statistics and Probability Letters}
 
@@
 %% `Elsevier LaTeX' style
-\bibliographystyle{elsarticle-num}
+%\bibliographystyle{elsarticle-num}
+\bibliographystyle{plain}
 %%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{frontmatter}
+%\begin{frontmatter}
 
-\title{A Stopped Negative Binomial Distribution}
-\tnotetext[mytitlenote]{This research was supported by grants R01CA131301, 
-R01CA157749, R01CA148996, R01CA168733, and PC50CA196530 awarded by the 
-National Cancer Institute along with support from the Yale Comprehensive Cancer 
-Center and the Yale Center for Outcomes Research. We would also like 
-to thank Rick Landin at Ignyta Inc. for his suggestions.}
+\title{Clinical Trial Design Using A Stopped Negative Binomial Distribution}
+\maketitle
+%\tnotetext[mytitlenote]{This research was supported by grants R01CA131301, 
+%R01CA157749, R01CA148996, R01CA168733, and PC50CA196530 awarded by the 
+%National Cancer Institute along with support from the Yale Comprehensive Cancer 
+%Center and the Yale Center for Outcomes Research. We would also like 
+%to thank Rick Landin at Ignyta Inc. for his suggestions.}
 
@@
 %% or include affiliations in footnotes:
-\author{Michelle DeVeaux}
-\ead{michelle.deveaux@yale.edu}
+%\author{Michelle DeVeaux}
+%\ead{michelle.deveaux@yale.edu}
 
-\author{Michael J. Kane$^*$\footnote{* Corresponding author}}
-\ead{michael.kane@yale.edu}
+%\author{Michael J. Kane$^*$\footnote{* Corresponding author}}
+%\ead{michael.kane@yale.edu}
 
-\author{Daniel Zelterman}
-\ead{daniel.zelterman@yale.edu}
+%\author{Daniel Zelterman}
+%\ead{daniel.zelterman@yale.edu}
 
-\address{Department of Biostatistics\\ School of Epidemiology and Public Health\\ Yale University, New Haven, CT}
+%\address{Department of Biostatistics\\ School of Epidemiology and Public Health\\ Yale University, New Haven, CT}
 
@@
 
-\begin{keyword}
-discrete distribution\sep curtailed sampling
-\end{keyword}
+%\begin{keyword}
+%discrete distribution\sep curtailed sampling
+%\end{keyword}
 
-\end{frontmatter}
+%\end{frontmatter}
 
@@
 17 patients
-are enrolled and treated. The binomial probability of a patient 
+are enrolled and treated. Suppose the Bernoulli probability of a patient 
 responding to treatment is $p=0.2$ 
@@
 hypothesis is not 
-rejected and the treatment is deemed ineffective.
+rejected and the treatment is deemed ineffective. The trial is 
+realized as a sequence of Bernoulli($p$) samples stopping when either
+the specified number of responders or non-responders is reached.
 
@@
 
-Let us assume each patient outcome can be modeled as an independent,
-identically distributed Bernoulli($p$) random variable. The trial is realized
-as a sequence of Bernoulli samples stopping when either a
-specified number of responders or non-responders has been reached.
-
 \begin{figure}[bp!]
 \centering
-\includegraphics[width=0.9\textwidth]{KanePlot.pdf}
+\includegraphics[width=\textwidth]{KanePlot.pdf}
 \caption{
@@
 successful outcomes. The horizontal axis counts the number of patients 
-enrolled. The horizontal and vertical boundaries represent
+enrolled. The horizontal and tilted vertical boundaries represent
 endpoints for the trial. In this case, a seventh response was reached on
@@
 \centering
-\subfloat[enrollment distribution]{\includegraphics[width=0.45\textwidth]{snb-first-plot.pdf}}
-\hfill
-\subfloat[mean and variance varying $p$]{\includegraphics[width=0.45\textwidth]{mean-and-variance.pdf}}
+\includegraphics[width=\textwidth]{snb-first-plot.pdf}
 \caption{
-The distribution of the sample size in a trial that stops after seven 
-patients respond to treatment
-or 11 patients do not. Panel (a) shows the distribution when
-$p = 0.2$. The mean and variance of the distribution varying $p$ 
-between zero and one are shown in panel (b).
+The distribution of the sample size in a trial that stops after seven patients
+respond to treatment or 11 do not when $p=0.2$.
 }
-\label{fig:exp-and-var}
+\label{fig:kane_viz}
 \end{figure}
 
-More generally the distribution of the number of trial enrollments is shown in 
-Fig.~\ref{fig:exp-and-var} (a). There is relatively little probability mass
+\begin{figure}[bp!]
+\centering
+\includegraphics[width=\textwidth]{mean-and-variance.pdf}
+\caption{
+The effect of varying $p$ between zero and one when the trial stops after
+seven patients respond or 11 do not.
+}
+\label{fig:kane_viz}
+\end{figure}
+
+%\begin{figure}[bp!]
+%\centering
+%\subfloat[enrollment distribution]{\includegraphics[width=0.45\textwidth]{snb-first-plot.pdf}}
+%\hfill
+%\subfloat[mean and variance varying $p$]{\includegraphics[width=0.45\textwidth]{mean-and-variance.pdf}}
+%\caption{
+%The distribution of the sample size in a trial that stops after seven 
+%patients respond to treatment
+%or 11 patients do not. Panel (a) shows the distribution when
+%$p = 0.2$. The mean and variance of the distribution varying $p$ 
+%between zero and one are shown in panel (b).
+%}
+%\label{fig:exp-and-var}
+%\end{figure}
+
+The distribution of the number of trial enrollments is shown in 
+Fig.~\ref{fig:kane_viz}. There is relatively little probability mass
 for values of sample size equal to seven through 10 since $p$ is small and it 
@@
 %non-responders before reaching 2 responders.
-Fig.~\ref{fig:exp-and-var} (b) shows the expected value and variance for the
+Fig.~\ref{fig:kane_viz} shows the expected value and variance for the
 number of trial enrollments varying $p$ between zero and one. When $p$ is
@@
 $\,b_i = 1$ corresponds to a patient responding to treatment.
+
 Let $s$ and $t$ be positive integers.  Define the SNB random
@@
 is the negative binomial probability mass. 
+The CDF of the negative binomial is commonly expressed in terms of the
+{\em regularized incomplete beta function} \citep{Olver2010}. Using standard
+notation, (\ref{eqn:N}) is written as $\mathcal{I}_{1-p}(t, s)$.
 
@@
 and $t-1$ unsuccessful outcomes multiplied by the probability of a final
-unsuccessful outcome at time $k$.
+unsuccessful outcome at time $k$. That is,
+\begin{align*}
+p {k-1 \choose s=1} p^{s-1} & (1-p)^{k-s} + (1-p) {k-1 \choose k-t} p^{k-t} (1-p)^{t-1} \\ 
+& = S(k, p, s) + S(k, 1-p, t)
+\end{align*}
+as claimed in (\ref{eqn:pmf}).
 
@@
 
-To show that (\ref{eqn:pmf}) sums to one, define
+To show (\ref{eqn:pmf}) sums to one, define
 %\begin{align} \label{eqn:sum_proof}
@@
 
-The CDF of the negative binomial is commonly expressed in terms of the
-{\em regularized incomplete beta function} \citep{Olver2010}. Using standard
-notation, the first summation in (\ref{eqn:transformed_sum})
-is written as $\mathcal{I}_{1-p}(t, s)$.
-This function satisfies 
+The regularized incomplete beta function satisfies
 $\mathcal{I}_p(s, t) = 1-\mathcal{I}_{1-p}(t, s)$ \citep{Uppuluri1967}. Then
@@
 \end{align*}
-This completes the proof that (\ref{eqn:pmf}) is the distribution of the
+This completes the proof (\ref{eqn:pmf}) is the distribution of the
 stopping time and is a valid probability mass function.
@@
 \caption{Different shapes of the SNB distribution with parameters 
-($p$, $s$, $t$), as given. Black indicates mass contributed by reaching
+($p$, $s$, $t$), as given. Black areas indicate the mass contributed by reaching
 $s$ responders before $t$ non-responders. Grey indicates
@@
 the smallest number of outcomes necessary to 
-observe at least $s$ responders {\em and} $s$ non-responders and is equivalent
+observe at least $s$ responders {\em and} $s$ non-responders. This is equivalent
 to a translated version of the riff-shuffle.
@@
 %to the probability of at most $s-1$ successes in a binomial distribution. 
-There is an equivalence between the probability of reaching an endpoint 
+There is also an equivalence between the probability of reaching an endpoint 
 in the SNB model and the tail probability of the binomial distribution.
-That is, the probability that the number of responders is at least $s$ in the 
-binomial model is the same as the probability the treatment succeeds 
-(reaches $s$) in the SNB 
-model.
+Specifically, the probability that the number of responders is at least 
+$s$ in the binomial model is the same as the probability the treatment succeeds 
+(reaches $s$) in the SNB model.
 \begin{prop} \label{binomial_tail}
@@
 to the number of responders at the end of the trial. Let 
-$B$ be distributed binomial with size $n=s+t-1$ and response probability
-$p$. Then
+$B$ be distributed binomial with index parameter $n=s+t-1$ and response 
+probability $p$. Then
 \begin{equation}
@@
 \end{equation*}
-%\begin{align*}
-%\mathbb{P}[B \geq s] &= \sum_{k=s}^{s+t-1} {n \choose k} p^k (1-p)^{n-k} \\
-%  &= 1 - \sum_{k=0}^{s-1} {n \choose k} p^k (1-p)^{n-k} \\
-%  &= 1 - \mathcal{I}_{1-p}.
-%\end{align*}
-where $\mathcal{I}_{1-p}(s, t)$ is the {\em regularized incomplete beta} 
-function. The corresponding success probability is
+The corresponding success probability is
 \begin{equation} \label{eqn:success}
@@
 \end{align*}
-Since the regularized incomplete beta function's subscript parameter 
-has support on zero 
-to one, we have $0 \leq qe^x < 1$. This implies
-$x < -\log(q)$.
+Since the $p$ parameter in $\mathcal{I}_p$ has support on zero 
+to one, we have $0 \leq pe^x < 1$. This implies $x < -\log(p)$.
 A similar expression can be derived for the second summation in 
 (\ref{eqn:first_sum}) and results in
-the constraint $x < -\log(p)$.
+the constraint $x < -\log(1-p)$.
 \end{proof}
 
-The SNB's ability to approximate the geometric, normal, gamma, and Poisson
-distributions follow from it generalizing the negative binomial distribution. 
+Depending on the parameter values the SNB can be approximated by 
+the geometric, normal, gamma, and Poisson
+distributions.
+
 \begin{prop}
@@
 \begin{equation*}
-\mathbb{E} e^{xY} \rightarrow \left( \frac{pe^x}{1-qe^x} \right)^s
+lim_{t\to\infty} \ \mathbb{E} e^{xY} = \left( \frac{pe^x}{1-qe^x} \right)^s
 \end{equation*}
-as $t \rightarrow \infty$. The analogous result holds when 
-$s \rightarrow \infty$.
+as. The analogous result holds when $s \rightarrow \infty$.
 \end{prop}
 \begin{proof}
-The second regularized incomplete beta term in (\ref{eqn:mgf}) can be written
+The second incomplete beta function in (\ref{eqn:mgf}) can be written
 in terms of a cumulative binomial distribution
@@
 where $B$ is distributed as
-Binomial($t-k$, $pe^x$). By Chebychev's inequality %\cite{Hoeffding1963}
+Binomial($t-k$, $pe^x$). From Chebychev's inequality %\cite{Hoeffding1963}
 it follows that
@@
 \end{equation}
-As $t$ gets large $\mathcal{I}_{1-pe^x}(t, s)$ goes to zero
-Likewise, $\mathcal{I}_{1-qe^x}(s, t)$ goes
-to one. The proof follows by realizing 
+As $t$ gets large $\mathcal{I}_{1-pe^x}(t, s)$ tends to zero
+and $\mathcal{I}_{1-qe^x}(s, t)$ approaches 
+one. The proof follows by realizing 
 \begin{equation*}
@@
 
-When $s=1$ (and $t$ is still large) the SNB's MGF is approximately 
+When $s=1$ and $t$ is large the SNB's MGF is approximately 
 the same as that of the
@@
 binomial can therefore be seen as a sum of i.i.d. geometric distributions.
-For an appropriately large number of samples, the central limit theorem
+For an appropriately large number of distributions, the central limit theorem
 yields a normal approximation.
 
-Drawing connections to the gamma and Poisson distributions are more complicated.
-However a connection to the gamma distribution well-studied problem in the 
+Drawing connections to the gamma and Poisson distributions can also be shown.
+A connection to the gamma distribution well-studied problem in the 
 literature (see \cite{Best1974,Ord1968,Guenther1972} for examples). 
 A connection to the Poisson appears in \cite{Anscombe1950} where
-it is shown that if the mean of a Poisson is proportional
+it is shown that when the mean of a Poisson is proportional
 to a chi-square distribution with $2k$ degrees of freedom
@@
 for appropriately large $s$ and $t$ and $p = 0.5$. In this
-case, the SNB can be viewed as identical, negative binomials
+case, the SNB can be viewed as identical, negative binomial distributions
 approximating a normal and truncated at the median.
@@
 
-Let us consider the case where the rate parameter is distributed 
+Let us consider the Bayesian setting where the rate parameter is distributed 
 as Beta($\alpha$, $\beta$) and denoted $P$.
-The prior times the likelihood is given by the function
+The posterior distribution of $P$ is proportional to the 
+likelihood, given by the function
+%f_P(p, \alpha, \beta ) f_{Y|P}(p, k, s, t) &= 
 \begin{align} \label{eqn:ptl}
-f_P(p, \alpha, \beta ) f_{Y|P}(p, k, s, t) &= 
+f_{P|Y}(p, k, s, t, & \alpha, \beta) \propto
   \frac{ {k-1 \choose s-1} }{B(\alpha, \beta)} p^{\alpha +s -1} 
-    (1-p)^{k+\beta-s-1} + \\
-  & \frac{ {k-1 \choose t-1} }{B(\alpha, \beta)} p^{k+\alpha -t -1} 
+    (1-p)^{k+\beta-s-1} \\
+  & + \frac{ {k-1 \choose t-1} }{B(\alpha, \beta)} p^{k+\alpha -t -1} 
     (1-p)^{\beta+t-1} \nonumber
@@
 where $0 \leq p \leq 1$, $s \leq k \leq s+k-1$, and $t \leq k \leq s+k-1$.
-This result can be found directly by multiplying the probability mass function
-in (\ref{eqn:pmf}) by the density of the Beta distribution.
 
-The predictive distribution of the SNB can be found by integrating
+The predictive distribution of the SNB can be found as by integrating
 $p$ over the interval zero to one and applying the definition of the 
@@
 \begin{align} \label{eqn:predictive}
-f_{Y}(k, s, &t, \alpha, \beta) = 
+f_{Y}(&k, s, t, \alpha, \beta) = 
   \int_0^1 f_P(p | \alpha, \beta )  f_{Y|P}(p, k, s, t) dp \nonumber \\ 
@@
 
-%\begin{proof}
-%For notational simplicity, assume that $k$ is in the range
-%$\min(s,t) \leq k \leq s+t-1$. When 
-%this is not the case appropriate terms should be removed as dictated by the indicator functions.
-%\begin{align*}
-%f(k | s, t, \alpha, \beta) = \frac{1}{B(\alpha, \beta)} & \int_0^1 {k-1 \choose s-1} p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} + \\
-% & {k-1 \choose t-1} p^{k-t+\alpha-1}\left(1-p\right)^{t+\beta-1} dp \\
-%= \frac{1}{B(\alpha, \beta)}  {k-1 \choose s-1} & \int_0^1  p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} dp + \\
-% & \frac{1}{B(\alpha, \beta)} {k-1 \choose t-1} \int_0^1  p^{k-t+\alpha-1}\left(1-p\right)^{t+\beta-1} dp.
-%\end{align*}
-%The result in (\ref{eqn:posterior}) follows by the definition of the 
-%beta function.
-%\end{proof}
+\section{Powering Trials using Confidence Sets}
 
-%To better understand the predictive distribution, let us once again return to 
-%our example where $s=7$, $t=11$. However, we assume $p$ has a beta distribution
-%with shape parameters $\alpha = 0.2 c$ and $\beta = 0.8 c$ for any $c > 0$.
-%This parameterization has mean 0.2 for all $c > 0$ and variance 
-%$0.16 / (c+1)$.
-%
-%
-%\begin{figure}[h!]
-%\centering
-%\subfloat[mean]{\includegraphics[width=0.5\textwidth]{bayesian-sample-expectation.pdf}}
-%\hfill
-%\subfloat[variance]{\includegraphics[width=0.5\textwidth]{bayesian-sample-variance.pdf}}
-%\caption{
-%The mean and variance of the predictive SNB with $s=7$, $t=11$
-%with $\mathbb{E}\ p = 0.2$ varying the uncertainty 
-%in the distribution of $p$. A larger value of $c$
-%corresponds to a beta variance.
-%}
-%\label{fig:bayesian-sample-size}
-%\end{figure}
+The previous section shows the posterior distribution $P$ at each of the 
+endpoints in the domain is defined by a Beta distribution. This fact can be 
+used to provide power calculations based on confidence sets with specified 
+minimum response rate. Let $\mathcal{P}_s(s,t)$ be the set of random variables
+associated success. Then a trial whose minimum
+response rate is $p_{min}$ at a significance level of at least $\alpha_C$ is 
+one where each distribution is at least $p_{min}$ with probability 
+$1-\alpha_C$.  That is,
+\begin{equation} \label{pwr_signif}
+  \mathbb{P} \left[X > p_{min} \right] > (1 - \alpha_C) 
+  \ \ \forall X \in \mathcal{P}_s(s,t)
+\end{equation}
+Letting $\mathcal{Q}_t(s,t)$ be the set of random variables associated with
+failure, a trial whose power is at least $1-\beta_C$ has
+\begin{equation*} \label{pwr}
+  \mathbb{P} \left[X < p_{min} \right] < \beta_C
+  \ \ \forall X \in \mathcal{Q}_t(s,t).
+\end{equation*}
 
-%As $c$ increases the 
-%predictive distribution sample statistics converge to corresponding 
-%statistics for the SNB with $p = 0.2$. 
-%For estimation it may be 
-%preferable to select shape parameters based on the median or mode of the 
-%beta distribution since they are less sensitive to distributional uncertainty.
+\begin{prop} \label{alpha_c}
+Let $P_k$ be the posterior distribution of reaching the success endpoint
+at the $k$th enrollment with density
+\begin{equation*}
+f_k(p) = \frac{ {k-1 \choose s-1} }{B(\alpha, \beta)} p^{\alpha +s -1} 
+    (1-p)^{k+\beta-s-1}.
+\end{equation*}
+If $0 \leq \alpha_C \leq 1 - 1/k$ then 
+\begin{equation}
+\mathbb{P} \left[P_k \geq \alpha_C \right]  \leq
+  \mathbb{P} \left[P_{k-1} \geq \alpha_C \right].
+\end{equation}
+\end{prop}
+\begin{proof}
+Consider the ratio of the posterior distributions
+\begin{align*}
+\frac{f_k(p)}{f_{k+1}(p)} &= 
+  \frac{\frac{ {k-1 \choose s-1} }{B(\alpha, \beta)} p^{\alpha +s -1} 
+    (1-p)^{k+\beta-s-1}}
+    {\frac{ {k \choose s-1} }{B(\alpha, \beta)} p^{\alpha +s -1} 
+    (1-p)^{k+\beta-s}} \\
+  &= \frac{ p^{\alpha +s -1} (1-p)^{k+\beta-s-1} }
+    {k p^{\alpha +s -1} (1-p)^{k+\beta-s}} \\
+  &= \frac{1}{k(1-p)}.
+\end{align*}
+This implies that the density $f_{k+1}(p)$ is greater than $f_k(p)$ for
+all values of $0 \leq p \leq 1 - 1/k$. This implies the
+integral of $f_{k+1}(p)$ must also be greater than $f_k(p)$ over a similar
+interval. The result follows by realizing that this relationship is reversed
+over the complement of the interval.
+\end{proof}
 
-%%It may be noted that the compound 
-%%distribution's mean
-%%approaches the deterministic limit slowly. This
-%%is because when $c$ is small, the distribution of $p$ is heavily skewed. 
-%%As $c$ increases this effect is minimized.
-%When the distribution of $p$ is centered close to zero or one the mean 
-%estimate approach the deterministic limit relatively slowly in the uncertainty
-%parameter $c$.  
-%\section{Discussion}
-%
-%The SNB can be used to extend clinical trial methodology
-%in at least two areas. First, it can be used to construct nested criteria
-%for early stopping and efficacy. Current approaches are often inspired by
-%the Simon 2-stage Phase II clinical trial \cite{Simon1989} where a
-%a small number of patients are enrolled for an initial stage. For patients,
-%success or failure is assessed. If at least a pre-specified number of
-%successes are observed then the trial enters a second stage with more
-%patients. If a sufficient number of successes are not observed then the
-%trial fails.
+The maximum significance of a trial is 
+$\mathbb{P} \left[P_{s+t-1} \leq \alpha_C \right]$ Proposition \ref{alpha_c}
+guarantees the significance for all other success endpoints will be smaller.
+A similar calculation can be performed for power to show that the minimum
+power is $\mathbb{P} \left[Q_{s+t-1} \leq 1 - \beta_C \right]$ where
+$Q_{s+t-1}$ is the posterior distribution of reaching a trial failure
+at the $s+t-1$ step.
 
-%The SNB is being used to generalize the Simon design by providing separate
-%criteria for futility and efficacy.
-%The Simon design tests for a single response probability and time-to-event
-%parameter for both the first
-%and second stage. The SNB is being used to generalize this design by 
-%allowing these parameters to vary for the two stages, and even allow
-%information about patients in the first stage to be used in the second
-%stage. Early results show the design provides
-%both early testing for futility along with the ability
-%to make statistically sound decisions sooner than traditional trials.
+These two results imply a relatively simple procedure for finding the sample
+size for trials of this type. A minimum response rate, maximum significance,
+and minimum power are selected. The task is to find the values of $s$ and
+$t$ that fulfill these requirements given the minimum response. Start with 
+$s = t = 1$. If the maximum significance criterion is not met then increase
+$s$ by one. If the minimum power criterion is not met then increase $t$ by
+one. Repeat the procedure until both criteria have been met. The sample
+size is then determined by the calculation $s+t-1$.
 
-%A second avenue for methodological innovation comes directly from
-%Corollary \ref{conditional_distribution} and Proposition \ref{prop:bayesian}.
-%Our knowledge of the response probability $p$ comes from published reports
-%yielding a prior distribution.
-%While a trial is being conducted the response probabilities are 
-%updated as new results are received. The number of individual successes
-%and failure are used to learn the shape characteristics of the corresponding
-%beta distribution and resulting posterior
-%distribution. An estimate of the probability of reaching the success (and
-%failure) endpoints can be provided at each step of the trial.
+\section{A Bayesian Sequential Probability Ratio Test}
 
-%There are also several avenues for extending the SNB itself. First is the 
-%development of a stopped negative multinomial distribution allowing
-%for more than two endpoints. This extension would have clinical trial
-%applications including dose escalation where the outcomes are response
-%to a therapy, failure to respond, and an adverse outcome endpoint where
-%an enrollee is removed from the trial with neither response or non-response
-%due to side effects.
+Keeping $s$ and $t$ constant for all time steps in the trial provides 
+mathematical convenience since the distribution can be calculated in closed 
+form. It should be noted though that any set of endpoints where every endpoint 
+can be reached defines a valid probability distribution. This fact allows us 
+to define new stopping criteria when defining success and failure endpoints.
 
-%A second, more theoretical extension is to the asymptotic setting. Under
-%proper reparameterization, the process from which the SNB is derived 
-%can be shown to converge to a Brownian motion. In this case, we conjecture
-%that the SNB can be defined in terms of the L\'{e}vy-Bachelier 
-%\citep{Bachelier1900} formula
-%thereby relating the SNB to the classic result in stochastic calculus.
+One such set of stopping criteria is based on the confidence sets of the 
+response rate from the previous section. Proposition (\ref{alpha_c}) shows 
+$\mathbb{P_k} [P_k \geq \alpha_c] \leq \mathbb{P}[P_{k-1} \geq \alpha_c]$. 
+That is, the probability the response rate is at least $p_{min}$ is increasing 
+in $-k$. This observation implies we can guarantee that the response rate is 
+greater than $p_min$ before reaching $s$ when a greater proportion of 
+successful outcomes are observed.
 
-\section*{References}
+Therefore, we can define a new set of success and failure endpoints that stop 
+a trial as soon as the response distribution is at least $\alpha_c$ or the 
+criteria cannot be met, based on the sample size. This design is analogous to 
+the Sequential Probability Ratio Test (SPRT) \cite{Wald1945} in that the 
+stopping criteria is defined by the ratio responders to non-responders, 
+through the beta prior. It has two distinct advantages though. First, unlike 
+the SPRT it has a maximum sample size, similar to the MaxSPRT 
+\cite{Kulldorff2011}, making it appropriate for small-sample studies. Second, 
+the stopping criteria are based on both a minimum response rate as well as the 
+degree of certainty, $1-\alpha$, that this response rate has been attained.
+
+\section{The Approximate Distribution for Two Linear Boundaries}
+
+Consider an alternative visualization of the sample path counting the 
+number of responders and non-responders in the vertical and horizontal 
+direction respectively, as in Figure \ref{fig:zplot}. This alternative
+formulation can be rotated and scaled to form a random walk on the integers.
+A functional central limit can be then applied when the endpoints are 
+sufficiently far from the origin and the process converges to a Brownian 
+motion. The hitting time of the success and failure endpoints is under 
+this reparameterization is a mixture of normal distributions.
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=\textwidth]{zplot.pdf}
+\caption{
+An alternative visualization of the hypothetical trial.
+}
+\label{fig:zplot}
+\end{figure}
+
+For a response rate $p$, the expected direction of each step in the sample 
+path is $\tan^{-1}(p)$ radians from the vertical direction.
+If the graph is rotated $\pi/2- \Theta$ radians then the 
+expected value of the process is zero at any time.
+Let $\Theta = tan^{-1}(p)$. Then, after rotation, the process moves 
+$\sin \Theta$ units in the horizontal direction and $\cos \Theta$ units in 
+the positive vertical direction for each response. Each non-response 
+corresponds to $\cos (\Theta)$ units in the horizontal direction and 
+$\sin \Theta$ units in the negative vertical direction. The rotation is shown 
+pictorially in Figure \ref{fig:theta_rot}.
+
+\begin{figure}[bp!] 
+\centering
+\subfloat[Before rotation]{\includegraphics[width=0.45\textwidth]{proc.pdf}}
+\hfill
+\subfloat[After rotation]{\includegraphics[width=0.45\textwidth]{proc-rot.pdf}}
+\caption{
+The direction of a single step in the binomial process before (a) and after
+rotation (b).
+}
+\label{fig:theta_rot}
+\end{figure}
+
+After the rotation, a response step is normalized by $\sin(\pi/2-\Theta)$ in the
+vertical direction and $\cos(\pi/2 - \Theta)$ in the horizontal direction.
+Likewise, a non-response step is normalized by $\sin(\Theta)$ and
+$\cos(\Theta)$ in the vertical and horizontal direction respectively. Each
+step is a Rademacher random variable. The
+resulting process forms a random walk on the integers. Let $m$ be the
+shortest path from the origin to any endpoint, let step $k=mt$
+the process, and let $R_i$ be a sample from a Rademacher random variable. Then
+it is well known that a c\'{a}dl\'{a}g version of the rescaled process on the
+natural filtration converges to a Brownian motion as $m$ gets large
+\begin{equation*}
+t \frac{1}{\sqrt{t m}} \sum_{i=0}^{t m} R_i \rightarrow
+B(t).
+\end{equation*}
+
+The density of the first time a Brownian motion hits a sloping line with
+intercept $a$ and is \cite{Bachelier1900}
+\begin{equation}
+f_{a,b}(t) = \frac{a}{t^{3/2}} \phi \left(\frac{a + bt}{\sqrt{t}} \right)
+\end{equation}
+where $\phi(t)$ is the density of the standard normal distribution at $t$.
+Therefore, for success and failure endpoints defined by lines 
+$a + bt$ and $a' + b't$ with support $[s_0 \leq t \leq s_m]$ and 
+$[r_0 \leq t \leq r_m]$ respectively, the density of the SNB approximation
+is 
+\begin{equation}
+f(t) = \frac{ f_{a,b}(t)[s_0 \leq t \leq s_m] + 
+  f_{a',b'}(t)[r_0 \leq t \leq r_m]}{
+  \int_{s_0}^{s_m} f_{a,b}(x) dx + \int_{r_0}^{r_m} f_{a',b'}(x) dx}.
+\end{equation}
 

### 2017-02-27 stat-and-prob-letters/s.r

@@ -77,11 +77,11 @@ approx_plots_start = list(
     labs(x="", y="", title="gamma approximation (0.06, 3, 175)") + theme_bw()+
     scale_fill_grey(),
   dsnb_stack_plot(0.5, 25, 25) + scale_x_discrete(breaks=seq(25, 50, by=5))+
-    labs(x="", y="", title="lower half normal / riff-shuffle  (0.5, 25, 25)") + 
+    labs(x="", y="", title="lower half normal (0.5, 25, 25)") + 
     theme_bw() + scale_fill_grey(),
   dsnb_stack_plot(p=0.45, 25, 25) + 
     scale_x_discrete(breaks=seq(25, 49, by=3))+
-    labs(x="", y="", title="riff-shuffle (0.45, 25, 25)") + 
+    labs(x="", y="", title="upper truncated normal (0.45, 25, 25)") + 
     theme_bw() + scale_fill_grey())
 
 library(foreach)
@@ -252,3 +252,121 @@ p = ggplot(x, aes(x=p, y=value, fill=Outcome)) +
   theme_bw() + scale_fill_grey(start=0.8, end=0.2) + xlab(expression(x)) +
   ylab(TeX("f_{P|Y}(x | k, s, t)")) 
 ggsave("beta-mixture.pdf", p, width=8, height=5)
+
+
+library(grid)
+library(latex2exp)
+
+theta = pi / 3
+uo = cos(theta)
+ua = sin(theta)
+do = ua
+da = uo
+
+sx = 1/(tan(theta) +1)
+sy = -sx + 1
+
+p1 = ggplot() +
+  geom_segment(mapping=aes(x=0, y=0, xend=1, yend=0)) + 
+  geom_segment(mapping=aes(x=0, y=0, xend=0, yend=1)) +
+  geom_segment(mapping=aes(x=0, y=0, xend=sx, yend=sy),
+    arrow=arrow(length=unit(.1, "inches"))) +
+  geom_curve(mapping=aes(x=sx/2.5, y=sy/2.5, xend=0, yend=1/3),
+    curvature=0.3) +
+  geom_curve(mapping=aes(x=1/4.5, y=0, xend=sx/3.5, yend=sy/3.5),
+    curvature=0.35) +
+  annotate("text", x=0.15, y=0.5, 
[…]

### 2017-07-12 stat-and-prob-letters/snb-paper.tex
Lots of update.

@@
 rejected and the treatment is deemed ineffective. The trial is 
-realized as a sequence of Bernoulli($p$) samples stopping when either
-the specified number of responders or non-responders is reached.
+realized as a sequence of indiependent Bernoulli($p$) samples stopping when 
+either the specified number of responders or non-responders is reached.
 
@@
 }
-\label{fig:kane_viz}
+\label{fig:kane_viz_1}
 \end{figure}
@@
 }
-\label{fig:kane_viz}
+\label{fig:kane_viz_2}
 \end{figure}
@@
 The distribution of the number of trial enrollments is shown in 
-Fig.~\ref{fig:kane_viz}. There is relatively little probability mass
+Fig.~\ref{fig:kane_viz_1}. There is relatively little probability mass
 for values of sample size equal to seven through 10 since $p$ is small and it 
@@
 %non-responders before reaching 2 responders.
-Fig.~\ref{fig:kane_viz} shows the expected value and variance for the
+Fig.~\ref{fig:kane_viz_2} shows the expected value and variance for the
 number of trial enrollments varying $p$ between zero and one. When $p$ is
@@
 {\em regularized incomplete beta function} \citep{Olver2010}. Using standard
-notation, (\ref{eqn:N}) is written as $\mathcal{I}_{1-p}(t, s)$.
+notation, (\ref{eqn:N}) is written as $\mathcal{I}_{1-p}(t, s)$ and has
+the property $\mathcal{I}_{1-p}(t, s) = 1 - \mathcal{I}_p(s, t)$
+\citep{Uppuluri1967}.
 
@@
 \end{equation*}
-At each step a patient's outcome is measured. In Fig.~\ref{fig:kane_viz} 
+At each step a patient's outcome is measured. In Fig.~\ref{fig:kane_viz_1} 
 we consider a graphical illustration of the plot $X_k$ against
@@
 process continues until either $X_k = s$ or $X_k = k-t$ corresponding to the
-success and failure boundaries in Fig. \ref{fig:kane_viz}, respectively.
+success and failure boundaries in Fig. \ref{fig:kane_viz_2}, respectively.
 
@@
 \end{equation*}
-If we substitute $i=k-s$ in the first summation and $j=k-t$ in the second then
+Substitute $i=k-s$ in the first summation and $j=k-t$ in the second then
 $R$ can be written as the cumulative distribution function (CDF) of two
@@
 
-The regularized incomplete beta function satisfies
-$\mathcal{I}_p(s, t) = 1-\mathcal{I}_{1-p}(t, s)$ \citep{Uppuluri1967}. Then
+Then
 \begin{align*}
@@
 \begin{equation*}                                    %   (1)   
-\mathbb{P}[Y=s+j\ |\ t \text{ is large\ }]        \label{nb1.eq}          
+\mathbb{P}[Y=s+j ]        \label{nb1.eq}          
   = {{s+j-1}\choose{s-1}} p^s (1-p)^j
@@
 This implies that the density $f_{k+1}(p)$ is greater than $f_k(p)$ for
-all values of $0 \leq p \leq 1 - 1/k$. This implies the
+all values of $0 \leq p \leq 1 - 1/k$. The
 integral of $f_{k+1}(p)$ must also be greater than $f_k(p)$ over a similar
@@
 
-Keeping $s$ and $t$ constant for all time steps in the trial provides 
+Keeping $s$ and $t$ constant for all time steps in a trial provides 
 mathematical convenience since the distribution can be calculated in closed 
 form. It should be noted though that any set of endpoints where every endpoint 
-can be reached defines a valid probability distribution. This fact allows us 
+can be reached by the process $\mathbf X$ defines a valid probability 
+distribution. This fact allows us 
 to define new stopping criteria when defining success and failure endpoints.
@@
 response rate from the previous section. Proposition (\ref{alpha_c}) shows 
-$\mathbb{P_k} [P_k \geq \alpha_c] \leq \mathbb{P}[P_{k-1} \geq \alpha_c]$. 
+$\mathbb{P} [P_k \geq \alpha_c] \leq \mathbb{P}[P_{k-1} \geq \alpha_c]$. 
 That is, the probability the response rate is at least $p_{min}$ is increasing 
 in $-k$. This observation implies we can guarantee that the response rate is 
-greater than $p_min$ before reaching $s$ when a greater proportion of 
+greater than $p_{min}$ before reaching $s$ when a greater proportion of 
 successful outcomes are observed.
@@
 degree of certainty, $1-\alpha$, that this response rate has been attained.
+
+As an example, consider the design of the hypothetical trial from the
+first section. Assume that we are testing for a response rate of at least
+0.2 and a uniform prior on the response rate then, by Proposition 
+\ref{alpha_c}, the design implies a confidence of at least 98\% 
+($\alpha_c = 0.02)$. That is, at each of the success endpoints, 
+we believe the response rate is at least 0.2 with 98\% confidence. This also 
+implies that $\beta = 2.24526 * 10^{-5}$.
+Figure \ref{fig:bsprt} shows the success and failure boundaries for the 
+SNB trial design from Section 1 along with the Bayesian SPRT analogue.
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=\textwidth]{bsprt.pdf}
+\caption{
+Success and failure boundaries for the Bayesian SPRT analogue (BSPRT) and SNB
+trial designs of the hypothetical trial.
+}
+\label{fig:bsprt}
+\end{figure}
 

### 2017-07-12 stat-and-prob-letters/s.r

@@ -3,50 +3,6 @@ library(ggplot2)
 library(tidyr)
 library(latex2exp)
 
-## kplot
-
-#kplot = function(flips, s, t) {
-#  if (!is.list(flips)) {
-#    d = flips_to_kplot_df(flips)
-#    p = qplot(k, path, data = d, geom = "line") +
-#      scale_x_continuous(breaks = 0:(t + s), limits = c(0, t + s - 1)) +
-#      scale_y_continuous(breaks = 0:s, limits=c(0, s+0.15)) +
-#      geom_segment(x=s, y=s, xend=(t+s-1), yend=s, linetype=2) +
-#      geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, linetype=2)
-##      geom_segment(x=s, y=s, xend=(t+s-1), yend=s, color="green", linetype=1) +
-##      geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, col="red")
-#  } else {
-#    flip_set = lapply(flips, flips_to_kplot_df)
-#    for (i in 1:length(flip_set)) {
-#      flip_set[[i]]$num = as.factor(i)
-#      flip_set[[i]]$k = jitter(flip_set[[i]]$k)
-#      flip_set[[i]]$k[flip_set[[i]]$k < 0] = 0
-#    }
-#    d = Reduce(rbind, flip_set)[, -(4:5)]
-#    p = qplot(k, path, data = d, geom = "path", group = num) +
-#        scale_x_continuous(breaks=0:(t+s), limits = c(0, t+s)) +
-#        geom_segment(x = s, y = s, xend = (t + s - 1), yend = s,
-#                     linetype=2) +
-#        geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, linetype=2)
-##        geom_segment(x = s, y = s, xend = (t + s - 1), yend = s,
-##                     color = "green") +
-##        geom_segment(x=t, y=0, xend=(s+t-1), yend=s-1, col="red")
-#    p
-#  }
-#  p
-#}
-
-#flips_to_kplot_df = function (flips) {
-#  d = data.frame(k = 0:length(flips))
-#  d$head = c(0, cumsum(flips))
-#  d$tail = c(0, cumsum(1 - flips))
-#  d$headEnd = c(d$head[-1], NA)
[…]

### 2017-08-10 sii/deveaux-kane-zelterman.tex
Draft sent to SII.

% Template for the submittion to:
%   Statistics and Its Interface      [sii]
%
% Author: In this template, the places where you need to add information
%         (or delete line) are indicated by {???}.  Mostly the information
%         required is obvious, but some explanations are given in lines starting
% Author:
% All other lines should be ignored.  After editing, there should be
% no instances of ??? after this line.

% journal options: sii
\documentclass[sii]{ipart}

\usepackage{amsthm,amsfonts,amsmath}
\usepackage{graphicx}
\usepackage{bbm}
\usepackage{subfig}
\usepackage{lipsum}
\usepackage[numbers,square]{natbib}
\RequirePackage{hyperref}

% will be filled by editor:
\pubyear{0000}
\volume{0}
\issue{0}
\firstpage{1}
\lastpage{10}
\arxiv{arXiv:1508.01264 }

% put your definitions there:
\startlocaldefs
\DeclareMathOperator*{\argmin}{argmin}
\newcommand*{\argminl}{\argmin\limits}
\newtheorem{prop}{Proposition}
\newtheorem{corollary}{Corollary}

\endlocaldefs

\begin{document}

\begin{frontmatter}

% "Title of the Paper"
\title{Clinical Trial Design Using A Stopped Negative Binomial Distribution}
\runtitle{Stopped Negative Binomial}

%\thankstext{T1}{This research was supported by grants R01CA131301, 
%R01CA157749, R01CA148996, R01CA168733, and PC50CA196530 awarded by the 
%National Cancer Institute along with support from the Yale Comprehensive Cancer 
%Center and the Yale Center for Outcomes Research. The authors would also like 
%to thank Rick Landin at Ignyta Inc. for his suggestions.}

% indicate corresponding author with \corref{}
% \author{\fnms{John} \snm{Smith}\thanksref{t2}\corref{}\ead[label=e1]{smith@foo.com}\ead[label=e2,url]{www.foo.com}}
% \address{line 1\\ line 2\\ \printead{e1}\\ \printead{e2}}

\begin{aug}
\author{\fnms{Michelle} \snm{DeVeaux} \ead[label=e1]{michelle.deveaux@yale.edu}}
\address{Department of Biostatistics\\ School of Epidemiology and Public Health \\ Yale University, New Haven, CT \\ \printead{e1}}
\and
\author{\fnms{Michael J.} \snm{Kane}\thanksref{t2}\corref{} \ead[label=e2]{michael.kane@yale.edu}}
\address{Department of Biostatistics\\ School of Epidemiology and Public Health \\ Yale University, New Haven, CT \\ \printead{e2}}
\and
\author{\fnms{Daniel} \snm{Zelterman} \ead[label=e3]{daniel.zelterman@yale.edu}}
\address{Department of Biostatistics\\ School of Epidemiology and Public Health \\ Yale University, New Haven, CT \\ \printead{e3}}
%\address{???\printead{e2}}

\thankstext{t2}{Corresponding author.}

\runauthor{DeVeaux, Kane, and Zelterman}

\affiliation{Yale University}
\end{aug}

\begin{abstract}
We introduce a discrete distribution suggested by curtailed
sampling rules common in early-stage clinical trials unifying the
negative, minimum negative (riff-shuffle), and
maximum negative binomial distributions. We derive the
distribution of the smallest number of independent and identically
distributed Bernoulli trials needed to observe either $s$ successes 
or $t$ failures. This report provides a closed-form expression for the 
mass function, moment generating function, and provides connections to other,
standard distributions.
\end{abstract}

\begin{keyword}[class=AMS]
\kwd[Primary ]{62E15}
%\kwd{}
\kwd[; secondary ]{62P10}
\end{keyword}

\begin{keyword}
\kwd{discrete distributions}
\kwd{stopped negative binomial distribution}
\kwd{early-stage clinical trials}
\kwd{curtailed clinical trials}
\end{keyword}

% history:
% \received{\smonth{1} \syear{0000}}
%\tableofcontents

\end{frontmatter}

\section{Introduction and Motivation}

Consider a prototypical early phase, single-arm clinical trial in which 17 patients
are enrolled and treated. The trial is modeled as a sequence of independent Bernoulli($p$) samples. Suppose the Bernoulli probability of a patient 
responding to treatment is $p=0.2$ under the null hypothesis that the treatment is not any more effective than the current standard of care.
If seven or more patients out of these 17 respond to the treatment then we 
reject this hypothesis and claim the treatment has successfully showed superiority at 
a significance level of $0.1$.  If fewer than seven respond then the null 
hypothesis is not rejected and the treatment is said to have failed to show superiority. The trial ends when 
either seven responders or 11 non-responders are observed.

If all 17 patients are enrolled at once, as in the classic
design, then the sample size is 17. However, in most clinical trials the
patients are enrolled sequentially over time.
In the present example, observing seven
patients responding to treatment ends the trial and so the number of 
enrollees required could be as small as seven. Similarly 11
observed treatment non-responses would also end the trial. This sampling mechanism, in
which the experiment ends as soon as either predefined endpoint is reached, is
called {\em curtailed sampling}. Under curtailed sampling, the range of the
sample size for this trial is seven through 17.

\begin{figure*}
\centering
\includegraphics[width=0.8\textwidth]{snb-first-plot.pdf}
\caption{
The distribution of the sample size in a trial that stops after either seven patients
respond to treatment or 11 do not when $p=0.2$.
}
\label{fig:kane_viz_1}
\end{figure*}

\begin{figure*}[htbp!]
\centering
\includegraphics[width=\textwidth]{mean-and-variance.pdf}
\caption{
The mean and variance of the sample size for the probability of success $p$ 
when the trial stops after $s=7$ patients respond or $t=11$ fail to respond.
}
\label{fig:kane_viz_2}
\end{figure*}

The distribution of the number of trial enrollments is shown in 
Figure~\ref{fig:kane_viz_1}. There is relatively little probability mass
for values of sample sizes from seven through 10 since $p$ is small and it 
is unlikely the treatment will succeed quickly.
Figure~\ref{fig:kane_viz_2} shows the expected value and variance for the
number of trial enrollments varying the success rate $p$ between zero and one. When $p$ is small then the treatment
is likely to fail shortly after the $11^{\text{th}}$ enrollment.
When $p$ is large then the treatment is more likely to succeed and the 
number of enrollees approaches seven from above. 
When $p=0$ or $1$ then the sequence of resposes is deterministic and the 
variance is zero.
Values of $p$ between zero and one change the mixing proportions of 
the two endpoints. When $p$ is close to zero, most of the variance
is contributed from the failure endpoint. The saddle around $p=0.25$ results 
from a trade-off between the success and failure endpoints.

In the rest of this work, we derive the distribution of the number of 
enrollees needed
to observe either $s$ successes or $t$ failures. We refer to this distribution
as the Stopped Negative Binomial (SNB). 
In Section 2 we derive the distribution function and explores its properties.
Section 3 derives the moment generating function.
Section 4 describes the likelihood function.
Section 5 relates the distribution to other standard
distributions and connects the SNB tail probability to the binomial tail 
probability. Section 6 considers single arm, early-phase trial design with the 
SNB.

\section{Probability Mass Function}
\label{notation.section}

Let $\,b_1, b_2, \ldots \,$ denote a sequence of independent, identically
distributed, Bernoulli random variables with $\mathbb{P}[b_i=1]=p$ and
$\mathbb{P}[b_i = 0] = 1-p$, for
probability parameter $0\leq p \leq 1$. In the clinical trial setting
$\,b_i = 1$ corresponds to the $i$th patient responding to treatment.

Let $s$ and $t$ be positive integers.  Define the SNB random
variable $Y$ as the smallest
integer value such that $\,\{b_1, \ldots , b_Y\}\,$ contains {\em either}
$\,s\,$ responders {\em or} $\,t\,$ non-responders. That is, the SNB 
distribution of $Y$ is the smallest integer such that either
$\sum_i^Y b_i = s$ or $\sum_i^Y 1-b_i = t$.

The sequence of Bernoulli random variables can be modeled as the process 
$\mathbf{X} = \left\{X(k) : k = 0,1,... \right\}$
with $X(0)=0$ and
\begin{equation*} \label{eqn:proc}
X_{k+1} = X_k + b_{k+1} \ \mathbbm{1}_{\{ k-t < X_k < s\}}.
\end{equation*}
At each step a patient's outcome is measured. The process continues until either 
$X_k = s$ or $X_k = k-t$ corresponding to the success and failure boundaries.

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{KanePlot.pdf}
\caption{
A hypothetical realization of the trial.
}
\label{fig:kane_viz}
\end{figure*}

In Figure~\ref{fig:kane_viz} provides a graphical illustration of $X_k$ against
$k$ for one possible realization where $s=7$ and $t=11$. The vertical axis denotes the number of
successful outcomes. The vertical axis counts the number of responders observed.
The horizontal axis counts the number of patients 
enrolled. It is strictly increasing in sequence size and the failure boundary is tilted so that no more than $t$ (in this case 11) failure can occur. The horizontal and tilted boundaries represent the two
endpoints for the trial. In this case, the seventh response was reached on
the $15^{\text{th}}$ enrollment.
Since the success boundary is reached, we would reject the null hypothesis in this example.

We next derive the mass function of $Y$. The distribution of $\,Y\,$ has support on integer values in the range
\begin{equation*}               
     \min(s,t) \leq \; Y \;\leq s+t-1  \label{range.y.eq}.
\end{equation*}
In Proposition \ref{stopping_time} below, we show the probability mass function of $Y$ is
\begin{align} \label{eqn:pmf}
\mathbb{P} [Y=k] =& S(k, p, s) \ \mathbbm{1}_{\{s \leq k \leq s+t-1\}} \ + \nonumber \\
  & S(k, 1-p, t) \ \mathbbm{1}_{\{ t \leq k \leq s+t-1 \}}
\end{align}
where $\mathbbm{1}_{\{f\}}$ is the {\em indicator function}, taking the value 
of one if $f$ is true and zero otherwise, and
\begin{equation} \label{eqn:N}
S(k, p, s) = {k-1 \choose s-1} p^s (1-p)^{k-s} 
\end{equation}
is a translated version of the negative binomial probability mass.

The negative binomial cumulative distribution function can be expressed in terms of the
{\em regularized incomplete beta function} \citep{Olver2010} 
\begin{equation*}
\mathcal{I}_{1-p}(l - s + 1, s) = \sum_{k=s}^l S(k, p, s) 
\end{equation*}
for all integer values $l\,$ satisfying $s \leq l \leq s+t-1$. It has property \citep{Uppuluri1967}
\begin{equation} \label{ribf}
\mathcal{I}_{1-p}(l-s+1, s) = 1 - \mathcal{I}_p(s, l-s+1).
\end{equation}

\begin{prop} \label{stopping_time}
The distribution of the stopping time
\begin{equation*}
Y = \argminl_k \left[X_k \geq s \cup X_k \leq k-t \right]
\end{equation*}
is given at (\ref{eqn:pmf}).
\end{prop}
\begin{proof}

The endpoint $X_k = s$ can only occur if $X_{k-1} = s-1$ followed by a success. That is, 
\begin{align} \label{prop11}
\mathbb{P} [X_k =s ] \;&=\; p \, \mathbb{P} [X_{k-1} = s-1 ] \nonumber \\
 &=\; p\, {k-1 \choose s-1} p^{s-1} (1-p)^{k-s} \nonumber \\
 &=\; {k-1 \choose s-1} p^{s} (1-p)^{k-s}.
\end{align}
This expression is given
in (\ref{eqn:N}). 
Similarly, the probability a given realization reaches the endpoint $X_k = k-t$
satisfies 
\begin{align} \label{prop12}
\mathbb{P} [X_k = k-t ] \;&=\; (1-p) \, \mathbb{P} [X_{k-1} = k-t ] \nonumber \\
 &=\; (1-p)\, {k-1 \choose t-1} (1-p)^{t-1} p^{k-t} \nonumber \\
&=\; {k-1 \choose t-1} (1-p)^{t} p^{k-t}.
\end{align}
The result follows by summing (\ref{prop11}) and (\ref{prop12}).

To show (\ref{eqn:pmf}) sums to one, define
\begin{equation*} 
R = \sum_{k=s}^{s+t-1} S(k, p, s) + \sum_{k=t}^{s+t-1} S(k, 1-p, t).
\end{equation*}
Substitute $i=k-s$ in the first summation and $j=k-t$ in the second. Then
$R$ can be written as the cumulative distribution function (CDF) of two
negative binomial distributions:
\begin{align*}
R = \sum_{i=0}^{t-1} &{i+s-1 \choose i} p^s (1-p)^i +
\sum_{j=0}^{s-1}  {j+t-1 \choose j} p^j  (1-p)^t \\
   &= 1-\mathcal{I}_p(s, t) + 1 - \mathcal{I}_{1-p}(t, s) \\
   &= 1
\end{align*}
using (\ref{ribf}). This completes the proof (\ref{eqn:pmf}) is the distribution 
of the stopping time and is a valid probability mass function.
\end{proof}

We next consider an interim analysis of a clinical trial after $s'$ 
patients have responded to treatment 
and $t'$ failed to respond for $s' < s$ and $t' < t$.
\begin{corollary} \label{conditional_distribution}
The number of subsequent enrollments needed 
to reach either $s$ or $t$ endpoints behaves as SNB($p$, $s-s'$, $t-t'$).
\end{corollary}
Having observed $s'$ responders and $t'$ non-responders, there are $s-s'$ 
additional responders needed to reach the success endpoint and $t-t'$ additional 
non-responders needed to reach the failure endpoint.

\begin{figure*}
\begin{center}
\includegraphics[width=0.8\textwidth]{likelihood.pdf}
\end{center}
\caption{Shapes of the likelihood function for given values of $Y$ with 
$s=7$ and $t=11$.
\label{likelihood.fig}}
\end{figure*}

\section{The Moment Generating Function} \label{sect:mgf}

\begin{prop} Let $Y$ be distributed SNB with parameters $p$, $s$, and $t$.
Then the moment generating function (MGF) of $Y$ is
\begin{align} \label{eqn:mgf}
\mathbb{E}~e^{xY} = &\left(\frac{p e^x}{1 - qe^x}\right)^s 
  \mathcal{I}_{1-qe^x} (s, t) \ + \nonumber \\
  & \left(\frac{qe^x}{1-pe^x}\right)^t \mathcal{I}_{1-pe^x}(t, s)
\end{align}
for $q = 1-p$ and is defined for 
$x < \min \left\{\log(1/p), \log(1/q) \right\}$.
\end{prop}
The moment generating function for the SNB is calculated in a manner similar to 
that of two negative binomial distributions. Appendix 1 provides a proof for the derivation.

\begin{prop}
The MGF of the SNB converges to that of the negative binomial when either
$s$ or $t$ gets large. That is
\begin{equation*}
lim_{t\to\infty} \ \mathbb{E} e^{xY} = \left( \frac{pe^x}{1-qe^x} \right)^s
\end{equation*}
as. The analogous result holds when $s \rightarrow \infty$.
\end{prop}
\begin{proof}
The second incomplete beta function in (\ref{eqn:mgf}) can be written
in terms of a cumulative binomial distribution
\begin{equation*}
\mathcal{I}_{1-pe^x}(t, s) = \mathbb{P}\left[ B \leq s-1 \right]
\end{equation*}
where $B$ is distributed as
Binomial($t-k$, $pe^x$). From Chebychev's inequality %\cite{Hoeffding1963}
it follows that
\begin{equation} \label{eqn:hoeffding}
\mathbb{P}\left[ B \leq s-1 \right] \leq 
  \frac{ (t-k) pe^x (1-pe^x) }{ \left(s - (t-k)pe^x\right)^2 }
\end{equation}
As $t$ gets large $\mathcal{I}_{1-pe^x}(t, s)$ tends to zero
and $\mathcal{I}_{1-qe^x}(s, t)$ approaches 
one. The proof follows by realizing 
\begin{equation*}
0 < \frac{qe^x}{1-pe^x} < 1
\end{equation*}
over the support of $x$.
\end{proof}

\section{The Likelihood Function} \label{sect:likelihood}

Let $Y$ be distributed SNB($p, s, t$). Then, the likelihood of $Y$ is proportional to a mixture of Beta distributions.
\begin{equation*} \label{eqn:likelihood}
\mathcal{L} (p | s, t, Y) \propto B_1 \, \mathbbm{1}_{\{s \leq Y_1\}} + B_2 \, \mathbbm{1}_{\{t \leq Y_1 \}}
\end{equation*}
where $B_1 =$ Beta$\left(s, Y - s\right)$ and $B_2 =$ Beta$\left(Y - t, t\right)$ and
\begin{equation*}
\text{Beta}\left( \alpha, \beta \right) = \frac{\Gamma(\alpha+\beta)}{\Gamma(\alpha)\, \Gamma(\beta)} 
p^{\alpha-1} (1-p)^{\beta-1}.
\end{equation*}

\begin{prop} \label{mode_order}
The mode of $B_1$ occurs at a value of $p$ greater than that of $B_2$ in the likelihood function. 
\end{prop}
\begin{proof}
The mode of the Beta($\alpha,\, \beta$) distribution is $(\alpha-1) / (\alpha+\beta-2)$. Plugging in the shape parameters of $B_1$ and $B_2$ into the expression of the mode, the proposition is equivalent to showing
\begin{equation*}
\frac{s-1}{Y - 2} > \frac{Y - t- 1}{Y -2 }
\end{equation*}
which is true when $s > Y - t$. The maximum value of the right hand of the inequality occurs when $Y_1 = s+t-1$ and the inequality is equivalent to $s > s-1$, which is true.
\end{proof}
\begin{prop}
The difference between the modes of $B_1$ and $B_2$ is bounded by
\begin{equation*}
\textrm{MODE}(B_1) - \textrm{MODE}(B_2) 
\; \geq \; \frac{1}{s+t-3}.
\end{equation*}
\begin{proof}
Proposition \ref{mode_order} shows that the mode of $B_1$ is greater than that of $B_2$. The difference
between the two can be expressed as
\begin{equation*}
\frac{s-1}{Y-2} - \frac{Y-t-1}{Y-2} = \frac{s+t-Y}{Y-2}.
\end{equation*}
This function is strictly increasing in $Y$ over its support and obtains its minimum at $s+t-1$. The result follows.
\end{proof}
\end{prop}

As an example, the likelihood function for $Y = 7, 11, 13$ and 17 is shown in
Figure~\ref{likelihood.fig}. When $Y = 7$ the trial must have ended in success and the likelihood function concentrates near 1. The success and
failure endpoints can be reached for any value of $Y \geq 11$. When 
$Y = 11$ we see a bimodal likelihood function which one mode, at $p=0.6$
provided by the success endpoint and the other, at $p=0$, from the
failure endpoint where no responses are observed. Similarly, when $Y_1=13$
we see contributions from both the success and failure endpoints but
the two modes are converging. At $Y=17$ the endpoints contributed likelihoods with similar modes and the result is unimodal.

After an endpoint has been reached, the resulting conditional likelihood is either $B_1$ or $B_2$, depending if the trial was a success or failure. However, when the endpoint is not known, such as the planning phase of a trial, unintuitive situations may arise. Since the likelihood is bimodal, there are even settings where we may reject the null despite a poorer alternative. Suppose in the hypothetical trial $p_0=0.25$ and $p_1=0.5$, the outcome is unknown, and the trial completes after 11 enrollees, as shown in Figure~\ref{likelihood.fig} labeled $Y=11$. Since the null is at a ``trough'' in the likelihood we fail to reject even though it is closer to the most likely value of $p=0$.

\section{Connections and Approximations to Other Distributions}

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{shapes.pdf}
\end{center}
\caption{Different shapes of the SNB mass function with parameters 
($p$, $s$, $t$) given. Black areas indicate the mass contributed by reaching
$s$ responders before $t$ non-responders. Grey indicates
mass contributed by reaching $t$ non-responders first. \label{shapes.fig}}
\end{figure*}

Examples of different 
shapes of the SNB are shown in in Figure~\ref{shapes.fig} varying parameters $p$, 
$s$, and $t$. The SNB distribution is a generalization of the negative 
binomial. As a result, the SNB can approximate other distributions in the same
manner as the negative binomial. When $t$ is large then $Y-s$ has a 
negative binomial distribution with
\begin{equation*}                                    %   (1)   
\mathbb{P}[Y=s+j ]        \label{nb1.eq}          
  = {{s+j-1}\choose{s-1}} p^s (1-p)^j
\end{equation*}
for $\,j=0, 1,\ldots\,$. A similar statement can be made when $s$ is large
and $t$ is moderate. As a result, with proper parameter choice, the SNB
can mimic other probability distributions in a manner similar to 
those described in \cite{Best1974} and \cite{Peizer1968}. As a generalization of the negative binomial distribution the 
SNB inherits the ability to approximate other distributions. For example,
when $s=1$ and $t \rightarrow \infty$, the SNB($p$, $s$, $t$) converges
to an negative binomial distribution with index parameter $s$ and rate parameter $p$. When $s=1$, this is the geometric distribution. The connection between the negative binomial and the gamma distribution
are well-studied in the literature (see \cite{Best1974,Ord1968,Guenther1972} for examples) as well the connection to the Poisson \cite{Anscombe1950}.

The SNB generalizes both the minimum (riff-shuffle) and maximum negative
binomial distributions up to a translation of the support.
For the special case of $\,s=t,$ the SNB distribution is the
maximum negative binomial \cite{Johnson2005,Zelterman2005,Zhang2000} - 
the smallest number of outcomes necessary to 
observe at least $s$ responders {\em and} $s$ non-responders. This is equivalent to a 
translated version of the riff-shuffle or minimum negative binomial 
distribution~\citep{Johnson2005,Uppuluri1967}.

There is also an equivalence between the probability of reaching an endpoint 
in the SNB model and the tail probability of the binomial distribution.
Specifically, the probability the number of responders is at least 
$s$ in the binomial model is the same as the probability the treatment succeeds 
(reaches $s$) in the SNB model.
\begin{prop} \label{binomial_tail}
Let $Y$ be distributed as SNB($p$, $s$, $t$) and let $X_Y$ correspond
to the number of responders at the end of the trial. Let 
$B$ be distributed binomial with index parameter $n=s+t-1$ and response 
probability $p$. Then
\begin{equation}
\mathbb{P}[B \geq s] = \mathbb{P} [X_Y = s].
\end{equation}
\end{prop}
\begin{proof}
The binomial tail probability is
\begin{equation*}
\mathbb{P}[B \geq s] = 1 - \mathcal{I}_{1-p}(s, t)
\end{equation*}
The corresponding success probability is
\begin{equation} \label{eqn:success}
\mathbb{P} [X_Y = s] 
  = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s}.
\end{equation}
Let $i=k-s$. Since
\begin{equation*}
{i+s-1 \choose s-1} = {i+s-1 \choose i},
\end{equation*}
the summation in (\ref{eqn:success}) can be rewritten as
\begin{align*}
\mathbb{P} [X_Y = s] &= \sum_{i=0}^{t-1} {i+s-1 \choose i} p^s (1-p)^i\\
  &= 1 - \mathcal{I}_{1-p}(t, s)
\end{align*}
completing the proof.
\end{proof}

\begin{figure*}
\centering
\subfloat[SNB distribution]{\includegraphics[width=0.5\textwidth]{snb_density.pdf}}
\hfill
\subfloat[binomial distribution]{\includegraphics[width=0.5\textwidth]{bin_density.pdf}}
\caption{
SNB(0.2, 7, 11) with mass contributed from 
7 responders (black) or 11 non-responders (grey) along with 
Binomial(0.2, 17) with at least 2 responders (black) or fewer (grey).
}
\label{fig:snb_bin_compare}
\end{figure*}

To illustrate this result, let us return to our initial example
where $s=7$, $t=11$, and $p=0.2$.  The probability masses in
Figure~\ref{fig:snb_bin_compare} represented in 
black are equal in panels (a) and (b) as are the masses in grey.
The probability $s$
responders are reached in the SNB process is the same as the binomial 
probability of at least seven responders. Likewise, the probability $t$ 
non-responders reached in the SNB process is the same as the binomial
probability of zero through six responders.

\section{Trial Design with the SNB}

\begin{figure*}
\centering
\subfloat[ROC Curve]{\includegraphics[width=0.45\textwidth]{power-vs-significance.pdf}}
\hfill
\subfloat[Expected Sample Size]{\includegraphics[width=0.45\textwidth]{expected-sample-size.pdf}}
\caption{
The ROC and expected sample sizes of all trial designs where the maximum
number of patients is 17, $p=0.2$ under the null, and $p=0.4$ under
the alternative. The numerical values in the ROC curve indicate the number
of responses $s$ required to reach the success endpoint.
}
\label{fig:all_sample_sizes}
\end{figure*}

Consider the problem of designing a trial using curtailed sampling and the SNB. 
Assume that the maximum number of patients $n$ is given along with the null 
and alternative hypotheses. In this case, the parameter $s$ tells how many 
responses are needed to reach the success endpoint, can vary between 1 and 
$n-1$. For fixed $s$ and $n$  the value of $t$ is determined by the relation 
$t=n-s+1$. The ROC curves for all trial designs in the prototype, 
where $n=17$, $p=0.2$ under the null, and $p=0.4$ under the alternative, are 
shown in Figure~\ref{fig:all_sample_sizes} (a). In designing these trials, 
small significance and large power values are attained by either increasing 
the value of $n$ or increasing the difference between the alternative and null 
response rates. Figure~\ref{fig:all_sample_sizes} (b) shows the expected 
number of enrollees for each of these trials as a function of the number of 
responses required to reach the success endpoint.
The curve reaches its maximum of 15 patients when $s=5$. The prototype 
design ($p=0.2$, $s=7$, and $t=11$) has an expected sample size of 14 enrollees.

\section{Appendix: Proof of the MGF of the SNB}
Let $Y$ be distributed SNB with parameters $p$, $s$, and $t$.
Then the moment generating function (MGF) of $Y$ is
\begin{equation} 
\mathbb{E}~e^{xY} = \left(\frac{p e^x}{1 - qe^x}\right)^s 
  \mathcal{I}_{1-qe^x} (s, t) + \left(\frac{qe^x}{1-pe^x}\right)^t 
  \mathcal{I}_{1-pe^x}(t, s)
\end{equation}
for $q = 1-p$ and is defined for 
$x < \min \left\{\log(1/p), \log(1/q) \right\}$.
\begin{proof}
The MGF of the SNB is, by definition:
\begin{equation*}
\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s q^{k-s} e^{kx} 
  + \sum_{k=t}^{s+t-1} {k-1 \choose t-1} p^{k-t} q^t e^{kx}
\end{equation*}
and can be rewritten as:
\begin{align} \label{eqn:first_sum}
\mathbb{E}~e^{xY} =& 
  \sum_{k=s}^{s+t-1}{k-1 \choose s-1} (pe^x)^{s} (qe^x)^{k-s} \ + \nonumber \\
  & \sum_{k=t}^{s+t-1}{k-1 \choose t-1} (qe^x)^t (pe^x)^{k-t}.
\end{align}
The first summation in (\ref{eqn:first_sum}) satisfies
\begin{align*}
\sum_{k=s}^{s+t-1} & {k-1 \choose s-1} (pe)^{sx} (qe^x)^{k-s} \\ 
  &= \left(\frac{pe^x}{1 - qe^x}\right)^s \ \ \sum_{k=s}^{s+t-1} {k-1 \choose s-1} 
    (qe^x)^{k-s} (1-qe^x)^s \\
  &= \left(\frac{pe^x}{1 - qe^x}\right)^s \mathcal{I}_{1-qe^x}(s, t).
\end{align*}
Since the $p$ parameter in $\mathcal{I}_p$ has domain zero 
to one, we have $0 \leq pe^x < 1$. This implies $x < -\log(p)$.
A similar expression can be derived for the second summation in 
(\ref{eqn:first_sum}) and results in
the constraint $x < -\log(1-p)$.
\end{proof}

\section*{Acknowledgements}
This research was supported by grants R01CA131301, 
R01CA157749, R01CA148996, R01CA168733, and PC50CA196530 awarded by the 
National Cancer Institute along with support from the Yale Comprehensive Cancer 
Center and the Yale Center for Outcomes Research. We would also like 
to thank Rick Landin at Ignyta Inc. for his suggestions.

\bibliographystyle{acm}
\bibliography{mybibfile}

\end{document}

### 2017-08-10 stat-and-prob-letters/s.r

@@ -157,7 +157,7 @@ ggsave("bayesian-sample-variance.pdf")
 
 dsnb_plot(p, s, t) + ylab("Probability") + theme(text = element_text(size=15)) +
   theme_bw()
-ggsave("snb-first-plot.pdf")
+ggsave("../sii/snb-first-plot.pdf", width=8, height=5)
 
 ps = seq(0, 1, by=0.01)
 moments = foreach(p=ps, .combine=rbind) %do% {
@@ -441,10 +441,15 @@ lls = foreach(y=Y, .combine=cbind) %do% {
 colnames(lls) = Y
 lldf = as.data.frame(cbind(p, lls))
 llldf = gather(lldf, Y, "Likelihood", -p)
-p = ggplot(llldf, aes(x=p, y=Likelihood, group=Y, color=Y)) + 
-  geom_line(size=2) +
-  xlab(expression(italic(p))) + ylab("Likelihood") + 
-  labs(color=expression(Y[1])) + theme_bw() # +  scale_color_grey()
+p = ggplot(llldf, aes(x=p, y=Likelihood, group=Y)) + 
+  geom_line(size=1.5) +
+  xlab(expression(italic(p))) + ylab("Likelihood") +
+  geom_text(data=NULL, x=0.08, y=0.75, label="Y=11") +
+  geom_text(data=NULL, x=0.92, y=0.75, label="Y=7") +
+  geom_text(data=NULL, x=0.2, y=0.27, label="Y=13") +
+  geom_text(data=NULL, x=0.45, y=0.21, label="Y=17") +
+  theme_bw() + scale_fill_grey()
+  # labs(color=expression(Y[1])) + theme_bw() # +  scale_color_grey()
 
 ggsave("likelihood.pdf", p, width=11, height=7)
 
@@ -561,23 +566,27 @@ designs_long = gather(designs, `Design Feature`, value, Significance:Power)
 
 ggplot(designs_long, aes(x=s, y=value, color=`Design Feature`)) + 
   geom_line() + ylab("Probability") + 
-  xlab("Number of Responses to Stop the Trial") + theme_bw() +
+  xlab("Number of Responses to Stop the Trial (s)") + theme_bw() +
   scale_x_continuous(breaks=seq_len(n-1))
 
 ggsave("all-hypothetical-designs.pdf")
 
 ggplot(designs, aes(x=s, y=ess)) +
   geom_line() + ylab("Expected Sample Size Under the Null") +
[…]

### 2018-02-12 sii/deveaux-kane-zelterman.tex
SII accepted the paper.

@@
 \begin{aug}
-\author{\fnms{Michelle} \snm{DeVeaux} \ead[label=e1]{michelle.deveaux@yale.edu}}
-\address{Department of Biostatistics\\ School of Epidemiology and Public Health \\ Yale University, New Haven, CT \\ \printead{e1}}
+\author{\fnms{Michelle} \snm{DeVeaux} \ead[label=e1]{michelle.deveaux@regeneron.com}}
+\address{Regeneron Pharmaceuticals Inc.\\ Tarrytown, NY\\ \printead{e1}}
 \and
@@
 We introduce a discrete distribution suggested by curtailed
-sampling rules common in early-stage clinical trials unifying the
-negative, minimum negative (riff-shuffle), and
-maximum negative binomial distributions. We derive the
+sampling rules common in early-stage clinical trials. We derive the
 distribution of the smallest number of independent and identically
@@
 In the present example, observing seven
-patients responding to treatment ends the trial and so the number of 
-enrollees required could be as small as seven. Similarly 11
-observed treatment non-responses would also end the trial. This sampling mechanism, in
+successful patients ends the trial and so the number of enrollees required
+could be as small as seven. Similarly 11
+observed treatment failures would also end the trial. This sampling mechanism, in
 which the experiment ends as soon as either predefined endpoint is reached, is
@@
 sample size for this trial is seven through 17.
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.8\textwidth]{snb-first-plot.pdf}
-\caption{
-The distribution of the sample size in a trial that stops after either seven patients
-respond to treatment or 11 do not when $p=0.2$.
-}
-\label{fig:kane_viz_1}
-\end{figure*}
-
-\begin{figure*}[htbp!]
-\centering
-\includegraphics[width=\textwidth]{mean-and-variance.pdf}
-\caption{
-The mean and variance of the sample size for the probability of success $p$ 
-when the trial stops after $s=7$ patients respond or $t=11$ fail to respond.
-}
-\label{fig:kane_viz_2}
-\end{figure*}
 
@@
 number of enrollees approaches seven from above.
-When $p=0$ or $1$ then the sequence of resposes is deterministic and the 
-variance is zero.
+
+When $p=0$ or $1$ then the processes are deterministic and variance is zero.
 Values of $p$ between zero and one change the mixing proportions of 
@@
 Section 4 describes the likelihood function.
-Section 5 relates the distribution to other standard
+Section 5 describes the posterior and predictive probability distributions. Section 6 relates the distribution to other standard
 distributions and connects the SNB tail probability to the binomial tail 
-probability. Section 6 considers single arm, early-phase trial design with the 
-SNB.
+probability. Section 7 shows how to design a trial using the SNB in a prototypical setting.
 
@@
 distribution of $Y$ is the smallest integer such that either
-$\sum_i^Y b_i = s$ or $\sum_i^Y 1-b_i = t$.
+$\sum_i^Y b_i = s$ or $\sum_i^Y (1-b_i) = t$.
 
 The sequence of Bernoulli random variables can be modeled as the process 
-$\mathbf{X} = \left\{X(k) : k = 0,1,... \right\}$
-with $X(0)=0$ and
+$\mathbf{X} = \left\{X_k : k = 0,1,... \right\}$
+with $X_0=0$ and
 \begin{equation*} \label{eqn:proc}
@@
 \end{equation*}
-At each step a patient's outcome is measured. The process continues until either 
-$X_k = s$ or $X_k = k-t$ corresponding to the success and failure boundaries.
-
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{KanePlot.pdf}
-\caption{
-A hypothetical realization of the trial.
-}
-\label{fig:kane_viz}
-\end{figure*}
+where $\mathbbm{1}_{\{f\}}$ is the {\em indicator function}, taking the value 
+of one if $f$ is true and zero otherwise, and
+At each step a patient's outcome is measured. If it is a response, the response count increases, otherwise it stays the same. The process continues until either 
+$X_k = s$ or $X_k = k-t$ corresponding to the success and failure boundaries
 
@@
 
-We next derive the mass function of $Y$. The distribution of $\,Y\,$ has support on integer values in the range
+We next derive the probability mass function of $Y$. The distribution of $\,Y\,$ has support on integer values in the range
 \begin{equation*}               
@@
 \begin{align} \label{eqn:pmf}
-\mathbb{P} [Y=k] =& S(k, p, s) \ \mathbbm{1}_{\{s \leq k \leq s+t-1\}} \ + \nonumber \\
-  & S(k, 1-p, t) \ \mathbbm{1}_{\{ t \leq k \leq s+t-1 \}}
+\mathbb{P} [Y=k] =& \ S(k, p, s) \ \mathbbm{1}_{\{s \leq k \leq s+t-1\}} \nonumber \\
+  & + S(k, 1-p, t) \ \mathbbm{1}_{\{ t \leq k \leq s+t-1 \}}
 \end{align}
-where $\mathbbm{1}_{\{f\}}$ is the {\em indicator function}, taking the value 
-of one if $f$ is true and zero otherwise, and
+where
 \begin{equation} \label{eqn:N}
@@
 \end{align}
-The result follows by summing (\ref{prop11}) and (\ref{prop12}).
+The result (\ref{eqn:pmf}) follows by summing (\ref{prop11}) and (\ref{prop12}).
 
@@
 \end{align*}
-using (\ref{ribf}). This completes the proof (\ref{eqn:pmf}) is the distribution 
-of the stopping time and is a valid probability mass function.
+using (\ref{ribf}). This completes the proof that (\ref{eqn:pmf}) is the distribution 
+of the stopping time and it is a valid probability mass function.
 \end{proof}
@@
 
-\begin{figure*}
-\begin{center}
-\includegraphics[width=0.8\textwidth]{likelihood.pdf}
-\end{center}
-\caption{Shapes of the likelihood function for given values of $Y$ with 
-$s=7$ and $t=11$.
-\label{likelihood.fig}}
-\end{figure*}
-
 \section{The Moment Generating Function} \label{sect:mgf}
@@
 Then the moment generating function (MGF) of $Y$ is
-\begin{align} \label{eqn:mgf}
-\mathbb{E}~e^{xY} = &\left(\frac{p e^x}{1 - qe^x}\right)^s 
-  \mathcal{I}_{1-qe^x} (s, t) \ + \nonumber \\
-  & \left(\frac{qe^x}{1-pe^x}\right)^t \mathcal{I}_{1-pe^x}(t, s)
-\end{align}
+\begin{equation} \label{eqn:mgf}
+\mathbb{E}~e^{xY} = \left(\frac{p e^x}{1 - qe^x}\right)^s 
+  \mathcal{I}_{1-qe^x} (s, t) + \left(\frac{qe^x}{1-pe^x}\right)^t 
+  \mathcal{I}_{1-pe^x}(t, s)
+\end{equation}
 for $q = 1-p$ and is defined for 
@@
 
+In this section we derive the likelihood function for the SNB for a single trial.
+In early-stage clinical trial only a single trial is performed, usually because of resource constraints, and the object of interest is $p$ parameter, which determines if a therapy will be delivered to the market.
+A multi-sample version is less relevant for this use case but is represented as a product of mixtures of Beta distributions.
+Deriving it's theoretical characteristics not straight-forward.
+
 Let $Y$ be distributed SNB($p, s, t$). Then, the likelihood of $Y$ is proportional to a mixture of Beta distributions.
 \begin{equation*} \label{eqn:likelihood}
-\mathcal{L} (p | s, t, Y) \propto B_1 \, \mathbbm{1}_{\{s \leq Y_1\}} + B_2 \, \mathbbm{1}_{\{t \leq Y_1 \}}
+\mathcal{L} (p | s, t, Y) = B_1 \, \mathbbm{1}_{\{s \leq Y_1\}} + B_2 \, \mathbbm{1}_{\{t \leq Y_1 \}}
 \end{equation*}
@@
 
+\section{The Posterior and Predictive Probability Distribution} \label{bayesian_extension}
+
+Consider the Bayesian setting where $Y$ is an SNB($P$, $s$, $t$) distribution and the rate parameter, $P$ 
+is distributed as Beta($\alpha$, $\beta$). The posterior distribution $P|Y$ is proportional to the 
+likelihood, given by the function
+%f_P(p, \alpha, \beta ) f_{Y|P}(p, k, s, t) &= 
+\begin{align} \label{eqn:ptl}
+f_{P|Y}(p, k, s, t, \alpha, \beta) \; \propto \ &
+  \ \frac{ {k-1 \choose s-1} }{B(\alpha, \beta)} p^{\alpha +s -1} 
+    (1-p)^{k+\beta-s-1} \\
+  & + \frac{ {k-1 \choose t-1} }{B(\alpha, \beta)} p^{k+\alpha -t -1} 
+    (1-p)^{\beta+t-1} \nonumber
+\end{align}
+where $0 \leq p \leq 1$ and $\min(s, t) \leq k \leq s+k-1$.
+
+The predictive distribution of the SNB can be found as by integrating
+$p$ over the interval zero to one and applying the definition of the 
+beta function.
+\begin{align} \label{eqn:predictive}
+f_{Y}(&k, s, t, \alpha, \beta) = 
+  \int_0^1 f_P(p | \alpha, \beta )  f_{Y|P}(p, k, s, t) dp \\ 
+  = & {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} 
+     \nonumber \\
+  & + {k-1 \choose t-1} 
+    \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} \nonumber
+\end{align}
+If both $\alpha$ and $\beta$ are non-negative integers then the predictive
+distribution is a mixture of negative-hypergeometric distributions. 
+\begin{align*} \label{eqn:hypergeo}
+f_{Y}(k, s, t, \alpha, \beta) = & 
+  \frac{ {k - 1 \choose s - 1 } 
+         {\alpha + \beta \choose \alpha} }{
+         {\alpha + \beta + k - 1 \choose \alpha + s}}
+  \frac{\alpha}{\alpha + \beta}
+  \frac{ \beta }{ k-s+\beta } \\ & +
+  \frac{ {k - 1 \choose t - 1} 
+         {\alpha + \beta \choose \beta} 
+         }{
+         {\alpha + \beta + k -1 \choose t + \beta}
+       }
+  \frac{ \beta }{ \alpha + \beta}
+  \frac{ \alpha }{ k-t + \alpha}
+\end{align*}
+The ratio of combinations in the first term can be interpreted as the
+probability of $s-1$ responders from $k-1$ patients in $\alpha + s$ draws
+from a population size of $\alpha + \beta + k - 1$. This value is multiplied
+by $\alpha / (\alpha + \beta)$, the expected response rate of the prior. 
+The final term in the product weights the prior based on the number
+of non-responders ($k-s$). Terms in the second summand are interpreted similarly
+for non-responders.
+
+The ratio of (\ref{eqn:ptl}) divided by (\ref{eqn:predictive}) gives the 
+posterior distribution of $P$. It is a mixture of beta distributions. The
+mixing parameter depend on the endpoints ($s$ and $t$), the number of enrollees needed to reach an endpoint ($k$), and the prior parameters ($\alpha$ and
+$\beta$).
+
 \section{Connections and Approximations to Other Distributions}
-
-\begin{figure*}
-\begin{center}
-\includegraphics[width=\textwidth]{shapes.pdf}
-\end{center}
-\caption{Different shapes of the SNB mass function with parameters 
-($p$, $s$, $t$) given. Black areas indicate the mass contributed by reaching
-$s$ responders before $t$ non-responders. Grey indicates
-mass contributed by reaching $t$ non-responders first. \label{shapes.fig}}
-\end{figure*}
 
@@
 
-\begin{figure*}
-\centering
-\subfloat[SNB distribution]{\includegraphics[width=0.5\textwidth]{snb_density.pdf}}
-\hfill
-\subfloat[binomial distribution]{\includegraphics[width=0.5\textwidth]{bin_density.pdf}}
-\caption{
-SNB(0.2, 7, 11) with mass contributed from 
-7 responders (black) or 11 non-responders (grey) along with 
-Binomial(0.2, 17) with at least 2 responders (black) or fewer (grey).
-}
-\label{fig:snb_bin_compare}
-\end{figure*}
-
 To illustrate this result, let us return to our initial example
@@
 
-\begin{figure*}
-\centering
-\subfloat[ROC Curve]{\includegraphics[width=0.45\textwidth]{power-vs-significance.pdf}}
-\hfill
-\subfloat[Expected Sample Size]{\includegraphics[width=0.45\textwidth]{expected-sample-size.pdf}}
-\caption{
-The ROC and expected sample sizes of all trial designs where the maximum
-number of patients is 17, $p=0.2$ under the null, and $p=0.4$ under
-the alternative. The numerical values in the ROC curve indicate the number
-of responses $s$ required to reach the success endpoint.
-}
-\label{fig:all_sample_sizes}
-\end{figure*}
+Consider the problem of designing a trial using curtailed sampling and the SNB.
+Assume that the maximum number of patients $n$ is given along with the null and alternative hypotheses.
+In this case, the parameter $s$ tells how many responses are needed to reach the success endpoint, can vary between 1 and $n-1$.
+For fixed $s$ and $n$  the value of $t$ is determined by the relation $t=n-s+1$.
+The ROC curves for all trial designs in the prototype, where $n=17$, $p=0.2$ under the null, and $p=0.4$ under the alternative, are shown in Figure~\ref{fig:all_sample_sizes_power_vs_signif}.
+In designing these trials, small significance and large power values are attained by either increasing the value of $n$ or increasing the difference between the alternative and null response rates.
+Figure~\ref{fig:all_sample_sizes} shows the expected number of enrollees for each of these trials as a function of the number of responses required to reach the success endpoint.
 
-Consider the problem of designing a trial using curtailed sampling and the SNB. 
-Assume that the maximum number of patients $n$ is given along with the null 
-and alternative hypotheses. In this case, the parameter $s$ tells how many 
-responses are needed to reach the success endpoint, can vary between 1 and 
-$n-1$. For fixed $s$ and $n$  the value of $t$ is determined by the relation 
-$t=n-s+1$. The ROC curves for all trial designs in the prototype, 
-where $n=17$, $p=0.2$ under the null, and $p=0.4$ under the alternative, are 
-shown in Figure~\ref{fig:all_sample_sizes} (a). In designing these trials, 
-small significance and large power values are attained by either increasing 
-the value of $n$ or increasing the difference between the alternative and null 
-response rates. Figure~\ref{fig:all_sample_sizes} (b) shows the expected 
-number of enrollees for each of these trials as a function of the number of 
-responses required to reach the success endpoint.
-The curve reaches its maximum of 15 patients when $s=5$. The prototype 
-design ($p=0.2$, $s=7$, and $t=11$) has an expected sample size of 14 enrollees.
+The expected number of enrollees for each trial is shown in Figure~\ref{fig:all_sample_sizes}.
+The curve reaches its maximum of 15 patients when $s=5$.
+The prototype design ($p=0.2$, $s=7$, and $t=11$) has an expected sample size of 14 enrollees.
 
-\section{Appendix: Proof of the MGF of the SNB}
-Let $Y$ be distributed SNB with parameters $p$, $s$, and $t$.
-Then the moment generating function (MGF) of $Y$ is
-\begin{equation} 
-\mathbb{E}~e^{xY} = \left(\frac{p e^x}{1 - qe^x}\right)^s 
-  \mathcal{I}_{1-qe^x} (s, t) + \left(\frac{qe^x}{1-pe^x}\right)^t 
-  \mathcal{I}_{1-pe^x}(t, s)
-\end{equation}
-for $q = 1-p$ and is defined for 
-$x < \min \left\{\log(1/p), \log(1/q) \right\}$.
-\begin{proof}
+\section*{Acknowledgements}
+This research was supported by grants R01CA131301, 
+R01CA157749, R01CA148996, R01CA168733, and PC50CA196530 awarded by the 
+National Cancer Institute along with support from the Yale Comprehensive Cancer 
+Center and the Yale Center for Outcomes Research. We would also like 
+to thank Rick Landin at LaJolla Pharmaceutical for his suggestions.
+
+\section*{APPENDIX: PROOF OF PROPOSITION 2}
+%Let $Y$ be distributed SNB with parameters $p$, $s$, and $t$.
+%Then the moment generating function (MGF) of $Y$ is
+%\begin{equation} 
+%\mathbb{E}~e^{xY} = \left(\frac{p e^x}{1 - qe^x}\right)^s 
+%  \mathcal{I}_{1-qe^x} (s, t) + \left(\frac{qe^x}{1-pe^x}\right)^t 
+%  \mathcal{I}_{1-pe^x}(t, s)
+%\end{equation}
+%for $q = 1-p$ and is defined for 
+%$x < \min \left\{\log(1/p), \log(1/q) \right\}$.
+%\begin{proof}
 The MGF of the SNB is, by definition:
-\begin{equation*}
-\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s q^{k-s} e^{kx} 
-  + \sum_{k=t}^{s+t-1} {k-1 \choose t-1} p^{k-t} q^t e^{kx}
-\end{equation*}
+\begin{align*}
+\mathbb{E}~e^{xY} = & \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s q^{k-s} e^{kx} \\
+  & + \sum_{k=t}^{s+t-1} {k-1 \choose t-1} p^{k-t} q^t e^{kx}
+\end{align*}
 and can be rewritten as:
 \begin{align} \label{eqn:first_sum}
-\mathbb{E}~e^{xY} =& 
-  \sum_{k=s}^{s+t-1}{k-1 \choose s-1} (pe^x)^{s} (qe^x)^{k-s} \ + \nonumber \\
-  & \sum_{k=t}^{s+t-1}{k-1 \choose t-1} (qe^x)^t (pe^x)^{k-t}.
+\mathbb{E}~e^{xY} = & \sum_{k=s}^{s+t-1}{k-1 \choose s-1} (pe^x)^{s} (qe^x)^{k-s} \nonumber \\
+  & + \sum_{k=t}^{s+t-1}{k-1 \choose t-1} (qe^x)^t (pe^x)^{k-t}.
 \end{align}
@@
 \begin{align*}
-\sum_{k=s}^{s+t-1} & {k-1 \choose s-1} (pe)^{sx} (qe^x)^{k-s} \\ 
-  &= \left(\frac{pe^x}{1 - qe^x}\right)^s \ \ \sum_{k=s}^{s+t-1} {k-1 \choose s-1} 
+\sum_{k=s}^{s+t-1} &{k-1 \choose s-1}  (pe)^{sx} (qe^x)^{k-s}  \\
+  & = \left(\frac{pe^x}{1 - qe^x}\right)^s \ \ \sum_{k=s}^{s+t-1} {k-1 \choose s-1} 
     (qe^x)^{k-s} (1-qe^x)^s \\
-  &= \left(\frac{pe^x}{1 - qe^x}\right)^s \mathcal{I}_{1-qe^x}(s, t).
+  & = \left(\frac{pe^x}{1 - qe^x}\right)^s \mathcal{I}_{1-qe^x}(s, t).
 \end{align*}
@@
 the constraint $x < -\log(1-p)$.
-\end{proof}
-
-\section*{Acknowledgements}
-This research was supported by grants R01CA131301, 
-R01CA157749, R01CA148996, R01CA168733, and PC50CA196530 awarded by the 
-National Cancer Institute along with support from the Yale Comprehensive Cancer 
-Center and the Yale Center for Outcomes Research. We would also like 
-to thank Rick Landin at Ignyta Inc. for his suggestions.
+%\end{proof}
 
@@
 
+\pagebreak
+
+\onecolumn
+
+\section*{Figures}
+
+\begin{figure}[htbp!]
+\centering
+\includegraphics[width=\textwidth]{snb-first-plot.pdf}
+\caption{
+The distribution of the sample size in a trial that stops after either seven patients
+respond to treatment or 11 do not when $p=0.2$.
+}
+\label{fig:kane_viz_1}
+\end{figure}
+
+\pagebreak
+
+\begin{figure}[htbp!]
+\centering
+\includegraphics[width=\textwidth]{mean-and-variance.pdf}
+\caption{
+The mean and variance of the sample size for the probability of success $p$ 
+when the trial stops after $s=7$ patients respond or $t=11$ fail to respond.
+}
+\label{fig:kane_viz_2}
+\end{figure}
+
+\pagebreak
+
+\begin{figure}[htbp!]
+\centering
+\includegraphics[width=\textwidth]{KanePlot.pdf}
+\caption{
+A hypothetical realization of the trial.
+}
+\label{fig:kane_viz}
+\end{figure}
+
+\pagebreak
+
+\begin{figure}[htbp!]
+\begin{center}
+\includegraphics[width=\textwidth]{likelihood.pdf}
+\end{center}
+\caption{Shapes of the likelihood function for given values of $Y$ with 
+$s=7$ and $t=11$.
+\label{likelihood.fig}}
+\end{figure}
+
+\pagebreak
+
+\begin{figure}[htbp!]
+\begin{center}
+\includegraphics[width=0.8\textwidth]{shapes.pdf}
+\end{center}
+\caption{Different shapes of the SNB mass function with parameters 
+($p$, $s$, $t$) given. Black areas indicate the mass contributed by reaching
+$s$ responders before $t$ non-responders. Grey indicates
+mass contributed by reaching $t$ non-responders first. \label{shapes.fig}}
+\end{figure}
+
+\pagebreak
+
+\begin{figure}[htbp!]
+\centering
+\subfloat[SNB distribution]{\includegraphics[width=0.5\textwidth]{snb_density.pdf}}
+\hfill
+\subfloat[binomial distribution]{\includegraphics[width=0.5\textwidth]{bin_density.pdf}}
+\caption{
+SNB(0.2, 7, 11) with mass contributed from 
+7 responders (black) or 11 non-responders (grey) along with 
+Binomial(0.2, 17) with at least 2 responders (black) or fewer (grey).
+}
+\label{fig:snb_bin_compare}
+\end{figure}
+
+\pagebreak
+
+\begin{figure}[htbp!]
+\centering
+\includegraphics[width=\textwidth]{power-vs-significance.pdf}
+\caption{
+ROC curve of the all trial designs where the maximum number of patients is 17, $p=0.2$ under the null and $p=0.4$ under the alternative.
+The numerical values indicate the number of responses $s$ required reach the success endpoint.
+}
+\label{fig:all_sample_sizes_power_vs_signif}
+\end{figure}
+
+\pagebreak
+
+\begin{figure}[htbp!]
+\centering
+\includegraphics[width=\textwidth]{expected-sample-size.pdf}
+\caption{
+The expected sample size all trial designs where the maximum number
+of patients is 17 and $p=0.2$.}
+\label{fig:all_sample_sizes}
+\end{figure}
+
+\pagebreak
+
 \end{document}

### 2018-02-23 sii/deveaux-kane-zelterman.tex
Next version... Hopefully this goes to print.

@@
 
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{snb-first-plot.pdf}
+\caption{
+The distribution of the sample size in a trial that stops after either seven patients
+respond to treatment or 11 do not when $p=0.2$.
+}
+\label{fig:kane_viz_1}
+\end{figure}
+
 The distribution of the number of trial enrollments is shown in 
@@
 number of enrollees approaches seven from above.
+
+\begin{figure*}
+\centering
+\includegraphics[width=0.7\textwidth]{mean-and-variance.pdf}
+\caption{
+The mean and variance of the sample size for the probability of success $p$ 
+when the trial stops after $s=7$ patients respond or $t=11$ fail to respond.
+}
+\label{fig:kane_viz_2}
+\end{figure*}
 
@@
 of one if $f$ is true and zero otherwise, and
-At each step a patient's outcome is measured. If it is a response, the response count increases, otherwise it stays the same. The process continues until either 
+At each step a patient's outcome is measured. If it is a response, then
+the response count increases, otherwise it stays the same. The process 
+continues until either 
 $X_k = s$ or $X_k = k-t$ corresponding to the success and failure boundaries
+
+\begin{figure*}%[htbp!]
+\centering
+\includegraphics[width=0.7\textwidth]{KanePlot.pdf}
+\caption{
+A hypothetical realization of the trial.
+}
+\label{fig:kane_viz}
+\end{figure*}
 
@@
 
+\begin{figure*}%[htbp!]
+\centering
+\includegraphics[width=0.7\textwidth]{likelihood.pdf}
+\caption{Shapes of the likelihood function for given values of $Y$ with 
+$s=7$ and $t=11$.}
+\label{likelihood.fig}
+\end{figure*}
+
 After an endpoint has been reached, the resulting conditional likelihood is either $B_1$ or $B_2$, depending if the trial was a success or failure.
@@
 
+\begin{figure*}
+\centering
+\includegraphics[width=0.7\textwidth]{shapes.pdf}
+\caption{Different shapes of the SNB mass function with parameters 
+($p$, $s$, $t$) given. Black areas indicate the mass contributed by reaching
+$s$ responders before $t$ non-responders. Grey indicates
+mass contributed by reaching $t$ non-responders first.}
+\label{fig:shapes}
+\end{figure*}
+
 Examples of different 
-shapes of the SNB are shown in in Figure~\ref{shapes.fig} varying parameters $p$, 
+shapes of the SNB are shown in in Figure~\ref{fig:shapes} varying parameters $p$, 
 $s$, and $t$. The SNB distribution is a generalization of the negative 
@@
 
+\begin{figure}[htbp!]
+\centering
+\subfloat[SNB distribution]{\includegraphics[width=0.5\textwidth]{snb_density.pdf}}
+\hfill
+\subfloat[binomial distribution]{\includegraphics[width=0.5\textwidth]{bin_density.pdf}}
+\caption{
+SNB(0.2, 7, 11) with mass contributed from 
+7 responders (black) or 11 non-responders (grey) along with 
+Binomial(0.2, 17) with at least 2 responders (black) or fewer (grey).
+}
+\label{fig:snb_bin_compare}
+\end{figure}
+
 To illustrate this result, let us return to our initial example
@@
 
+\begin{figure}[htbp!]
+\centering
+\includegraphics[width=0.4\textwidth]{power-vs-significance.pdf}
+\caption{
+ROC curve of the all trial designs where the maximum number of patients is 17, $p=0.2$ under the null and $p=0.4$ under the alternative.
+The numerical values indicate the number of responses $s$ required reach the success endpoint.
+}
+\label{fig:all_sample_sizes_power_vs_signif}
+\end{figure}
+
 The expected number of enrollees for each trial is shown in Figure~\ref{fig:all_sample_sizes}.
@@
 The prototype design ($p=0.2$, $s=7$, and $t=11$) has an expected sample size of 14 enrollees.
+
+\begin{figure}[htbp!]
+\centering
+\includegraphics[width=0.4\textwidth]{expected-sample-size.pdf}
+\caption{
+The expected sample size all trial designs where the maximum number
+of patients is 17 and $p=0.2$.}
+\label{fig:all_sample_sizes}
+\end{figure}
 
@@
 
-\pagebreak
-
-\onecolumn
-
-\section*{Figures}
-
-\begin{figure}[htbp!]
-\centering
-\includegraphics[width=\textwidth]{snb-first-plot.pdf}
-\caption{
-The distribution of the sample size in a trial that stops after either seven patients
-respond to treatment or 11 do not when $p=0.2$.
-}
-\label{fig:kane_viz_1}
-\end{figure}
-
-\pagebreak
-
-\begin{figure}[htbp!]
-\centering
-\includegraphics[width=\textwidth]{mean-and-variance.pdf}
-\caption{
-The mean and variance of the sample size for the probability of success $p$ 
-when the trial stops after $s=7$ patients respond or $t=11$ fail to respond.
-}
-\label{fig:kane_viz_2}
-\end{figure}
-
-\pagebreak
-
-\begin{figure}[htbp!]
-\centering
-\includegraphics[width=\textwidth]{KanePlot.pdf}
-\caption{
-A hypothetical realization of the trial.
-}
-\label{fig:kane_viz}
-\end{figure}
-
-\pagebreak
-
-\begin{figure}[htbp!]
-\begin{center}
-\includegraphics[width=\textwidth]{likelihood.pdf}
-\end{center}
-\caption{Shapes of the likelihood function for given values of $Y$ with 
-$s=7$ and $t=11$.
-\label{likelihood.fig}}
-\end{figure}
-
-\pagebreak
-
-\begin{figure}[htbp!]
-\begin{center}
-\includegraphics[width=0.8\textwidth]{shapes.pdf}
-\end{center}
-\caption{Different shapes of the SNB mass function with parameters 
-($p$, $s$, $t$) given. Black areas indicate the mass contributed by reaching
-$s$ responders before $t$ non-responders. Grey indicates
-mass contributed by reaching $t$ non-responders first. \label{shapes.fig}}
-\end{figure}
-
-\pagebreak
-
-\begin{figure}[htbp!]
-\centering
-\subfloat[SNB distribution]{\includegraphics[width=0.5\textwidth]{snb_density.pdf}}
-\hfill
-\subfloat[binomial distribution]{\includegraphics[width=0.5\textwidth]{bin_density.pdf}}
-\caption{
-SNB(0.2, 7, 11) with mass contributed from 
-7 responders (black) or 11 non-responders (grey) along with 
-Binomial(0.2, 17) with at least 2 responders (black) or fewer (grey).
-}
-\label{fig:snb_bin_compare}
-\end{figure}
-
-\pagebreak
-
-\begin{figure}[htbp!]
-\centering
-\includegraphics[width=\textwidth]{power-vs-significance.pdf}
-\caption{
-ROC curve of the all trial designs where the maximum number of patients is 17, $p=0.2$ under the null and $p=0.4$ under the alternative.
-The numerical values indicate the number of responses $s$ required reach the success endpoint.
-}
-\label{fig:all_sample_sizes_power_vs_signif}
-\end{figure}
-
-\pagebreak
-
-\begin{figure}[htbp!]
-\centering
-\includegraphics[width=\textwidth]{expected-sample-size.pdf}
-\caption{
-The expected sample size all trial designs where the maximum number
-of patients is 17 and $p=0.2$.}
-\label{fig:all_sample_sizes}
-\end{figure}
-
-\pagebreak
-
 \end{document}

### 2018-02-23 stat-and-prob-letters/s.r

@@ -72,8 +72,8 @@ n = s+t-1
 plot.new()
 d = as.data.frame(dsnb_stacked(min(s,t):(t+s-1), p = p, s = s, t = t))
 names(d)[2:3] = c("success", "failure")
-d = gather(d, x)
-names(d)[2] = "Outcome"
+d = gather(d, Outcome, value, -x)
+#names(d)[2] = "Outcome"
 d$Outcome = factor(d$Outcome)
 d$Outcome = relevel(d$Outcome, "failure")
 ggplot(data = d, aes(x = factor(x), y = value, fill = Outcome)) + 
@@ -157,7 +157,7 @@ ggsave("bayesian-sample-variance.pdf")
 
 dsnb_plot(p, s, t) + ylab("Probability") + theme(text = element_text(size=15)) +
   theme_bw()
-ggsave("../sii/snb-first-plot.pdf", width=8, height=5)
+ggsave("../sii/snb-first-plot.pdf")
 
 ps = seq(0, 1, by=0.01)
 moments = foreach(p=ps, .combine=rbind) %do% {
@@ -176,14 +176,16 @@ p = ggplot(m, aes(x=p, y=var)) + geom_line() + ylab("Distribution Variance") +
   theme(text = element_text(size=15))
 ggsave("dist-var.pdf", p)
 
-y = gather(m, p, value)
+y = gather(m, type, value, -p)
 names(y) = c("p", "type", "value")
 y$type[y$type=='exp'] = "Mean"
 y$type[y$type=='var'] = "Variance"
 p = ggplot(y, aes(x=p, y=value)) + geom_line() + 
   facet_grid(type ~ ., scales="free") + ylab("") +
   theme(text = element_text(size=15)) + theme_bw()
-ggsave("mean-and-variance.pdf")
+ggsave("../sii/mean-and-variance.pdf", width=8, height=5)
+
+stop("here")
 
 # Assume the uniform prior.
 pp = function(x, k, s, t) {
@@ -199,8 +201,8 @@ pp = function(x, k, s, t) {
 }
 
 x = pp(seq(0, 1, by=0.01), 15, s, t)
-x = gather(x, p)
-names(x) = c("p", "Outcome", "value")
+x = gather(x, Outcome, value, -p)
+#names(x) = c("p", "Outcome", "value")
 x$Outcome = relevel(factor(x$Outcome), "failure")
[…]
